### 2011-01-26 Documentation/ubercal.tex
euclid: start documentation
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17294 03a3cea6-2c03-0410-ac9b-9271b1c66c29

\documentclass[12pt,a4paper,twoside]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{color}
\usepackage{listings}

\author{David Hogg, Rory Holmes}
\title{Ubercalibration Simulations \\(Euclid)}
\date{Jan 2011}

\makeatletter
\def\maketitle{%
  \null
  \thispagestyle{empty}%
  \vfill
  \begin{center}\leavevmode
    \normalfont
    {\LARGE \@title\par}%
    \vskip 1cm
    {\Large \@author\par}%
    \vskip 1cm
    {\Large \@date\par}%
  \end{center}%
  \vfill
  \null
  \cleardoublepage
  }
\makeatother

\begin{document}
\maketitle

\section{Introduction}
These simulations...

\section{General}

\subsection{Parameter File}

Simulation parameters are stored in the parameter.py file.

\subsection{General Functions}
The functions.py file contains two routine for converting between magnitudes and fluxes. ..

\section{Sky Catalog}
The

\section{Camera Model}

\subsection{Sky to Focal Plane Coordinate Transformations}

\subsection{Uncertainty Inverse-Variance}

\subsection{Catalog to Measured Flux Conversion}

\section{Ubercalibration}
The ubercalibration procedure works in two sub-steps: (i) the star fluxes are refined based on the latest flat-field estimate (\textit{s\_step}) and (ii) the flat-field parameters are refined based on the latest star flux estimates (\textit{q\_step}). We iterate these steps until the parameters converge. In our notation $j$ refers to the iteration step.

Our flat-field is defined by a set of parameters $\vec{q}$. After the $j$th iteration, the flat-field parameters will be indexed as $\vec{q_j}$. The individual star measurements are defined as $c_i$ (``counts''), where each $i$ corresponds to a exposure number $n$, a specific star with ID number $k$ and a focal plane position $\vec{x}$.

XX Our model is

\begin{equation}
c_i = f(\vec{x_i} | \vec{q_j}) . s_{kj} + e_{ij}
\end{equation}

where $s_k$ is the true flux from star $k$ and our error is drawn from the Normal Distribution $e_{ij} = N(e|0,{\sigma_i}^2)$.

\subsection{Star Flux Refinement: \textbf{\textit{s\_step}}}
@j, for each star

\begin{equation}
\chi^2 = \sum_{i touching k} \frac{(c_i-f_{ij}s_{kj})^2}{{\sigma_i}^2}
\end{equation}

\begin{equation}
\frac{d\chi^2}{d s_{kj}} = \sum_{i touching k} \frac{-2 f_{ij} (c_i-f_{ij}s_{kj+1})}{{\sigma_i}^2} = 0
\end{equation}

\begin{equation}
\Rightarrow \sum_{i touching k} \frac{f_{ij} c_i}{{\sigma_i}^2}= \sum_{i touching k} \frac{f_{ij}^2 s_{kj+1}}{{\sigma_i}^2}
\end{equation}

\begin{equation}
\Rightarrow s_{kj+1} = \frac{\sum_{i touching k} \frac{f_{ij} c_i}{{\sigma_i}^2}}/{\sum_{i touching k} \frac{f_{ij}^2 s_{kj+1}}{{\sigma_i}^2}}
\end{equation}

The 
The measurements of each star are refined based on

On the $j$th iteration, the following proce

\subsection{Flat-Field Refinement: \textbf{\textit{p\_step}}}

\end{document}

### 2011-01-26 camera.py
euclid: test change
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17295 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -1,6 +1,6 @@
 #!/usr/bin/env python
 # -*- coding: utf-8 -*-
-
+#
 import numpy as np
 import math
 import matplotlib.pylab as plt

### 2011-01-26 camera.py
euclid: test change
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17296 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -1,6 +1,6 @@
 #!/usr/bin/env python
 # -*- coding: utf-8 -*-
-#
+
 import numpy as np
 import math
 import matplotlib.pylab as plt

### 2011-01-26 Documentation/ubercal.tex
euclid: documentation update
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17297 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \subsection{Star Flux Refinement: \textbf{\textit{s\_step}}}
-@j, for each star
+The flux estimates, $s_k$, for each star can be considered individually.
+At step $j$ we can refine a star's flux estimate (to $s_{kj+1}$) by minimizing the error function with the latest flat-field parameters:
 
 \begin{equation}
-\chi^2 = \sum_{i touching k} \frac{(c_i-f_{ij}s_{kj})^2}{{\sigma_i}^2}
+\chi^2_{k} = \sum_{i \in \mathcal{O}(k)} \frac{(c_i-f_{ij}s_{kj})^2}{{\sigma_i}^2}
 \end{equation}
@@
 \begin{equation}
-\frac{d\chi^2}{d s_{kj}} = \sum_{i touching k} \frac{-2 f_{ij} (c_i-f_{ij}s_{kj+1})}{{\sigma_i}^2} = 0
+\frac{d\chi^2_{k}}{d s_{kj}} = \sum_{i \in \mathcal{O}(k)} \frac{-2 f_{ij} (c_i-f_{ij}s_{kj+1})}{{\sigma_i}^2} = 0
 \end{equation}
@@
 \begin{equation}
-\Rightarrow \sum_{i touching k} \frac{f_{ij} c_i}{{\sigma_i}^2}= \sum_{i touching k} \frac{f_{ij}^2 s_{kj+1}}{{\sigma_i}^2}
+\Rightarrow \sum_{i \in \mathcal{O}(k)} \frac{f_{ij} c_i}{{\sigma_i}^2}= \sum_{i \in \mathcal{O}(k)} \frac{f_{ij}^2 s_{kj+1}}{{\sigma_i}^2}
 \end{equation}
@@
 \begin{equation}
-\Rightarrow s_{kj+1} = \frac{\sum_{i touching k} \frac{f_{ij} c_i}{{\sigma_i}^2}}/{\sum_{i touching k} \frac{f_{ij}^2 s_{kj+1}}{{\sigma_i}^2}}
+\Rightarrow s_{kj+1} = {\sum_{i \in \mathcal{O}(k)} \frac{f_{ij} c_i}{{\sigma_i}^2}}/{\sum_{i \in \mathcal{O}(k)} \frac{f_{ij}^2}{{\sigma_i}^2}}
 \end{equation}
 
-The 
-The measurements of each star are refined based on
+\subsection{Flat-Field Refinement: \textbf{\textit{p\_step}}}
+The flat-field parameters can now be refined with the latest star flux estimates. Minimise error function with respect to the...
 
-On the $j$th iteration, the following proce
+Consider all measurements
 
-\subsection{Flat-Field Refinement: \textbf{\textit{p\_step}}}
+\begin{equation}
+\chi^2 = \sum_{all~i} \frac{(c_i-f_{ij}s_{kj})^2}{{\sigma_i}^2}
+\end{equation}
+
+\begin{equation}
+f_{ij} = \sum_{l = 1}^L q_{lj} g_i(\vec{x_i})
+\end{equation}
+
+\begin{equation}
+\chi^2 = \sum_{all~i} \frac{(c_i- s_{kj} \sum_{l = 1}^L q_{lj} g_i(\vec{x_i}))^2}{{\sigma_i}^2}
+\end{equation}
+
+\begin{equation}
+\frac{d\chi^2}{dq_{lj}} = \sum_{all~i} \frac{-2 g_l(\vec{x_i}) s_{kj+1} (c_i- s_{kj} \sum_{l = 1}^L q_{lj} g_i(\vec{x_i}))}{{\sigma_i}^2} = 0
+\end{equation}
+
+\begin{equation}
+\sum_{all~i} \frac{g_l(\vec{x_i}) s_{kj+1} c_i}{{\sigma_i}^2} = \sum_{all~i} \frac{g_l(\vec{x_i}) s_{kj+1}^2 \sum_{l = 1}^L q_{lj+1} g_l(\vec{x_i})}{{\sigma_i}^2}
+\end{equation}
 

### 2011-01-26 functions.py
euclid: documentation update
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17301 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -14,3 +14,8 @@ def flux_uncertainty_variance(mag, eta):
   sky_error= 0.1*mag2flux(parameters.mag_at_ten_sigma())
   return (np.power(sky_error,2) + (eta**2)*np.power(mag2flux(mag),2))
   
+def evaluate_flat_field(focal_position, q, plots, verbose):
+  return 1
+
+def normalize_flat_field(q):
+  return q

### 2011-01-27 functions.py
euclid: program update
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17310 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -15,7 +15,32 @@ def flux_uncertainty_variance(mag, eta):
   return (np.power(sky_error,2) + (eta**2)*np.power(mag2flux(mag),2))
   
 def evaluate_flat_field(focal_position, q, plots, verbose):
-  return 1
+  # Function returns flat field at focal plane position specified. The function also calculates the order of the flat field equation from the number of flat field parameters
+  
+  # Calculate required order
+  order = int(np.around(np.sqrt(0.25+2*len(q))-1.5))
+  
+  # Calculate flat-field
+  l = 0
+  total = 0
+  for n in range(order+1):
+    for m in range(n+1):
+      total += q[l] * (focal_position[0]**(n-m)) * (focal_position[0]**m)
+      l += 1 
+  
+  # Normalize the flat field
+  q = normalize_flat_field(q)
+  
+  return total # Return flat-field at specified [x,y] focal plane position
 
 def normalize_flat_field(q):
   return q
+
+def s_step(obs_cat, q, plots, verbose):
+  # First step in ubercalibration. This function calculates the new flux estimates.
+  # obs_cat = observation_catalog = [pointing ID, star ID, observed_flux, observed_invvar, focal plane x, focal plane y]
+  
+  # Calculate flat-field based on source positions and current flat field parameters 
+  
+  
+  return 1

### 2011-01-27 master.py

@@ -21,8 +21,8 @@ import survey
 os.system('rm ./Figures/Camera_Images/*.png')
 os.system('rm ./Figures/*.png')
 
-plots = '' #None
-verbose = '' #None
+plots = None
+verbose = None
 
 #*************************************************************
 #******************* Generate Sky Catalog ********************
@@ -45,8 +45,16 @@ sky_catalog = god.create_catalog(M, m_min, m_max, powerlaw, sky, limits, seed, p
 #************************ Survey Sky *************************
 #*************************************************************
 
-data = survey.survey(sky_catalog, "A.txt", plots=plots, verbose=verbose) 
+observation_catalog = survey.survey(sky_catalog, "A.txt", plots=plots, verbose=verbose) 
+# observed_catalog = [pointing ID, star ID, observed_flux, observed_invvar, focal plane x, focal plane y]
 
+#*************************************************************
+#********************* Ubercalibration ***********************
+#*************************************************************
+
+p = [1,0,0,0,0]
+
+test = functions.s_step(observation_catalog,p,plots=plots, verbose=verbose)
 
 #*************************************************************
 #*********************** Health Checks ***********************

### 2011-01-27 survey.py

@@ -23,5 +23,23 @@ def survey(catalog, survey_file, plots, verbose):
     # data_dic[pointing ID] = [star ID, observed_flux, observed_invvar, focal_position]
   if verbose != None: print "...done!"
   
-  return data_dic
-  # data_dic[pointing ID] = [star ID, observed_flux, observed_invvar, focal_position]
+  # Rearrange observations into an observation_catalog
+  if verbose != None: print "Rearranging observation catalog"
+  array_size = 0
+  for i in range(0,len(data_dic)):
+    array_size = array_size + len(data_dic[i][:,0])
+  
+  if verbose != None: print "Observation catalog requires %d rows)" % array_size
+  
+  # Copying dictionary into array
+  observation_catalog = np.zeros((array_size,6))
+  count = 0
+  for i in range(0,len(data_dic)):
+    single_exposure = data_dic[i]
+    for j in range(0,len(single_exposure[:,0])):
+      observation_catalog[count,0] = i
+      observation_catalog[count,1:6] = single_exposure[j,:]
+      count = count +1
+      
+  return observation_catalog
+  # observed_catalog = [pointing ID, star ID, observed_flux, observed_invvar, focal plane x, focal plane y]

### 2011-01-27 Documentation/ubercal.tex
euclid: documentation update
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17300 03a3cea6-2c03-0410-ac9b-9271b1c66c29
euclid: documentation update
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17301 03a3cea6-2c03-0410-ac9b-9271b1c66c29
euclid: documentation
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17311 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \usepackage{listings}
+\usepackage{graphicx}
+\usepackage{upgreek}
 
@@
 \section{Sky Catalog}
-The
+We create a fake sky, based on realistic star densities, between a magnitude of 19 to 22.5.
+The upper and lower limits are defined based the saturation and 10$\upsigma$ limits for a single Euclid exposure.
+Star's are generated with random coordinates (within the sky region being investigated) and with a random magnitude according to the following powerlaw:
+\begin{equation}
+\frac{1}{N} \frac{dN}{dm} = A + B m
+\end{equation}
+
+The \textit{Transformation Method} is used generate random numbers with the appropriate probability density function.
+
+\begin{equation}
+\log_{10} \frac{\Delta N}{\Delta m} = A + B m
+\end{equation}
+
+\begin{equation}
+\frac{d N}{d m} =10^ {A + B m}
+\end{equation}
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{source_mag_histogram.png}
+\end{center}
+\caption{A histogram of the generated star magnitudes. The number of stars is given by... XX \label{fig:source_magnitude}}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{gods_sky.png}
+\end{center}
+\caption{A small subset of the simulated sky catalog produced. \label{fig:gods_sky}}
+\end{figure}
+
+The random number generate is given the same seed, so that the same simulated sky is produced each time the routine is called.
 
@@
 \subsection{Uncertainty Inverse-Variance}
-
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{flux_uncertainty_variance.png}
+\end{center}
+\caption{XX UPDATE \label{fig:flux_uncertainty}}
+\end{figure}
 \subsection{Catalog to Measured Flux Conversion}
@@
 \subsection{Flat-Field Refinement: \textbf{\textit{p\_step}}}
-The flat-field parameters can now be refined with the latest star flux estimates. Minimise error function with respect to the...
-
-Consider all measurements
+The flat-field parameters can now be refined with the latest star flux estimates. To do this we minimize the error function for all measurements with respect to the flat-field parameters.
 
 \begin{equation}
-\chi^2 = \sum_{all~i} \frac{(c_i-f_{ij}s_{kj})^2}{{\sigma_i}^2}
+\chi^2 = \sum_{i} \frac{(c_i-f_{ij}s_{kj+1})^2}{{\sigma_i}^2}
 \end{equation}
@@
 \begin{equation}
-f_{ij} = \sum_{l = 1}^L q_{lj} g_i(\vec{x_i})
+f_{ij} = \sum_{l = 1}^L q_{lj} g_l(\vec{x_i})
 \end{equation}
@@
 \begin{equation}
-\chi^2 = \sum_{all~i} \frac{(c_i- s_{kj} \sum_{l = 1}^L q_{lj} g_i(\vec{x_i}))^2}{{\sigma_i}^2}
+\chi^2 = \sum_{i} \frac{(c_i- s_{kj+1} \sum_{l = 1}^L q_{lj} g_l(\vec{x_i}))^2}{{\sigma_i}^2}
 \end{equation}
@@
 \begin{equation}
-\frac{d\chi^2}{dq_{lj}} = \sum_{all~i} \frac{-2 g_l(\vec{x_i}) s_{kj+1} (c_i- s_{kj} \sum_{l = 1}^L q_{lj} g_i(\vec{x_i}))}{{\sigma_i}^2} = 0
+\frac{d\chi^2}{dq_{lj}} = \sum_{i} \frac{-2 g_l(\vec{x_i}) s_{kj+1} (c_i- s_{kj+1} \sum_{l' = 1}^{L'} q_{l'j} g_{l'}(\vec{x_i}))}{{\sigma_i}^2} = 0
 \end{equation}
@@
 \begin{equation}
-\sum_{all~i} \frac{g_l(\vec{x_i}) s_{kj+1} c_i}{{\sigma_i}^2} = \sum_{all~i} \frac{g_l(\vec{x_i}) s_{kj+1}^2 \sum_{l = 1}^L q_{lj+1} g_l(\vec{x_i})}{{\sigma_i}^2}
+\sum_{i} \frac{g_l(\vec{x_i}) s_{kj+1} c_i}{{\sigma_i}^2} = \sum_{i} \frac{g_l(\vec{x_i}) s_{kj+1}^2 \sum_{l' = 1}^{L'} q_{l'j+1} g_{l'}(\vec{x_i})}{{\sigma_i}^2}
 \end{equation}

### 2011-01-27 camera.py
euclid: add s_step
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17317 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -9,7 +9,7 @@ import parameters
 import functions
 import god
 
-def camera(catalog, pointing, orientation, plots, verbose):
+def camera(catalog, pointing, orientation, plots=None, verbose=None):
   # Measure the sources within the camera's FoV
   # catalog = [Star ID, magnitude, α, β]
   # pointing = [α, β] # telescope's pointing "sky" coordinates

### 2011-01-27 functions.py

@@ -14,33 +14,47 @@ def flux_uncertainty_variance(mag, eta):
   sky_error= 0.1*mag2flux(parameters.mag_at_ten_sigma())
   return (np.power(sky_error,2) + (eta**2)*np.power(mag2flux(mag),2))
   
-def evaluate_flat_field(focal_position, q, plots, verbose):
+def evaluate_flat_field(x, y, q):
   # Function returns flat field at focal plane position specified. The function also calculates the order of the flat field equation from the number of flat field parameters
   
   # Calculate required order
   order = int(np.around(np.sqrt(0.25+2*len(q))-1.5))
-  
+  assert(len(q) == ((order + 1) * (order + 2) / 2)) # assert to break code if condition not met
+ 
   # Calculate flat-field
   l = 0
   total = 0
   for n in range(order+1):
     for m in range(n+1):
-      total += q[l] * (focal_position[0]**(n-m)) * (focal_position[0]**m)
+      total += q[l] * (x**(n-m)) * (y**m)
       l += 1 
-  
-  # Normalize the flat field
-  q = normalize_flat_field(q)
   
   return total # Return flat-field at specified [x,y] focal plane position
 
 def normalize_flat_field(q):
   return q
 
-def s_step(obs_cat, q, plots, verbose):
+def s_step(obs_cat, q):
   # First step in ubercalibration. This function calculates the new flux estimates.
   # obs_cat = observation_catalog = [pointing ID, star ID, observed_flux, observed_invvar, focal plane x, focal plane y]
   
   # Calculate flat-field based on source positions and current flat field parameters 
   
+  ff = evaluate_flat_field(obs_cat[:,4], obs_cat[:,5], q)
+  fcss = ff * obs_cat[:,2] * obs_cat[:,3]  
+  ffss = ff * ff * obs_cat[:,3]
+  star_ID = np.around(obs_cat[:,1]).astype('int')
+  max_star = np.max(star_ID)
+  s = np.zeros(max_star)
+  s_invvar = np.zeros(max_star)
+  for ID in range(max_star):
+    indx = (star_ID == ID)
+    denominator = np.sum(ffss[indx])
+    s_invvar[ID] = denominator
+    if denominator > 0.:
+      s[ID] = np.sum(fcss[indx])/denominator
+  return (s, s_invvar)
   
-  return 1
+#def q_step(obs_cat, s, order):
+  
+  

### 2011-01-27 god.py

@@ -22,7 +22,7 @@ def generate_magnitudes(m_min,m_max,powerlaw,size):
   return m
   
   
-def create_catalog(M, m_min, m_max, powerlaw, sky, limits, seed, plots, verbose):
+def create_catalog(M, m_min, m_max, powerlaw, sky, limits, seed, plots=None, verbose=None):
   # M = density of stars 
   # m_min = minimum magnitude (i.e. saturation limit)
   # m_max = maximum magnitude limit per dither (i.e. sensitivity limit 10σ)

### 2011-01-27 master.py

@@ -52,9 +52,13 @@ observation_catalog = survey.survey(sky_catalog, "A.txt", plots=plots, verbose=v
 #********************* Ubercalibration ***********************
 #*************************************************************
 
-p = [1,0,0,0,0]
+q = [1,0,0,0,0,0]
 
-test = functions.s_step(observation_catalog,p,plots=plots, verbose=verbose)
+s, s_invvar = functions.s_step(observation_catalog,q)
+
+order = 2
+
+#q, q_invvar = functions.q_star(observation_catalog, s, order)
 
 #*************************************************************
 #*********************** Health Checks ***********************

### 2011-01-27 parameters.py

@@ -15,7 +15,7 @@ def m_max():
 
 def density_of_stars():
   # density of stars (all magnitudes)
-  return 1000
+  return 100
 
 def powerlaw():
   #B in log10(dN/dm) = A + B*m

### 2011-01-27 survey.py

@@ -3,7 +3,7 @@
 import numpy as np
 import camera
 
-def survey(catalog, survey_file, plots, verbose):
+def survey(catalog, survey_file, plots=None, verbose=None):
   # Routing loads the survey from file and then calls the camera function accordingly.
   
   if verbose != None: print "Loading survey..."

### 2011-01-27 Documentation/ubercal.tex
euclid: add s_step
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17317 03a3cea6-2c03-0410-ac9b-9271b1c66c29
euclid: doc update, include ubercal s_step
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17321 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \usepackage{upgreek}
+\usepackage{subfigure}
 
@@
 
-\section{General}
-
-\subsection{Parameter File}
-
-Simulation parameters are stored in the parameter.py file.
-
-\subsection{General Functions}
-The functions.py file contains two routine for converting between magnitudes and fluxes. ..
-
 \section{Sky Catalog}
@@
 \begin{equation}
-\frac{1}{N} \frac{dN}{dm} = A + B m
+\frac{1}{N} \frac{dN}{dm} = a + b m
 \end{equation}
 
-The \textit{Transformation Method} is used generate random numbers with the appropriate probability density function.
+The following equation is used to generate random star magnitudes with the appropriate probability density distribution function (see Appendix for derivation):
 
 \begin{equation}
-\log_{10} \frac{\Delta N}{\Delta m} = A + B m
+m = \frac{1}{b} \log{(\left[ 10^{b m_{max}} - 10^{b m_{min}} \right] r + 10^{b m_{min}})}
 \end{equation}
 
+\noindent{}where r is a random number in the range [0.0, 1.0). See Fig. \ref{fig:catalog}.
+
+\begin{figure}[ht]
+\centering
+\subfigure[A histogram of the generated stars' magnitudes.]{
+\includegraphics[width=0.45\textwidth]{source_mag_histogram.png}
+\label{fig:source_magnitude}
+}
+\subfigure[A small subset of the simulated sky catalog shown on the fake sky.]{
+\includegraphics[width=0.45\textwidth]{gods_sky.png}
+\label{fig:gods_sky}
+}
+\label{fig:catalog}
+\end{figure}
+
+The random number generator is given the same seed, so that the same simulated sky is produced each time the routine is called.
+
+\section{Single Exposure}
+With a given pointing and orientation, we transform the sky catalog into focal plane coordinates and find all the stars falling within the field-of-view.
+We do not simulate images, instead we trivially convert the actual star fluxes into measured fluxes.
+
+\subsection{Uncertainty Variance}
+We assume that the Euclid exposures will be background limited and that we will hit an upper limit on the signal-to-noise for bright sources. We model the uncertainty on the flux as:
+
 \begin{equation}
-\frac{d N}{d m} =10^ {A + B m}
+\sigma_{\text{f}}^{2} = \alpha^{2} + \eta^{2} f^{2} 
 \end{equation}
 
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=\textwidth]{source_mag_histogram.png}
-\end{center}
-\caption{A histogram of the generated star magnitudes. The number of stars is given by... XX \label{fig:source_magnitude}}
-\end{figure}
+\noindent{}where $\alpha$ and $\eta$ are both constants (see Fig. \ref{fig:flux_uncertainty}).
 
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=\textwidth]{gods_sky.png}
-\end{center}
-\caption{A small subset of the simulated sky catalog produced. \label{fig:gods_sky}}
-\end{figure}
-
-The random number generate is given the same seed, so that the same simulated sky is produced each time the routine is called.
-
-\section{Camera Model}
-
-\subsection{Sky to Focal Plane Coordinate Transformations}
-
-\subsection{Uncertainty Inverse-Variance}
 \begin{figure}[ht]
@@
 \end{center}
-\caption{XX UPDATE \label{fig:flux_uncertainty}}
+\caption{The flux uncertainty used as a function of star magnitude.\label{fig:flux_uncertainty}}
 \end{figure}
 \subsection{Catalog to Measured Flux Conversion}
+The measured flux from a star depends on its focal plane position ($\vec{x_i}$), God's flat-field (with parameters $\vec{q}$) and the star's true flux ($s$).
+
+\begin{equation}
+c_i = f(\vec{x_i} | \vec{q}) . s_{k} + e_{i}
+\end{equation}
+
+\noindent{}An error, drawn from the Normal Distribution $e_{ij} = N(e|0,{\sigma_i}^2)$, is also introduced to the measurement.
 
@@
 \begin{equation}
-\Rightarrow s_{kj+1} = {\sum_{i \in \mathcal{O}(k)} \frac{f_{ij} c_i}{{\sigma_i}^2}}/{\sum_{i \in \mathcal{O}(k)} \frac{f_{ij}^2}{{\sigma_i}^2}}
+\Rightarrow s_{kj+1} = {\sum_{i \in \mathcal{O}(k)} \frac{f_{ij} c_i}{{\sigma_i}^2}}/{\sum_{i \in \mathcal{O}(k)} \frac{f_{ij}^2}{{\sigma_i}^2}} \label{eqn:s_step}
 \end{equation}
+
+The standard uncertainty inverse variance on $s_{kj+1}$ is given by the denominator in Eqn \ref{eqn:s_step}.
 
@@
 
+\appendix
+\section{Random Magnitude Generation}
+XX DAVID: I have included this so that I remember how the formula was derived...
+
+\noindent{}The Transformation Method is used generate random numbers with the appropriate probability density function.
+
+\begin{equation}
+\frac{1}{N} \frac{dN}{dm} = a + b m
+\end{equation}
+
+\noindent{}The probability a star is generated with a magnitude $m$ in the range [$m_0$,$m_1$) is therefore:
+
+\begin{equation}
+p(m) = \frac{\dfrac{d N}{d m}}{\int_{m_{0}}^{m_{1}}\dfrac{d N}{d m} dm} = A \frac{d N}{d m}
+\end{equation}
+
+\noindent{}Using the Transformation Method, we find:
+
+\begin{eqnarray}
+r (m) & = & \int_{m_{0}}^{m} p(m') dm'\\
+& = & A \int_{m_{0}}^{m} 10^{a+b m'} dm'\\
+& = & B \int_{m_{0}}^{m} e^{b m' \ln{10}} dm'\\
+& = & \frac{B}{b \ln{10}} \left[ 10^{b m} - 10^{b m_{0}} \right] \label{eqn:Bover}
+\end{eqnarray}
+
+\noindent{}Since the probability of generating a magnitude between $m_0$ and $m_1$ is 1:
+
+\begin{equation}
+B = \frac{b \ln{10}}{\left[ 10^{b m_{1}} - 10^{b m_{0}} \right]} \label{eqn:B}
+\end{equation}
+
+\noindent{}Substituting Eqn.
+\ref{eqn:B} into Eqn.
+\ref{eqn:Bover} and rearranging we find that stars can be produced according to the appropriate probability density distribution function with the equation:
+
+\begin{equation}
+m = \frac{1}{b} \log{(\left[ 10^{b m_{max}} - 10^{b m_{min}} \right] r + 10^{b m_{min}})}
+\end{equation}
+\noindent{}where r is a random number in the range [0.0, 1.0).
+
 \end{document}

### 2011-01-27 god.py

@@ -58,24 +58,38 @@ def create_catalog(M, m_min, m_max, powerlaw, sky, limits, seed, plots=None, ver
     
   if plots != None:
     # Plot portion of sky
+    fontsize = 25
     if verbose != None: print "Plotting portion of sky..."
     plt.figure()
     plt.plot(catalog[0:1000,2],catalog[0:1000,3],'ko',markersize=2)
-    plt.xlabel(ur'$\alpha$', fontsize=20)
-    plt.ylabel(ur'$\beta$', fontsize=20)
-    plt.title("God's Sky")
-    plt.savefig("Figures/gods_sky.png")
+    plt.xlabel(ur'$\alpha$', fontsize=fontsize)
+    plt.ylabel(ur'$\beta$', fontsize=fontsize)
+    plt.title("God's Sky", fontsize=fontsize)
+    ax = plt.gca()
+    for tick in ax.xaxis.get_major_ticks():
+      tick.label1.set_fontsize(fontsize)
+    for tick in ax.yaxis.get_major_ticks():
+      tick.label1.set_fontsize(fontsize)
+    plt.savefig("Figures/gods_sky.png",bbox_inches='tight',pad_inches=0.)
     if verbose != None: print "...done!"
     
     # Histogram of source magnitude
     plt.figure()
+    fontsize = 25
     bin=np.arange(m_min,m_max,0.05)
     if verbose != None: print "Plotting histogram of source magnitude..."
     plt.hist(catalog[:,1],bins=bin, log=True)
-    plt.title("%i Sources in Sample" % len(catalog[:,0]))
-    plt.xlabel("Source Magnitude")
-    plt.ylabel("log(N)")
-    plt.savefig("Figures/source_mag_histogram.png")
+    plt.title("%i Sources in Sample" % len(catalog[:,0]), fontsize=fontsize)
+    plt.xlabel("Source Magnitude", fontsize=fontsize)
+    plt.ylabel("log(N)", fontsize=fontsize)
+    ax = plt.gca()
+    for tick in ax.xaxis.get_major_ticks():
+      tick.label1.set_fontsize(fontsize)
+    for tick in ax.yaxis.get_major_ticks():
+      tick.label1.set_fontsize(fontsize)
+
+    
+    plt.savefig("Figures/source_mag_histogram.png",bbox_inches='tight',pad_inches=0.)
     if verbose != None: print "...done!"
   
   return catalog

### 2011-01-27 master.py

@@ -54,7 +54,7 @@ observation_catalog = survey.survey(sky_catalog, "A.txt", plots=plots, verbose=v
 
 q = [1,0,0,0,0,0]
 
-s, s_invvar = functions.s_step(observation_catalog,q)
+#s, s_invvar = functions.s_step(observation_catalog,q)
 
 order = 2
 

### 2011-01-27 parameters.py

@@ -15,7 +15,7 @@ def m_max():
 
 def density_of_stars():
   # density of stars (all magnitudes)
-  return 100
+  return 1000
 
 def powerlaw():
   #B in log10(dN/dm) = A + B*m

### 2011-01-31 Documentation/ubercal.tex
euclid: document update
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17373 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \usepackage{subfigure}
+\usepackage{verbatim}
 
@@
 
-\section{Introduction}
-These simulations...
-
 \section{Sky Catalog}
@@
 The upper and lower limits are defined based the saturation and 10$\upsigma$ limits for a single Euclid exposure.
-Star's are generated with random coordinates (within the sky region being investigated) and with a random magnitude according to the following powerlaw:
+Stars are generated with random coordinates (within the sky region being investigated) and with a random magnitude according to the following powerlaw:
 \begin{equation}
@@
 }
-\subfigure[A small subset of the simulated sky catalog shown on the fake sky.]{
+\subfigure[A small subset of the simulated stars in the fake sky.]{
 \includegraphics[width=0.45\textwidth]{gods_sky.png}
@@
 
-The random number generator is given the same seed, so that the same simulated sky is produced each time the routine is called.
+The random number generator is given the same seed, so that the same simulated sky is produced each time the simulation is run.
 
 \section{Single Exposure}
-With a given pointing and orientation, we transform the sky catalog into focal plane coordinates and find all the stars falling within the field-of-view.
+With a given pointing and orientation, we transform the sky catalog into focal plane coordinates and find all the stars falling within the instrument's field-of-view.
+The default field-of-view size is the same as the Euclid NISP instrument's.
 We do not simulate images, instead we trivially convert the actual star fluxes into measured fluxes.
@@
 \end{figure}
+
 \subsection{Catalog to Measured Flux Conversion}
@@
 \noindent{}An error, drawn from the Normal Distribution $e_{ij} = N(e|0,{\sigma_i}^2)$, is also introduced to the measurement.
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{camera.png}
+\end{center}
+\caption{A single exposure of the sky with the pointing coordinates and the field orientation given in the title.
+Left: Plot of the entire fake sky, with the instrument's field-of-view shown in black and the measured sources highlighted in red.
+Right: The distribution of the highlighted sources on the instrument's focal plane.
+\label{fig:camera}}
+\end{figure}
 
@@
 
-XX Our model is
+Our model is
 
@@
 
+\section{Initial Results}
+\subsection{Surveys}
+To test the code we considered two extreme survey examples.
+The first, called Survey A, consists of a regular grid of pointings over the sky with a 2.5\% overlap between neighboring fields.
+Nine passes are performed over the sky and each time the grids are aligned exactly.
+The second survey, Survey D, has the same number of fields as Survey A, but the pointing and orientation of these fields are random.
+These two surveys can be seen in Fig.
+\ref{fig:surveys}.
+
+\begin{figure}[ht]
+\centering
+\subfigure{
+\includegraphics[width=0.47\textwidth]{A.pdf}
+}
+\subfigure{
+\includegraphics[width=0.47\textwidth]{D.pdf}
+}
+\caption{Left: The A Survey has regular pointings over the sky with a 2.5\% overlap between adjacent fields.
+Nine passes are performed over the sky and each time the grids are aligned exactly.
+Right: The D Survey has the same number of pointings as the A Survey, but the pointings and orientations are random.\label{fig:surveys}}
+\end{figure}
+
+\subsection{Ubercalibration Performance}
+
+The difference in ubercalibration method's ability to fit the flat field and the star magnitudes for the two different surveys is striking.
+With the random survey (D) we can get very close to the actual flat field in a small number of ubercalibration iterations (see Fig \ref{fig:D}).
+The well ordered survey (A), on the other hand, does not converge to the correct flat field even with 1,000 ubercal iterations.
+Even when we start with almost the correct flat-field parameters (those found with the D survey - shown in Fig.
+\ref{fig:D}), the ubercalibration procedure does not converge.
+In fact, with each iteration solution drifts away from the actual flat-field.
+Fig \ref{fig:A} shows the fitted flat-field from Survey A after 1,000 with the initial parameters shown in Fig \ref{fig:D}.
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{D_15_ff.png}
+\end{center}
+\caption{The fitted flat field compared to the actual flat field found with Survey D. \label{fig:D}}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{A_1000_ff.png}
+\end{center}
+\caption{The fitted flat field compared to the actual flat field calculated found with Survey A.
+The ubercalibration procedure was started with almost exact flat field parameters (those found from Survey D - see Fig.
+\ref{fig:D}) and has drifted away from the actual solution.
+\label{fig:A}}
+\end{figure}
+
 \appendix
+\begin{comment}
 \section{Random Magnitude Generation}
@@
 \noindent{}where r is a random number in the range [0.0, 1.0).
-
+\end{comment}
 \end{document}

### 2011-01-31 functions.py

@@ -96,7 +96,7 @@ def single_image(sky_catalog, pointing, orientation, plots=None, verbose=None):
     plt.ylim(y_min,y_max)
     filename = "Figures/Camera_Images/%s_alpha_%.1lf_beta_%.1lf_rot_%.1lf.png" % (plots, pointing[0],pointing[1], orientation)
     print filename
-    plt.savefig(filename)
+    plt.savefig(filename,bbox_inches='tight',pad_inches=0.5)
     plt.clf()
 
   return measured_catalog
@@ -296,7 +296,7 @@ def plot_flat_fields(our_q, iteration_number,plots=None):
     filename = 'Figures/Flat_Fields/%s_0%d_ff.png' % (plots,iteration_number)
   else:
     filename = 'Figures/Flat_Fields/%s_%d_ff.png' % (plots, iteration_number)
-  plt.savefig(filename)
+  plt.savefig(filename,bbox_inches='tight',pad_inches=0.5)
   plt.clf()  
 
 def sign(x):

### 2011-01-31 generate_survey.py

@@ -9,8 +9,6 @@ import matplotlib.pylab as plt
 FoV = parameters.FoV()
 sky_limits = parameters.sky_limits()
 
-#def plot_survey(survey):
-  # Transform focal plane into sky
 def plot_survey(survey, filename):
   x_min = -FoV[0]/2; y_min = -FoV[1]/2
   x_max = FoV[0]/2; y_max = FoV[1]/2
@@ -20,13 +18,13 @@ def plot_survey(survey, filename):
   for image in survey:
     alpha, beta = functions.fp2sky(x,y,image[1:3], image[3])
     plt.plot(alpha,beta,'k-',alpha=0.25)
-  plt.xlabel(r"$\alpha$")
-  plt.ylabel(r"$\beta$")
+  plt.xlabel(r"$\alpha$", fontsize=25)
+  plt.ylabel(r"$\beta$", fontsize=25)
   plt.title(r"%i Pointings" % len(survey[:,0]))
   plt.xlim(sky_limits[0]-FoV[0], sky_limits[1]+FoV[0])
   plt.axis('equal')
   plt.ylim(sky_limits[2]-FoV[1], sky_limits[3]+FoV[1])
-  plt.savefig(filename)
+  plt.savefig(filename,bbox_inches='tight',pad_inches=0.)
   return
 
 def generate_uniform_survey(Ncovering):

### 2011-01-31 master.py

@@ -27,9 +27,9 @@ os.system('rm ./Figures/Flat_Fields/*.gif')
 if __name__ == "__main__":
   for strategy in ['D']: #['A', 'D']:
 
-    catalog_plots = None
+    catalog_plots = ''#None
     survey_plots = None#strategy
-    ff_plots = 'all'#None # None, 'all'
+    ff_plots = None # None, 'all'
     verbose = None
     plots = None
 

### 2011-02-01 functions.py
euclid: add coverage plot, add average ff function
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17383 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -164,7 +164,7 @@ def ubercalibration(observation_catalog,sky_catalog, strategy,ff_plots = None):
   stop_condition = 1e-6
   chi2 = 1e9
   old_chi2 = 1e10
-  iteration_number = 1
+  iteration_number = 0
   while (abs(chi2 - old_chi2) > stop_condition):
     temp_chi2 = chi2
     s, s_invvar = s_step(observation_catalog,q)
@@ -176,7 +176,7 @@ def ubercalibration(observation_catalog,sky_catalog, strategy,ff_plots = None):
     rms = rms_error(s[indx],sky_catalog.flux[indx])
     print "%i: f(x,y) =%.2f%s%.2fx%s%.2fy%s%.2fxx%s%.2fxy%s%.2fyy   RMS = %.6f %%   chi2 = %0.2f (%i)" % (iteration_number+1, abs(q[0]),  sign(q[1]), abs(q[1]), sign(q[2]), abs(q[2]), sign(q[3]), abs(q[3]), sign(q[4]), abs(q[4]), sign(q[5]), abs(q[5]), rms, chi2, observation_catalog.size)
 
-    if (ff_plots == 'all') or (abs(chi2 - old_chi2) < stop_condition): plot_flat_fields(q, (iteration_number), plots=strategy)
+    if (ff_plots == 'all') or (abs(chi2 - old_chi2) < stop_condition): plot_flat_fields(q, (iteration_number), strategy=strategy)
     
     if ff_plots == 'all' and (abs(chi2 - old_chi2) < stop_condition): 
       os.system(("convert -delay 20 -loop 0 ./Figures/Flat_Fields/%s*.png ./Figures/Flat_Fields/%s_00_animation.gif" % (strategy,strategy)))
@@ -199,7 +199,7 @@ def evaluate_flat_field(x, y, q):
   assert(len(q) == ((order + 1) * (order + 2) / 2))
   g = evaluate_flat_field_functions(x, y, order)
   return np.dot(g, q)
-
+  
 def normalize_flat_field(q):
   return (q / q[0])
 
@@ -237,18 +237,32 @@ def q_step(obs_cat, s, order, iteration_number, plots=None):
 def rms_error(flux_estimate, true_flux):
   return 100*np.sqrt(np.mean((flux_estimate-true_flux)/true_flux)**2)
 
+def average_over_ff(func, q):
[…]

### 2011-02-01 generate_survey.py

@@ -9,7 +9,7 @@ import matplotlib.pylab as plt
 FoV = parameters.FoV()
 sky_limits = parameters.sky_limits()
 
-def plot_survey(survey, filename):
+def plot_survey(survey, survey_name):
   x_min = -FoV[0]/2; y_min = -FoV[1]/2
   x_max = FoV[0]/2; y_max = FoV[1]/2
   x = np.array([x_min, x_min, x_max, x_max, x_min])
@@ -20,10 +20,11 @@ def plot_survey(survey, filename):
     plt.plot(alpha,beta,'k-',alpha=0.25)
   plt.xlabel(r"$\alpha$", fontsize=25)
   plt.ylabel(r"$\beta$", fontsize=25)
-  plt.title(r"%i Pointings" % len(survey[:,0]))
+  plt.title(r"Survey %s: %i Pointings" % (survey_name, len(survey[:,0])))
   plt.xlim(sky_limits[0]-FoV[0], sky_limits[1]+FoV[0])
   plt.axis('equal')
   plt.ylim(sky_limits[2]-FoV[1], sky_limits[3]+FoV[1])
+  filename = './Figures/%s.pdf' % survey_name
   plt.savefig(filename,bbox_inches='tight',pad_inches=0.)
   return
 
@@ -83,9 +84,9 @@ def generate_random_survey(N):
 
 if __name__ == "__main__":
   xA = generate_uniform_survey(9)
-  plot_survey(xA,'Figures/A.pdf')
+  plot_survey(xA,'A')
   np.savetxt('A.txt',xA)
   xD = generate_random_survey(len(xA))
-  plot_survey(xD,'Figures/D.pdf')
+  plot_survey(xD,'D')
   np.savetxt('D.txt',xD)
   

### 2011-02-01 god.py

@@ -10,8 +10,7 @@ import functions as f
 def flat_field_parameters():
   return np.array([1, 0, 0, -0.2, 0, -0.5])
 
-def flat_field(x,y):
-  par = flat_field_parameters()
+def flat_field(x,y,par = flat_field_parameters()):
   return (par[0] + par[1]*x + par[2]*y + par[3]*x**2 + par[4]*x*y + par[5]*y**2) 
 
 def generate_magnitudes(m_min,m_max,powerlaw,size):

### 2011-02-01 master.py

@@ -27,11 +27,12 @@ os.system('rm ./Figures/Flat_Fields/*.gif')
 if __name__ == "__main__":
   for strategy in ['D']: #['A', 'D']:
 
-    catalog_plots = ''#None
-    survey_plots = None#strategy
+    catalog_plots = None
+    survey_plots = None # strategy
+    coverage_plots = None # strategy
     ff_plots = None # None, 'all'
     verbose = None
-    plots = None
+    health_plots = None
 
     #********************************************************
     #*************** Generate Sky Catalog *******************
@@ -51,6 +52,7 @@ if __name__ == "__main__":
     observation_catalog = f.survey(sky_catalog, survey_file, plots=survey_plots, verbose=verbose) 
         # observation_catalog: *.size, *.pointing_ID, *.star_ID, *.flux, *.invvar, *.x, *.y
 
+    if coverage_plots != None: f.coverage(observation_catalog, strategy)
 
     #********************************************************
     #********************* Ubercalibration ******************
@@ -61,7 +63,7 @@ if __name__ == "__main__":
     #*********************** Health Checks ******************
     #********************************************************
 
-    if plots != None:
+    if health_plots != None:
       # Flux Uncertainty variance check
       if verbose != None: print "Plotting flux uncertainty variance..."
       eta = p.eta()

### 2011-02-01 parameters.py

@@ -15,7 +15,7 @@ def m_max():
 
 def density_of_stars():
   # density of stars (all magnitudes)
-  return 50
+  return 5
 
 def powerlaw():
   #B in log10(dN/dm) = A + B*m
@@ -28,3 +28,6 @@ def FoV():
 def sky_limits():
   # limits of spacial/angular coordinates 
   return [0.0,10.0,0.0,10.0] #[α_min, α_max, β_min, β_max]
+  
+def ff_samples():
+  return [1000,1000]

### 2011-02-01 functions.py
euclid: normalize ff with means
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17384 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -201,7 +201,9 @@ def evaluate_flat_field(x, y, q):
   return np.dot(g, q)
   
 def normalize_flat_field(q):
-  return (q / q[0])
+  fit_mean = average_over_ff(evaluate_flat_field, q)
+  god_mean = average_over_ff(god.flat_field, god.flat_field_parameters())
+  return (q * god_mean/fit_mean)#(q / q[0])
 
 def s_step(obs_cat, q):
   ff = evaluate_flat_field(obs_cat.x, obs_cat.y, q)

### 2011-02-01 master.py

@@ -25,12 +25,12 @@ os.system('rm ./Figures/Flat_Fields/*.png')
 os.system('rm ./Figures/Flat_Fields/*.gif')
 
 if __name__ == "__main__":
-  for strategy in ['D']: #['A', 'D']:
+  for strategy in ['A']: #['A', 'D']:
 
     catalog_plots = None
     survey_plots = None # strategy
     coverage_plots = None # strategy
-    ff_plots = None # None, 'all'
+    ff_plots = 'all' # None, 'all'
     verbose = None
     health_plots = None
 

### 2011-02-01 god.py
euclid: god's catalog plot update
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17385 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -42,29 +42,29 @@ def create_catalog(M, m_min, m_max, powerlaw, limits, seed, plots=None, verbose=
   #**********************************************************
     
   if plots != None:
+    fontsize = 20
+    plt.figure(figsize=(14,6))
+    st = plt.suptitle("Sky Catalog", fontsize=fontsize)
     # Plot portion of sky
-    fontsize = 25
-    if verbose != None: print "Plotting portion of sky..."
-    plt.figure()
-    plt.plot(catalog.alpha,catalog.beta,'ko',markersize=2)
+    plt.subplot(121)
+    if verbose != None: print "Plotting portion of sky and histogram..."
+    plt.plot(catalog.alpha,catalog.beta,'.k',markersize=1)
     plt.xlabel(ur'$\alpha$', fontsize=fontsize)
     plt.ylabel(ur'$\beta$', fontsize=fontsize)
-    plt.title("God's Sky", fontsize=fontsize)
+    #plt.title("God's Sky", fontsize=fontsize)
+    plt.xlim(limits[0] - (limits[1]-limits[0])*0.1, limits[1] + (limits[1]-limits[0])*0.1)
+    plt.ylim(limits[2] - (limits[3]-limits[2])*0.1, limits[3] + (limits[3]-limits[2])*0.1)
     ax = plt.gca()
     for tick in ax.xaxis.get_major_ticks():
       tick.label1.set_fontsize(fontsize)
     for tick in ax.yaxis.get_major_ticks():
       tick.label1.set_fontsize(fontsize)
-    plt.savefig("Figures/gods_sky.png",bbox_inches='tight',pad_inches=0.)
-    if verbose != None: print "...done!"
-    
     # Histogram of source magnitude
-    plt.figure()
-    fontsize = 25
+    plt.subplot(122)
     bin=np.arange(m_min,m_max,0.05)
     if verbose != None: print "Plotting histogram of source magnitude..."
     plt.hist(catalog.mag,bins=bin, log=True)
-    plt.title("%i Sources in Sample" % catalog.size, fontsize=fontsize)
+    #plt.title("%i Sources in Sample" % catalog.size, fontsize=fontsize)
     plt.xlabel("Source Magnitude", fontsize=fontsize)
     plt.ylabel("log(N)", fontsize=fontsize)
     ax = plt.gca()
@@ -72,7 +72,9 @@ def create_catalog(M, m_min, m_max, powerlaw, limits, seed, plots=None, verbose=
[…]

### 2011-02-01 master.py

@@ -27,10 +27,10 @@ os.system('rm ./Figures/Flat_Fields/*.gif')
 if __name__ == "__main__":
   for strategy in ['A']: #['A', 'D']:
 
-    catalog_plots = None
+    catalog_plots = ''#None
     survey_plots = None # strategy
     coverage_plots = None # strategy
-    ff_plots = 'all' # None, 'all'
+    ff_plots = None # None, 'all'
     verbose = None
     health_plots = None
 

### 2011-02-02 god.py
euclid: change epsilon_max to a parameter
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17408 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -6,7 +6,7 @@ import math
 import sys
 import matplotlib.pylab as plt
 import functions as f
-import parameters
+import parameters as p
 
 # magic numbers
 def flat_field_parameters():
@@ -103,7 +103,7 @@ def flat_field(x,y,par = flat_field_parameters()):
   ff = (par[0] + par[1]*x + par[2]*y + par[3]*x**2 + par[4]*x*y + par[5]*y**2)
   # detector part
   k = 6
-  fov = parameters.FoV()
+  fov = p.FoV()
   for nx in range(8):
     kx = nx * np.pi / fov[0]
     ckx = np.cos(kx * x)
@@ -134,7 +134,7 @@ class SourceCatalog:
       self.mag = generate_magnitudes(m_min,m_max,powerlaw,size)
       self.size = size
       self.flux = f.mag2flux(self.mag)
-      self.epsilon = np.random.uniform(0,1, size=size)
+      self.epsilon = np.random.uniform(0,p.epsilon_max, size=size)
       
 def create_catalog(M, m_min, m_max, powerlaw, limits, seed, plots=None, verbose=None):
   if verbose != None: print "Generating God's Catalog..."

### 2011-02-02 parameters.py

@@ -33,3 +33,5 @@ def ff_samples():
   return [100,100]
   
 flat_field_order = 9
+
+epsilon_max = 1

### 2011-02-02 Plotting_Data/coverage_plot.py
euclid: add survey coverage plot
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17412 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -1,0 +1,56 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+import numpy as np
+import matplotlib.pylab as plt
+import pickle
+import os
+
+plt.figure()
+fontsize = 20
+tick_fontsize = 12
+guide_frac = 0.9 # what coverage to draw guide line at
+
+# Select current or old pickle file
+filename = "coverage.p"
+strategies_filename = "./strategies.p"
+if os.path.exists(strategies_filename) != True:
+  filename += '_old'
+  strategies_filename += 'old'
+  plt.title('(Old Data)')
+
+sky_surveys = pickle.load(open(strategies_filename))
+
+for surveys in sky_surveys:
+  dic = pickle.load(open(('./'+surveys+'_'+filename)))
+  number_stars = dic['number_stars']
+  k = dic['k']
+  strategy = dic['strategy']
+
+  num_obs_of_star = np.bincount(k)
+  num_star_with_N_obs = np.bincount(num_obs_of_star)
+  # "Integrate" 
+  for i in range(1,len(num_star_with_N_obs)):
+    num_star_with_N_obs[i] += num_star_with_N_obs[i-1]
+    
+  fraction = (1.*num_star_with_N_obs)/number_stars
+
+  plt.plot(fraction, label = ('Survey '+strategy))
+  gt90 = np.where (fraction > guide_frac)
+  print gt90[0][0]
+  guide_line = np.array([[0,float(gt90[0][0]), float(gt90[0][0])], [fraction[gt90[0][0]], fraction[gt90[0][0]],0.]])
+  plt.plot(guide_line[0], guide_line[1],'k--', alpha = 0.25)
+  
+plt.xlabel(r"Number of Observations")
+plt.ylabel(r"Fraction of Sources Covered")
+plt.title('Survey Coverage')
+plt.legend(loc=4)
+filename = "./coverage.png"
+ax = plt.gca()
+for tick in ax.xaxis.get_major_ticks():
+  tick.label1.set_fontsize(tick_fontsize)
+for tick in ax.yaxis.get_major_ticks():
+  tick.label1.set_fontsize(tick_fontsize)
+plt.ylim(0.,1.1)
+plt.savefig(filename,bbox_inches='tight')#,pad_inches=0.5)
+plt.clf()

### 2011-02-02 functions.py

@@ -111,7 +111,7 @@ def single_image(sky_catalog, pointing, orientation, plots=None, verbose=None):
     plt.plot(alpha,beta,'k', linewidth=2)
     plt.xlabel(ur'$\alpha$', fontsize=20)
     plt.ylabel(ur'$\beta$', fontsize=20)
-    sky = parameters.sky_limits()
+    sky = parameters.sky_limits
     plt.xlim(sky[0],sky[1])
     plt.ylim(sky[2],sky[3])
     plt.axis('equal')
@@ -322,31 +322,16 @@ def plot_flat_fields(our_q, iteration_number,strategy=None):
   plt.savefig(filename,bbox_inches='tight',pad_inches=0.5)
   plt.clf()  
 
-def coverage(obs_cat, strategy):
-  limits = parameters.sky_limits()
-  num_stars = int(np.around(parameters.density_of_stars()*(limits[1]-limits[0]) * (limits[3]-limits[2])))
-  num_obs = np.zeros((num_stars,2))
-  for i in range(num_stars):
-    num_obs[i,0] = i
-    indx = np.where(obs_cat.k == i)
-    num_obs[i,1] = len(obs_cat.k[indx])
-  max_obs = int(np.around(np.max(num_obs[:,1])))
-  int_obs = np.zeros((max_obs+1,2))
-  
-  for i in range(max_obs+1):
-    int_obs[i,0] = i
-    indx = np.where(num_obs[:,1] <= i)
-    int_obs[i,1] = len(indx[0])
-  plt.figure()
-  plt.plot(int_obs[:,0],int_obs[:,1]/float(num_stars))
-  plt.xlabel(r"Number of Observations")
-  plt.ylabel(r"Fraction of Sources Covered")
-  plt.title('Survey %s Coverage' % strategy)
-  plt.ylim(0.,1.)
-  filename = "./Figures/%s_coverage.png" % strategy
-  plt.savefig(filename,bbox_inches='tight')#,pad_inches=0.5)
-  plt.clf()
-  return num_obs
+def coverage(obs_cat, strategy, verbose=None):
+  if verbose != None: print "Writing out coverage pickle..."
+  dic = {}
+  dic['number_stars'] = int(np.around(parameters.density_of_stars()*(parameters.sky_limits[1]-parameters.sky_limits[0]) * (parameters.sky_limits[3]-parameters.sky_limits[2])))
+  dic['k'] = obs_cat.k 
[…]

### 2011-02-02 master.py

@@ -6,6 +6,7 @@ import matplotlib.pylab as plt
 import math
 import os
 import glob
+import pickle
 # Set up LaTeX for plots
 from matplotlib import rc
 rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
@@ -16,6 +17,7 @@ import functions as f
 import god
 import parameters as p
 
+os.system('rm ./Plotting_Data/*.p_old')
 os.chdir("./Plotting_Data/")
 # Change all pickles to _old
 for files in glob.glob("*.p"):
@@ -30,9 +32,11 @@ os.system('rm ./Figures/*.pdf')
 os.system('rm ./Figures/Flat_Fields/*.png')
 os.system('rm ./Figures/Flat_Fields/*.gif')
 
+survey_strategies = ['D', 'A'] #['A', 'D']:
+pickle.dump(survey_strategies, open("./Plotting_Data/strategies.p", "wb" ))
 
 if __name__ == "__main__":
-  for strategy in ['D']: #['A', 'D']:
+  for strategy in survey_strategies: 
 
     catalog_plots = ''#None
     survey_plots = None # strategy
@@ -46,7 +50,7 @@ if __name__ == "__main__":
     #*************** Generate Sky Catalog *******************
     #********************************************************
     # Creates catalog of stars used for calibration
-    sky_catalog = god.create_catalog(p.density_of_stars(), p.m_min(), p.m_max(), p.powerlaw(), p.sky_limits(), seed = 1, plots = catalog_plots, verbose = verbose)
+    sky_catalog = god.create_catalog(p.density_of_stars(), p.m_min(), p.m_max(), p.powerlaw(), p.sky_limits, seed = 1, plots = catalog_plots, verbose = verbose)
         # powerlaw = B in log10(dN/dm) = A + B*m
         # sky_limits = [α_min, α_max, β_min, β_max]
         # returns sky_catalog = [Star ID, magnitude, α, β]
@@ -60,7 +64,7 @@ if __name__ == "__main__":
     observation_catalog = f.survey(sky_catalog, survey_file, plots=survey_plots, verbose=verbose) 
         # observation_catalog: *.size, *.pointing_ID, *.star_ID, *.flux, *.invvar, *.x, *.y
 
-    if coverage_plots != None: f.coverage(observation_catalog, strategy)
+    if coverage_plots != None: f.coverage(observation_catalog, strategy, verbose=verbose)
 
     if plot_invvar != None: f.invvar_saveout(observation_catalog)
     #********************************************************

### 2011-02-02 parameters.py

@@ -25,13 +25,11 @@ def FoV():
   # Camera's field-of-view
   return [0.7,0.7] # [Δα, Δβ]
   
-def sky_limits():
-  # limits of spacial/angular coordinates 
-  return [0.0,10.0,0.0,10.0] #[α_min, α_max, β_min, β_max]
-  
 def ff_samples():
   return [100,100]
   
 flat_field_order = 9
 
 epsilon_max = 1
+
+sky_limits = [0.0,10.0,0.0,10.0] #[α_min, α_max, β_min, β_max]

### 2011-02-03 Plotting_Data/camera_image_plot.py
euclid: update document
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17426 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -1,0 +1,91 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+import numpy as np
+import matplotlib.pylab as plt
+import pickle
+import os
+
+
+plt.figure(figsize=(13,6))
+fontsize = 20
+tick_fontsize = 12
+
+# Select current or old pickle file
+filename = "./camera_image.p"
+if os.path.exists(filename) != True:
+  filename += '_old'
+  plt.title('(Old Data)')
+
+# Load data from pickle file
+dic = pickle.load(open(filename))
+x = dic['measured_catalog.x'] 
+y = dic['measured_catalog.y']
+alpha = dic['sky_catalog.alpha'] 
+beta = dic['sky_catalog.beta']
+sky = dic['sky'] 
+pointing = dic['pointing']
+orientation = dic['orientation']
+fp_x = dic['fp_x']
+fp_y = dic['fp_y']
+fp_alpha = dic['fp_alpha']
+fp_beta = dic['fp_beta']
+inside_FoV = dic['inside_FoV']
+
+
+title = ur'$\alpha$ = %.1lf, $\beta$ = %.1lf, $\theta$ = %.1lf$^\circ$' % (pointing[0],pointing[1], orientation)
+plt.suptitle(title, fontsize=fontsize)
+plt.subplot(121)
+plt.plot(alpha,beta,'.', markersize=2)
+plt.plot(alpha[inside_FoV],beta[inside_FoV],'r.',markersize=2)  
+plt.plot(fp_alpha,fp_beta,'k', linewidth=2)
+plt.xlabel(ur'$\alpha$', fontsize=fontsize)
+plt.ylabel(ur'$\beta$', fontsize=fontsize)
+ax = plt.gca()
+for tick in ax.xaxis.get_major_ticks():
+  tick.label1.set_fontsize(tick_fontsize)
+for tick in ax.yaxis.get_major_ticks():
+  tick.label1.set_fontsize(tick_fontsize)
+dalpha_sky = sky[1] - sky[0]
+dbeta_sky = sky[3] - sky[2]
+buffer_sky = 0.1
+plt.xlim(sky[0] - buffer_sky*dalpha_sky, sky[1] + buffer_sky*dalpha_sky)
+plt.ylim(sky[2] - buffer_sky*dbeta_sky, sky[3] + buffer_sky*dbeta_sky)
+
[…]

### 2011-02-03 Plotting_Data/coverage_plot.py

@@ -16,7 +16,7 @@ filename = "coverage.p"
 strategies_filename = "./strategies.p"
 if os.path.exists(strategies_filename) != True:
   filename += '_old'
-  strategies_filename += 'old'
+  strategies_filename += '_old'
   plt.title('(Old Data)')
 
 sky_surveys = pickle.load(open(strategies_filename))
@@ -35,11 +35,11 @@ for surveys in sky_surveys:
     
   fraction = (1.*num_star_with_N_obs)/number_stars
 
-  plt.plot(fraction, label = ('Survey '+strategy))
+  plt.plot(fraction, label = ('Strategy '+strategy))
   gt90 = np.where (fraction > guide_frac)
   print gt90[0][0]
   guide_line = np.array([[0,float(gt90[0][0]), float(gt90[0][0])], [fraction[gt90[0][0]], fraction[gt90[0][0]],0.]])
-  plt.plot(guide_line[0], guide_line[1],'k--', alpha = 0.25)
+  #plt.plot(guide_line[0], guide_line[1],'k--', alpha = 0.25)
   
 plt.xlabel(r"Number of Observations")
 plt.ylabel(r"Fraction of Sources Covered")

### 2011-02-03 Plotting_Data/invvar_plot.py

@@ -32,7 +32,7 @@ sort_reported_invvar = sort[:,1]
 sort_counts = sort[:,0]
 
 plt.plot(counts, np.log10((1/true_invvar)/counts**2),'r.', markersize = 1., label = "Actual Variance")
-plt.plot(sort_counts, np.log10((1/sort_reported_invvar)/sort_counts**2),'k', label = "Reported Variance")
+plt.plot(sort_counts, np.log10((1/sort_reported_invvar)/sort_counts**2),'k', label = "Assumed Variance")
 plt.xlabel(r'$c_i$', fontsize=fontsize)
 plt.ylabel(ur'$\log_{10}(\frac{{\sigma_i}^2}{c_i^2})$',fontsize=fontsize)#, rotation='horizontal')
 plt.legend()

### 2011-02-03 functions.py

@@ -96,37 +96,7 @@ def single_image(sky_catalog, pointing, orientation, plots=None, verbose=None):
   if verbose != None: print "Measuring stars within FoV..."
   measured_catalog = MeasuredCatalog(camera_catalog, inside_FoV)
   if verbose != None: print "...done!"
-    
-  if plots != None:
-    # Plot sky (highlight extracted sources), with FoV
-    plt.figure(2000, figsize=(13,6))
-    title = ur'$\alpha$ = %.1lf, $\beta$ = %.1lf, $\theta$ = %.1lf$^\circ$' % (pointing[0],pointing[1], orientation)
-    plt.suptitle(title, fontsize=20)
-    plt.subplot(121)
-    x = np.array([x_min, x_min, x_max, x_max, x_min])
-    y = np.array([y_min, y_max, y_max, y_min, y_min])
-    alpha, beta = fp2sky(x,y,pointing, orientation)
-    plt.plot(sky_catalog.alpha,sky_catalog.beta,'o', markersize=2)
-    plt.plot(sky_catalog.alpha[inside_FoV],sky_catalog.beta[inside_FoV],'ro',markersize=2)
-    plt.plot(alpha,beta,'k', linewidth=2)
-    plt.xlabel(ur'$\alpha$', fontsize=20)
-    plt.ylabel(ur'$\beta$', fontsize=20)
-    sky = pdic['sky_limits']
-    plt.xlim(sky[0],sky[1])
-    plt.ylim(sky[2],sky[3])
-    plt.axis('equal')
-    # Plot sources on focal plane
-    plt.subplot(122)
-    plt.plot(measured_catalog.x,measured_catalog.y,'o', markersize=2)
-    plt.xlabel(ur'$x$', fontsize=20)
-    plt.ylabel(ur'$y$', fontsize=20)
-    plt.xlim(x_min,x_max)
-    plt.ylim(y_min,y_max)
-    filename = "Figures/Camera_Images/%s_alpha_%.1lf_beta_%.1lf_rot_%.1lf.png" % (plots, pointing[0],pointing[1], orientation)
-    print filename
-    plt.savefig(filename,bbox_inches='tight',pad_inches=0.5)
-    plt.clf()
-
+  if plots != None: save_camera(sky_catalog, measured_catalog, inside_FoV, pointing, orientation, verbose = verbose)
   return measured_catalog
   # measured_sources  *.size, *.k, *.flux, *.invvar, *.x, *.y
 
[…]

### 2011-02-03 generate_survey.py

@@ -3,29 +3,28 @@
 
 import numpy as np
 import random
-import parameters
+import tparameters as p
 import functions
 import matplotlib.pylab as plt
-FoV = parameters.FoV()
-sky_limits = parameters.sky_limits()
+
+pdic = p.dic
+FoV = pdic['FoV']
+sky_limits = pdic['sky_limits']
 
 def plot_survey(survey, survey_name):
   x_min = -FoV[0]/2; y_min = -FoV[1]/2
   x_max = FoV[0]/2; y_max = FoV[1]/2
   x = np.array([x_min, x_min, x_max, x_max, x_min])
   y = np.array([y_min, y_max, y_max, y_min, y_min])
-  plt.clf()
   for image in survey:
     alpha, beta = functions.fp2sky(x,y,image[1:3], image[3])
     plt.plot(alpha,beta,'k-',alpha=0.25)
   plt.xlabel(r"$\alpha$", fontsize=25)
   plt.ylabel(r"$\beta$", fontsize=25)
-  plt.title(r"Survey %s: %i Pointings" % (survey_name, len(survey[:,0])))
+  plt.title(r"Strategy %s: %i Pointings" % (survey_name, len(survey[:,0])))
   plt.xlim(sky_limits[0]-FoV[0], sky_limits[1]+FoV[0])
   plt.axis('equal')
   plt.ylim(sky_limits[2]-FoV[1], sky_limits[3]+FoV[1])
-  filename = './Figures/%s.pdf' % survey_name
-  plt.savefig(filename,bbox_inches='tight',pad_inches=0.)
   return
 
 def generate_uniform_survey(Ncovering):
@@ -83,10 +82,14 @@ def generate_random_survey(N):
   return x
 
 if __name__ == "__main__":
+  plt.figure(figsize=(15,6))
+  plt.subplot(121)
   xA = generate_uniform_survey(9)
   plot_survey(xA,'A')
   np.savetxt('A.txt',xA)
   xD = generate_random_survey(len(xA))
+  plt.subplot(122)
   plot_survey(xD,'D')
   np.savetxt('D.txt',xD)
+  plt.savefig("./Plotting_Data/surveys.png",bbox_inches='tight',pad_inches=0.)
   

### 2011-02-03 master.py

@@ -40,29 +40,29 @@ def clean_up_old_files():
     newfilename = files+'_old'
     os.rename(files,newfilename)
   os.chdir("..")
-
+  
   # Remove all old plots
   os.system('rm ./Figures/Camera_Images/*.png')
   os.system('rm ./Figures/*.png')
   os.system('rm ./Figures/*.pdf')
   os.system('rm ./Figures/Flat_Fields/*.png')
   os.system('rm ./Figures/Flat_Fields/*.gif')
-
   
-survey_strategies = ['D'] #['A', 'D']:
-pickle.dump(survey_strategies, open("./Plotting_Data/strategies.p", "wb" ))
 
 if __name__ == "__main__":
-  os.system('mkdir -p ./Plotting_Data/')
   os.system('mkdir -p ./Figures/Camera_Images/')
   os.system('mkdir -p ./Figures/Flat_Fields/')
-  pdic, mod_dic = init_func()
-  clean_up_old_files()
   
+  clean_up_old_files()
+  pdic, mod_dic = init_func()
+
+  survey_strategies = ['D','A'] #['A', 'D']:
+  pickle.dump(survey_strategies, open("./Plotting_Data/strategies.p", "wb" ))
+
   for strategy in survey_strategies: 
 
     catalog_plots = ''#None
-    survey_plots = None # strategy
+    survey_plots = ''#None # strategy
     
     coverage_plots = strategy # strategy
     ff_plots = 'all' # None, 'all'
@@ -74,7 +74,7 @@ if __name__ == "__main__":
     #*************** Generate Sky Catalog *******************
     #********************************************************
     # Creates catalog of stars used for calibration
-    sky_catalog = god.create_catalog(seed = 1, plots = catalog_plots, verbose = verbose)
+    sky_catalog = god.create_catalog(seed = 2, plots = catalog_plots, verbose = verbose)
         # powerlaw = B in log10(dN/dm) = A + B*m
         # sky_limits = [α_min, α_max, β_min, β_max]
         # returns sky_catalog = [Star ID, magnitude, α, β]

### 2011-02-03 ugly_wrapper.py

@@ -9,7 +9,7 @@ import matplotlib.pylab as plt
 os.system('rm ./test.txt')
 
 parameter = 'epsilon_max'
-par_range = np.arange(0,1,0.1)
+par_range = np.arange(0,1,0.2)
 
 for i in range(len(par_range)):
   command = './master.py %s %s' % (parameter, par_range[i])

### 2011-02-04 generate_survey.py
euclid: generate survey bug fix
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17448 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -4,7 +4,6 @@
 import numpy as np
 import random
 import tparameters as p
-import functions
 import matplotlib.pylab as plt
 
 # XX Dangerous hack!!
@@ -12,13 +11,19 @@ pdic = p.dic
 FoV = pdic['FoV']
 sky_limits = pdic['sky_limits']
 
+def fp2sky(x, y, pointing, orientation):
+  theta = - orientation*np.pi/180 # telescope rotates NOT sky
+  alpha = x * np.cos(theta) + y * np.sin(theta) + pointing[0]
+  beta = -x * np.sin(theta) + y * np.cos(theta) + pointing[1]
+  return alpha, beta
+
 def plot_survey(survey, survey_name):
   x_min = -FoV[0]/2; y_min = -FoV[1]/2
   x_max = FoV[0]/2; y_max = FoV[1]/2
   x = np.array([x_min, x_min, x_max, x_max, x_min])
   y = np.array([y_min, y_max, y_max, y_min, y_min])
   for image in survey:
-    alpha, beta = functions.fp2sky(x,y,image[1:3], image[3])
+    alpha, beta = fp2sky(x,y,image[1:3], image[3])
     plt.plot(alpha,beta,'k-',alpha=0.25)
   plt.xlabel(r"$\alpha$", fontsize=25)
   plt.ylabel(r"$\beta$", fontsize=25)
@@ -118,5 +123,5 @@ if __name__ == "__main__":
   plt.subplot(224)
   plot_survey(xD,'D')
   np.savetxt('D.txt',xD)
-  plt.savefig("./Plotting_Data/surveys.png",bbox_inches='tight',pad_inches=0.)
+  plt.savefig("./surveys.png",bbox_inches='tight',pad_inches=0.)
   

### 2011-02-04 plotting_function.py
euclid: plotting function bug fix
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@17449 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -5,6 +5,7 @@ import numpy as np
 import matplotlib.pylab as plt
 import pickle
 import os
+import sys
 import string
 from matplotlib import rc
 rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
@@ -16,8 +17,11 @@ rc('text', usetex=True)
 # Data directory
 dir_path = 'temp/density_of_stars/'
 parameter_values = os.listdir(dir_path)
-if 'simulation_parameters.p' in parameter_values:
-  parameter_values.remove('simulation_parameters.p')
+temp = 1*parameter_values
+for isfile in temp:
+  if os.path.isfile(dir_path + isfile):
+    parameter_values.remove(isfile)
+
 
 # General Plotting Parameters
 fontsize = 20

### 2011-06-16 functions.py
euclid: add deep survey simulations
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@18670 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -135,7 +135,8 @@ def ubercalibration(observation_catalog,sky_catalog, strategy,modified_parameter
   chi2 = 1e9
   old_chi2 = 1e10
   iteration_number = 0
-  next_plot_iteration = 1
+  next_plot_iteration = 1 
+  saveout_flat_fields(q, iteration_number, strategy=strategy)
   while ((abs(chi2 - old_chi2) > stop_condition) and (iteration_number < 258)):
     iteration_number += 1
     temp_chi2 = chi2

### 2011-06-16 generate_deep.py

@@ -1,0 +1,92 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+import numpy as np
+import tparameters as p
+import matplotlib.pylab as plt
+
+# XX Dangerous hack!!
+pdic = p.dic
+FoV = pdic['FoV']
+sky_limits = pdic['sky_limits']
+field_overlap = 0.02
+dither_overlap = 0.98
+no_dithers = 4
+square_num = 5.
+passes = 12 # per year
+angle_between_passes = 30
+
+def single_pass(offset, theta):
+  # Create Center Square
+  x_low = -FoV[0]*(square_num+1)*0.5*(1-field_overlap)
+  x_high = -x_low
+  y_low = -FoV[1]*(square_num+1)*0.5*(1-field_overlap)
+  y_high = -y_low
+  x_centers = np.linspace(y_low,y_high, num = (square_num+2), endpoint=True)
+  y_centers = np.linspace(y_low,y_high, num = (square_num+2), endpoint=True)
+  X_centers,Y_centers = np.meshgrid(x_centers[1:-1],y_centers[1:-1])
+  # Add Edges
+  for ii in range(int(square_num-1)):
+    X_centers = np.append(X_centers, x_centers[0])
+    Y_centers = np.append(Y_centers, y_centers[0]+(ii+1.5) * FoV[1])
+    X_centers = np.append(X_centers, x_centers[-1])
+    Y_centers = np.append(Y_centers, y_centers[0]+(ii+1.5) * FoV[1])
+    X_centers = np.append(X_centers, x_centers[0]+(ii+1.5) * FoV[0])
+    Y_centers = np.append(Y_centers, y_centers[0])
+    X_centers = np.append(X_centers, x_centers[0]+(ii+1.5) * FoV[0])
+    Y_centers = np.append(Y_centers, y_centers[-1])
+  # Create Dithers
+  X = np.zeros((no_dithers*len(X_centers), 2))
+  for ii in range(len(X_centers)):
+    dx = 0.5*(1-dither_overlap) * FoV[0]
+    dy = 0.5*(1-dither_overlap) * FoV[1]
+    add_sub = np.array([[-1,-1,1,1], [-1,1,-1,1]])
+    for jj in range(len(add_sub[0,:])):
[…]

### 2011-06-16 generate_survey.py

@@ -102,7 +102,7 @@ def generate_random_survey(N):
   return x
 
 if __name__ == "__main__":
-  number_passes = 3
+  number_passes = 20
   plt.figure(figsize=(15,15))
   plt.subplot(221)
   xA = generate_uniform_survey(number_passes)

### 2011-06-16 god.py

@@ -139,13 +139,14 @@ class SourceCatalog:
       self.flux = f.mag2flux(self.mag)
       self.epsilon = np.random.uniform(0,pdic['epsilon_max'], size=size)
       
-def create_catalog(seed, plots=None, verbose=False):
+def create_catalog(plots=None, verbose=False):
   if verbose: print "Generating God's Catalog..."
   M = pdic['density_of_stars']
   m_min = pdic['m_min']
   m_max = pdic['m_max']
   powerlaw = pdic['powerlaw']
   limits = pdic['sky_limits']
+  seed = pdic['seed']
   
   # Calculate total number of stars in catalog
   number_stars = M * (limits[1]-limits[0])*(limits[3]-limits[2])

### 2011-06-16 master.py

@@ -43,7 +43,7 @@ if __name__ == "__main__":
     os.system(('mkdir -p %s/%s' % (directory_path, survey_strategies[ii])))
     
     # Creates catalog of stars used for calibration
-    sky_catalog = god.create_catalog(seed = 103, plots = sky_catalog_plots, verbose = verbose)
+    sky_catalog = god.create_catalog(plots = sky_catalog_plots, verbose = verbose)
         # sky_catalog: *.star_ID, *.mag, *.alpha, *.beta, *.size
 
     # Survey sky

### 2011-06-16 plotting_function.py

@@ -214,7 +214,7 @@ def plot_flat_fields(ff_filename, strategy):
   
   plt.subplot(121)
   plt.suptitle('Survey %s' % strategy, fontsize = fontsize)
-  plt.title(r"Flat-Fields (God's = Black; Fitted = Red) Iteration: %i" % (iteration_number))
+  plt.title(r"Flat-Fields (True = Black; Fitted = Red) Iteration: %i" % (iteration_number))
   plt.xlabel(r"$\alpha$")
   plt.ylabel(r"$\beta$")
 
@@ -232,8 +232,8 @@ def plot_flat_fields(ff_filename, strategy):
   
   # Plot residual in flat-field
   plt.subplot(122)
-  plt.title(r"Residual (Fit - God) in Flat-Field (\%)")
-  a = plt.imshow((100*(our_ff-god_ff)/god_ff),extent=(-FoV[0]/2,FoV[0]/2,-FoV[1]/2,FoV[1]/2), vmin = -1,vmax = 1, cmap='gray')
+  plt.title(r"Residual (Fit - True) in Flat-Field (\%)")
+  a = plt.imshow((100*(our_ff-god_ff)/god_ff),extent=(-FoV[0]/2,FoV[0]/2,-FoV[1]/2,FoV[1]/2), vmin = -0.5,vmax = 0.5, cmap='gray')
   plt.colorbar(a,shrink=0.7)
   plt.xlabel(r"$\alpha$")
   plt.ylabel(r"$\beta$")
@@ -287,7 +287,7 @@ if __name__ == "__main__":
       png_dir = ('%s/%s/' % (new_dir_path, survey_strategies[ii]))
       out_dir = ('%s/' % (new_dir_path))
       print "...animating..."
-      command = ('convert -delay 20 -loop 0 %s*.png %s%s_animation.gif' % (png_dir, out_dir, 
+      command = ('convert -delay 50 -loop 0 %s*.png %s%s_animation.gif' % (png_dir, out_dir, 
       survey_strategies[ii]))
       os.system(command)
       print "...done!"

### 2011-06-16 tparameters.py

@@ -2,15 +2,15 @@
 # -*- coding: utf-8 -*-
 
 dic = {
-  'survey_strategies'     :     ['C','D'], #['D', 'C', 'B', 'A']
+  'survey_strategies'     :     ['deep'], #['D', 'C', 'B', 'A']
   
   'mag_at_ten_sigma'    :     22.5,
 
   'eta'                 :     0.01,
   
-  'm_min'               :     19.5,
+  'm_min'               :     17, #19.5,
   
-  'm_max'               :     22.5,
+  'm_max'               :     19, #22.5,
   
   'density_of_stars'    :     5, # all magnitudes
   
@@ -20,10 +20,12 @@ dic = {
   
   'ff_samples'          :     [100, 100],
   
-  'flat_field_order'    :     9,
+  'flat_field_order'    :     15,
   
   'epsilon_max'         :     1,
   
-  'sky_limits'          :     [0.0, 10.0, 0.0, 10.0], #[α_min, α_max, β_min, β_max]
+  'sky_limits'          :     [-5.0, 5.0, -5.0, 5.0], #[α_min, α_max, β_min, β_max]
+  
+  'seed'                :     1 # Random number seed
   
   } 

### 2011-06-16 ugly_wrapper.py

@@ -24,10 +24,10 @@ ff_directory = directory_path+'/Flat_Fields/'
 # Single run, or many runs?
 modify_parameter = True
 if modify_parameter:
-  parameter = 'eta'
-  par_low = 0.
-  par_high = 1.
-  steps = 5
+  parameter = 'density_of_stars'
+  par_low = 100
+  par_high = 110.
+  steps = 1
   parameter_values = np.around(np.linspace(par_low, par_high, num=steps),decimals=1)
   simulation_parameters[parameter] = parameter_values
 else:

### 2011-07-03 Documentation/cross_cal_paper.tex
euclid: re-draft paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@18942 03a3cea6-2c03-0410-ac9b-9271b1c66c29

% TO DO:
% Plot catalog (histogram and sample sky)
% Select brightess sources
% Principle component analysis???!!!!

% Introduction
% add coverage plot
% finish ubercal in matrix form

%% Use whichever is most appropriate for your purposes.
%%
%%\documentclass[12pt,preprint]{aastex}
%% manuscript produces a one-column, double-spaced document:
\documentclass[manuscript]{aastex}
%% preprint2 produces a double-column, single-spaced document:
%% \documentclass[preprint2]{aastex}

%% Sometimes a paper's abstract is too long to fit on the
%% title page in preprint2 mode. When that is the case,
%% use the longabstract style option.

%% \documentclass[preprint2,longabstract]{aastex}

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{holmes@mpia.de}

%% You can insert a short comment on the title page using the command below.

\slugcomment{Version 0.1}

\shorttitle{Optimizing Large Scale Surveys for Photometric Calibration}
\shortauthors{Holmes et al.}

\begin{document}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire.

\title{Optimizing Large Scale Surveys for Photometric Calibration, \\
    The Euclid Dark Energy Mission as an Example}

%% Use \author, \affil, and the \and command to format
%% author and affiliation information.
%% Note that \email has replaced the old \authoremail command
%% from AASTeX v4.0. You can use \email to mark an email address
%% anywhere in the paper, not just in the front matter.
%% As in the title, use \\ to force line breaks.

\author{Rory Holmes\altaffilmark{1}, David W. Hogg\altaffilmark{1,2} and H.-W. Rix\altaffilmark{1}}

%% Notice that each of these authors has alternate affiliations, which
%% are identified by the \altaffilmark after each name.  Specify alternate
%% affiliation information with \altaffiltext, with one command per each
%% affiliation.

\altaffiltext{1}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, Heidelberg, 69117, Germany.}
\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA.}

%% Mark off your abstract in the ``abstract'' environment. In the manuscript
%% style, abstract will output a Received/Accepted line after the
%% title and affiliation information. No date will appear since the author
%% does not have this information. The dates will be filled in by the
%% editorial office after submission.

\begin{abstract}
XX
\end{abstract}

%% Keywords should appear after the \end{abstract} command. The uncommented
%% example has been keyed in ApJ style. See the instructions to authors
%% for the journal to which you are submitting your paper to determine
%% what keyword punctuation is appropriate.

\keywords{XX, XX, XX}

\section{Introduction}
Astronomers tend to think in terms of taking science data and calibration data separately. The former is used for science and the latter is used to set instrument parameters like flat-field, bias and so-on. But typically far more photons are collected during science exposures; are these not incredibly constraining on the calibration themselves? Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (XX ref) (SDSS) much more calibration information was obtained from the science data than the calibration data. But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: good redundancy is required in the data stream, and a redundancy of a very specific kind.

In this paper we will show that this distinction between science and calibration data no longer applies in the large survey mission, both ground and space based, currently being planned and developed. We show that, as long as the survey strategy is designed appropriately, the calibration information within the science data can be used to accurately constrain the relative photometric response of an imaging instrument. This retrospective, cross-calibration of the dataset is a very powerful calibration method; future large scale imaging surveys should be optimized with this in mind.

We concentrate on catalog level simulations of different survey strategies and assess how well the resultant dataset can be used to constrain the relative response of the imaging instrument it came from. Although we do not model the imaging instrument on a pixel scale, we do include realistic measurement uncertainties within the simulations.

Although the results we present here are general to all large scale imaging surveys, in Section XX we focus on the proposed Euclid Deep Field Survey. This survey has been optimized, within other constraints, based on the results from this work.

\section{Simulating the Measured Source Database}
To assess how well this cross-calibration performs with different observation strategies, we first constructed the required simulation machinery to take an observing strategy as an input, generate a realistic sky, perform the survey - with appropriate measurement uncertainties - and cross-calibration the resulting dataset. The following subsections detail the main methods and assumptions used in our simulations.

\subsection{Sky Catalog}
We first create a catalog of fake sources, both stars and galaxies, based on realistic densities within the magnitude range $m_{min} < m < m_{max}$. We choose these limits based on the properties of single exposure of the current configuration of Euclid NIR Imaging Channel (NIP): $m_{min} = 17$, the approximate saturation limit, and $m_{max} = 22$, the 10$\sigma$ limit. Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes distributed according to the following powerlaw:
\begin{displaymath}
\log_{10} \frac{dN}{d\Omega\,dm} = a + b\,m
\end{displaymath}
Where $a$ sets the total source density and $b_{star} = 0.25$ steller and galaxy populations reported in XX. We make no distinction between stars and galaxies in our simulations. The random number generator is given the same seed, so that the same simulated sky is produced each time the simulation is run. This ensures that the differences observed are not a result of random fluctuations. We relate source magnitudes, $m$, to source fluxes, $s$, simply by: $m = 22.5 - 2.5\log_{10}(s)$. We therefore do not consider the imaging instrument's absolute response.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{coverage.png}
\end{center}
\caption{An example catalog of fake sources used within the simulations. The sources are shown at their position on a portion of the sky (left) and as a histogram of their magnitudes (right).
\label{fig:sky}}
\end{figure}

\subsection{Single Exposure}
With a given camera pointing $(\alpha, \beta)$ and camera orientation $(\theta)$, we transform the we transform the sky catalog into focal plane coordinates and find all the stars falling within the instrument's field-of-view. For the instrument's field-of-view, we use the current size of the Euclid NIR imager's: $0.7 \times 0.7$~deg$^{2}$. We do not simulate images, instead we trivially convert the true source fluxes into measured counts $(c)$ with an instrument response model $(f)$ and a measurement uncertainty model. A typical exposure is depicted in Figure \ref{fig:camera}.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{camera_image.png}
\end{center}
\caption{A single exposure of the sky with the pointing coordinates $(\alpha = 9.2^\circ, \beta = 2.2^\circ)$ and an orientation of $\theta = 175.4^\circ$. Left: Plot of the entire fake sky, with the instrument's field-of-view shown in black and the measured sources highlighted in red. Right: The distribution of the highlighted sources on the instrument's focal plane. \label{fig:camera}}
\end{figure}

\subsubsection{Catalog to Measured Counts Conversion}
For a measurement $(i)$ the counts recorded from a source ($k$) depends on true instrument response ($f(\vec{q}_{true})$), which is a function of focal plane position $(\vec{x_i})$, and the source's true flux ($s$):

\begin{displaymath}
c_i = f(\vec{x_i} | \vec{q}_{true}) . s_{k} + e_{i}
\end{displaymath}

\noindent{}An error, drawn from the Normal Distribution $N(e|0,{\sigma_i}^2)$, is also introduced to the measurement. In this analysis, we do not consider temporal variations in the instrument's response.

\subsubsection{Noise Model}
\label{sec:noise_model}
We assume that the Euclid exposures will be background limited and that, for systematic reasons, we will hit an upper limit on the signal-to-noise for bright sources. We assume that the uncertainty variance on the counts are:

\begin{displaymath}
\sigma_{{ass}}^{2} = \alpha^{2} + \eta^{2}\, c^{2} 
\end{displaymath}
\noindent{}where $\alpha$ and $\eta$ are both constants. In fact, we complicate the model by applying an extra term to the count's uncertainty variance, which we do not take into account in our analysis:

\begin{displaymath}
\sigma_{{act}}^{2} = (1 + \epsilon) \,\alpha^{2} + \eta^{2}\, c^{2} 
\end{displaymath}
\noindent{}where $\epsilon$ is a random number, in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$. An example of the assumed compared to the true uncertainty variances is shown in Fig. \ref{fig:count_uncertainty}.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{invvar_plot.png}
\end{center}
\caption{The assumed and the actual variances as a function of counts.\label{fig:count_uncertainty}}
\end{figure}

\subsubsection{Instrument Response Model}
We simulate the instrument response $f(\vec{x_i} | \vec{q}_{true})$ variations on both large and small scales. We model large scale structure, such as the residual from a flat-field calibration source, as a second order polynomial. We model small scale variations in the instrument response with XX sine and cosine contributions. We intentionally do not for the instrument response with a model of the same form, instead we fit with a XX order polynomial only.

\subsection{Cross-Calibrating the Source Dataset}
The next step is to calibrate the simulated dataset generated for the survey being investigated. We consider the same cross-calibration procedure used with a number of ground based surveys, such as SDSS XX. This cross-calibration, termed ``ubercalibration´´ in SDSS terminology is an iterative procedure with two sub-steps: (i) the source fluxes are refined based on the latest instrument response estimate (\textit{s\_step}) and (ii) the instrument response is then refined based on the latest star flux estimates (\textit{q\_step}). We iterate these steps until the solution converges. There is a degeneracy in the problem, as both the true value of the instrument response and the source magnitudes are unknown. With this method we are therefore only able to constrain the relative instrument response. We could not, for example, know if the sources are fainter or if the instrument response is lower. This is not an issue though, as absolute standards are always used to constrain the absolute instrument response. We define a relative instrument response with a set of parameters $\vec{q}$. The individual source measurements are defined as $c_i$ (``counts''), where each $i$ corresponds to a exposure number $n$, a specific star ($k$) and a focal plane position $\vec{x}$.

\noindent{}We solve for the following \textit{incomplete} (see Section \ref{sec:noise_model}) model.
\begin{displaymath}
c_i = f(\vec{x_i} | \vec{q}) . s_{k} + e_{i}
\end{displaymath}
where $s_k$ is the true flux from star $k$ and our error is drawn from the Normal Distribution $e_{i} = N(e|0,{\sigma_i}^2)$.

\noindent{}The two steps in the cross-calibration procedure, called the \textit{s\_step} and the \textit{q\_step}, are detailed in the following two sections:

\subsubsection{Source Flux Refinement (\textit{s\_step})}
The flux estimates of each source, $s_k$, are considered individually. At each \textit{s\_step} we can refine a sources's flux estimate by minimizing the error function with the latest instrument response parameters:

\begin{displaymath}
\chi^2_{k} = \sum_{i \in \mathcal{O}(k)} \frac{(c_i-f_{i}s_{k})^2}{{\sigma_i}^2}
\end{displaymath}

\begin{displaymath}
\frac{d\chi^2_{k}}{d s_{k}} = \sum_{i \in \mathcal{O}(k)} \frac{-2 f_{i} (c_i-f_{i}s_{k+1})}{{\sigma_i}^2} = 0
\end{displaymath}

\begin{displaymath}
s'_{k} \leftarrow \left[{\sum_{i \in \mathcal{O}(k)}  \frac{f_{i}^2}{{\sigma_i}^2}} \right]^{-1}  \left[ {\sum_{i \in \mathcal{O}(k)} \frac{f_{i} c_i}{{\sigma_i}^2}} \right]
\end{displaymath}

The standard uncertainty variance on the new star flux estimate, $s'_{k}$, is given by:

\begin{displaymath}
\sigma'^2_k = \left[{\sum_{i \in \mathcal{O}(k)}  \frac{f_{i}^2}{{\sigma_i}^2}} \right]
\end{displaymath}

\subsubsection{Instrument Response Refinement: \textbf{\textit{p\_step}}}
The instrument response parameters can now be refined with the latest star flux estimates. To do this we minimize the total error function for all measurements with respect to the instrument response parameters.

\begin{displaymath}
\chi^2 = \sum_{k} \chi^2_k
\end{displaymath}
where:
\begin{displaymath}
\chi^2_k = \sum_{i} \frac{(c_i-f_{i}s'_{k})^2}{{\sigma_i}^2}
\end{displaymath}
The instrument response $f_i$ is parameterized as:
\begin{displaymath}
f_{i} = \sum_{l = 1}^L q_{lj} g_l(\vec{x_i})
\end{displaymath}
So the total error function can be rewritten as:
\begin{displaymath}
\chi^2 = \sum_{i} \frac{(c_i- s'_{k} \sum_{l = 1}^L q_{lj} g_l(\vec{x_i}))^2}{{\sigma_i}^2}
\end{displaymath}
The error function is the minimized with respect to the instrument response parameters $q_{lj}$:
\begin{displaymath}
\frac{d\chi^2}{dq_{lj}} = \sum_{i} \frac{-2 g_l(\vec{x_i}) s'_{k} (c_i- s'_{k} \sum_{l' = 1}^{L'} q_{l'} g_{l'}(\vec{x_i}))}{{\sigma_i}^2} = 0
\end{displaymath}

\begin{displaymath}
\sum_{i} \frac{g_l(\vec{x_i}) s'_{k} c_i}{{\sigma_i}^2} = \sum_{i} \frac{g_l (\vec{x_i}) s'^2_{k} \sum_{l' = 1}^{L'} q_{l'} g_{l'} (\vec{x_i})} {{\sigma_i}^2}
\end{displaymath}

These two steps are iterated until the solution converges, or until, it is clear that a solution will not be found.

\subsubsection{Badness}
To assess the performance of the cross-calibration procedure we define a ``badness'' quantity, which quantifies the difference between the true and fitted instrument response in single number. The focal plane is sampled on a $300 \times 300$ grid and the badness is defined as the root-mean-squared difference between the true and fitted instrument response at these positions. This provides a more complete description of the cross-calibration performance than the RMS of the fitted star fluxes, especially in the case - such as Euclid - in which repeat observations of a deep field will be used to calibrate a wide survey.

\section{Results}
\subsection{Simple Survey Strategies}
In this section we consider four simple, very different, survey strategies.

\subsection{Surveys Strategies}
We consider two extreme survey examples. The first, called Strategy A, consists of a regular grid of pointings over the sky with a small overlap between neighboring fields. Nine passes are performed over the sky and each time the grids are aligned exactly. The second survey, Strategy D, has the same number of fields as Strategy A, but the pointing and orientation of these fields are random. These two surveys can be seen in Fig. \ref{fig:surveys}.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{coverage.png}
\end{center}
\caption{Left: The A Strategy has regular pointings over the sky with a small overlap between adjacent fields. Nine passes are performed over the sky and each time the grids are aligned exactly. Right: The D Strategy has the same number of pointings as the A Strategy, but the pointings and orientations are random.\label{fig:surveys}}
\end{figure}

The coverage of each strategy is shown in Fig. \ref{fig:coverage}.

\subsection{Ubercalibration Performance}

The difference in ubercalibration method's ability to fit the flat field and the star magnitudes for the two different Strategies is striking. With the random Strategy (D) we can get very close to the actual flat field in a small number of ubercalibration iterations (see Fig \ref{fig:D}). The well ordered Strategy (A), on the other hand, does not converge to the correct flat field even with 1,000 ubercal iterations. Even when we start with almost the correct flat-field parameters (those found with the D Strategy - shown in Fig. \ref{fig:D}), the ubercalibration procedure does not converge. In fact, with each iteration solution drifts away from the actual flat-field. Fig \ref{fig:A} shows the fitted flat-field from Strategy A after 1,000 with the initial parameters shown in Fig \ref{fig:D}.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{D_15_ff.png}
\end{center}
\caption{The fitted flat field compared to the actual flat field found with Strategy D. \label{fig:D}}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{A_1000_ff.png}
\end{center}
\caption{The fitted flat field compared to the actual flat field calculated found with Strategy A. The ubercalibration procedure was started with almost exact flat field parameters (those found from Strategy D - see Fig. \ref{fig:D}) and has drifted away from the actual solution. \label{fig:A}}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.47\textwidth]{coverage.png}
\end{center}
\caption{The coverage of both of the surveys investigated. \label{fig:coverage}}
\end{figure}

\section{Discussion}
small changes to survey that greatly improve coverage plots
random strategy can be defined so that the telescope does not make large steps in pointing, orientation
have't addressed time dependence, how do these surveys with time will be different for different assumptions on temporal variability. 
is the small scale power we put in realistic?
- survey D is much better than survey C for astrometric self calibration 
we haven't thought about gaps
\end{document}

%The following displaymath is used to generate random star magnitudes with the appropriate probability density distribution function:
%
%\begin{displaymath}
%m = \frac{1}{b} \log_{10}{(\left[ 10^{b m_{max}} - 10^{b m_{min}} \right] r + 10^{b m_{min}})}
%\end{displaymath}
%
%\noindent{}where $r$ is a random number in the range [0.0, 1.0). The catalog is shown in Fig. \ref{fig:sky}.

### 2011-07-03 god.py

@@ -65,10 +65,7 @@ def create_catalog(params, out_dir, plots=None, verbose=False):
   catalog = SourceCatalog(params, out_dir, number_stars, m_min, m_max,powerlaw, limits, seed)
   if verbose: print "...done!"
 
-  #**********************************************************
-  #********************* Plot Catalog ***********************
-  #**********************************************************
-    
+  # Print out catalog for plotting
   if plots:
     if verbose: print "Saving source catalog data..."
     pickle_dic = {}

### 2011-07-03 Documentation/cross_cal_paper.tex
euclid: paper update, plotting function update: nicer figures
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@18943 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \begin{center}
-\includegraphics[width=\textwidth]{coverage.png}
+\includegraphics[width=\textwidth]{survey.png}
 \end{center}
@@
 An example of the assumed compared to the true uncertainty variances is shown in Fig.
-\ref{fig:count_uncertainty}.
+\ref{fig:invvar}.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{invvar_plot.png}
+\includegraphics[width=\textwidth]{invvar.png}
 \end{center}
-\caption{The assumed and the actual variances as a function of counts.\label{fig:count_uncertainty}}
+\caption{This plot highlights the difference between the assumed (solid line) and the actual (black points) measurement uncertainty variances.\label{fig:invvar}.}
 \end{figure}
@@
 
-\subsubsection{Instrument Response Refinement: \textbf{\textit{p\_step}}}
+\subsubsection{Instrument Response Refinement (\textit{p\_step})}
 The instrument response parameters can now be refined with the latest star flux estimates.
@@
 In this section we consider four simple, very different, survey strategies.
+We investigate how well the cross-calibration procedure can constrain the relative instrument response with these surveys.
+We label these strategies A to D.
+The different strategies are summarized in Tab.
+\ref{tab:surveys} and are shown in Fig.
+\ref{fig:surveys}.
+Strategy A is the simplest strategy; the field is regularly tiled with a small (XX\%) overlap between adjacent pointings.
+The pointings in the 12 passes over the same field are exactly aligned.
+Survey B is the same as A, but with each pass over the field the orientation of the telescope is rotated by $30^\circ$.
+Survey C is more complex.
+The first pass over the field is the same as in Survey A, with $12 \times 12$ pointings.
+In the next pass, one of the pointings in the $\alpha$ directions is removed and one is added in the $\beta$, so our grid of pointings is now $13 \times 11$.
+The overlap in now XX\% and XX\% in the $\alpha$ and $\beta$ directions respectively.
+In the third pass over the field, the change to the point grid, compared to the first pass, is reversed and we image the field on a $11 \times 13$ grid.
+These three passes are then repeated four times.
+Survey D is quasi-random.
+The orientation of each pointing is random.
+The pointing positions are quasi-random: the pointing center is randomly assigned within a $0.7 \times 0.7$ box whose center lies on the pointing grid in Survey A.
+This is done to ensure a uniform coverage of the field.
 
-\subsection{Surveys Strategies}
-We consider two extreme survey examples.
-The first, called Strategy A, consists of a regular grid of pointings over the sky with a small overlap between neighboring fields.
-Nine passes are performed over the sky and each time the grids are aligned exactly.
-The second survey, Strategy D, has the same number of fields as Strategy A, but the pointing and orientation of these fields are random.
-These two surveys can be seen in Fig.
-\ref{fig:surveys}.
+\begin{table}
+\begin{tabular}{|c|p{5.5cm}|p{3cm}|p{3cm}|}
+\hline \textbf{Survey Name} & \textbf{Pointing Center} & \textbf{Pointing Orientation} & \textbf{Number of Pointings}\\ 
+\hline \textbf{A} & Uniform Grid ($12\times12$) & Constant & 1728\\ 
+\hline \textbf{B} & Uniform Grid ($12\times12$) & 30$^\circ$ Rotation Between Passes & 1728 \\ 
+\hline \textbf{C} & Pass 1:  Uniform Grid ($12\times12$) Pass 2:  Uniform Grid ($1§\times11$) Pass 3:  Uniform Grid ($13\times11$)& Constant & 1720 \\ 
+\hline \textbf{D} & Quasi-random & Random & 1728 \\ 
+\hline 
+\end{tabular}   
+\caption{A summary of the four simple survey strategies investigated.\label{tab:surveys}}
+\end{table}
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{coverage.png}
+\includegraphics[width=\textwidth]{simple_surveys.png}
 \end{center}
-\caption{Left: The A Strategy has regular pointings over the sky with a small overlap between adjacent fields.
-Nine passes are performed over the sky and each time the grids are aligned exactly.
-Right: The D Strategy has the same number of pointings as the A Strategy, but the pointings and orientations are random.\label{fig:surveys}}
+\caption{The four simple survey strategies described in the text and summarized in Tab. \ref{tab:surveys}. Surveys A, B and D have 1728 pointings and survey C has 1720 pointings.\label{fig:surveys}}
 \end{figure}
 
-The coverage of each strategy is shown in Fig. \ref{fig:coverage}.
+\begin{table}
+\begin{tabular}{|c|p{4cm}|c|c|}
+\hline Survey & Number of Iterations Until Converges & Instrument Response Fit Badness & RMS Source Error \\ 
+\hline A & no convergence & - & - \\ 
+\hline B & no converges & - & - \\ 
+\hline C &  &  &  \\ 
+\hline D &  &  &  \\ 
+\hline 
+\end{tabular}
+\caption{The performance of the cross-calibration of the datasets with the survey strategies detailed in Tab. \ref{tab:surveys}\label{fig:simple_results}}
+\end{table}
 
-\subsection{Ubercalibration Performance}
-
-The difference in ubercalibration method's ability to fit the flat field and the star magnitudes for the two different Strategies is striking.
-With the random Strategy (D) we can get very close to the actual flat field in a small number of ubercalibration iterations (see Fig \ref{fig:D}).
-The well ordered Strategy (A), on the other hand, does not converge to the correct flat field even with 1,000 ubercal iterations.
-Even when we start with almost the correct flat-field parameters (those found with the D Strategy - shown in Fig.
-\ref{fig:D}), the ubercalibration procedure does not converge.
-In fact, with each iteration solution drifts away from the actual flat-field.
-Fig \ref{fig:A} shows the fitted flat-field from Strategy A after 1,000 with the initial parameters shown in Fig \ref{fig:D}.
+We can draw a strong conclusion from these simple survey strategies: small and regular overlaps between adjacent pointings is essentially useless for this kind of cross-calibration.
+The simple strategies, where the sky is just tiled, should be avoided in large scale imaging surveys.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{D_15_ff.png}
+\includegraphics[width=\textwidth]{fit_example.png}
 \end{center}
@@
 
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=\textwidth]{A_1000_ff.png}
-\end{center}
-\caption{The fitted flat field compared to the actual flat field calculated found with Strategy A.
-The ubercalibration procedure was started with almost exact flat field parameters (those found from Strategy D - see Fig.
-\ref{fig:D}) and has drifted away from the actual solution.
-\label{fig:A}}
-\end{figure}
-
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=0.47\textwidth]{coverage.png}
-\end{center}
-\caption{The coverage of both of the surveys investigated. \label{fig:coverage}}
-\end{figure}
-
-\section{Discussion}
-small changes to survey that greatly improve coverage plots
-random strategy can be defined so that the telescope does not make large steps in pointing, orientation
-have't addressed time dependence, how do these surveys with time will be different for different assumptions on temporal variability. 
-is the small scale power we put in realistic?
-- survey D is much better than survey C for astrometric self calibration 
-we haven't thought about gaps
 \end{document}

### 2011-07-03 default_parameters.py

@@ -2,7 +2,7 @@
 # -*- coding: utf-8 -*-
 
 dic = {
-  'survey_strategies'     :     ['deep', 'D'], #['D', 'C', 'B', 'A']
+  'survey_strategies'     :     ['D'], #['D', 'C', 'B', 'A']
   
   'mag_at_ten_sigma'    :     22.5,
 
@@ -24,7 +24,7 @@ dic = {
   
   'epsilon_max'         :     1,
   
-  'sky_limits'          :     [-5.0, 5.0, -5.0, 5.0], #[α_min, α_max, β_min, β_max]
+  'sky_limits'          :     [-4.0, 4.0, -4.0, 4.0], #[α_min, α_max, β_min, β_max]
   
   'seed'                :     1 # Random number seed
   

### 2011-07-03 generate_survey.py

@@ -3,13 +3,17 @@
 
 import numpy as np
 import random
-import tparameters as p
+import default_parameters as p
 import matplotlib.pylab as plt
 
-# XX Dangerous hack!!
 pdic = p.dic
 FoV = pdic['FoV']
 sky_limits = pdic['sky_limits']
+
+number_fonts = 20
+label_fontsize = 25
+plt.rc('xtick', labelsize=number_fonts) 
+plt.rc('ytick', labelsize=number_fonts) 
 
 def fp2sky(x, y, pointing, orientation):
   theta = - orientation*np.pi/180 # telescope rotates NOT sky
@@ -25,12 +29,12 @@ def plot_survey(survey, survey_name):
   for image in survey:
     alpha, beta = fp2sky(x,y,image[1:3], image[3])
     plt.plot(alpha,beta,'k-',alpha=0.25)
-  plt.xlabel(r"$\alpha$", fontsize=25)
-  plt.ylabel(r"$\beta$", fontsize=25)
-  plt.title(r"Strategy %s: %i Pointings" % (survey_name, len(survey[:,0])))
+  #plt.title(r"Strategy %s: %i Pointings" % (survey_name, len(survey[:,0])))
+  print ('Survey %s: %d pointings' % (survey_name, len(survey[:,0])))
+  plt.text(sky_limits[0]-1*FoV[0], sky_limits[3]+0.8*FoV[1], ('('+survey_name+')'), fontsize = label_fontsize)
   plt.axis('equal')
-  plt.xlim(sky_limits[0]-FoV[0], sky_limits[1]+FoV[0])
-  plt.ylim(sky_limits[2]-FoV[1], sky_limits[3]+FoV[1])
+  plt.xlim(sky_limits[0]-2*FoV[0], sky_limits[1]+2*FoV[0])
+  plt.ylim(sky_limits[2]-2*FoV[1], sky_limits[3]+2*FoV[1])
   return
 
 def uniform_center_list(nx,ny):
@@ -102,26 +106,36 @@ def generate_random_survey(N):
   return x
 
 if __name__ == "__main__":
-  number_passes = 20
+  plt.clf()
+  number_passes = 12
   plt.figure(figsize=(15,15))
   plt.subplot(221)
   xA = generate_uniform_survey(number_passes)
+  plt.ylabel(r"$\beta$", fontsize= label_fontsize+5)
+  plt.gca().set_xticklabels([])
   plot_survey(xA,'A')
[…]

### 2011-07-03 plot.py

@@ -27,11 +27,11 @@ else:
   sys.exit()
 
 # General Plotting Parameters
-fontsize = 20
-tick_fontsize = 12
-double_fig_width = 14
-single_fig_width = 6
-single_fig_height = 6
+fontsize = 40
+tick_fontsize = 24
+double_fig_width = 28
+single_fig_width = 12
+single_fig_height = 12
 
 
 def plot_flat_fields(params, ff_filename, strategy):
@@ -45,30 +45,42 @@ def plot_flat_fields(params, ff_filename, strategy):
   god_ff = dic['god_ff']
   iteration_number = dic['iteration_number']
   plt.subplot(121)
-  plt.suptitle('Survey %s' % strategy, fontsize = fontsize)
-  plt.title(r"Flat-Fields (True = Black; Fitted = Red) Iteration: %i" % (iteration_number))
-  plt.xlabel(r"$\alpha$")
-  plt.ylabel(r"$\beta$")
+  #plt.suptitle('Survey %s' % strategy, fontsize = fontsize)
+  plt.title(r"Flat-Fields (True = Black; Fitted = Red) Iteration: %i" % (iteration_number), fontsize = fontsize-10)
+  plt.xlabel(r"$\alpha$", fontsize = fontsize)
+  plt.ylabel(r"$\beta$", fontsize = fontsize)
   # Find parameters for contour plot
   god_ff_max = np.max(god_ff)
   god_ff_min = np.min(god_ff)
   # XX magic number
   levels = np.arange(0.5,1.5,0.01)
-  CS = plt.contour(X,Y,god_ff,levels ,colors='k')
-  plt.clabel(CS, fontsize=9, inline=1)
-  CS2 = plt.contour(X,Y,our_ff,levels,colors='r',alpha=0.5)
-
+  CS = plt.contour(X,Y,god_ff,levels ,colors='k', linewidths = 3)
+  plt.clabel(CS, fontsize=tick_fontsize, inline=1)
+  CS2 = plt.contour(X,Y,our_ff,levels,colors='r',alpha=0.5, linewidths = 3)
+  ax = plt.gca()
+  for tick in ax.xaxis.get_major_ticks():
+    tick.label1.set_fontsize(tick_fontsize)
+  for tick in ax.yaxis.get_major_ticks():
+    tick.label1.set_fontsize(tick_fontsize)
[…]

### 2011-07-03 Documentation/cross_cal_paper.tex
euclid: paper update, plotting function update: nicer figures
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@18944 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 
+The difference between the performance of the cross-calibration between Surveys A and B and Surveys C and D is striking.
+With the former, the solution does not converge.
+With the later, a fit to the relative instrument response is found after a small number of iterations XX.
+In both cases, the cross-calibration procedure is able to reproduce the actual instrument response to a high accuracy, with Survey D performing slightly better.
+
 We can draw a strong conclusion from these simple survey strategies: small and regular overlaps between adjacent pointings is essentially useless for this kind of cross-calibration.

### 2011-07-05 Documentation/cross_cal_paper.tex
euclid: paper update
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@18954 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \newcommand{\myemail}{holmes@mpia.de}
+\newcommand{\true}{\mathrm{true}}
 
@@
 This ensures that the differences observed are not a result of random fluctuations.
-We relate source magnitudes, $m$, to source fluxes, $s$, simply by: $m = 22.5 - 2.5\log_{10}(s)$.
-We therefore do not consider the imaging instrument's absolute response.
+We relate source magnitudes, $m$, to source fluxes, $s$, simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies.
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{survey.png}
+\includegraphics[width=\textwidth]{temp.png}
 \end{center}
@@
 \end{center}
-\caption{A single exposure of the sky with the pointing coordinates $(\alpha = 9.2^\circ, \beta = 2.2^\circ)$ and an orientation of $\theta = 175.4^\circ$.
-Left: Plot of the entire fake sky, with the instrument's field-of-view shown in black and the measured sources highlighted in red.
-Right: The distribution of the highlighted sources on the instrument's focal plane.
+\caption{A single exposure of the fake sky.
+Left: Plot of the entire fake sky, with the instrument's field-of-view shown.
+Right: The resultant distribution of the highlighted sources on the instrument's focal plane.
 \label{fig:camera}}
@@
 
-\subsubsection{Catalog to Measured Counts Conversion}
-For a measurement $(i)$ the counts recorded from a source ($k$) depends on true instrument response ($f(\vec{q}_{true})$), which is a function of focal plane position $(\vec{x_i})$, and the source's true flux ($s$):
+\subsubsection{Measurement Model}
+For a measurement $i$ the counts recorded from a source $k$ depends on true instrument response $f_{\true}$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s$:
 
 \begin{displaymath}
-c_i = f(\vec{x_i} | \vec{q}_{true}) . s_{k} + e_{i}
+c_i = f_{\true}(\vec{x_i}) \, s_{k, \true} + e_{i}
 \end{displaymath}
 
-\noindent{}An error, drawn from the Normal Distribution $N(e|0,{\sigma_i}^2)$, is also introduced to the measurement.
+\noindent{}An error, drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$, is also introduced to the measurement.
 In this analysis, we do not consider temporal variations in the instrument's response.
@@
 We assume that the Euclid exposures will be background limited and that, for systematic reasons, we will hit an upper limit on the signal-to-noise for bright sources.
-We assume that the uncertainty variance on the counts are:
+We complicate the model by applying an extra term to the count's uncertainty variance, which we do not take into account in our analysis:
+\begin{displaymath}
+\sigma_{i, \true}^{2} = (1 + \epsilon_k) \, \alpha^{2} + \eta^{2}\, [ f_\true(\vec{x_i}) \, s_{k, \true} ]^2
+\end{displaymath}
+\noindent{}where $\epsilon$ is a random number, in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$.
 
+We do not take into account this final term in our analysis and we therefoer assume that the uncertainty variance on the counts are:
 \begin{displaymath}
-\sigma_{{ass}}^{2} = \alpha^{2} + \eta^{2}\, c^{2} 
+\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2}\, c^{2}_i
 \end{displaymath}
@@
 
-\begin{displaymath}
-\sigma_{{act}}^{2} = (1 + \epsilon) \,\alpha^{2} + \eta^{2}\, c^{2} 
-\end{displaymath}
-\noindent{}where $\epsilon$ is a random number, in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$.
-An example of the assumed compared to the true uncertainty variances is shown in Fig.
-\ref{fig:invvar}.
+An example of the assumed compared to the true uncertainty variances is shown in Fig. \ref{fig:invvar}.
 
@@
 \subsubsection{Instrument Response Model}
-We simulate the instrument response $f(\vec{x_i} | \vec{q}_{true})$ variations on both large and small scales.
+We simulate the instrument response $f_\true(\vec{x_i} | \vec{q}_{\true})$ variations on both large and small scales.
 We model large scale structure, such as the residual from a flat-field calibration source, as a second order polynomial.
 We model small scale variations in the instrument response with XX sine and cosine contributions.
-We intentionally do not for the instrument response with a model of the same form, instead we fit with a XX order polynomial only.
+We intentionally do not for the instrument response with a model of the same form.
 
@@
 \end{displaymath}
-where $s_k$ is the true flux from star $k$ and our error is drawn from the Normal Distribution $e_{i} = N(e|0,{\sigma_i}^2)$.
+where $s_k$ is a model flux for source $k$ and the error $e_i$ is drawn from the Normal Distribution $N(e|0,{\sigma_i}^2)$. We model the instrument response as a XX order polynominal.
 
@@
 \subsubsection{Source Flux Refinement (\textit{s\_step})}
-The flux estimates of each source, $s_k$, are considered individually.
+The model fluxes of each source, $s_k$, are considered individually.
 At each \textit{s\_step} we can refine a sources's flux estimate by minimizing the error function with the latest instrument response parameters:
@@
 
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{density_of_stars.png}
+\end{center}
+
+\end{figure}
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{epsilon_max.png}
+\end{center}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{m_min.png}
+\end{center}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{mag_at_10_sigma.png}
+\end{center}
+\end{figure}
+
 \end{document}

### 2011-07-05 plot.py

@@ -105,8 +105,8 @@ def camera_image(params, camera_filename):
   #title = ur'$\alpha$ = %.1lf, $\beta$ = %.1lf, $\theta$ = %.1lf$^\circ$' % (pointing[0],pointing[1], orientation)
   #plt.suptitle(title, fontsize=fontsize)
   plt.subplot(121)
-  plt.plot(alpha,beta,'.', markersize=5)
-  plt.plot(alpha[inside_FoV],beta[inside_FoV],'r.',markersize=5)  
+  plt.plot(alpha,beta,'k.', markersize=5)
+  plt.plot(alpha[inside_FoV],beta[inside_FoV],'k.',markersize=5)  
   plt.plot(fp_alpha,fp_beta,'k', linewidth=4)
   plt.xlabel(ur'$\alpha$', fontsize=fontsize)
   plt.ylabel(ur'$\beta$', fontsize=fontsize)
@@ -123,7 +123,7 @@ def camera_image(params, camera_filename):
 
   # Plot sources on focal plane
   plt.subplot(122)
-  plt.plot(x,y,'r.', markersize=20)
+  plt.plot(x,y,'k.', markersize=20)
   plt.plot(fp_x, fp_y, 'k', linewidth=6)
   plt.xlabel(ur'$x$', fontsize=fontsize)
   plt.ylabel(ur'$y$', fontsize=fontsize)

### 2011-07-21 functions.py
euclid: clean directory
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19088 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -219,7 +219,7 @@ def q_step(params, obs_cat, s, order, iteration_number, plots=None):
 #*****************************************************
 
 def compare_flats(a, Z_true, g, X, Y):
-  error = np.sqrt(np.mean((Z_true - np.dot(g,a))**2))
+  error = (np.sum((Z_true - np.dot(g,a))**2))
   #print error
   return error
 
@@ -240,7 +240,7 @@ def bestfit_ff(params, out_dir):
   a[5] = 0.5
   print "Fitting god's flat-field with basis..."
   fitted_parameters = sci.fmin_bfgs(compare_flats, a, args = (god_ff, g, x, y),
-gtol = 1e-16, maxiter = 1e16)
+gtol = 0, maxiter = 1e16)
   print "Fitted Parameters: ", fitted_parameters
   "...done!"
   fitted_parameters = normalize_flat_field(params, fitted_parameters)

### 2011-07-24 Documentation/cross_cal_paper.tex
euclid: clean directory
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19088 03a3cea6-2c03-0410-ac9b-9271b1c66c29
euclid: correct typo in paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19132 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=\textwidth]{temp.png}
-\end{center}
-\caption{An example catalog of fake sources used within the simulations. The sources are shown at their position on a portion of the sky (left) and as a histogram of their magnitudes (right).
-\label{fig:sky}}
-\end{figure}
+%\begin{figure}[ht]
+%\begin{center}
+%\includegraphics[width=\textwidth]{temp.png}
+%\end{center}
+%\caption{An example catalog of fake sources used within the simulations. The sources are shown at their position on a portion of the sky (left) and as a histogram of their magnitudes (right).
+%\label{fig:sky}}
+%\end{figure}
 
@@
 
-We do not take into account this final term in our analysis and we therefoer assume that the uncertainty variance on the counts are:
+We do not take into account this final term in our analysis and we therefore assume that the uncertainty variance on the counts are:
 \begin{displaymath}

### 2011-10-11 paper/self_calibration_I.tex
euclid: begin paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19733 03a3cea6-2c03-0410-ac9b-9271b1c66c29

\documentclass[manuscript]{aastex}
\newcommand{\vdag}{(v)^\dagger}

\shorttitle{Optimizing Large Scale Imaging Surveys for Relative Photometric Calibration}
\shortauthors{Holmes et al.}

%% This is the end of the preamble.  Indicate the beginning of the
%% paper itself with \begin{document}.

\begin{document}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire.

\title{Optimizing Large Scale Imaging Surveys for Relative Photometric Calibration, \\
    the Euclid Dark Energy Mission as an Example}

%% Use \author, \affil, and the \and command to format
%% author and affiliation information.
%% Note that \email has replaced the old \authoremail command
%% from AASTeX v4.0. You can use \email to mark an email address
%% anywhere in the paper, not just in the front matter.
%% As in the title, use \\ to force line breaks.

\author{R. Holmes\altaffilmark{1} and H.-W. Rix\altaffilmark{1}}
\affil{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, Heidelberg, 69117, Germany.}

\and

\author{D. W. Hogg\altaffilmark{2}}
\affil{2 Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA.}

\begin{abstract}
In this paper we show that, with due care given to the observing strategy, it will be possible to accurately constrain the relative photometric calibration of instruments used in large imaging surveys with the science data alone. We create end-to-end simulations of an imaging survey, which produces simulated datasets from mock observations of a synthetic sky -- with realistic measurement uncertainties and a complex instrument response -- according to defined survey strategies. We then use a self-calibration technique to recover the relative instrument response by fitting an model that best explains the survey dataset, based on the multiple observations of (non-varying) sources at different focal plane positions. By considering four simple survey strategies we find that, with a correct redundancy built into the survey strategy, it is possible to accurately constrain the relative photometric response of an imaging instrument, and therefore the relative calibration of the resulting dataset. The majority of the remaining post self-calibration errors are due to the limitations in the basis used to model the relative instrument response. We find that returning the same sources to very different focal plane positions is the key property of a survey strategy that is required for an accurate calibration. From the results of this study, we are able to highlight an important point for those considering the design of large scale imaging surveys: depart from a regular tiling of the sky and return the same sources to very different focal plane positions. 
\end{abstract}

\keywords{Relative Photometric Calibration: Imaging Survey: Euclid}

%% From the front matter, we move on to the body of the paper.
%% In the first two sections, notice the use of the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below. We have
%% chosen the first three characters of the first author's name plus
%% the last two numeral of the year of publication as our KEY for
%% each reference.

%% Authors who wish to have the most important objects in their paper
%% linked in the electronic edition to a data center may do so by tagging
%% their objects with \objectname{} or \object{}.  Each macro takes the
%% object name as its required argument. The optional, square-bracket 
%% argument should be used in cases where the data center identification
%% differs from what is to be printed in the paper.  The text appearing 
%% in curly braces is what will appear in print in the published paper. 
%% If the object name is recognized by the data centers, it will be linked
%% in the electronic edition to the object data available at the data centers  
%%
%% Note that for sources with brackets in their names, e.g. [WEG2004] 14h-090,
%% the brackets must be escaped with backslashes when used in the first
%% square-bracket argument, for instance, \object[\[WEG2004\] 14h-090]{90}).
%%  Otherwise, LaTeX will issue an error.

\section{Introduction}
Astronomers tend to think in terms of taking science data and calibration data separately. The former is used for science and the latter is used to set instrument parameters like flat-field, bias and so-on. But typically far more photons are collected during science exposures; are these not incredibly constraining on the calibration themselves? Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (XX ref) (SDSS) much more calibration information was obtained from the science data than the calibration data. But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: good redundancy is required in the data stream, and a redundancy of a very specific kind.

In this paper we will show that this distinction between science and calibration data no longer applies in the large survey mission, both ground and space based, currently being planned and developed. We show that, as long as the survey strategy is designed appropriately, the calibration information within the science data can be used to accurately constrain the relative photometric response of an imaging instrument. This retrospective, cross-calibration of the dataset is a very powerful calibration method; future large scale imaging surveys should be optimized with this in mind.

We concentrate on catalog level simulations of different survey strategies and assess how well the resultant dataset can be used to constrain the relative response of the imaging instrument it came from. Although we do not model the imaging instrument on a pixel scale, we do include realistic measurement uncertainties within the simulations.

Although the results we present here are general to all large scale imaging surveys, in Section XX we focus on the proposed Euclid Deep Field Survey. This survey has been optimized, within other constraints, based on the results from this work.

\section{Observations}

%% In a manner similar to \objectname authors can provide links to dataset
%% hosted at participating data centers via the \dataset{} command.  The
%% second curly bracket argument is printed in the text while the first
%% parentheses argument serves as the valid data set identifier.  Large
%% lists of data set are best provided in a table (see Table 3 for an example).
%% Valid data set identifiers should be obtained from the data center that
%% is currently hosting the data.
%%
%% Note that AASTeX interprets everything between the curly braces in the 
%% macro as regular text, so any special characters, e.g. "#" or "_," must be 
%% preceded by a backslash. Otherwise, you will get a LaTeX error when you 
%% compile your manuscript.  Special characters do not 
%% need to be escaped in the optional, square-bracket argument.

All our observations were short direct exposures with CCD's.  We also have
a random {\it Chandra} data set \dataset{ADS/Sa.ASCA\#X/86008020} and a
neat \dataset[ADS/Sa.HST#Y0Q70101T]{HST FOS spectrum} that readers can access
via the links in the electronic edition.  Unfortunately this has nothing 
whatsoever to do with this research.  At Lick Observatory we used a TI 500$\times$500 
chip and a GEC 575$\times$385, on the 1-m Nickel reflector.  The only
filter available at Lick was red.  At CTIO we used a GEC 575$\times$385, with
$B, V,$ and $R$ filters, and an RCA 512$\times$320, with $U, B, V, R,$ and $I$
filters, on the 1.5-m reflector. In the CTIO observations we tried to
concentrate on the shortest practicable wavelengths; but faintness, reddening,
and poor short-wavelength sensitivity often kept us from observing in $U$ or
even in $B$. All four cameras had scales of the order of 0.4 arcsec/pixel, and
our field sizes were around 3 arcmin.

The CCD images are unfortunately not always suitable, for very poor
clusters or for clusters with large cores.  Since the latter are easily
studied by other means, we augmented our own CCD profiles by collecting
from the literature a number of star-count profiles 
\citep{kin68,pet76,har84,ort85}, as well as photoelectric profiles 
\citep{kin66,kin75} and electronographic profiles \citep{kro84}.
In a few cases we judged normality by eye estimates on one of the Sky
Surveys.

%% In this section, we use  the \subsection command to set off
%% a subsection.  \footnote is used to insert a footnote to the text.

%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.

%% This section also includes several of the displayed math environments
%% mentioned in the Author Guide.

\section{Helicity Amplitudes}

It has been realized that helicity amplitudes provide a convenient means
for Feynman diagram\footnote{Footnotes can be inserted like this.}
evaluations.  These amplitude-level techniques
are particularly convenient for calculations involving many Feynman
diagrams, where the usual trace techniques for the amplitude
squared becomes unwieldy.  Our calculations use the helicity techniques
developed by other authors \cite[]{hag86}; we briefly summarize below.

\subsection{Formalism} \label{bozomath}

%% The equation environment wil produce a numbered display equation.

A tree-level amplitude in $e^+e^-$ collisions can be expressed in
terms of fermion strings of the form
\begin{equation}
\bar v(p_2,\sigma_2)P_{-\tau}\hat a_1\hat a_2\cdots
\hat a_nu(p_1,\sigma_1) ,
\end{equation}
where $p$ and $\sigma$ label the initial $e^{\pm}$ four-momenta
and helicities $(\sigma = \pm 1)$, $\hat a_i=a^\mu_i\gamma_\nu$
and $P_\tau=\frac{1}{2}(1+\tau\gamma_5)$ is a chirality projection
operator $(\tau = \pm1)$.  The $a^\mu_i$ may be formed from particle
four-momenta, gauge-boson polarization vectors or fermion strings with
an uncontracted Lorentz index associated with final-state fermions.

%% The \notetoeditor{TEXT} command allows the author to communicate
%% information to the copy editor.  This information will appear as a
%% footnote on the printed copy for the manuscript style file.  Nothing will
%% appear on the printed copy if the preprint or
%% preprint2 style files are used.

%% The eqnarray environment produces multi-line display math. The end of
%% each line is marked with a \\. Lines will be numbered unless the \\
%% is preceded by a \nonumber command.
%% Alignment points are marked by ampersands (&). There should be two
%% ampersands (&) per line.

In the chiral \notetoeditor{Figures 1 and 2 should appear side-by-side in
print} representation the $\gamma$ matrices are expressed
in terms of $2\times 2$ Pauli matrices $\sigma$ and the unit matrix 1 as
\begin{eqnarray}
\gamma^\mu  & = &
 \left(
\begin{array}{cc}
0 & \sigma^\mu_+ \\
\sigma^\mu_- & 0
\end{array}     \right) ,
 \gamma^5= \left(
\begin{array}{cc}
-1 &   0\\
0 &   1
\end{array}     \right)  , \nonumber \\
\sigma^\mu_{\pm}  & = &   ({\bf 1} ,\pm \sigma) , \nonumber
\end{eqnarray}
giving
\begin{equation}
\hat a= \left(
\begin{array}{cc}
0 & (\hat a)_+\\
(\hat a)_- & 0
\end{array}\right), (\hat a)_\pm=a_\mu\sigma^\mu_\pm ,
\end{equation}
The spinors are expressed in terms of two-component Weyl spinors as
\begin{equation}
u=\left(
\begin{array}{c}
(u)_-\\
(u)_+
\end{array}\right), v={\bf (}\vdag_+{\bf ,}   \vdag_-{\bf )} .
\end{equation}

%% Putting eqnarrays or equations inside the mathletters environment groups
%% the enclosed equations by letter. For instance, the eqnarray below, instead
%% of being numbered, say, (4) and (5), would be numbered (4a) and (4b).
%% LaTeX the paper and look at the output to see the results.

The Weyl spinors are given in terms of helicity eigenstates
$\chi_\lambda(p)$ with $\lambda=\pm1$ by
\begin{mathletters}
\begin{eqnarray}
u(p,\lambda)_\pm & = & (E\pm\lambda|{\bf p}|)^{1/2}\chi_\lambda(p) , \\
v(p,\lambda)_\pm & = & \pm\lambda(E\mp\lambda|{\bf p}|)^{1/2}\chi
_{-\lambda}(p)
\end{eqnarray}
\end{mathletters}

%% This section contains more display math examples, including unnumbered
%% equations (displaymath environment). The last paragraph includes some
%% examples of in-line math featuring a couple of the AASTeX symbol macros.

\section{Floating material and so forth}

%% The displaymath environment will produce the same sort of equation as
%% the equation environment, except that the equation will not be numbered
%% by LaTeX.

Consider a task that computes profile parameters for a modified
Lorentzian of the form
\begin{equation}
I = \frac{1}{1 + d_{1}^{P (1 + d_{2} )}}
\end{equation}
where
\begin{displaymath}
d_{1} = \sqrt{ \left( \begin{array}{c} \frac{x_{1}}{R_{maj}}
\end{array} \right) ^{2} +
\left( \begin{array}{c} \frac{y_{1}}{R_{min}} \end{array} \right) ^{2} }
\end{displaymath}
\begin{displaymath}
d_{2} = \sqrt{ \left( \begin{array}{c} \frac{x_{1}}{P R_{maj}}
\end{array} \right) ^{2} +
\left( \begin{array}{c} \case{y_{1}}{P R_{min}} \end{array} \right) ^{2} }
\end{displaymath}
\begin{displaymath}
x_{1} = (x - x_{0}) \cos \Theta + (y - y_{0}) \sin \Theta
\end{displaymath}
\begin{displaymath}
y_{1} = -(x - x_{0}) \sin \Theta + (y - y_{0}) \cos \Theta
\end{displaymath}

In these expressions $x_{0}$,$y_{0}$ is the star center, and $\Theta$ is the
angle with the $x$ axis.  Results of this task are shown in table~\ref{tbl-1}.
It is not clear how these sorts of analyses may affect determination of
 $M_{\sun}$, but the assumption is that the alternate results
should be less than 90\degr\ out of phase with previous values.
We have no observations of \ion{Ca}{2}.
Roughly \slantfrac{4}{5} of the electronically submitted abstracts
for AAS meetings are error-free.

%% If you wish to include an acknowledgments section in your paper,
%% separate it off from the body of the text using the \acknowledgments
%% command.

%% Included in this acknowledgments section are examples of the
%% AASTeX hypertext markup commands. Use \url without the optional [HREF]
%% argument when you want to print the url directly in the text. Otherwise,
%% use either \url or \anchor, with the HREF as the first argument and the
%% text to be printed in the second.

\acknowledgments

We are grateful to V. Barger, T. Han, and R. J. N. Phillips for
doing the math in section~\ref{bozomath}.
More information on the AASTeX macros package is available \\ at
\url{http://www.aas.org/publications/aastex}.
For technical support, please write to
\email{aastex-help@aas.org}.

%% To help institutions obtain information on the effectiveness of their
%% telescopes, the AAS Journals has created a group of keywords for telescope
%% facilities. A common set of keywords will make these types of searches
%% significantly easier and more accurate. In addition, they will also be
%% useful in linking papers together which utilize the same telescopes
%% within the framework of the National Virtual Observatory.
%% See the AASTeX Web site at http://aastex.aas.org/
%% for information on obtaining the facility keywords.

%% After the acknowledgments section, use the following syntax and the
%% \facility{} macro to list the keywords of facilities used in the research
%% for the paper.  Each keyword will be checked against the master list during
%% copy editing.  Individual instruments or configurations can be provided 
%% in parentheses, after the keyword, but they will not be verified.

{\it Facilities:} \facility{Nickel}, \facility{HST (STIS)}, \facility{CXO (ASIS)}.

%% Appendix material should be preceded with a single \appendix command.
%% There should be a \section command for each appendix. Mark appendix
%% subsections with the same markup you use in the main body of the paper.

%% Each Appendix (indicated with \section) will be lettered A, B, C, etc.
%% The equation counter will reset when it encounters the \appendix
%% command and will number appendix equations (A1), (A2), etc.

\appendix

\section{Appendix material}

Consider once again a task that computes profile parameters for a modified
Lorentzian of the form
\begin{equation}
I = \frac{1}{1 + d_{1}^{P (1 + d_{2} )}}
\end{equation}
where
\begin{mathletters}
\begin{displaymath}
d_{1} = \frac{3}{4} \sqrt{ \left( \begin{array}{c} \frac{x_{1}}{R_{maj}}
\end{array} \right) ^{2} +
\left( \begin{array}{c} \frac{y_{1}}{R_{min}} \end{array} \right) ^{2} }
\end{displaymath}
\begin{equation}
d_{2} = \case{3}{4} \sqrt{ \left( \begin{array}{c} \frac{x_{1}}{P R_{maj}}
\end{array} \right) ^{2} +
\left( \begin{array}{c} \case{y_{1}}{P R_{min}} \end{array} \right) ^{2} }
\end{equation}
\begin{eqnarray}
x_{1} & = & (x - x_{0}) \cos \Theta + (y - y_{0}) \sin \Theta \\
y_{1} & = & -(x - x_{0}) \sin \Theta + (y - y_{0}) \cos \Theta
\end{eqnarray}
\end{mathletters}

For completeness, here is one last equation.
\begin{equation}
e = mc^2
\end{equation}

%% The reference list follows the main body and any appendices.
%% Use LaTeX's thebibliography environment to mark up your reference list.
%% Note \begin{thebibliography} is followed by an empty set of
%% curly braces.  If you forget this, LaTeX will generate the error
%% "Perhaps a missing \item?".
%%
%% thebibliography produces citations in the text using \bibitem-\cite
%% cross-referencing. Each reference is preceded by a
%% \bibitem command that defines in curly braces the KEY that corresponds
%% to the KEY in the \cite commands (see the first section above).
%% Make sure that you provide a unique KEY for every \bibitem or else the
%% paper will not LaTeX. The square brackets should contain
%% the citation text that LaTeX will insert in
%% place of the \cite commands.

%% We have used macros to produce journal name abbreviations.
%% AASTeX provides a number of these for the more frequently-cited journals.
%% See the Author Guide for a list of them.

%% Note that the style of the \bibitem labels (in []) is slightly
%% different from previous examples.  The natbib system solves a host
%% of citation expression problems, but it is necessary to clearly
%% delimit the year from the author name used in the citation.
%% See the natbib documentation for more details and options.

\begin{thebibliography}{}
\bibitem[Auri\`ere(1982)]{aur82} Auri\`ere, M.  1982, \aap,
    109, 301
\bibitem[Canizares et al.(1978)]{can78} Canizares, C. R.,
    Grindlay, J. E., Hiltner, W. A., Liller, W., \&
    McClintock, J. E.  1978, \apj, 224, 39
\bibitem[Djorgovski \& King(1984)]{djo84} Djorgovski, S.,
    \& King, I. R.  1984, \apjl, 277, L49
\bibitem[Hagiwara \& Zeppenfeld(1986)]{hag86} Hagiwara, K., \&
    Zeppenfeld, D.  1986, Nucl.Phys., 274, 1
\bibitem[Harris \& van den Bergh(1984)]{har84} Harris, W. E.,
    \& van den Bergh, S.  1984, \aj, 89, 1816
\bibitem[H\`enon(1961)]{hen61} H\'enon, M.  1961, Ann.d'Ap., 24, 369
\bibitem[Heiles \& Troland(2003)]{heiles03} Heiles, C. \& Troland, T. H., 2003, \apjs, preprint doi:10.1086/381753
\bibitem[Kim, Ostricker, \& Stone(2003)]{kim03} Kim, W.-T.,  Ostriker, E., \& Stone, J. M., 2003, \apj, 599, 1157
\bibitem[King(1966)]{kin66}  King, I. R.  1966, \aj, 71, 276
\bibitem[King(1975)]{kin75}  King, I. R.  1975, Dynamics of
    Stellar Systems, A. Hayli, Dordrecht: Reidel, 1975, 99
\bibitem[King et al.(1968)]{kin68}  King, I. R., Hedemann, E.,
    Hodge, S. M., \& White, R. E.  1968, \aj, 73, 456
\bibitem[Kron et al.(1984)]{kro84} Kron, G. E., Hewitt, A. V.,
    \& Wasserman, L. H.  1984, \pasp, 96, 198
\bibitem[Lynden-Bell \& Wood(1968)]{lyn68} Lynden-Bell, D.,
    \& Wood, R.  1968, \mnras, 138, 495
\bibitem[Newell \& O'Neil(1978)]{new78} Newell, E. B.,
    \& O'Neil, E. J.  1978, \apjs, 37, 27
\bibitem[Ortolani et al.(1985)]{ort85} Ortolani, S., Rosino, L.,
    \& Sandage, A.  1985, \aj, 90, 473
\bibitem[Peterson(1976)]{pet76} Peterson, C. J.  1976, \aj, 81, 617
\bibitem[Rudnick et al.(2003)]{rudnick03} Rudnick, G. et al., 2003, \apj, 599, 847
\bibitem[Spitzer(1985)]{spi85} Spitzer, L.  1985, Dynamics of
    Star Clusters, J. Goodman \& P. Hut, Dordrecht: Reidel, 109
\bibitem[Treu et al.(2003)]{treu03} Treu, T. et al., 2003, \apj, 591, 53
\end{thebibliography}

\clearpage

%% Use the figure environment and \plotone or \plottwo to include
%% figures and captions in your electronic submission.
%% To embed the sample graphics in
%% the file, uncomment the \plotone, \plottwo, and
%% \includegraphics commands
%%
%% If you need a layout that cannot be achieved with \plotone or
%% \plottwo, you can invoke the graphicx package directly with the
%% \includegraphics command or use \plotfiddle. For more information,
%% please see the tutorial on "Using Electronic Art with AASTeX" in the
%% documentation section at the AASTeX Web site, http://aastex.aas.org/
%%
%% The examples below also include sample markup for submission of
%% supplemental electronic materials. As always, be sure to check
%% the instructions to authors for the journal you are submitting to
%% for specific submissions guidelines as they vary from
%% journal to journal.

%% This example uses \plotone to include an EPS file scaled to
%% 80% of its natural size with \epsscale. Its caption
%% has been written to indicate that additional figure parts will be
%% available in the electronic journal.

\begin{figure}
\epsscale{.80}
%\plotone{f1.eps}
\caption{Derived spectra for 3C138 \citep[see][]{heiles03}. Plots for all sources are available
in the electronic edition of {\it The Astrophysical Journal}.\label{fig1}}
\end{figure}

\clearpage

%% Here we use \plottwo to present two versions of the same figure,
%% one in black and white for print the other in RGB color
%% for online presentation. Note that the caption indicates
%% that a color version of the figure will be available online.
%%

\begin{figure}
%\plottwo{f2.eps}{f2_color.eps}
\caption{A panel taken from Figure 2 of \citet{rudnick03}. 
See the electronic edition of the Journal for a color version 
of this figure.\label{fig2}}
\end{figure}

%% This figure uses \includegraphics to scale and rotate the still frame
%% for an mpeg animation.

\begin{figure}
%\includegraphics[angle=90,scale=.50]{f3.eps}
\caption{Animation still frame taken from \citet{kim03}.
This figure is also available as an mpeg
animation in the electronic edition of the
{\it Astrophysical Journal}.}
\end{figure}

%% If you are not including electonic art with your submission, you may
%% mark up your captions using the \figcaption command. See the
%% User Guide for details.
%%
%% No more than seven \figcaption commands are allowed per page,
%% so if you have more than seven captions, insert a \clearpage
%% after every seventh one.

%% Tables should be submitted one per page, so put a \clearpage before
%% each one.

%% Two options are available to the author for producing tables:  the
%% deluxetable environment provided by the AASTeX package or the LaTeX
%% table environment.  Use of deluxetable is preferred.
%%

%% Three table samples follow, two marked up in the deluxetable environment,
%% one marked up as a LaTeX table.

%% In this first example, note that the \tabletypesize{}
%% command has been used to reduce the font size of the table.
%% We also use the \rotate command to rotate the table to
%% landscape orientation since it is very wide even at the
%% reduced font size.
%%
%% Note also that the \label command needs to be placed
%% inside the \tablecaption.

%% This table also includes a table comment indicating that the full
%% version will be available in machine-readable format in the electronic
%% edition.

\clearpage

\begin{deluxetable}{ccrrrrrrrrcrl}
\tabletypesize{\scriptsize}
\rotate
\tablecaption{Sample table taken from \citet{treu03}\label{tbl-1}}
\tablewidth{0pt}
\tablehead{
\colhead{POS} & \colhead{chip} & \colhead{ID} & \colhead{X} & \colhead{Y} &
\colhead{RA} & \colhead{DEC} & \colhead{IAU$\pm$ $\delta$ IAU} &
\colhead{IAP1$\pm$ $\delta$ IAP1} & \colhead{IAP2 $\pm$ $\delta$ IAP2} &
\colhead{star} & \colhead{E} & \colhead{Comment}
}
\startdata
0 & 2 & 1 & 1370.99 & 57.35    &   6.651120 &  17.131149 & 21.344$\pm$0.006  & 2
4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2
3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2
4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2
5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2
3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2
6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2
5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 & 22.189$\pm$0.011  & 2
4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 & edge \\
\enddata
%% Text for table notes should follow after the \enddata but before
%% the \end{deluxetable}. Make sure there is at least one \tablenotemark
%% in the table for each \tablenotetext.
\tablecomments{Table \ref{tbl-1} is published in its entirety in the 
electronic edition of the {\it Astrophysical Journal}.  A portion is 
shown here for guidance regarding its form and content.}
\tablenotetext{a}{Sample footnote for table~\ref{tbl-1} that was generated
with the deluxetable environment}
\tablenotetext{b}{Another sample footnote for table~\ref{tbl-1}}
\end{deluxetable}

%% If you use the table environment, please indicate horizontal rules using
%% \tableline, not \hline.
%% Do not put multiple tabular environments within a single table.
%% The optional \label should appear inside the \caption command.

\clearpage

\begin{table}
\begin{center}
\caption{More terribly relevant tabular information.\label{tbl-2}}
\begin{tabular}{crrrrrrrrrrr}
\tableline\tableline
Star & Height & $d_{x}$ & $d_{y}$ & $n$ & $\chi^2$ & $R_{maj}$ & $R_{min}$ &
\multicolumn{1}{c}{$P$\tablenotemark{a}} & $P R_{maj}$ & $P R_{min}$ &
\multicolumn{1}{c}{$\Theta$\tablenotemark{b}} \\
\tableline
1 &33472.5 &-0.1 &0.4  &53 &27.4 &2.065  &1.940 &3.900 &68.3 &116.2 &-27.639\\
2 &27802.4 &-0.3 &-0.2 &60 &3.7  &1.628  &1.510 &2.156 &6.8  &7.5 &-26.764\\
3 &29210.6 &0.9  &0.3  &60 &3.4  &1.622  &1.551 &2.159 &6.7  &7.3 &-40.272\\
4 &32733.8 &-1.2\tablenotemark{c} &-0.5 &41 &54.8 &2.282  &2.156 &4.313 &117.4 &78.2 &-35.847\\
5 & 9607.4 &-0.4 &-0.4 &60 &1.4  &1.669\tablenotemark{c}  &1.574 &2.343 &8.0  &8.9 &-33.417\\
6 &31638.6 &1.6  &0.1  &39 &315.2 & 3.433 &3.075 &7.488 &92.1 &25.3 &-12.052\\
\tableline
\end{tabular}
%% Any table notes must follow the \end{tabular} command.
\tablenotetext{a}{Sample footnote for table~\ref{tbl-2} that was
generated with the \LaTeX\ table environment}
\tablenotetext{b}{Yet another sample footnote for table~\ref{tbl-2}}
\tablenotetext{c}{Another sample footnote for table~\ref{tbl-2}}
\tablecomments{We can also attach a long-ish paragraph of explanatory
material to a table.}
\end{center}
\end{table}

%% If the table is more than one page long, the width of the table can vary
%% from page to page when the default \tablewidth is used, as below.  The
%% individual table widths for each page will be written to the log file; a
%% maximum tablewidth for the table can be computed from these values.
%% The \tablewidth argument can then be reset and the file reprocessed, so
%% that the table is of uniform width throughout. Try getting the widths
%% from the log file and changing the \tablewidth parameter to see how
%% adjusting this value affects table formatting.

%% The \dataset{} macro has also been applied to a few of the objects to
%% show how many observations can be tagged in a table.

\clearpage

\begin{deluxetable}{lrrrrcrrrrr}
\tablewidth{0pt}
\tablecaption{Literature Data for Program Stars}
\tablehead{
\colhead{Star}           & \colhead{V}      &
\colhead{b$-$y}          & \colhead{m$_1$}  &
\colhead{c$_1$}          & \colhead{ref}    &
\colhead{T$_{\rm eff}$}  & \colhead{log g}  &
\colhead{v$_{\rm turb}$} & \colhead{[Fe/H]} &
\colhead{ref}}
\startdata
HD 97 & 9.7& 0.51& 0.15& 0.35& 2 & \nodata & \nodata & \nodata & $-1.50$ & 2 \\
& & & & & & 5015 & \nodata & \nodata & $-1.50$ & 10 \\
\dataset[ADS/Sa.HST#O6H04VAXQ]{HD 2665} & 7.7& 0.54& 0.09& 0.34& 2 & \nodata & \nodata & \nodata & $-2.30$ & 2 \\
& & & & & & 5000 & 2.50 & 2.4 & $-1.99$ & 5 \\
& & & & & & 5120 & 3.00 & 2.0 & $-1.69$ & 7 \\
& & & & & & 4980 & \nodata & \nodata & $-2.05$ & 10 \\
HD 4306 & 9.0& 0.52& 0.05& 0.35& 20, 2& \nodata & \nodata & \nodata & $-2.70$ & 2 \\
& & & & & & 5000 & 1.75 & 2.0 & $-2.70$ & 13 \\
& & & & & & 5000 & 1.50 & 1.8 & $-2.65$ & 14 \\
& & & & & & 4950 & 2.10 & 2.0 & $-2.92$ & 8 \\
& & & & & & 5000 & 2.25 & 2.0 & $-2.83$ & 18 \\
& & & & & & \nodata & \nodata & \nodata & $-2.80$ & 21 \\
& & & & & & 4930 & \nodata & \nodata & $-2.45$ & 10 \\
HD 5426 & 9.6& 0.50& 0.08& 0.34& 2 & \nodata & \nodata & \nodata & $-2.30$ & 2 \\
\dataset[ADS/Sa.HST#O5F654010]{HD 6755} & 7.7& 0.49& 0.12& 0.28& 20, 2& \nodata & \nodata & \nodata & $-1.70$ & 2 \\
& & & & & & 5200 & 2.50 & 2.4 & $-1.56$ & 5 \\
& & & & & & 5260 & 3.00 & 2.7 & $-1.67$ & 7 \\
& & & & & & \nodata & \nodata & \nodata & $-1.58$ & 21 \\
& & & & & & 5200 & \nodata & \nodata & $-1.80$ & 10 \\
& & & & & & 4600 & \nodata & \nodata & $-2.75$ & 10 \\
\dataset[ADS/Sa.HST#O56D06010]{HD 94028} & 8.2& 0.34& 0.08& 0.25& 20 & 5795 & 4.00 & \nodata & $-1.70$ & 22 \\
& & & & & & 5860 & \nodata & \nodata & $-1.70$ & 4 \\
& & & & & & 5910 & 3.80 & \nodata & $-1.76$ & 15 \\
& & & & & & 5800 & \nodata & \nodata & $-1.67$ & 17 \\
& & & & & & 5902 & \nodata & \nodata & $-1.50$ & 11 \\
& & & & & & 5900 & \nodata & \nodata & $-1.57$ & 3 \\
& & & & & & \nodata & \nodata & \nodata & $-1.32$ & 21 \\
HD 97916 & 9.2& 0.29& 0.10& 0.41& 20 & 6125 & 4.00 & \nodata & $-1.10$ & 22 \\
& & & & & & 6160 & \nodata & \nodata & $-1.39$ & 3 \\
& & & & & & 6240 & 3.70 & \nodata & $-1.28$ & 15 \\
& & & & & & 5950 & \nodata & \nodata & $-1.50$ & 17 \\
& & & & & & 6204 & \nodata & \nodata & $-1.36$ & 11 \\
\cutinhead{This is a cut-in head}
+26\arcdeg2606& 9.7&0.34&0.05&0.28&20,11& 5980 & \nodata & \nodata &$<-2.20$ & 19 \\
& & & & & & 5950 & \nodata & \nodata & $-2.89$ & 24 \\
+26\arcdeg3578& 9.4&0.31&0.05&0.37&20,11& 5830 & \nodata & \nodata & $-2.60$ & 4 \\
& & & & & & 5800 & \nodata & \nodata & $-2.62$ & 17 \\
& & & & & & 6177 & \nodata & \nodata & $-2.51$ & 11 \\
& & & & & & 6000 & 3.25 & \nodata & $-2.20$ & 22 \\
& & & & & & 6140 & 3.50 & \nodata & $-2.57$ & 15 \\
+30\arcdeg2611& 9.2&0.82&0.33&0.55& 2 & \nodata & \nodata & \nodata & $-1.70$ & 2 \\
& & & & & & 4400 & 1.80 & \nodata & $-1.70$ & 12 \\
& & & & & & 4400 & 0.90 & 1.7 & $-1.20$ & 14 \\
& & & & & & 4260 & \nodata & \nodata & $-1.55$ & 10 \\
+37\arcdeg1458& 8.9&0.44&0.07&0.22&20,11& 5296 & \nodata & \nodata & $-2.39$ & 11 \\
& & & & & & 5420 & \nodata & \nodata & $-2.43$ & 3 \\
+58\arcdeg1218&10.0&0.51&0.03&0.36& 2 & \nodata & \nodata & \nodata & $-2.80$ & 2 \\
& & & & & & 5000 & 1.10 & 2.2 & $-2.71$ & 14 \\
& & & & & & 5000 & 2.20 & 1.8 & $-2.46$ & 5 \\
& & & & & & 4980 & \nodata & \nodata & $-2.55$ & 10 \\
+72\arcdeg0094&10.2&0.31&0.09&0.26&12 & 6160 & \nodata & \nodata & $-1.80$ & 19 \\
\sidehead{I'm a side head:}
G5--36 & 10.8& 0.40& 0.07& 0.28& 20 & \nodata & \nodata & \nodata & $-1.19$ & 21 \\
G18--54 & 10.7& 0.37& 0.08& 0.28& 20 & \nodata & \nodata & \nodata & $-1.34$ & 21 \\
G20--08 & 9.9& 0.36& 0.05& 0.25& 20,11& 5849 & \nodata & \nodata & $-2.59$ & 11 \\
& & & & & & \nodata & \nodata & \nodata & $-2.03$ & 21 \\
G20--15 & 10.6& 0.45& 0.03& 0.27& 20,11& 5657 & \nodata & \nodata & $-2.00$ & 11 \\
& & & & & & 6020 & \nodata & \nodata & $-1.56$ & 3 \\
& & & & & & \nodata & \nodata & \nodata & $-1.58$ & 21 \\
G21--22 & 10.7& 0.38& 0.07& 0.27& 20,11& \nodata & \nodata & \nodata & $-1.23$ & 21 \\
G24--03 & 10.5& 0.36& 0.06& 0.27& 20,11& 5866 & \nodata & \nodata & $-1.78$ & 11 \\
& & & & & & \nodata & \nodata & \nodata & $-1.70$ & 21 \\
G30--52 & 8.6& 0.50& 0.25& 0.27& 11 & 4757 & \nodata & \nodata & $-2.12$ & 11 \\
& & & & & & 4880 & \nodata & \nodata & $-2.14$ & 3 \\
G33--09 & 10.6& 0.41& 0.10& 0.28& 20 & 5575 & \nodata & \nodata & $-1.48$ & 11 \\
G66--22 & 10.5& 0.46& 0.16& 0.28& 11 & 5060 & \nodata & \nodata & $-1.77$ & 3 \\
& & & & & & \nodata & \nodata & \nodata & $-1.04$ & 21 \\
G90--03 & 10.4& 0.37& 0.04& 0.29& 20 & \nodata & \nodata & \nodata & $-2.01$ & 21 \\
LP 608--62\tablenotemark{a} & 10.5& 0.30& 0.07& 0.35& 11 & 6250 & \nodata &
\nodata & $-2.70$ & 4 \\
\enddata
\tablenotetext{a}{Star LP 608--62 is also known as BD+1\arcdeg 2341p.  We will
make this footnote extra long so that it extends over two lines.}
%% You can append references to a table using the \tablerefs command.
\tablerefs{
(1) Barbuy, Spite, \& Spite 1985; (2) Bond 1980; (3) Carbon et al. 1987;
(4) Hobbs \& Duncan 1987; (5) Gilroy et al. 1988: (6) Gratton \& Ortolani 1986;
(7) Gratton \& Sneden 1987; (8) Gratton \& Sneden (1988); (9) Gratton \& Sneden 1991;
(10) Kraft et al. 1982; (11) LCL, or Laird, 1990; (12) Leep \& Wallerstein 1981;
(13) Luck \& Bond 1981; (14) Luck \& Bond 1985; (15) Magain 1987;
(16) Magain 1989; (17) Peterson 1981; (18) Peterson, Kurucz, \& Carney 1990;
(19) RMB; (20) Schuster \& Nissen 1988; (21) Schuster \& Nissen 1989b;
(22) Spite et al. 1984; (23) Spite \& Spite 1986; (24) Hobbs \& Thorburn 1991;
(25) Hobbs et al. 1991; (26) Olsen 1983.}
\end{deluxetable}

%% Tables may also be prepared as separate files. See the accompanying
%% sample file table.tex for an example of an external table file.
%% To include an external file in your main document, use the \input
%% command. Uncomment the line below to include table.tex in this
%% sample file. (Note that you will need to comment out the \documentclass,
%% \begin{document}, and \end{document} commands from table.tex if you want
%% to include it in this document.)

%% \input{table}

%% The following command ends your manuscript. LaTeX will ignore any text
%% that appears after it.

\end{document}

%%
%% End of file `sample.tex'.

### 2011-10-13 paper/self_calibration_I.tex
euclid .
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19734 03a3cea6-2c03-0410-ac9b-9271b1c66c29
euclid: only time for v minor addition to paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19750 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \begin{abstract}
-In this paper we show that, with due care given to the observing strategy, it will be possible to accurately constrain the relative photometric calibration of instruments used in large imaging surveys with the science data alone.
-We create end-to-end simulations of an imaging survey, which produces simulated datasets from mock observations of a synthetic sky -- with realistic measurement uncertainties and a complex instrument response -- according to defined survey strategies.
+In this paper we show that, with due care given to the observing strategy, it will be possible to accurately constrain the relative photometric calibration of instruments used in large imaging surveys with their science data alone.
+We create end-to-end simulations of an imaging survey, which produces simulated datasets from mock observations of a synthetic sky according to a defined survey strategy.
+These catalog based simulations include realistic measurement uncertainties and a complex, position dependent instrument response.
 We then use a self-calibration technique to recover the relative instrument response by fitting an model that best explains the survey dataset, based on the multiple observations of (non-varying) sources at different focal plane positions.
@@
 We find that returning the same sources to very different focal plane positions is the key property of a survey strategy that is required for an accurate calibration.
-From the results of this study, we are able to highlight an important point for those considering the design of large scale imaging surveys: depart from a regular tiling of the sky and return the same sources to very different focal plane positions.
+From the results of this work, we are able to highlight an important point for those considering the design of large scale imaging surveys: depart from a regular tiling of the sky and return the same sources to very different focal plane positions.
 \end{abstract}
@@
 Astronomers tend to think in terms of taking science data and calibration data separately.
-The former is used for science and the latter is used to set instrument parameters like flat-field, bias and so-on.
+The former is used for science and the latter is used to constrain instrument parameters like the flat-field, the dark currents and so-on.
 But typically far more photons are collected during science exposures; are these not incredibly constraining on the calibration themselves?
-Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (XX ref) (SDSS) much more calibration information was obtained from the science data than the calibration data.
+Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS) \citep{pad08} much more calibration information was obtained from the science data than the calibration data.
 But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: good redundancy is required in the data stream, and a redundancy of a very specific kind.
+
+In this paper, we argue that the next generation of large surveys should have their observation strategies optimized from the very start with calibration in mind.
+
+The question then just becomes: what makes a survey strategy good for the retrospective self-calibration of the dataset?
+In this paper, we focus on the relative photometric calibration of a typical imaging instrument only.
+For instrument parameters, we consider an instrument similar to the near-infrared photometric channel proposed for the European Space Agency's Euclid Dark Energy Mission.
+
+and not just as an afterthought (as with SDSS), it will be possible to constrain calibration parameters to a high degree of accuracy.
+
+turn to the planning of large imaging surveys, and not the retrospective
+
+We feel, an
+
+Why survey mission is unique etc
+
+In this paper we summarize the end-to-end simulation used
 
@@
 
-Although the results we present here are general to all large scale imaging surveys, in Section XX we focus on the proposed Euclid Deep Field Survey.
-This survey has been optimized, within other constraints, based on the results from this work.
-
-\section{Observations}
-
-%% In a manner similar to \objectname authors can provide links to dataset
-%% hosted at participating data centers via the \dataset{} command.  The
-%% second curly bracket argument is printed in the text while the first
-%% parentheses argument serves as the valid data set identifier.  Large
-%% lists of data set are best provided in a table (see Table 3 for an example).
-%% Valid data set identifiers should be obtained from the data center that
-%% is currently hosting the data.
-%%
-%% Note that AASTeX interprets everything between the curly braces in the 
-%% macro as regular text, so any special characters, e.g. "#" or "_," must be 
-%% preceded by a backslash. Otherwise, you will get a LaTeX error when you 
-%% compile your manuscript.  Special characters do not 
-%% need to be escaped in the optional, square-bracket argument.
-
-All our observations were short direct exposures with CCD's.  We also have
-a random {\it Chandra} data set \dataset{ADS/Sa.ASCA\#X/86008020} and a
-neat \dataset[ADS/Sa.HST#Y0Q70101T]{HST FOS spectrum} that readers can access
-via the links in the electronic edition.  Unfortunately this has nothing 
-whatsoever to do with this research.  At Lick Observatory we used a TI 500$\times$500 
-chip and a GEC 575$\times$385, on the 1-m Nickel reflector.  The only
-filter available at Lick was red.  At CTIO we used a GEC 575$\times$385, with
-$B, V,$ and $R$ filters, and an RCA 512$\times$320, with $U, B, V, R,$ and $I$
-filters, on the 1.5-m reflector. In the CTIO observations we tried to
-concentrate on the shortest practicable wavelengths; but faintness, reddening,
-and poor short-wavelength sensitivity often kept us from observing in $U$ or
-even in $B$. All four cameras had scales of the order of 0.4 arcsec/pixel, and
-our field sizes were around 3 arcmin.
-
-The CCD images are unfortunately not always suitable, for very poor
-clusters or for clusters with large cores.  Since the latter are easily
-studied by other means, we augmented our own CCD profiles by collecting
-from the literature a number of star-count profiles 
-\citep{kin68,pet76,har84,ort85}, as well as photoelectric profiles 
-\citep{kin66,kin75} and electronographic profiles \citep{kro84}.
-In a few cases we judged normality by eye estimates on one of the Sky
-Surveys.
-
-%% In this section, we use  the \subsection command to set off
-%% a subsection.  \footnote is used to insert a footnote to the text.
-
-%% Observe the use of the LaTeX \label
-%% command after the \subsection to give a symbolic KEY to the
-%% subsection for cross-referencing in a \ref command.
-%% You can use LaTeX's \ref and \label commands to keep track of
-%% cross-references to sections, equations, tables, and figures.
-%% That way, if you change the order of any elements, LaTeX will
-%% automatically renumber them.
-
-%% This section also includes several of the displayed math environments
-%% mentioned in the Author Guide.
-
-\section{Helicity Amplitudes}
-
-It has been realized that helicity amplitudes provide a convenient means
-for Feynman diagram\footnote{Footnotes can be inserted like this.}
-evaluations.  These amplitude-level techniques
-are particularly convenient for calculations involving many Feynman
-diagrams, where the usual trace techniques for the amplitude
-squared becomes unwieldy.  Our calculations use the helicity techniques
-developed by other authors \cite[]{hag86}; we briefly summarize below.
-
-\subsection{Formalism} \label{bozomath}
-
-%% The equation environment wil produce a numbered display equation.
-
-A tree-level amplitude in $e^+e^-$ collisions can be expressed in
-terms of fermion strings of the form
-\begin{equation}
-\bar v(p_2,\sigma_2)P_{-\tau}\hat a_1\hat a_2\cdots
-\hat a_nu(p_1,\sigma_1) ,
-\end{equation}
-where $p$ and $\sigma$ label the initial $e^{\pm}$ four-momenta
-and helicities $(\sigma = \pm 1)$, $\hat a_i=a^\mu_i\gamma_\nu$
-and $P_\tau=\frac{1}{2}(1+\tau\gamma_5)$ is a chirality projection
-operator $(\tau = \pm1)$.  The $a^\mu_i$ may be formed from particle
-four-momenta, gauge-boson polarization vectors or fermion strings with
-an uncontracted Lorentz index associated with final-state fermions.
-
-%% The \notetoeditor{TEXT} command allows the author to communicate
-%% information to the copy editor.  This information will appear as a
-%% footnote on the printed copy for the manuscript style file.  Nothing will
-%% appear on the printed copy if the preprint or
-%% preprint2 style files are used.
-
-%% The eqnarray environment produces multi-line display math. The end of
-%% each line is marked with a \\. Lines will be numbered unless the \\
-%% is preceded by a \nonumber command.
-%% Alignment points are marked by ampersands (&). There should be two
-%% ampersands (&) per line.
-
-In the chiral \notetoeditor{Figures 1 and 2 should appear side-by-side in
-print} representation the $\gamma$ matrices are expressed
-in terms of $2\times 2$ Pauli matrices $\sigma$ and the unit matrix 1 as
-\begin{eqnarray}
-\gamma^\mu  & = &
- \left(
-\begin{array}{cc}
-0 & \sigma^\mu_+ \\
-\sigma^\mu_- & 0
-\end{array}     \right) ,
- \gamma^5= \left(
-\begin{array}{cc}
--1 &   0\\
-0 &   1
-\end{array}     \right)  , \nonumber \\
-\sigma^\mu_{\pm}  & = &   ({\bf 1} ,\pm \sigma) , \nonumber
-\end{eqnarray}
-giving
-\begin{equation}
-\hat a= \left(
-\begin{array}{cc}
-0 & (\hat a)_+\\
-(\hat a)_- & 0
-\end{array}\right), (\hat a)_\pm=a_\mu\sigma^\mu_\pm ,
-\end{equation}
-The spinors are expressed in terms of two-component Weyl spinors as
-\begin{equation}
-u=\left(
-\begin{array}{c}
-(u)_-\\
-(u)_+
-\end{array}\right), v={\bf (}\vdag_+{\bf ,}   \vdag_-{\bf )} .
-\end{equation}
-
-%% Putting eqnarrays or equations inside the mathletters environment groups
-%% the enclosed equations by letter. For instance, the eqnarray below, instead
-%% of being numbered, say, (4) and (5), would be numbered (4a) and (4b).
-%% LaTeX the paper and look at the output to see the results.
-
-The Weyl spinors are given in terms of helicity eigenstates
-$\chi_\lambda(p)$ with $\lambda=\pm1$ by
-\begin{mathletters}
-\begin{eqnarray}
-u(p,\lambda)_\pm & = & (E\pm\lambda|{\bf p}|)^{1/2}\chi_\lambda(p) , \\
-v(p,\lambda)_\pm & = & \pm\lambda(E\mp\lambda|{\bf p}|)^{1/2}\chi
-_{-\lambda}(p)
-\end{eqnarray}
-\end{mathletters}
-
-%% This section contains more display math examples, including unnumbered
-%% equations (displaymath environment). The last paragraph includes some
-%% examples of in-line math featuring a couple of the AASTeX symbol macros.
-
-\section{Floating material and so forth}
-
-%% The displaymath environment will produce the same sort of equation as
-%% the equation environment, except that the equation will not be numbered
-%% by LaTeX.
-
-Consider a task that computes profile parameters for a modified
-Lorentzian of the form
-\begin{equation}
-I = \frac{1}{1 + d_{1}^{P (1 + d_{2} )}}
-\end{equation}
-where
-\begin{displaymath}
-d_{1} = \sqrt{ \left( \begin{array}{c} \frac{x_{1}}{R_{maj}}
-\end{array} \right) ^{2} +
-\left( \begin{array}{c} \frac{y_{1}}{R_{min}} \end{array} \right) ^{2} }
-\end{displaymath}
-\begin{displaymath}
-d_{2} = \sqrt{ \left( \begin{array}{c} \frac{x_{1}}{P R_{maj}}
-\end{array} \right) ^{2} +
-\left( \begin{array}{c} \case{y_{1}}{P R_{min}} \end{array} \right) ^{2} }
-\end{displaymath}
-\begin{displaymath}
-x_{1} = (x - x_{0}) \cos \Theta + (y - y_{0}) \sin \Theta
-\end{displaymath}
-\begin{displaymath}
-y_{1} = -(x - x_{0}) \sin \Theta + (y - y_{0}) \cos \Theta
-\end{displaymath}
-
-In these expressions $x_{0}$,$y_{0}$ is the star center, and $\Theta$ is the
-angle with the $x$ axis.  Results of this task are shown in table~\ref{tbl-1}.
-It is not clear how these sorts of analyses may affect determination of
- $M_{\sun}$, but the assumption is that the alternate results
-should be less than 90\degr\ out of phase with previous values.
-We have no observations of \ion{Ca}{2}.
-Roughly \slantfrac{4}{5} of the electronically submitted abstracts
-for AAS meetings are error-free.
+\section{Discussion}
+In this paper, we have considered the optimization of survey strategies for calibrating the relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained for this method.
+For example, the optical distortion of an instrument could also be constrained with such a method, although we concede that the properties of a survey strategy that makes them good for one calibration, may not be the same as that for another; ultimately a trade-off would need to be performed.
 
@@
 \acknowledgments
-
-We are grateful to V. Barger, T. Han, and R. J. N. Phillips for
-doing the math in section~\ref{bozomath}.
-More information on the AASTeX macros package is available \\ at
-\url{http://www.aas.org/publications/aastex}.
-For technical support, please write to
-\email{aastex-help@aas.org}.
+\section{Acknowledgments}
+We are grateful to
 
@@
 
-{\it Facilities:} \facility{Nickel}, \facility{HST (STIS)}, \facility{CXO (ASIS)}.
-
 %% Appendix material should be preceded with a single \appendix command.
@@
 \appendix
-
-\section{Appendix material}
-
-Consider once again a task that computes profile parameters for a modified
-Lorentzian of the form
-\begin{equation}
-I = \frac{1}{1 + d_{1}^{P (1 + d_{2} )}}
-\end{equation}
-where
-\begin{mathletters}
-\begin{displaymath}
-d_{1} = \frac{3}{4} \sqrt{ \left( \begin{array}{c} \frac{x_{1}}{R_{maj}}
-\end{array} \right) ^{2} +
-\left( \begin{array}{c} \frac{y_{1}}{R_{min}} \end{array} \right) ^{2} }
-\end{displaymath}
-\begin{equation}
-d_{2} = \case{3}{4} \sqrt{ \left( \begin{array}{c} \frac{x_{1}}{P R_{maj}}
-\end{array} \right) ^{2} +
-\left( \begin{array}{c} \case{y_{1}}{P R_{min}} \end{array} \right) ^{2} }
-\end{equation}
-\begin{eqnarray}
-x_{1} & = & (x - x_{0}) \cos \Theta + (y - y_{0}) \sin \Theta \\
-y_{1} & = & -(x - x_{0}) \sin \Theta + (y - y_{0}) \cos \Theta
-\end{eqnarray}
-\end{mathletters}
-
-For completeness, here is one last equation.
-\begin{equation}
-e = mc^2
-\end{equation}
 
@@
 \begin{thebibliography}{}
-\bibitem[Auri\`ere(1982)]{aur82} Auri\`ere, M.  1982, \aap,
-    109, 301
-\bibitem[Canizares et al.(1978)]{can78} Canizares, C. R.,
-    Grindlay, J. E., Hiltner, W. A., Liller, W., \&
-    McClintock, J. E.  1978, \apj, 224, 39
-\bibitem[Djorgovski \& King(1984)]{djo84} Djorgovski, S.,
-    \& King, I. R.  1984, \apjl, 277, L49
-\bibitem[Hagiwara \& Zeppenfeld(1986)]{hag86} Hagiwara, K., \&
-    Zeppenfeld, D.  1986, Nucl.Phys., 274, 1
-\bibitem[Harris \& van den Bergh(1984)]{har84} Harris, W. E.,
-    \& van den Bergh, S.  1984, \aj, 89, 1816
-\bibitem[H\`enon(1961)]{hen61} H\'enon, M.  1961, Ann.d'Ap., 24, 369
-\bibitem[Heiles \& Troland(2003)]{heiles03} Heiles, C. \& Troland, T. H., 2003, \apjs, preprint doi:10.1086/381753
-\bibitem[Kim, Ostricker, \& Stone(2003)]{kim03} Kim, W.-T.,  Ostriker, E., \& Stone, J. M., 2003, \apj, 599, 1157
-\bibitem[King(1966)]{kin66}  King, I. R.  1966, \aj, 71, 276
-\bibitem[King(1975)]{kin75}  King, I. R.  1975, Dynamics of
-    Stellar Systems, A. Hayli, Dordrecht: Reidel, 1975, 99
-\bibitem[King et al.(1968)]{kin68}  King, I. R., Hedemann, E.,
-    Hodge, S. M., \& White, R. E.  1968, \aj, 73, 456
-\bibitem[Kron et al.(1984)]{kro84} Kron, G. E., Hewitt, A. V.,
-    \& Wasserman, L. H.  1984, \pasp, 96, 198
-\bibitem[Lynden-Bell \& Wood(1968)]{lyn68} Lynden-Bell, D.,
-    \& Wood, R.  1968, \mnras, 138, 495
-\bibitem[Newell \& O'Neil(1978)]{new78} Newell, E. B.,
-    \& O'Neil, E. J.  1978, \apjs, 37, 27
-\bibitem[Ortolani et al.(1985)]{ort85} Ortolani, S., Rosino, L.,
-    \& Sandage, A.  1985, \aj, 90, 473
-\bibitem[Peterson(1976)]{pet76} Peterson, C. J.  1976, \aj, 81, 617
-\bibitem[Rudnick et al.(2003)]{rudnick03} Rudnick, G. et al., 2003, \apj, 599, 847
-\bibitem[Spitzer(1985)]{spi85} Spitzer, L.  1985, Dynamics of
-    Star Clusters, J. Goodman \& P. Hut, Dordrecht: Reidel, 109
-\bibitem[Treu et al.(2003)]{treu03} Treu, T. et al., 2003, \apj, 591, 53
+\bibitem[Padmanabhan et al.(2008)]{pad08} {Padmanabhan, N., Schlegel, D.~J., Finkbeiner, D.~P., Barentine, J.~C., Blanton, M.~R., Brewington, H.~J., Gunn, J.~E., Harvanek, M., Hogg, D.~W., Ivezi{\'c}, {\v Z}., Johnston, D., Kent, S.~M., Kleinman, S.~J., Knapp, G.~R., Krzesinski, J., Long, D., Neilsen, Jr., E.~H., Nitta, A., Loomis, C., Lupton, R.~H., Roweis, S., Snedden, S.~A., Strauss, M.~A., \& Tucker, D.~L.
+2008, \apj, 674, 1217}
 \end{thebibliography}
-
 \clearpage
@@
 
-\begin{figure}
-\epsscale{.80}
-%\plotone{f1.eps}
-\caption{Derived spectra for 3C138 \citep[see][]{heiles03}. Plots for all sources are available
-in the electronic edition of {\it The Astrophysical Journal}.\label{fig1}}
-\end{figure}
-
-\clearpage
-
 %% Here we use \plottwo to present two versions of the same figure,
@@
 
-\begin{figure}
-%\plottwo{f2.eps}{f2_color.eps}
-\caption{A panel taken from Figure 2 of \citet{rudnick03}. 
-See the electronic edition of the Journal for a color version 
-of this figure.\label{fig2}}
-\end{figure}
-
 %% This figure uses \includegraphics to scale and rotate the still frame
 %% for an mpeg animation.
-
-\begin{figure}
-%\includegraphics[angle=90,scale=.50]{f3.eps}
-\caption{Animation still frame taken from \citet{kim03}.
-This figure is also available as an mpeg
-animation in the electronic edition of the
-{\it Astrophysical Journal}.}
-\end{figure}
 
@@
 \clearpage
-
-\begin{table}
-\begin{center}
-\caption{More terribly relevant tabular information.\label{tbl-2}}
-\begin{tabular}{crrrrrrrrrrr}
-\tableline\tableline
-Star & Height & $d_{x}$ & $d_{y}$ & $n$ & $\chi^2$ & $R_{maj}$ & $R_{min}$ &
-\multicolumn{1}{c}{$P$\tablenotemark{a}} & $P R_{maj}$ & $P R_{min}$ &
-\multicolumn{1}{c}{$\Theta$\tablenotemark{b}} \\
-\tableline
-1 &33472.5 &-0.1 &0.4  &53 &27.4 &2.065  &1.940 &3.900 &68.3 &116.2 &-27.639\\
-2 &27802.4 &-0.3 &-0.2 &60 &3.7  &1.628  &1.510 &2.156 &6.8  &7.5 &-26.764\\
-3 &29210.6 &0.9  &0.3  &60 &3.4  &1.622  &1.551 &2.159 &6.7  &7.3 &-40.272\\
-4 &32733.8 &-1.2\tablenotemark{c} &-0.5 &41 &54.8 &2.282  &2.156 &4.313 &117.4 &78.2 &-35.847\\
-5 & 9607.4 &-0.4 &-0.4 &60 &1.4  &1.669\tablenotemark{c}  &1.574 &2.343 &8.0  &8.9 &-33.417\\
-6 &31638.6 &1.6  &0.1  &39 &315.2 & 3.433 &3.075 &7.488 &92.1 &25.3 &-12.052\\
-\tableline
-\end{tabular}
-%% Any table notes must follow the \end{tabular} command.
-\tablenotetext{a}{Sample footnote for table~\ref{tbl-2} that was
-generated with the \LaTeX\ table environment}
-\tablenotetext{b}{Yet another sample footnote for table~\ref{tbl-2}}
-\tablenotetext{c}{Another sample footnote for table~\ref{tbl-2}}
-\tablecomments{We can also attach a long-ish paragraph of explanatory
-material to a table.}
-\end{center}
-\end{table}
 

### 2011-10-23 paper/self_calibration_I.tex
euclid: working on intro and simulation part of paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19786 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \newcommand{\vdag}{(v)^\dagger}
+
+\usepackage{mathtools}
+\usepackage{upgreek}
+
+\newcommand{\true}{\text{true}}
+\newcommand{\fit}{\text{fit}}
+\newcommand{\basis}{\text{best}}
+\newcommand{\dd}{\text{d}}
 
@@
 But typically far more photons are collected during science exposures; are these not incredibly constraining on the calibration themselves?
-Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS) \citep{pad08} much more calibration information was obtained from the science data than the calibration data.
+Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS) much more calibration information was obtained from the science data than the calibration data \citep{pad08}.
 But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: good redundancy is required in the data stream, and a redundancy of a very specific kind.
 
-In this paper, we argue that the next generation of large surveys should have their observation strategies optimized from the very start with calibration in mind.
+In this paper, we argue that the next generation of large surveys should have their observation strategies optimized from the very start with this ``self-calibration'' in mind.
+We focus here on the relative photometric calibration of a typical imaging instrument only, although we also suspect that similar techniques could be used to constrain many other calibration parameters.
+This technique utilizes the multiple measurements of (non-varying) sources at different focal plane positions and at different times within the survey to constrain the relative instrument response by requiring that they yield the same count values.
+Through end-to-end -- catalog level -- survey simulations\footnote{All of the code used in this work is publicly available at XX.}, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.
 
-The question then just becomes: what makes a survey strategy good for the retrospective self-calibration of the dataset?
-In this paper, we focus on the relative photometric calibration of a typical imaging instrument only.
-For instrument parameters, we consider an instrument similar to the near-infrared photometric channel proposed for the European Space Agency's Euclid Dark Energy Mission.
+These simulations include a complex, position dependent instrument response; through mock observations of a synthetic sky, according to a defined survey strategy, we construct a realistic survey dataset.
+We then fit a (different) instrument response model that best describes this dataset.
+By comparing the \textit{fitted} instrument response to the \textit{true} instrument response we are able to assess the performance of the self-calibration procedure with different survey strategies.
+There is a degeneracy in this problem: the self-calibration procedure is only able to fit for a \textit{relative} instrument response.
+We have no way of know, for example, if all the sources are uniformly fainter, or if the instrument sensitivity is uniformly lower.
 
-and not just as an afterthought (as with SDSS), it will be possible to constrain calibration parameters to a high degree of accuracy.
+In Section \ref{sec:survey} we introduce the simulation chain constructed to produce the realistic survey datasets.
+Section \ref{sec:self_cal} goes onto to detail the self-calibration procedure.
+In Section \ref{sec:simple} we focus on four simple survey strategies, which allow us to draw conclusions on the performance of the self-calibration procedure with different survey properties.
+We do not produced pixelated images; instead we concentrate on catalog level simulations with realistic measurement uncertainties.
+Complex effects are included in the simulations that are not  precisely modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the dataset.
 
-turn to the planning of large imaging surveys, and not the retrospective
+\section{Survey Dataset Simulations}
+\label{sec:survey}
+We have constructed an end-to-end simulation chain that produces a realistic imaging dataset from a specified survey strategy.
+Simulation parameters are kept intentionally flexible, so that the sensitivity of the self-calibration procedure to different parameters can be investigated.
+The simulations are split into a number of steps.
+First, a synthetic sky is generated based on the specified simulation parameters (density of sources, magnitude ranges).
+With a given pointing, single mock observations are performed on this sky, again here with tunable simulation parameters, such as the instrument response model and the the noise model.
+The final step is the build up of the survey wide dataset from multiple single observations taken according to the prescribed survey strategy.
+In this section, we detail the assumptions and methods used in each of these steps.
 
-We feel, an
+\subsection{The Synthetic Sky}
+We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and $10\upsigma{}$ limits of deep, space-based, near-infrared survey.
+Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes $m$ distributed according to
+\begin{eqnarray}
+\log_{10} \frac{\dd N}{\dd m \, \dd \Omega} = a + b\,m + c\,m^2 \label{eqn:power_law} \quad , 
+\end{eqnarray}
+where $\dfrac{\dd N}{\dd m \, \dd  \Omega}$  is the density of sources $N$ per unit magnitude $m$ and per unit solid angle $\Omega$, and $a$, $b$ and $c$ are model parameters.
+Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in Windhorst et al.
+(2011) \cite{source_densities} only.
+These parameters were $a = -13.05$, $b = 1.25$ and $c = -0.02$.
+We intentionally omit the stellar population, so that the conclusions drawn from these simulations are independent of the final position of the survey area on the sky.
+The source magnitudes $m$ are related to the source fluxes $s$ simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies (nmgy).
 
-Why survey mission is unique etc
+\subsection{A Single Exposure}
+FoV
 
-In this paper we summarize the end-to-end simulation used
+\subsection{The Complete Survey}
 
-In this paper we will show that this distinction between science and calibration data no longer applies in the large survey mission, both ground and space based, currently being planned and developed.
-We show that, as long as the survey strategy is designed appropriately, the calibration information within the science data can be used to accurately constrain the relative photometric response of an imaging instrument.
-This retrospective, cross-calibration of the dataset is a very powerful calibration method; future large scale imaging surveys should be optimized with this in mind.
+\section{The Self-Calibration Technique}
+\label{sec:self_cal}
 
-We concentrate on catalog level simulations of different survey strategies and assess how well the resultant dataset can be used to constrain the relative response of the imaging instrument it came from.
-Although we do not model the imaging instrument on a pixel scale, we do include realistic measurement uncertainties within the simulations.
+\section{The Simple Survey Strategies}
+\label{sec:simple}
 

### 2011-10-24 paper/self_calibration_I.tex
euclid: paper update - single exposure, survey simulation, self-cal, metrics
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19791 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 
-\title{Optimizing Large Scale Imaging Surveys for Relative Photometric Calibration, \\
-    the Euclid Dark Energy Mission as an Example}
+\title{Optimizing Large Scale Imaging Surveys for Relative Photometric Calibration}
 
@@
 Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in Windhorst et al.
-(2011) \cite{source_densities} only.
+(2011) \citep{win11} only.
 These parameters were $a = -13.05$, $b = 1.25$ and $c = -0.02$.
@@
 The source magnitudes $m$ are related to the source fluxes $s$ simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies (nmgy).
+Due to computational reasons, we only select the brightest sources within the survey area, up to a source density $d$, for the self-calibration procedure.
 
 \subsection{A Single Exposure}
-FoV
+\label{sec:single_exposure}
+With a camera pointing $(\alpha, \beta)$ and camera orientation $\theta$ this synthetic sky is transformed into focal plane coordinates and all of the sources falling within the instrument's field-of-view are found.
+In our simulations, we use a large instrument field-of-view of $0.76~ \text{deg} \times 0.72~\text{deg}$; a size consistent with the next generation of large survey imagers.
+An example of a single pointing exposure is shown in Figure \ref{fig:single_image}.
+
+\subsubsection{Measured Count Rates}
+We do not consider pixelated images in these simulations; instead the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position dependent instrument response model $f_\true$ and a measurement noise model.
+For a measurement $i$ the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s_{k,\true}$
+\begin{eqnarray*}
+c_i = f_{\true}(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} + e_{i} \quad ,
+\end{eqnarray*}
+where $\vec{q}_\true$ are the parameters defining the \textit{true} instrument response, and $e_i$ is a noise contribution drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$.
+
+\subsubsection{Noise Model}
+\label{sec:noise}
+To construct the noise model, the simulated exposures are assumed to be background limited and that, for systematic reasons, there is an upper limit on the signal-to-noise ratio of 500 for bright sources.
+The noise model is complicated further by applying an extra term $\epsilon_i$ to the count rates' uncertainty variance, which we intentionally do not take into account in the analysis in order to simulate systematic problems with the instrument noise model.
+The \textit{true} noise model is therefore
+\begin{eqnarray}
+\sigma_{i, \true}^{2} = (1 + \epsilon_i) \, \alpha^{2} + \eta^{2}\, [ f_\true(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} ]^2 \quad , \label{eqn:noise}
+\end{eqnarray}
+where $\alpha$ and $\eta$ are both constants and $\epsilon_i$ is a random number, in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$.
+The $m = 22$ mag 10$\upsigma$ detection limit introduced previously and the 500 limit on the signal-to-noise ratio are used to set $\alpha = 0.1585$ and $\eta = 0.0017$.
+The $\epsilon_i$ contribution is not taken into account in the analysis stage and therefore the uncertainty variances on the count rates are assumed to be
+\begin{eqnarray*}
+\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2} \, c^{2}_i \quad 
+\end{eqnarray*}
+during the analysis stage.
+
+\subsubsection{The True Instrument Response Model}
+\label{sec:instrument_response_model}
+We construct a complex, position independent instrument response model $f_\true(\vec{x_i} | \vec{q}_{\true})$ from a superposition of large and small-scale variations:
+\begin{eqnarray*}
+f_\true(\vec{x_i} | \vec{q}_{\true, 1 \ldots 260}) = f_\text{large}(\vec{x_i} | \vec{q}_{\true, 1 \ldots 6}) + f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7 \ldots 260}) \quad ,
+\end{eqnarray*}
+where $\vec{x_i} = (x_i, y_i)$ is the focal plane position that the $k$th source falls at during the $i^\text{th}$ measurement and $\vec{q}_\true$ are the parameters defining the instrument response model.
+The large-scale instrument response $f_\text{large}(\vec{x_i} | \vec{q}_{\true, 1 \ldots 6})$ is modeled as a second order polynomial:
+\begin{eqnarray*}
+f_\text{large}(\vec{x_i} | \vec{q}_{\true, 1 \ldots 6}) = q_{\true, 1} + q_{\true, 2} \, x_i + q_{\true, 3} \, y_i + q_{\true, 4} \, x_i^2 + q_{\true, 5} \, x_i \, y_i  + q_{\true, 6} \, y_i^2  \quad .
+\end{eqnarray*}
+The small-scale instrument response, which is constructed from sine and cosine contributions, is superimposed on this large-scale instrument response.
+The small-scale instrument response $f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7 \ldots 260})$ is modeled as
+\begin{eqnarray*}
+f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7\ldots 260})  & = &  \sum_{a=0}^6 \sum_{b=0}^a \left[ q_{\true, 7+4b} \cos (k_x \, x_i) + q_{\true, 8+4b} \sin (k_x \, x_i) \right] \\
+& & \qquad \qquad \times \left[ q_{\true, 9+4b} \cos (k_y \, y_i) + q_{\true, 10+4b} \sin (k_y \, y_i) \right] \quad ,
+\end{eqnarray*}
+where
+\begin{eqnarray*}
+k_x & = & \dfrac{a \, \pi}{X} \quad ,\\
+k_y & = & \dfrac{b \, \pi}{Y} \quad ,\\
+\end{eqnarray*}
+with the physical focal plane dimensions $X$ and $Y$.
+In total, the instrument response model is parameterized with 260 parameters; an example can be seen in Figure \ref{fig:single_image}(right).
+It is this instrument response that we try and recover with the self-calibration procedure.
+Our final assumption in these simulations is that the instrument response is temporally stable.
 
 \subsection{The Complete Survey}
+In this work we are interested in simulating complete surveys, in order to identify the crucial characteristics of survey strategies that allows for the relative instrument response model to be accurately constrained from the resulting dataset.
+We therefore apply the single exposure procedure introduced in the previous section for each pointing specified in a defined survey strategy.
+The resultant source measurements are collaged into a survey wide dataset.
 
-\section{The Self-Calibration Technique}
+\section{Self-Calibrating the Survey Wide Dataset}
 \label{sec:self_cal}
+We self-calibrate the dataset generated in the survey simulations to recover the true instrument response applied and the true source fluxes.
+This self-calibration procedure has been successfully applied to ground based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}\footnote{Here this procedure is dubbed ``uber-calibration.''}.
+This iterative procedure comprises two steps: (1) a refinement of the source flux estimates based on the latest instrument response model and (2) a refinement of the instrument response model based on the updated source flux estimates.
+These steps are iterated until the system converges, or until it is clear that the system will not converge.
+There is a degeneracy in the problem, as both the true instrument response and the true source magnitudes are unknown.
+It is therefore only possible to calibrate the \textit{relative} instrument response and the \textit{relative} source fluxes.
+It is not possible to know, for example, if the sources are all fainter or if the instrument response is uniformly lower.
+This degeneracy can be broken through observations of standard absolute sources.
+
+\subsection{Fitted Measurement Model}
+To complicate the simulations, we use the self-calibration procedure to fit a model that is \textit{incomplete} in two ways.
+Firstly, the \textit{fitted} instrument response is modeled as an eighth order polynomial, and not the second order polynomial superimposed with sine and cosine contributions used to model the \textit{true} instrument response.
+Secondly, the assumed measurement uncertainty variances do not include the additional random measurement error $\epsilon_{i}$ introduced in Subsection \ref{sec:noise}.
+The incomplete measurement model is
+\begin{eqnarray*}
+c_i = f(\vec{x_i} | \vec{q}) \, s_{k} + e_{i} \quad ,
+\end{eqnarray*}
+where $c_i$ is the recorded count rate, $f(\vec{x_i} | \vec{q})$ is the fitted instrument response model at a focal plane position $\vec{x}_i$, $\vec{q}$ is a vector parameterizing the eighth order polynomial instrument response model, $s_k$ is the model source flux estimate and the error $e_i$ is drawn from the Normal Distribution $N(e|0,{\sigma_i}^2)$, such that
+\begin{eqnarray*}
+\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2}\, c^{2}_i \quad ,
+\end{eqnarray*} 
+where $\alpha$ and $\eta$ are the parameters set by the instruments $10\upsigma$ and saturation limits.
+The $\epsilon_i$ error contribution is intentionally not included in order to simulate systematic problems with the instrument noise model.
+
+\subsection{Step 1: Source Flux Refinement}
+The sources are considered individually in the first step of the self-calibration procedure; their flux estimates are refined based on the latest fitted instrument response parameters $\vec{q}$.
+An error function $\chi^2_{k}$ for all the measurements $i$ of a source $k$ ($i \in \mathcal{O}(k)$) is constructed:
+\begin{eqnarray*}
+\chi^2_{k} = \sum_{i \in \mathcal{O}(k)} \frac{(c_i-f_{i}(\vec{x_i} | \vec{q}) \, s_{k})^2}{{\sigma_i}^2} \quad ,
+\end{eqnarray*}
+where $c_i$ are the measured count rates, $f(\vec{x_i} | \vec{q})$ is the fitted instrument response model at a focal plane position $\vec{x}_i$ and $\sigma_i$ is the assumed noise model.
+A new estimate of the model source flux $s'_{k}$ is then found by minimizing the error function with respect to the old model source flux $s_{k}$:
+\begin{eqnarray*}
+\frac{d\chi^2_{k}}{d s_{k}} = \sum_{i \in \mathcal{O}(k)} \frac{-2 f_{i}(\vec{x_i} | \vec{q}) \, (c_i-f_{i}(\vec{x_i} | \vec{q}) \, s'_{k})}{{\sigma_i}^2} = 0 \quad ,
+\end{eqnarray*}
+\begin{eqnarray*}
+s'_{k} \leftarrow \left[{\sum_{i \in \mathcal{O}(k)}  \frac{f_{i}(\vec{x_i} | \vec{q})^2}{{\sigma_i}^2}} \right]^{-1}  \left[ {\sum_{i \in \mathcal{O}(k)} \frac{f_{i}(\vec{x_i} | \vec{q}) \, c_i}{{\sigma_i}^2}} \right] \quad .
+\end{eqnarray*}
+The standard uncertainty variance on the new source flux estimate $s'_{k}$ is given by
+\begin{eqnarray*}
+\sigma'^2_k = \left[{\sum_{i \in \mathcal{O}(k)}  \frac{f_{i}(\vec{x_i} | \vec{q})^2}{{\sigma_i}^2}} \right] \quad .
+\end{eqnarray*}
+
+\subsection{Step 2: Instrument Response Refinement}
+The instrument response parameters can now be refined with the latest source flux estimates. A error function for all the measurements of all the sources is constructed
+\begin{eqnarray*}
+\chi^2 = \sum_{k} \chi'^2_k \quad ,
+\end{eqnarray*}
+where
+\begin{eqnarray*}
+\chi'^2_k = \sum_{i \in \mathcal{O}(k)} \frac{(c_i-f_{i}(\vec{x_i} | \vec{q}) \, s'_{k})^2}{{\sigma_i}^2}  \quad .
+\end{eqnarray*}
+Recall that the fitted instrument response $f_{i}(\vec{x_i} | \vec{q})$ is modeled as an eight order polynomial. This can be expressed as
+\begin{eqnarray*}
+f_{i}(\vec{x_i} | \vec{q}) = \sum_{l = 1}^L q_{l} \, g_l(\vec{x_i})  \quad ,
+\end{eqnarray*}
+where $L = 45$ in this case. The total error function $\chi^2$ can be rewritten as
+\begin{eqnarray*}
+\chi^2 =\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{(c_i- s'_{k} \sum_{l = 1}^L q_{l} g_l(\vec{x_i}))^2}{{\sigma_i}^2}   \quad .
+\end{eqnarray*}
+To refine the instrument response model fit, this error function is minimized with respect to the instrument response model parameters $q_{l}$
+\begin{eqnarray*}
+\frac{d\chi^2}{dq_{l}} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{-2 g_l(\vec{x_i}) s'_{k} (c_i- s'_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'}(\vec{x_i}))}{{\sigma_i}^2} = 0  \quad ,
+\end{eqnarray*}
+\begin{eqnarray*}
+\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{{\sigma_i}^2} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l (\vec{x_i}) s'^2_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'} (\vec{x_i})} {{\sigma_i}^2}   \quad .
+\end{eqnarray*}
+It is now simpler to proceed in matrix notation. The following substitutions can be made
+\begin{eqnarray}
+b_l & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{\sigma_i^2}    \quad ,\\
+G_{ll'} & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{{s'}_{k}^2}{\sigma_i^2} g_l(x_i) g_{l'}(x_i)   \quad .
+\end{eqnarray}
+The matrix equation is then
+\begin{eqnarray*}
+\vec{b} = G \cdot \vec{q'}    \quad .
+\end{eqnarray*}
+The refined instrument response parameters are then found by
+\begin{eqnarray*}
+\vec{q'} \leftarrow G^{-1}  \cdot \vec{b}   \quad .
+\end{eqnarray*}
+These two steps are iterated until the solution converges to a final fit of the instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ and the source fluxes $s_{k,\fit}$, or until it is clear that a solution will not be found.
+
+\section{Metrics}
+\label{sec:metrics}
+To assess the performance of the self-calibration procedure with different survey strategies, it is necessary to quantify the quality of the final fitted solution.
+To do this we defined three quantities.
+The first is the root-mean-squared (RMS) error $S_\text{RMS}$ in the final fitted source fluxes $s_{k,\fit}$ compared to the true source fluxes $s_{k,\true}$ for all the $K$ sources:
+\begin{equation}
+S_\text{RMS} = \sqrt{\dfrac{1}{K} \sum_k^K \left( \dfrac{s_{k,\fit} - s_{k,\true}}{s_{k,\true}} \right)^2}   \quad .
+\end{equation}
+The other two metrics, called ``badnesses'', are defined as the RMS error between the final fitted instrument response and a reference instrument response sampled on a regular $500 \times 500$ grid across the focal plane.
+For the ``True Badness'' $B_\true$, the \textit{fitted} instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ is compared to the \textit{true} instrument response $f_\true(\vec{x} | \vec{q_\true})$ at the $J$ sample points
+\begin{equation}
+B_\true = \sqrt{\dfrac{1}{J} \sum_j^J \left(\dfrac{f_\fit(\vec{x_j} | \vec{q_\fit}) - f_\true(\vec{x_j} | \vec{q_\true})}{f_\true(\vec{x_j} | \vec{q_\true})}\right)^2}   \quad .
+\end{equation}
+The ``Best-in-Basis Badness'' $B_\basis$ compares the \textit{fitted} instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ to the \textit{best instrument response fit possible} $f_\basis(\vec{x_j} | \vec{q_\basis})$ with the basis used to describe the fitted model (in this case an eight order polynomial) at the $J$ sample points
+\begin{equation}
+B_\basis = \sqrt{\dfrac{1}{J} \sum_j^J \left( \dfrac{f_\fit(\vec{x_j} | \vec{q_\fit}) - f_\basis(\vec{x_j} | \vec{q_\basis})}{f_\basis(\vec{x_j} | \vec{q_\basis})} \right)^2}   \quad .
+\end{equation}
+The badnesses provide a more complete description of the self-calibration performance than the RMS error on the fitted sources' fluxes, as the RMS source error only applies to the bright sources within the survey selected for the self-calibration procedure.
 
@@
 \begin{thebibliography}{}
-\bibitem[Padmanabhan et al.(2008)]{pad08} {Padmanabhan, N., Schlegel, D.~J., Finkbeiner, D.~P., Barentine, J.~C., Blanton, M.~R., Brewington, H.~J., Gunn, J.~E., Harvanek, M., Hogg, D.~W., Ivezi{\'c}, {\v Z}., Johnston, D., Kent, S.~M., Kleinman, S.~J., Knapp, G.~R., Krzesinski, J., Long, D., Neilsen, Jr., E.~H., Nitta, A., Loomis, C., Lupton, R.~H., Roweis, S., Snedden, S.~A., Strauss, M.~A., \& Tucker, D.~L.
-2008, \apj, 674, 1217}
+\bibitem[Padmanabhan et al.
+(2008)]{pad08} {Padmanabhan, N., Schlegel, D.~J., Finkbeiner, D.~P., Barentine, J.~C., Blanton, M.~R., Brewington, H.~J., Gunn, J.~E., Harvanek, M., Hogg, D.~W., Ivezi{\'c}, {\v Z}., Johnston, D., Kent, S.~M., Kleinman, S.~J., Knapp, G.~R., Krzesinski, J., Long, D., Neilsen, Jr., E.~H., Nitta, A., Loomis, C., Lupton, R.~H., Roweis, S., Snedden, S.~A., Strauss, M.~A., \& Tucker, D.~L., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj, 674, 1217}
+\bibitem[Windhorst et al.
+(2011)]{win11}{Windhorst, R.~A., Cohen, S.~H., Hathi, N.~P., McCarthy, P.~J., Ryan, Jr., R.~E., Yan, H., Baldry, I.~K., Driver, S.~P., Frogel, J.~A., Hill, D.~T., Kelvin, L.~S., Koekemoer, A.~M., Mechtley, M., O'Connell, R.~W., Robotham, A.~S.~G., Rutkowski, M.~J., Seibert, M., Straughn, A.~N.,Tuffs, R.~J., Balick, B., Bond, H.~E., Bushouse, H., Calzetti, D., Crockett, M., Disney, M.~J., Dopita, M.~A., Hall, D.~N.~B., Holtzman, J.~A., Kaviraj, S., Kimble, R.~A., MacKenty, J.~W., Mutchler, M., Paresce, F., Saha, A., Silk, J.~I., Trauger, J.~T., Walker, A.~R., Whitmore, B.~C., Young, E.~T., The Hubble Space Telescope Wide Field Camera 3 Early Release Science Data: Panchromatic Faint Object Counts for 0.2-2 $\upmu$m Wavelength, 2011, XX.}
 \end{thebibliography}
@@
 %% edition.
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{./single_image.pdf}
+\end{center}
+\caption[A single camera pointing from the self-calibration simulations.]{A single exposure of the synthetic sky.
+Left: A plot of the bright sources within the synthetic sky used in the self-calibration procedure with the focal plane footprint overlaid.
+Right: The resultant distribution of the sources on the instrument's focal plane.
+The \textit{true} instrument model $f_{\true}(\vec{x_i} | \vec{q}_\true)$ is shown as contours.\label{fig:single_image}}
+\end{figure}
+
+\clearpage
+\begin{deluxetable}{lc}
+\tablewidth{0pt}
+\tablecaption{A summary of the tuneable parameters in these simulations and their fiducial values.\label{tab:parameters}}
+\tablehead{
+\colhead{Parameter} & \colhead{Fiducial Value}
+}
+\startdata
+Source Density -- Eqn. \ref{eqn:power_law} (deg$^{-2}$) & $a=-13.05, b=1.25, c=-0.02$ \\
+Survey Area (deg$^2$)& $8 \times 8$ \\
+Source Density (deg$^{-2}$) & $ d = 100 $ \\
+Saturation Limit (mag) & $m_\text{min} = 17$\\
+10$\upsigma$ Detection Limit (mag)& $m_\text{max} = 22$\\
+Field-of-View (deg$^{2}$)& $0.76 \times 0.72$ \\
+Noise Model -- Eqn. \ref{eqn:noise} & $\alpha = 0.1585$, $\eta = 0.0017$, $\epsilon_\text{max} = 1.0$\\
+Fitted Instrument Response Model & 8$^{\text{th}}$ order polynomial\\
+\enddata
+\end{deluxetable}
 

### 2011-10-28 paper/self_calibration_I.tex
euclid: add strategy description to paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19810 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 
-\shorttitle{Optimizing Large Scale Imaging Surveys for Relative Photometric Calibration}
+\shorttitle{Optimizing Large Scale Imaging Surveys for a Retrospective Relative Photometric Calibration}
 \shortauthors{Holmes et al.}
@@
 
-\title{Optimizing Large Scale Imaging Surveys for Relative Photometric Calibration}
+\title{Optimizing Large Scale Imaging Surveys for a Retrospective Relative Photometric Calibration}
 
@@
 
-\keywords{Relative Photometric Calibration: Imaging Survey: Euclid}
+\keywords{Relative Photometric Calibration: Imaging Survey: Survey Strategy}
 
@@
 Astronomers tend to think in terms of taking science data and calibration data separately.
-The former is used for science and the latter is used to constrain instrument parameters like the flat-field, the dark currents and so-on.
-But typically far more photons are collected during science exposures; are these not incredibly constraining on the calibration themselves?
-Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS) much more calibration information was obtained from the science data than the calibration data \citep{pad08}.
+The former is used for science and the latter is used to constrain instrument parameters, such as the instrument response, the dark currents and so-on.
+But typically far more photons are collected during science exposures themselves; are these not incredibly constraining on the calibration?
+Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS), much more calibration information was obtained from the science data than the calibration data \citep{pad08}.
 But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: good redundancy is required in the data stream, and a redundancy of a very specific kind.
 
-In this paper, we argue that the next generation of large surveys should have their observation strategies optimized from the very start with this ``self-calibration'' in mind.
-We focus here on the relative photometric calibration of a typical imaging instrument only, although we also suspect that similar techniques could be used to constrain many other calibration parameters.
-This technique utilizes the multiple measurements of (non-varying) sources at different focal plane positions and at different times within the survey to constrain the relative instrument response by requiring that they yield the same count values.
-Through end-to-end -- catalog level -- survey simulations\footnote{All of the code used in this work is publicly available at XX.}, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.
+In this paper, we argue that the next generation of large-scale imaging surveys should have their observation strategies optimized from the very start with this kind of ``self-calibration'' in mind.
+This work focuses on the relative photometric calibration of a typical imaging instrument only, although we also suspect that similar techniques could be used to constrain many other calibration parameters.
+The self-calibration technique utilizes the multiple measurements of (non-varying) sources, at different focal plane positions and at different times within the survey, to constrain the relative instrument response by requiring that the post-calibration measurements yield the same flux values.
+Through end-to-end, catalog level survey simulations\footnote{All of the code used in this work is publicly available at XX.}, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.
 
-These simulations include a complex, position dependent instrument response; through mock observations of a synthetic sky, according to a defined survey strategy, we construct a realistic survey dataset.
-We then fit a (different) instrument response model that best describes this dataset.
+We construct realistic survey datasets through mock observations of a synthetic sky according to a specified survey strategy.
+These simulations include a complex, position dependent instrument response for the imaging instrument.
+Through the self-calibration procedure, we then recover this instrument response by fitting a model that best describes the survey dataset.
 By comparing the \textit{fitted} instrument response to the \textit{true} instrument response we are able to assess the performance of the self-calibration procedure with different survey strategies.
 There is a degeneracy in this problem: the self-calibration procedure is only able to fit for a \textit{relative} instrument response.
-We have no way of know, for example, if all the sources are uniformly fainter, or if the instrument sensitivity is uniformly lower.
+We have no way to know, for example, if all the sources are uniformly fainter, or if the instrument sensitivity is uniformly lower.
 
@@
 \section{Survey Dataset Simulations}
-\label{sec:survey}
+\label{sec:survey_simulations}
 We have constructed an end-to-end simulation chain that produces a realistic imaging dataset from a specified survey strategy.
@@
 
-\section{The Simple Survey Strategies}
-\label{sec:simple}
+\section{Simple Survey Strategies}
+\label{sec:simple_surveys}
+In this paper, we consider four simple but very different survey strategies.
+We use the end-to-end dataset simulation and self-calibration chain, introduced in Sections \ref{sec:survey_simulations} and \ref{sec:self_cal}, to evaluate the performance of the self-calibration procedure with these strategies.
+
+\subsection{Survey Description}
+We label the four survey strategies -- which all cover the same patch of sky -- with the letters A to D.
+These survey strategies are summarized in Table \ref{tab:simple_surveys} and are shown in Figure \ref{fig:simple_surveys}.
+Strategy A is the simplest strategy; the field is regularly tiled with small overlaps between adjacent pointings ($\sim 5$~\% and $\sim 8$~\% in the camera pointing directions $\alpha$ and $\beta$ respectively).
+The pointings in the 12 passes over the same field are exactly aligned.
+Survey B is the same as A, but with each pass over the field the orientation of the telescope is rotated by $30^\circ$.
+Survey C is more complex.
+The first pass over the field is the same as in Survey A, with $12 \times 12$ pointings.
+In the next pass, one of the pointings in the $\alpha$ direction is removed and one is added in the $\beta$, so the resultant pointing grid is $13 \times 11$.
+In the third pass over the field this change is reversed and the field is measured on a $11 \times 13$ grid.
+These three passes are then repeated four times.
+The pointing positions in Survey D are quasi-random: the pointings are the same as with Survey A, but each has a random offset within [-0.35,0.35)~deg applied in both the $\alpha$ and $\beta$ directions.
+By fixing the pointings within these $0.7~\text{deg} \times 0.7~\text{deg}$ boxes, we ensure that the quasi random strategy has a uniform coverage of the field.
+The orientations of the pointings in Survey D are completely random.
+
+\subsection{Self-Calibration Performance}
 
@@
 \tablehead{
-\colhead{Parameter} & \colhead{Fiducial Value}
-}
+\colhead{Parameter} & \colhead{Fiducial Value}}
 \startdata
@@
 \clearpage
-
-\begin{deluxetable}{ccrrrrrrrrcrl}
-\tabletypesize{\scriptsize}
-\rotate
-\tablecaption{Sample table taken from \citet{treu03}\label{tbl-1}}
+\begin{deluxetable}{cccc}
 \tablewidth{0pt}
+\tablecaption{A summary of the four simple survey strategies considered in this paper. \label{tab:simple_surveys}}
 \tablehead{
-\colhead{POS} & \colhead{chip} & \colhead{ID} & \colhead{X} & \colhead{Y} &
-\colhead{RA} & \colhead{DEC} & \colhead{IAU$\pm$ $\delta$ IAU} &
-\colhead{IAP1$\pm$ $\delta$ IAP1} & \colhead{IAP2 $\pm$ $\delta$ IAP2} &
-\colhead{star} & \colhead{E} & \colhead{Comment}
-}
+\colhead{Survey Label} & \colhead{Pointing Center} & \colhead{Orientation} & \colhead{Number of Pointings}}
 \startdata
-0 & 2 & 1 & 1370.99 & 57.35    &   6.651120 &  17.131149 & 21.344$\pm$0.006  & 2
-4.385$\pm$0.016 & 23.528$\pm$0.013 & 0.0 & 9 & -    \\
-0 & 2 & 2 & 1476.62 & 8.03     &   6.651480 &  17.129572 & 21.641$\pm$0.005  & 2
-3.141$\pm$0.007 & 22.007$\pm$0.004 & 0.0 & 9 & -    \\
-0 & 2 & 3 & 1079.62 & 28.92    &   6.652430 &  17.135000 & 23.953$\pm$0.030  & 2
-4.890$\pm$0.023 & 24.240$\pm$0.023 & 0.0 & - & -    \\
-0 & 2 & 4 & 114.58  & 21.22    &   6.655560 &  17.148020 & 23.801$\pm$0.025  & 2
-5.039$\pm$0.026 & 24.112$\pm$0.021 & 0.0 & - & -    \\
-0 & 2 & 5 & 46.78   & 19.46    &   6.655800 &  17.148932 & 23.012$\pm$0.012  & 2
-3.924$\pm$0.012 & 23.282$\pm$0.011 & 0.0 & - & -    \\
-0 & 2 & 6 & 1441.84 & 16.16    &   6.651480 &  17.130072 & 24.393$\pm$0.045  & 2
-6.099$\pm$0.062 & 25.119$\pm$0.049 & 0.0 & - & -    \\
-0 & 2 & 7 & 205.43  & 3.96     &   6.655520 &  17.146742 & 24.424$\pm$0.032  & 2
-5.028$\pm$0.025 & 24.597$\pm$0.027 & 0.0 & - & -    \\
-0 & 2 & 8 & 1321.63 & 9.76     &   6.651950 &  17.131672 & 22.189$\pm$0.011  & 2
-4.743$\pm$0.021 & 23.298$\pm$0.011 & 0.0 & 4 & edge \\
+A  & Uniform Grid ($12\times12$) & $0^\circ$ & 1728 \\
+B      & Uniform Grid ($12\times12$)     & Each Pass: $\theta +30^\circ$& 1728 \\
+C   & Pass 1:  Uniform Grid ($12\times12$)  &  & 1720\\
+ & Pass 2:  Uniform Grid ($13§\times11$) & $0^\circ$&\\
+  & Pass 3:  Uniform Grid ($11§\times13$) & &\\
+D   & Quasi-Random  & Random & 1728\\
 \enddata
-%% Text for table notes should follow after the \enddata but before
-%% the \end{deluxetable}. Make sure there is at least one \tablenotemark
-%% in the table for each \tablenotetext.
-\tablecomments{Table \ref{tbl-1} is published in its entirety in the 
-electronic edition of the {\it Astrophysical Journal}.  A portion is 
-shown here for guidance regarding its form and content.}
-\tablenotetext{a}{Sample footnote for table~\ref{tbl-1} that was generated
-with the deluxetable environment}
-\tablenotetext{b}{Another sample footnote for table~\ref{tbl-1}}
 \end{deluxetable}
 
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{simple_surveys.pdf}
+\end{center}
+\caption{Focal plane footprints projected onto the synthetic sky according to the four simple survey strategies described in Section \ref{sec:simple_surveys} and summarized in Table \ref{tab:simple_surveys}.
+Surveys A, B and D have 1728 pointings and survey C has 1720 pointings.\label{fig:simple_surveys}}
+\end{figure}
 %% If you use the table environment, please indicate horizontal rules using
@@
 %% The optional \label should appear inside the \caption command.
-
-\clearpage
 
@@
 %% show how many observations can be tagged in a table.
-
-\clearpage
-
-\begin{deluxetable}{lrrrrcrrrrr}
-\tablewidth{0pt}
-\tablecaption{Literature Data for Program Stars}
-\tablehead{
-\colhead{Star}           & \colhead{V}      &
-\colhead{b$-$y}          & \colhead{m$_1$}  &
-\colhead{c$_1$}          & \colhead{ref}    &
-\colhead{T$_{\rm eff}$}  & \colhead{log g}  &
-\colhead{v$_{\rm turb}$} & \colhead{[Fe/H]} &
-\colhead{ref}}
-\startdata
-HD 97 & 9.7& 0.51& 0.15& 0.35& 2 & \nodata & \nodata & \nodata & $-1.50$ & 2 \\
-& & & & & & 5015 & \nodata & \nodata & $-1.50$ & 10 \\
-\dataset[ADS/Sa.HST#O6H04VAXQ]{HD 2665} & 7.7& 0.54& 0.09& 0.34& 2 & \nodata & \nodata & \nodata & $-2.30$ & 2 \\
-& & & & & & 5000 & 2.50 & 2.4 & $-1.99$ & 5 \\
-& & & & & & 5120 & 3.00 & 2.0 & $-1.69$ & 7 \\
-& & & & & & 4980 & \nodata & \nodata & $-2.05$ & 10 \\
-HD 4306 & 9.0& 0.52& 0.05& 0.35& 20, 2& \nodata & \nodata & \nodata & $-2.70$ & 2 \\
-& & & & & & 5000 & 1.75 & 2.0 & $-2.70$ & 13 \\
-& & & & & & 5000 & 1.50 & 1.8 & $-2.65$ & 14 \\
-& & & & & & 4950 & 2.10 & 2.0 & $-2.92$ & 8 \\
-& & & & & & 5000 & 2.25 & 2.0 & $-2.83$ & 18 \\
-& & & & & & \nodata & \nodata & \nodata & $-2.80$ & 21 \\
-& & & & & & 4930 & \nodata & \nodata & $-2.45$ & 10 \\
-HD 5426 & 9.6& 0.50& 0.08& 0.34& 2 & \nodata & \nodata & \nodata & $-2.30$ & 2 \\
-\dataset[ADS/Sa.HST#O5F654010]{HD 6755} & 7.7& 0.49& 0.12& 0.28& 20, 2& \nodata & \nodata & \nodata & $-1.70$ & 2 \\
-& & & & & & 5200 & 2.50 & 2.4 & $-1.56$ & 5 \\
-& & & & & & 5260 & 3.00 & 2.7 & $-1.67$ & 7 \\
-& & & & & & \nodata & \nodata & \nodata & $-1.58$ & 21 \\
-& & & & & & 5200 & \nodata & \nodata & $-1.80$ & 10 \\
-& & & & & & 4600 & \nodata & \nodata & $-2.75$ & 10 \\
-\dataset[ADS/Sa.HST#O56D06010]{HD 94028} & 8.2& 0.34& 0.08& 0.25& 20 & 5795 & 4.00 & \nodata & $-1.70$ & 22 \\
-& & & & & & 5860 & \nodata & \nodata & $-1.70$ & 4 \\
-& & & & & & 5910 & 3.80 & \nodata & $-1.76$ & 15 \\
-& & & & & & 5800 & \nodata & \nodata & $-1.67$ & 17 \\
-& & & & & & 5902 & \nodata & \nodata & $-1.50$ & 11 \\
-& & & & & & 5900 & \nodata & \nodata & $-1.57$ & 3 \\
-& & & & & & \nodata & \nodata & \nodata & $-1.32$ & 21 \\
-HD 97916 & 9.2& 0.29& 0.10& 0.41& 20 & 6125 & 4.00 & \nodata & $-1.10$ & 22 \\
-& & & & & & 6160 & \nodata & \nodata & $-1.39$ & 3 \\
-& & & & & & 6240 & 3.70 & \nodata & $-1.28$ & 15 \\
-& & & & & & 5950 & \nodata & \nodata & $-1.50$ & 17 \\
-& & & & & & 6204 & \nodata & \nodata & $-1.36$ & 11 \\
-\cutinhead{This is a cut-in head}
-+26\arcdeg2606& 9.7&0.34&0.05&0.28&20,11& 5980 & \nodata & \nodata &$<-2.20$ & 19 \\
-& & & & & & 5950 & \nodata & \nodata & $-2.89$ & 24 \\
-+26\arcdeg3578& 9.4&0.31&0.05&0.37&20,11& 5830 & \nodata & \nodata & $-2.60$ & 4 \\
-& & & & & & 5800 & \nodata & \nodata & $-2.62$ & 17 \\
-& & & & & & 6177 & \nodata & \nodata & $-2.51$ & 11 \\
-& & & & & & 6000 & 3.25 & \nodata & $-2.20$ & 22 \\
-& & & & & & 6140 & 3.50 & \nodata & $-2.57$ & 15 \\
-+30\arcdeg2611& 9.2&0.82&0.33&0.55& 2 & \nodata & \nodata & \nodata & $-1.70$ & 2 \\
-& & & & & & 4400 & 1.80 & \nodata & $-1.70$ & 12 \\
-& & & & & & 4400 & 0.90 & 1.7 & $-1.20$ & 14 \\
-& & & & & & 4260 & \nodata & \nodata & $-1.55$ & 10 \\
-+37\arcdeg1458& 8.9&0.44&0.07&0.22&20,11& 5296 & \nodata & \nodata & $-2.39$ & 11 \\
-& & & & & & 5420 & \nodata & \nodata & $-2.43$ & 3 \\
-+58\arcdeg1218&10.0&0.51&0.03&0.36& 2 & \nodata & \nodata & \nodata & $-2.80$ & 2 \\
-& & & & & & 5000 & 1.10 & 2.2 & $-2.71$ & 14 \\
-& & & & & & 5000 & 2.20 & 1.8 & $-2.46$ & 5 \\
-& & & & & & 4980 & \nodata & \nodata & $-2.55$ & 10 \\
-+72\arcdeg0094&10.2&0.31&0.09&0.26&12 & 6160 & \nodata & \nodata & $-1.80$ & 19 \\
-\sidehead{I'm a side head:}
-G5--36 & 10.8& 0.40& 0.07& 0.28& 20 & \nodata & \nodata & \nodata & $-1.19$ & 21 \\
-G18--54 & 10.7& 0.37& 0.08& 0.28& 20 & \nodata & \nodata & \nodata & $-1.34$ & 21 \\
-G20--08 & 9.9& 0.36& 0.05& 0.25& 20,11& 5849 & \nodata & \nodata & $-2.59$ & 11 \\
-& & & & & & \nodata & \nodata & \nodata & $-2.03$ & 21 \\
-G20--15 & 10.6& 0.45& 0.03& 0.27& 20,11& 5657 & \nodata & \nodata & $-2.00$ & 11 \\
-& & & & & & 6020 & \nodata & \nodata & $-1.56$ & 3 \\
-& & & & & & \nodata & \nodata & \nodata & $-1.58$ & 21 \\
-G21--22 & 10.7& 0.38& 0.07& 0.27& 20,11& \nodata & \nodata & \nodata & $-1.23$ & 21 \\
-G24--03 & 10.5& 0.36& 0.06& 0.27& 20,11& 5866 & \nodata & \nodata & $-1.78$ & 11 \\
-& & & & & & \nodata & \nodata & \nodata & $-1.70$ & 21 \\
-G30--52 & 8.6& 0.50& 0.25& 0.27& 11 & 4757 & \nodata & \nodata & $-2.12$ & 11 \\
-& & & & & & 4880 & \nodata & \nodata & $-2.14$ & 3 \\
-G33--09 & 10.6& 0.41& 0.10& 0.28& 20 & 5575 & \nodata & \nodata & $-1.48$ & 11 \\
-G66--22 & 10.5& 0.46& 0.16& 0.28& 11 & 5060 & \nodata & \nodata & $-1.77$ & 3 \\
-& & & & & & \nodata & \nodata & \nodata & $-1.04$ & 21 \\
-G90--03 & 10.4& 0.37& 0.04& 0.29& 20 & \nodata & \nodata & \nodata & $-2.01$ & 21 \\
-LP 608--62\tablenotemark{a} & 10.5& 0.30& 0.07& 0.35& 11 & 6250 & \nodata &
-\nodata & $-2.70$ & 4 \\
-\enddata
-\tablenotetext{a}{Star LP 608--62 is also known as BD+1\arcdeg 2341p.  We will
-make this footnote extra long so that it extends over two lines.}
-%% You can append references to a table using the \tablerefs command.
-\tablerefs{
-(1) Barbuy, Spite, \& Spite 1985; (2) Bond 1980; (3) Carbon et al. 1987;
-(4) Hobbs \& Duncan 1987; (5) Gilroy et al. 1988: (6) Gratton \& Ortolani 1986;
-(7) Gratton \& Sneden 1987; (8) Gratton \& Sneden (1988); (9) Gratton \& Sneden 1991;
-(10) Kraft et al. 1982; (11) LCL, or Laird, 1990; (12) Leep \& Wallerstein 1981;
-(13) Luck \& Bond 1981; (14) Luck \& Bond 1985; (15) Magain 1987;
-(16) Magain 1989; (17) Peterson 1981; (18) Peterson, Kurucz, \& Carney 1990;
-(19) RMB; (20) Schuster \& Nissen 1988; (21) Schuster \& Nissen 1989b;
-(22) Spite et al. 1984; (23) Spite \& Spite 1986; (24) Hobbs \& Thorburn 1991;
-(25) Hobbs et al. 1991; (26) Olsen 1983.}
-\end{deluxetable}
 

### 2011-10-31 paper/self_calibration_I.tex
euclid: minor edits to paper draft, start developement brach
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19826 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 
-In Section \ref{sec:survey} we introduce the simulation chain constructed to produce the realistic survey datasets.
+In Section \ref{sec:survey_simulations} we introduce the simulation chain constructed to produce the realistic survey datasets.
 Section \ref{sec:self_cal} goes onto to detail the self-calibration procedure.
-In Section \ref{sec:simple} we focus on four simple survey strategies, which allow us to draw conclusions on the performance of the self-calibration procedure with different survey properties.
+In Section \ref{sec:simple_surveys} we focus on four simple survey strategies, which allow us to draw conclusions on the performance of the self-calibration procedure with different survey properties.
 We do not produced pixelated images; instead we concentrate on catalog level simulations with realistic measurement uncertainties.
@@
 We have constructed an end-to-end simulation chain that produces a realistic imaging dataset from a specified survey strategy.
-Simulation parameters are kept intentionally flexible, so that the sensitivity of the self-calibration procedure to different parameters can be investigated.
+In this work we have kept the simulation parameters intentionally flexible, so that the sensitivity of the self-calibration procedure to different values can be investigated.
 The simulations are split into a number of steps.
-First, a synthetic sky is generated based on the specified simulation parameters (density of sources, magnitude ranges).
-With a given pointing, single mock observations are performed on this sky, again here with tunable simulation parameters, such as the instrument response model and the the noise model.
-The final step is the build up of the survey wide dataset from multiple single observations taken according to the prescribed survey strategy.
+The first of which is the generation of a synthetic sky.
+With a given pointing, single mock observations are performed on this sky.
+To build up a survey wide dataset, multiple single observations are performed according to the specified survey strategy.
 In this section, we detail the assumptions and methods used in each of these steps.
@@
 where $\dfrac{\dd N}{\dd m \, \dd  \Omega}$  is the density of sources $N$ per unit magnitude $m$ and per unit solid angle $\Omega$, and $a$, $b$ and $c$ are model parameters.
-Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in Windhorst et al.
-(2011) \citep{win11} only.
-These parameters were $a = -13.05$, $b = 1.25$ and $c = -0.02$.
+Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in \citep{win11} only.
+These parameters were found to be $a = -13.05$, $b = 1.25$ and $c = -0.02$.
 We intentionally omit the stellar population, so that the conclusions drawn from these simulations are independent of the final position of the survey area on the sky.
@@
 \label{sec:single_exposure}
-With a camera pointing $(\alpha, \beta)$ and camera orientation $\theta$ this synthetic sky is transformed into focal plane coordinates and all of the sources falling within the instrument's field-of-view are found.
+With a camera pointing $(\alpha, \beta)$ and camera orientation $\theta$ this synthetic sky is transformed into focal plane image.
 In our simulations, we use a large instrument field-of-view of $0.76~ \text{deg} \times 0.72~\text{deg}$; a size consistent with the next generation of large survey imagers.
@@
 \end{center}
-\caption[A single camera pointing from the self-calibration simulations.]{A single exposure of the synthetic sky.
-Left: A plot of the bright sources within the synthetic sky used in the self-calibration procedure with the focal plane footprint overlaid.
+\caption{A single exposure of the synthetic sky.
+Left: A plot of the bright sources within the synthetic sky used in the self-calibration procedure, with the focal plane footprint overlaid.
 Right: The resultant distribution of the sources on the instrument's focal plane.
-The \textit{true} instrument model $f_{\true}(\vec{x_i} | \vec{q}_\true)$ is shown as contours.\label{fig:single_image}}
+The \textit{true} instrument model $f_{\true}(\vec{x_i} | \vec{q}_\true)$ is shown with contours.\label{fig:single_image}}
 \end{figure}

### 2011-10-31 dev/default_parameters_test.py

@@ -1,0 +1,72 @@
+# TODO
+# 	Change alpha and beta
+# 	Add mode = general, density compare, universal plot
+
+# Parameters for self-calibration simulations
+{
+########################################################
+############## Simulation Parameters ###################
+########################################################
+
+# Simulation mode: 
+#		- 'general' = performs self-calibration for 
+#		- 'density_compare' = creates density plots comparing TWO parameters -- XX which?, uses 1st value for each parameter 
+#		- 'universal' = creates a universal plot?!
+	'simulation_mode' 		:			'general'
+# Survey Strategies to investigate  
+  'survey_strategies'   :     ['D', 'C', 'A'], #['D', 'C', 'B', 'A']
+
+########################################################
+################## God Parameters ######################
+########################################################
+
+# Random number generation seed for sky simulation
+  'seed'                :     1, # Random number seed
+
+# Density of Sources to include in the simulations (only brightest sources are used)
+  'density_of_sources'    :     np.linspace(1, 100, 10).astype('int'), # can run out of memory quite quickly! 100 is ok, 1000 can be run on broiler but is quite slow
+
+# Distribution of source magnitudes: log10(dN/dm) = A + B*mag + C*mag**2
+	'powerlaw_constants'  :     [np.array([-13.34863146, 1.25429311, -0.02122949]), np.array([1., 2., 3.])], # fit to Windhorst et al 2011 
+
+# Region of sky to survey
+	'sky_limits'          :     [np.array(-4.0, 4.0, -4.0, 4.0)], 
+  
+########################################################
+############### Instrument Parameters ##################
+########################################################
+
+# Saturation Limit
+	'm_min'               :     np.array([17]),
+
+# 10 sigma detection limit
+	'm_max'               :     np.array([22]),  
+
+# Instrument Field-of-View
+	'FoV'                 :     [np.array([0.76, 0.72])], 
+
+# Noise Model
+	'noise'								:			np.array([0.00173214, 0.1584655417]) # [eta, alpha]
+
+# Additional noise term not taken into account in the analysis
+   'epsilon_max'         :     np.array([1.]),
+  
+########################################################
+################ Analysis Parameters ###################
+########################################################
+
+# Number of samples of instrument response used during the badness calculations  
+  'ff_samples'          :     [300, 300],
+
+# What order polynomial to fit instrument response?
+  'flat_field_order'    :     8,
+  
[…]

### 2011-10-31 dev/fork_test.py

@@ -1,0 +1,48 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+# Todo:
+#	change density of stars ==> density_of_sources 
+#	
+
+import numpy as np
+import os
+import pickle
+import sys
+from multiprocessing import Pool
+import copy
+# Custom Modules
+
+def get_commandline_inputs(): # Read inputs from command line
+	if len(sys.argv) == 3:
+  		out_dir = sys.argv[1] # directory to save output data to - always needed
+  		if out_dir[-1] == '/': out_dir = out_dir[0:-1]
+  		param_file = sys.argv[2] # parameter file to use in simulation run
+	else:
+		print "Error - wrong number of command line arguements!"
+		sys.exit() 	
+  	return [out_dir, param_file]
+
+if __name__ == '__main__':
+	# Make output directory and load parameter dictionary
+	out_dir, param_file = get_commandline_inputs()
+	os.system(('mkdir -p %s' % out_dir))
+	f = open(param_file,'r') 
+	params = eval(f.read())
+
+	number_of_runs = 0
+	print "Parsing Parameter File: %s" % "test"
+	for survey_strategies in params['survey_strategies']:
+		for powerlaw_constants in params['powerlaw_constants']:
+			for density_of_sources in params['density_of_sources']:
+				print powerlaw_constants
+				number_of_runs += 1
+		
+	print params
+	print ("%d runs will now be launched..." % number_of_runs)
+
+# Do i need to create new god's? if so do, if not generate one with long filename
+
+# .god
+# .results
+# .params 

### 2011-11-04 paper/self_calibration_I.tex
euclid: trying to resolve conflict problem
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19859 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \newcommand{\basis}{\text{best}}
+\newcommand{\rms}{\text{RMS}}
 \newcommand{\dd}{\text{d}}
@@
 The $m = 22$ mag 10$\upsigma$ detection limit introduced previously and the 500 limit on the signal-to-noise ratio are used to set $\alpha = 0.1585$ and $\eta = 0.0017$.
-The $\epsilon_i$ contribution is not taken into account in the analysis stage and therefore the uncertainty variances on the count rates are assumed to be
+The $\epsilon_i$ contribution is not taken into account in the analysis and therefore the uncertainty variances on the count rates are assumed to be
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-during the analysis stage.
+during this stage.
 
@@
 with the physical focal plane dimensions $X$ and $Y$.
-In total, the instrument response model is parameterized with 260 parameters; an example can be seen in Figure \ref{fig:single_image}(right).
+In total, the instrument response model is parameterized with 260 parameters; an example can be seen in Figure \ref{fig:single_image}      (right).
 It is this instrument response that we try and recover with the self-calibration procedure.
@@
 \subsection{The Complete Survey}
-In this work we are interested in simulating complete surveys, in order to identify the crucial characteristics of survey strategies that allows for the relative instrument response model to be accurately constrained from the resulting dataset.
+In this work we are interested in simulating complete surveys, in order to identify the crucial characteristics of survey strategies that allow for the relative instrument response model to be accurately, retrospectively constrained from the resulting dataset.
 We therefore apply the single exposure procedure introduced in the previous section for each pointing specified in a defined survey strategy.
@@
 It is not possible to know, for example, if the sources are all fainter or if the instrument response is uniformly lower.
-This degeneracy can be broken through observations of standard absolute sources.
+In practice, this degeneracy can be broken through observations of standard absolute sources.
 
@@
 To do this we defined three quantities.
-The first is the root-mean-squared (RMS) error $S_\text{RMS}$ in the final fitted source fluxes $s_{k,\fit}$ compared to the true source fluxes $s_{k,\true}$ for all the $K$ sources:
+The first is the root-mean-squared (RMS) error $S_\text{RMS}$ in the final fitted source fluxes $s_{k,\fit}$ compared to the true source fluxes $s_{k,\true}$ for the $K$ sources used in the self-calibration procedure:
 \begin{equation}
@@
 \end{equation}
-The other two metrics, called ``badnesses'', are defined as the RMS error between the final fitted instrument response and a reference instrument response sampled on a regular $500 \times 500$ grid across the focal plane.
+The other two metrics, called ``badnesses'', are defined as the RMS error between the final fitted instrument response and a reference instrument response sampled on a regular $300 \times 300$ grid across the focal plane.
 For the ``True Badness'' $B_\true$, the \textit{fitted} instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ is compared to the \textit{true} instrument response $f_\true(\vec{x} | \vec{q_\true})$ at the $J$ sample points
@@
 The first pass over the field is the same as in Survey A, with $12 \times 12$ pointings.
-In the next pass, one of the pointings in the $\alpha$ direction is removed and one is added in the $\beta$, so the resultant pointing grid is $13 \times 11$.
+In the next pass, one of the pointings in the $\alpha$ direction is removed and one is added in the $\beta$ and the pointings are respaced uniformly, so the resultant pointing grid is $13 \times 11$.
 In the third pass over the field this change is reversed and the field is measured on a $11 \times 13$ grid.
@@
 \subsection{Self-Calibration Performance}
+The iterative self-calibration procedure converges to a final fitted solution for the Survey B, C and D datasets.
+With the Survey A dataset it does not converge, even when the system is started close to the optimum fit.
+The fitted instrument response solutions for the Survey D and A datasets are shown in Figures \ref{fig:D_result} and \ref{fig:A_result} respectively.
+The corresponding plot for the Survey B and C dataset has been omitted, due to its similarity to the Survey D dataset result.
+In these plots the final fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ is compared to the true instrument response $f_\true(\vec{x} | \vec{q_\true})$ and to the best possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$ (``best-in-basis'').
+The accuracy of the final fitted solutions are summarized, with the metrics introduced in Section \ref{sec:metrics}, for each of the Survey strategies in Table \ref{tab:simple_results}.
+
+FROM HERE NEEDS TO BE UPDATED
+
+The difference between the success of the self-calibration procedure with these four survey strategies is clear: with Surveys A no solution can be obtained, but with the Surveys B, C and D datasets the self-calibration procedure converges close to the correct solution.
+With the Survey C and D datasets the instrument response can be accurately fitted, close to the best fit possible with the basis used.
+The majority of the remaining errors in the fit come from the limitations of the basis used in the instrument response model.
+The calibration procedure with the Survey B dataset requires more iteration steps (4270) to converge to a final solution, and is less accurate than those found with the Survey C and D datasets, which require 28 and 10 iteration steps respectively.
+
+The results from this investigation into simple survey strategies can be used to draw conclusions about the properties of surveys that make them good, or bad, for retrospective self-calibration.
+From these examples, I conclude that regular survey strategies with little overlap between adjacent pointings (Survey A) -- such as the Euclid Wide Survey -- do not allow for the resulting dataset to be self-calibrated effectively.
+It is therefore necessary to depart from the standard, regularly tiled observing strategies in the Euclid Deep Field.
+The two best survey strategies (C, D) both return the same sources to very different focal plane positions, and this is the key property which makes their datasets good for self-calibration.
+By doing so, many different focal plane positions are connected to each other with observations of the same source.
+This result must be considered when the Deep Field survey strategy is defined and as such I have applied calibration requirements that capture this on the strategy (see Subsection \ref{sec:deep_req}).
+In addition, I also conclude that if the correct redundancy is built into the survey strategy, the self-calibration procedure will be able to accurately constrain the relative instrument response, and therefore the Deep Field relative photometric calibration could be transfered effectively to the Wide Survey.
+The majority of the remaining errors will come from the limitations of the basis used to fit the instrument response model.
 
@@
 \end{figure}
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{D_results.pdf}
+\end{center}
+\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey D dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+(a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response.
+(c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model.
+The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial).
+The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:D_result}}
+\end{figure}
+
+\clearpage
+\begin{deluxetable}{ccccc}
+\tablecaption{A summary of the quality of the final fit from the self-calibration procedure with the four simple survey strategies summarized in Table \ref{tab:simple_surveys}.
+The self-calibration procedure was run with only the brightest sources within the survey area; the source error $S_\text{RMS}$ corresponds to the measurements of these sources only and not all those within the survey.
+*Did not converge.\label{tab:simple_results}}
+\tablehead{
+\colhead{Survey Name} & \colhead{Iterations} & \colhead{Source Error $S_\rms$(\%)} & \colhead{True Badness $B_\true$ (\%)} & \colhead{Best-in-Basis Badness $basis$ (\%)} }
+\startdata
+A*  & 8192 & XX & XX & XX \\
+B  & 4160 & XX & XX & XX \\
+C  & 26 & XX & XX & XX \\
+D  & 9 & XX & XX & XX \\
+\enddata
+\end{deluxetable}
+
 %% If you use the table environment, please indicate horizontal rules using

### 2011-11-04 paper/self_calibration_I.tex
euclid: start adding rms, badness values from new results
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19862 03a3cea6-2c03-0410-ac9b-9271b1c66c29
euclid: adding results to paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19863 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 With the Survey A dataset it does not converge, even when the system is started close to the optimum fit.
-The fitted instrument response solutions for the Survey D and A datasets are shown in Figures \ref{fig:D_result} and \ref{fig:A_result} respectively.
-The corresponding plot for the Survey B and C dataset has been omitted, due to its similarity to the Survey D dataset result.
+The fitted instrument response solutions for the Surveys A to D datasets are shown in Figures \ref{fig:A_result} to \ref{fig:D_result}.
 In these plots the final fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ is compared to the true instrument response $f_\true(\vec{x} | \vec{q_\true})$ and to the best possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$ (``best-in-basis'').
@@
 FROM HERE NEEDS TO BE UPDATED
+
+The self-calibration procedure performs best with the Survey D dataset, followed by the Survey C dataset, then the Survey B dataset.
+The self-calibration procedure failed completely with the Survey A dataset and no solution was obtained.
+
+With the Survey C and D datasets, the instrument response is accurately recovered.
+As can be seen from Figs.
+\ref{fig:C_result} and \ref{fig:D_result} (top) the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level.
+That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the basis used.
+The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the Best-in-Basis Badness $B_\basis$ compared to the True Badness $B_\true$.
+The results from these two survey datasets are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the instrument response model fitted for.
+
+The results from the self-calibration procedure with the Survey A and B datasets are also interesting.
+
+tiling is crap
+
+parts well connected work well
+
+ when the fitted instrument response is compared to the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we can see that the final solution is close
+
+possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{D_results.pdf}
+\includegraphics[width=\textwidth]{A_8192_ff.pdf}
+\end{center}
+\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey A dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+The self-calibration procedure did not converged to the final solution, the results presented here correspond to the instrument response found after $8\,192$ iterations.
+(a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response.
+(c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model.
+The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial).
+The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:A_result}}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{B_4160_ff.pdf}
+\end{center}
+\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey B dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+The self-calibration procedure converged to the final solution after 4160 iterations.
+(a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response.
+(c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model.
+The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial).
+The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:B_result}}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{C_026_ff.pdf}
+\end{center}
+\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey C dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+The self-calibration procedure converged to the final solution after 26 iterations.
+(a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response.
+(c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model.
+The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial).
+The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:C_result}}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=\textwidth]{D_009_ff.pdf}
 \end{center}
 \caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey D dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+The self-calibration procedure converged to the final solution after 9 iterations.
 (a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response.
@@
 \tablehead{
-\colhead{Survey Name} & \colhead{Iterations} & \colhead{Source Error $S_\rms$(\%)} & \colhead{True Badness $B_\true$ (\%)} & \colhead{Best-in-Basis Badness $basis$ (\%)} }
+\colhead{Survey Name} & \colhead{Iterations} & \colhead{Source Error $S_\rms$(\%)} & \colhead{True Badness $B_\true$ (\%)} & \colhead{Best-in-Basis Badness $B_\basis$ (\%)} }
 \startdata
@@
 B  & 4160 & XX & XX & XX \\
-C  & 26 & XX & XX & XX \\
-D  & 9 & XX & XX & XX \\
+C  & 26 & 0.199 & 0.305 & 0.043 \\
+D  & 9 & 0.113 & 0.304 & 0.032 \\
 \enddata

### 2011-11-04 paper/self_calibration_I.tex
euclid: paper draft 0.8 complete
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19865 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 
-FROM HERE NEEDS TO BE UPDATED
-
 The self-calibration procedure performs best with the Survey D dataset, followed by the Survey C dataset, then the Survey B dataset.
 The self-calibration procedure failed completely with the Survey A dataset and no solution was obtained.
+The number of iterations of the self-calibration procedure required to converge to the final solution also decreases as the quality of the final fit increases: Survey D required 9 iterations, Survey C 26, Survey B 4160 and the procedure was iterated XX times with Survey A without reaching a solution.
 
 With the Survey C and D datasets, the instrument response is accurately recovered.
-As can be seen from Figs.
-\ref{fig:C_result} and \ref{fig:D_result} (top) the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level.
+As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top) the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level.
 That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the basis used.
@@
 
-The results from the self-calibration procedure with the Survey A and B datasets are also interesting.
-
-tiling is crap
-
-parts well connected work well
-
- when the fitted instrument response is compared to the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we can see that the final solution is close
-
-possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$
-
-The difference between the success of the self-calibration procedure with these four survey strategies is clear: with Surveys A no solution can be obtained, but with the Surveys B, C and D datasets the self-calibration procedure converges close to the correct solution.
-With the Survey C and D datasets the instrument response can be accurately fitted, close to the best fit possible with the basis used.
-The majority of the remaining errors in the fit come from the limitations of the basis used in the instrument response model.
-The calibration procedure with the Survey B dataset requires more iteration steps (4270) to converge to a final solution, and is less accurate than those found with the Survey C and D datasets, which require 28 and 10 iteration steps respectively.
-
-The results from this investigation into simple survey strategies can be used to draw conclusions about the properties of surveys that make them good, or bad, for retrospective self-calibration.
-From these examples, I conclude that regular survey strategies with little overlap between adjacent pointings (Survey A) -- such as the Euclid Wide Survey -- do not allow for the resulting dataset to be self-calibrated effectively.
-It is therefore necessary to depart from the standard, regularly tiled observing strategies in the Euclid Deep Field.
-The two best survey strategies (C, D) both return the same sources to very different focal plane positions, and this is the key property which makes their datasets good for self-calibration.
-By doing so, many different focal plane positions are connected to each other with observations of the same source.
-This result must be considered when the Deep Field survey strategy is defined and as such I have applied calibration requirements that capture this on the strategy (see Subsection \ref{sec:deep_req}).
-In addition, I also conclude that if the correct redundancy is built into the survey strategy, the self-calibration procedure will be able to accurately constrain the relative instrument response, and therefore the Deep Field relative photometric calibration could be transfered effectively to the Wide Survey.
-The majority of the remaining errors will come from the limitations of the basis used to fit the instrument response model.
+The Survey A and B datasets also give interesting results.
+With Survey A, the self-calibration procedure did not tend to a solution.
+This, in itself, is an important result, as the naive tiling of the sky currently being considered for many large surveys will result in datasets that cannot be retrospectively self-calibrated.
+Minor changes to survey strategies, such as those in Survey C, can drastically improve the performance of the self-calibration procedure.
+Without these changes, vast amounts of calibration information is simply lost.
+We can also draw conclusions from the results obtained with Survey C: here the self-calibration procedure converges close to the correct solutions, albeit not as close as the Surveys C and B.
+Here, it is the quality of the fit over the focal plane that is interesting.
+From the residual plot against the best-in-basis instrument response $f_\basis(\vec{x_j} | \vec{q_\basis})$ (Fig \ref{fig:B_result} bottom right), we can see that the quality of the fit is rotationally symmetric, with the self-calibration procedure able to fit the instrument response well at the edge of the focal plane, but poorly at the center.
+This, along with the success of the procedure with the Surveys C and D, gives an insight into the key property of a survey that makes it good for self-calibration.
+Namely, the survey should return the same sources to many, very different focal plane positions.
+This allows the self-calibration procedure to link the different parts of the focal plane with each other through the observations of the same sources.
+Surveys C and D do this and the self-calibration procedure performs well with their dataset.
+Survey A does not do this; it only connects the outer edges of the focal plane to each other.
+There are no redundant measurements connecting the majority of the focal plane to other positions.
+This conclusion in confirmed with Strategy B.
+Here, positions at the outside of the focal plane are well connected to other positions and therefore the self-calibration procedure can well fit the instrument response in these regions.
+In the center of the focal plane, on which the same sources always fall, there are insufficient redundant measurements to allow for a good fit.
+Returning the same sources to very different focal plane positions is critical for the self-calibration to work effectively.
 
 \section{Discussion}
-In this paper, we have considered the optimization of survey strategies for calibrating the relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained for this method.
+From the simulations presented in this paper, we are draw two firm conclusions about the self-calibration of large-scale imaging surveys.
+The first is that, in agreement with previous work, we show that this technique is very powerful, allowing the relative instrument response of an imaging instrument to be accurately retrospectively constrained from the science images alone.
+The second is that there must be the correct redundancy built into the survey strategy.
+The same sources must be imaged at very different focal plane positions; regular tiling of the sky is useless for this kind of calibration.
+
+In this paper, we have considered the optimization of survey strategies for calibrating the relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained from this method.
 For example, the optical distortion of an instrument could also be constrained with such a method, although we concede that the properties of a survey strategy that makes them good for one calibration, may not be the same as that for another; ultimately a trade-off would need to be performed.

### 2011-11-07 paper/self_calibration_I.tex
euclid: paper version 0.81, ready for David and Hans-Walter to read
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19872 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 With the Survey A dataset it does not converge, even when the system is started close to the optimum fit.
-The fitted instrument response solutions for the Surveys A to D datasets are shown in Figures \ref{fig:A_result} to \ref{fig:D_result}.
-In these plots the final fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ is compared to the true instrument response $f_\true(\vec{x} | \vec{q_\true})$ and to the best possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$ (``best-in-basis'').
+The fitted instrument response solutions from the Surveys A to D datasets are shown in Figures \ref{fig:A_result} to \ref{fig:D_result}.
+In these plots the final fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ is compared to the true instrument response $f_\true(\vec{x} | \vec{q_\true})$ and to the best possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$ (``Best-in-Basis'').
 The accuracy of the final fitted solutions are summarized, with the metrics introduced in Section \ref{sec:metrics}, for each of the Survey strategies in Table \ref{tab:simple_results}.
@@
 The self-calibration procedure failed completely with the Survey A dataset and no solution was obtained.
-The number of iterations of the self-calibration procedure required to converge to the final solution also decreases as the quality of the final fit increases: Survey D required 9 iterations, Survey C 26, Survey B 4160 and the procedure was iterated XX times with Survey A without reaching a solution.
+The number of iterations of the self-calibration procedure required to converge to the final solution also decreases as the quality of the final fit increases: Survey D required 9 iterations, Survey C 26, Survey B 4160 and the procedure was iterated 8192 times with Survey A without reaching a solution.
 
@@
 As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top) the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level.
-That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the basis used.
+That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the model basis used.
 The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the Best-in-Basis Badness $B_\basis$ compared to the True Badness $B_\true$.
-The results from these two survey datasets are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the instrument response model fitted for.
+The results from these two survey datasets are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the model used to represent the instrument response.
 
@@
 With Survey A, the self-calibration procedure did not tend to a solution.
-This, in itself, is an important result, as the naive tiling of the sky currently being considered for many large surveys will result in datasets that cannot be retrospectively self-calibrated.
+This -- in itself -- is an important result, as the naive tiling of the sky currently being considered for many large surveys will result in datasets that cannot be retrospectively self-calibrated.
 Minor changes to survey strategies, such as those in Survey C, can drastically improve the performance of the self-calibration procedure.
 Without these changes, vast amounts of calibration information is simply lost.
-We can also draw conclusions from the results obtained with Survey C: here the self-calibration procedure converges close to the correct solutions, albeit not as close as the Surveys C and B.
+We can also draw conclusions from the results obtained with Survey B: here the self-calibration procedure converges close to the correct solutions, albeit not as close as the Surveys C and D.
 Here, it is the quality of the fit over the focal plane that is interesting.
 From the residual plot against the best-in-basis instrument response $f_\basis(\vec{x_j} | \vec{q_\basis})$ (Fig \ref{fig:B_result} bottom right), we can see that the quality of the fit is rotationally symmetric, with the self-calibration procedure able to fit the instrument response well at the edge of the focal plane, but poorly at the center.
-This, along with the success of the procedure with the Surveys C and D, gives an insight into the key property of a survey that makes it good for self-calibration.
+This, along with the success of the procedure with the Surveys C and D datasets, gives an insight into the key property of a survey that makes it good for self-calibration.
 Namely, the survey should return the same sources to many, very different focal plane positions.
 This allows the self-calibration procedure to link the different parts of the focal plane with each other through the observations of the same sources.
-Surveys C and D do this and the self-calibration procedure performs well with their dataset.
+Surveys C and D do this, and the self-calibration procedure performs well with their dataset.
 Survey A does not do this; it only connects the outer edges of the focal plane to each other.
@@
 \caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey A dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
-The self-calibration procedure did not converged to the final solution, the results presented here correspond to the instrument response found after $8\,192$ iterations.
+The self-calibration procedure did not converged to the final solution and the results presented here correspond to the instrument response found after $8192$ iterations.
 (a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response.
@@
 \startdata
-A*  & 8192 & XX & XX & XX \\
-B  & 4160 & XX & XX & XX \\
+A*  & 8192 & 2.964 & 3.101 & 3.090 \\
+B  & 4160 & 0.299 & 0.379 & 0.232 \\
 C  & 26 & 0.199 & 0.305 & 0.043 \\

### 2011-11-29 paper/self_calibration_I.tex
euclid: paper after David comments
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19957 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 
-\shorttitle{Optimizing Large Scale Imaging Surveys for a Retrospective Relative Photometric Calibration}
+\shorttitle{Designing Large Scale Imaging Surveys for Retrospective Relative Photometric Calibration}
 \shortauthors{Holmes et al.}
@@
 
-\author{R. Holmes\altaffilmark{1} and H.-W. Rix\altaffilmark{1}}
+\author{Rory Holmes, David W. Hogg\altaffilmark{1} and Hans-Walter Rix}
 \affil{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, Heidelberg, 69117, Germany.}
-
-\and
-
-\author{D. W. Hogg\altaffilmark{2}}
-\affil{2 Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA.}
+\altaffiltext{1}{2 Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA.}
 
 \begin{abstract}
-In this paper we show that, with due care given to the observing strategy, it will be possible to accurately constrain the relative photometric calibration of instruments used in large imaging surveys with their science data alone.
+In this paper we show that, with a well chosen observing strategy, it is possible to accurately calibrate instruments used in large-scale imaging surveys with survey science data alone.
 We create end-to-end simulations of an imaging survey, which produces simulated datasets from mock observations of a synthetic sky according to a defined survey strategy.
-These catalog based simulations include realistic measurement uncertainties and a complex, position dependent instrument response.
+These catalog based simulations include realistic measurement uncertainties and a complex, position-dependent instrument response.
 We then use a self-calibration technique to recover the relative instrument response by fitting an model that best explains the survey dataset, based on the multiple observations of (non-varying) sources at different focal plane positions.
-By considering four simple survey strategies we find that, with a correct redundancy built into the survey strategy, it is possible to accurately constrain the relative photometric response of an imaging instrument, and therefore the relative calibration of the resulting dataset.
-The majority of the remaining post self-calibration errors are due to the limitations in the basis used to model the relative instrument response.
-We find that returning the same sources to very different focal plane positions is the key property of a survey strategy that is required for an accurate calibration.
-From the results of this work, we are able to highlight an important point for those considering the design of large scale imaging surveys: depart from a regular tiling of the sky and return the same sources to very different focal plane positions.
+By considering four simple survey strategies we find that, with appropriate redundancy built into the survey strategy, it is possible to accurately constrain the relative photometric response of an imaging instrument, and therefore the relative calibration of all the images and the relative fluxes of all the sources observed.
+The majority of the remaining post-self-calibration errors are due to the limitations in the basis used to model the relative instrument response.
+We find that returning the same sources to very different focal plane positions is the key property of any survey strategy designed for accurate calibration.
+The results of this work suggest the following advice for those considering the design of large scale imaging surveys: Do not use a regular tiling of the sky; instead return the same sources to very different focal plane positions.
 \end{abstract}
@@
 
-%% From the front matter, we move on to the body of the paper.
-%% In the first two sections, notice the use of the natbib \citep
-%% and \citet commands to identify citations.  The citations are
-%% tied to the reference list via symbolic KEYs. The KEY corresponds
-%% to the KEY in the \bibitem in the reference list below. We have
-%% chosen the first three characters of the first author's name plus
-%% the last two numeral of the year of publication as our KEY for
-%% each reference.
-
-%% Authors who wish to have the most important objects in their paper
-%% linked in the electronic edition to a data center may do so by tagging
-%% their objects with \objectname{} or \object{}.  Each macro takes the
-%% object name as its required argument. The optional, square-bracket 
-%% argument should be used in cases where the data center identification
-%% differs from what is to be printed in the paper.  The text appearing 
-%% in curly braces is what will appear in print in the published paper. 
-%% If the object name is recognized by the data centers, it will be linked
-%% in the electronic edition to the object data available at the data centers  
-%%
-%% Note that for sources with brackets in their names, e.g. [WEG2004] 14h-090,
-%% the brackets must be escaped with backslashes when used in the first
-%% square-bracket argument, for instance, \object[\[WEG2004\] 14h-090]{90}).
-%%  Otherwise, LaTeX will issue an error.
-
 \section{Introduction}
 Astronomers tend to think in terms of taking science data and calibration data separately.
-The former is used for science and the latter is used to constrain instrument parameters, such as the instrument response, the dark currents and so-on.
-But typically far more photons are collected during science exposures themselves; are these not incredibly constraining on the calibration?
+The former is used for scientific measurements of astronomical sources and the latter is used to constrain instrument parameters, such as the instrument response, the dark currents and so-on.
+But typically far more photons, or readout electrons, are collected during science exposures themselves; are these not incredibly constraining on the calibration?
 Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS), much more calibration information was obtained from the science data than the calibration data \citep{pad08}.
-But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: good redundancy is required in the data stream, and a redundancy of a very specific kind.
+But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: Good redundancy is required in the data stream, and a redundancy of a very specific kind.
 
-In this paper, we argue that the next generation of large-scale imaging surveys should have their observation strategies optimized from the very start with this kind of ``self-calibration'' in mind.
-This work focuses on the relative photometric calibration of a typical imaging instrument only, although we also suspect that similar techniques could be used to constrain many other calibration parameters.
-The self-calibration technique utilizes the multiple measurements of (non-varying) sources, at different focal plane positions and at different times within the survey, to constrain the relative instrument response by requiring that the post-calibration measurements yield the same flux values.
-Through end-to-end, catalog level survey simulations\footnote{All of the code used in this work is publicly available at XX.}, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.
+In this paper, we argue that the next generation of large-scale imaging surveys should have their observation strategies designed from the very start with this kind of ``self-calibration'' in mind.
+This work focuses on the relative photometric calibration of a typical imaging instrument only, although similar techniques could be used to constrain many other calibration parameters, such as the optical distortion, point-spread-functions and so-on.
+The self-calibration technique utilizes the multiple measurements of (non-varying) sources, at different focal plane positions and at different times within the survey, to constrain the relative instrument response by requiring that the post-calibration measurements yield consistent flux values.
+Through end-to-end, catalog-level survey simulations, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.
 
 We construct realistic survey datasets through mock observations of a synthetic sky according to a specified survey strategy.
-These simulations include a complex, position dependent instrument response for the imaging instrument.
-Through the self-calibration procedure, we then recover this instrument response by fitting a model that best describes the survey dataset.
+These simulations include a complex, position-dependent instrument response for the imaging instrument.
+Through the self-calibration procedure, we recover this instrument response by fitting a model that best describes the survey dataset.
 By comparing the \textit{fitted} instrument response to the \textit{true} instrument response we are able to assess the performance of the self-calibration procedure with different survey strategies.
@@
 
+In this work we do not produced pixelated images; instead we concentrate on catalog-level simulations with realistic measurement uncertainties.
+Complex effects are included in the simulations that are not  precisely modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the dataset.
+
 In Section \ref{sec:survey_simulations} we introduce the simulation chain constructed to produce the realistic survey datasets.
-Section \ref{sec:self_cal} goes onto to detail the self-calibration procedure.
+Section \ref{sec:self_cal} goes onto to detail the self-calibration procedure, with Section \ref{sec:metrics} summarizing the metrics we use to assess its performance.
 In Section \ref{sec:simple_surveys} we focus on four simple survey strategies, which allow us to draw conclusions on the performance of the self-calibration procedure with different survey properties.
-We do not produced pixelated images; instead we concentrate on catalog level simulations with realistic measurement uncertainties.
-Complex effects are included in the simulations that are not  precisely modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the dataset.
 
@@
 The simulations are split into a number of steps.
-The first of which is the generation of a synthetic sky.
+The first is the generation of a synthetic sky.
 With a given pointing, single mock observations are performed on this sky.
-To build up a survey wide dataset, multiple single observations are performed according to the specified survey strategy.
+To build up a full-survey dataset, multiple single observations are performed according to the specified survey strategy.
 In this section, we detail the assumptions and methods used in each of these steps.
@@
 where $\dfrac{\dd N}{\dd m \, \dd  \Omega}$  is the density of sources $N$ per unit magnitude $m$ and per unit solid angle $\Omega$, and $a$, $b$ and $c$ are model parameters.
-Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in \citep{win11} only.
+Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in \citet{win11} only.
 These parameters were found to be $a = -13.05$, $b = 1.25$ and $c = -0.02$.
-We intentionally omit the stellar population, so that the conclusions drawn from these simulations are independent of the final position of the survey area on the sky.
 The source magnitudes $m$ are related to the source fluxes $s$ simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies (nmgy).
-Due to computational reasons, we only select the brightest sources within the survey area, up to a source density $d$, for the self-calibration procedure.
+To limit the data, we only select the brightest sources within the survey area, up to a source density per unit solid angle $d$, for the self-calibration procedure.
 
@@
 \subsubsection{Measured Count Rates}
-We do not consider pixelated images in these simulations; instead the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position dependent instrument response model $f_\true$ and a measurement noise model.
+We do not consider pixelated images in these simulations; instead the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model.
 For a measurement $i$ the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s_{k,\true}$
@@
 \end{eqnarray}
-where $\alpha$ and $\eta$ are both constants and $\epsilon_i$ is a random number, in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$.
+where $\alpha$ and $\eta$ are both constants and $\epsilon_i$ is a random number, drawn uniformly in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$.
 The $m = 22$ mag 10$\upsigma$ detection limit introduced previously and the 500 limit on the signal-to-noise ratio are used to set $\alpha = 0.1585$ and $\eta = 0.0017$.
-The $\epsilon_i$ contribution is not taken into account in the analysis and therefore the uncertainty variances on the count rates are assumed to be
+The $\epsilon_i$ contribution is not taken into account in the analysis and therefore the uncertainty variances on the count rates are \textit{assumed} (incorrectly, but realistically) to be
 \begin{eqnarray*}
@@
 We therefore apply the single exposure procedure introduced in the previous section for each pointing specified in a defined survey strategy.
-The resultant source measurements are collaged into a survey wide dataset.
+The resultant source measurements are collated into a survey-wide dataset.
 
-\section{Self-Calibrating the Survey Wide Dataset}
+\section{Self-Calibrating the Survey-Wide Dataset}
 \label{sec:self_cal}
-We self-calibrate the dataset generated in the survey simulations to recover the true instrument response applied and the true source fluxes.
-This self-calibration procedure has been successfully applied to ground based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}\footnote{Here this procedure is dubbed ``uber-calibration.''}.
+We self-calibrate the dataset generated in the survey simulations to recover a best possible estimate of the true instrument response applied and the true source fluxes.
+This self-calibration procedure has been successfully applied to ground-based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}\footnote{Here this procedure is dubbed ``uber-calibration.''}.
 This iterative procedure comprises two steps: (1) a refinement of the source flux estimates based on the latest instrument response model and (2) a refinement of the instrument response model based on the updated source flux estimates.
@@
 It is not possible to know, for example, if the sources are all fainter or if the instrument response is uniformly lower.
-In practice, this degeneracy can be broken through observations of standard absolute sources.
+In practice, this degeneracy can be broken through observations of a small number of standard absolute sources.
 
@@
 The $\epsilon_i$ error contribution is intentionally not included in order to simulate systematic problems with the instrument noise model.
+
+We make one further assumption during this analysis phase: We assume that clearly variable sources have been removed from the source catalog, and that the remaining sources vary by less than $\epsilon_{max}$.
 
@@
 \end{eqnarray*}
-Recall that the fitted instrument response $f_{i}(\vec{x_i} | \vec{q})$ is modeled as an eight order polynomial. This can be expressed as
+Recall that the fitted instrument response $f_{i}(\vec{x_i} | \vec{q})$ is modeled as an eighth order polynomial. This can be expressed as
 \begin{eqnarray*}
@@
 The other two metrics, called ``badnesses'', are defined as the RMS error between the final fitted instrument response and a reference instrument response sampled on a regular $300 \times 300$ grid across the focal plane.
-For the ``True Badness'' $B_\true$, the \textit{fitted} instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ is compared to the \textit{true} instrument response $f_\true(\vec{x} | \vec{q_\true})$ at the $J$ sample points
+For the ``true badness'' $B_\true$, the \textit{fitted} instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ is compared to the \textit{true} instrument response $f_\true(\vec{x} | \vec{q_\true})$ at the $J$ sample points
 \begin{equation}
@@
 \end{equation}
-The ``Best-in-Basis Badness'' $B_\basis$ compares the \textit{fitted} instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ to the \textit{best instrument response fit possible} $f_\basis(\vec{x_j} | \vec{q_\basis})$ with the basis used to describe the fitted model (in this case an eight order polynomial) at the $J$ sample points
+The ``best-in-basis badness'' $B_\basis$ compares the \textit{fitted} instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ to the \textit{best instrument response fit possible} $f_\basis(\vec{x_j} | \vec{q_\basis})$ with the basis used to describe the fitted model (in this case an eight order polynomial) at the $J$ sample points
 \begin{equation}
@@
 \end{equation}
+The best-in-basis badness is always smaller than the true badness, as it does not include the errors associated with the limitation of the basis used for the fitted model.
 The badnesses provide a more complete description of the self-calibration performance than the RMS error on the fitted sources' fluxes, as the RMS source error only applies to the bright sources within the survey selected for the self-calibration procedure.
@@
 The iterative self-calibration procedure converges to a final fitted solution for the Survey B, C and D datasets.
-With the Survey A dataset it does not converge, even when the system is started close to the optimum fit.
+With the Survey A dataset it does not converge in any reasonable time period, even when the system is started close to the optimum fit.
 The fitted instrument response solutions from the Surveys A to D datasets are shown in Figures \ref{fig:A_result} to \ref{fig:D_result}.
-In these plots the final fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ is compared to the true instrument response $f_\true(\vec{x} | \vec{q_\true})$ and to the best possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$ (``Best-in-Basis'').
+In these plots the final fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ is compared to the true instrument response $f_\true(\vec{x} | \vec{q_\true})$ and to the best possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$ (``best-in-basis'').
 The accuracy of the final fitted solutions are summarized, with the metrics introduced in Section \ref{sec:metrics}, for each of the Survey strategies in Table \ref{tab:simple_results}.
@@
 That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the model basis used.
-The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the Best-in-Basis Badness $B_\basis$ compared to the True Badness $B_\true$.
+The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the best-in-basis badness $B_\basis$ compared to the true badness $B_\true$.
 The results from these two survey datasets are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the model used to represent the instrument response.
@@
 The Survey A and B datasets also give interesting results.
-With Survey A, the self-calibration procedure did not tend to a solution.
+With Survey A, the self-calibration procedure did not reach a solution.
 This -- in itself -- is an important result, as the naive tiling of the sky currently being considered for many large surveys will result in datasets that cannot be retrospectively self-calibrated.
@@
 From the residual plot against the best-in-basis instrument response $f_\basis(\vec{x_j} | \vec{q_\basis})$ (Fig \ref{fig:B_result} bottom right), we can see that the quality of the fit is rotationally symmetric, with the self-calibration procedure able to fit the instrument response well at the edge of the focal plane, but poorly at the center.
+
 This, along with the success of the procedure with the Surveys C and D datasets, gives an insight into the key property of a survey that makes it good for self-calibration.
@@
 
-In this paper, we have considered the optimization of survey strategies for calibrating the relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained from this method.
-For example, the optical distortion of an instrument could also be constrained with such a method, although we concede that the properties of a survey strategy that makes them good for one calibration, may not be the same as that for another; ultimately a trade-off would need to be performed.
+In this paper, we have considered the design of survey strategies for calibrating the relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained from this method.
+For example, the optical distortion of an instrument or the point-spread-function variation could also be constrained with such a method, although we concede that the properties of a survey strategy that makes them good for one calibration, may not be the same as that for another; ultimately a trade-off would need to be performed between the different calibration concerns and the other goals of the survey, such as sky coverage, variability monitoring and so-on.
 
-%% If you wish to include an acknowledgments section in your paper,
-%% separate it off from the body of the text using the \acknowledgments
-%% command.
-
-%% Included in this acknowledgments section are examples of the
-%% AASTeX hypertext markup commands. Use \url without the optional [HREF]
-%% argument when you want to print the url directly in the text. Otherwise,
-%% use either \url or \anchor, with the HREF as the first argument and the
-%% text to be printed in the second.
+All of the code used in this work is publicly available at XX.
 
@@
 We are grateful to
-
-%% To help institutions obtain information on the effectiveness of their
-%% telescopes, the AAS Journals has created a group of keywords for telescope
-%% facilities. A common set of keywords will make these types of searches
-%% significantly easier and more accurate. In addition, they will also be
-%% useful in linking papers together which utilize the same telescopes
-%% within the framework of the National Virtual Observatory.
-%% See the AASTeX Web site at http://aastex.aas.org/
-%% for information on obtaining the facility keywords.
-
-%% After the acknowledgments section, use the following syntax and the
-%% \facility{} macro to list the keywords of facilities used in the research
-%% for the paper.  Each keyword will be checked against the master list during
-%% copy editing.  Individual instruments or configurations can be provided 
-%% in parentheses, after the keyword, but they will not be verified.
-
-%% Appendix material should be preceded with a single \appendix command.
-%% There should be a \section command for each appendix. Mark appendix
-%% subsections with the same markup you use in the main body of the paper.
-
-%% Each Appendix (indicated with \section) will be lettered A, B, C, etc.
-%% The equation counter will reset when it encounters the \appendix
-%% command and will number appendix equations (A1), (A2), etc.
 
@@
 \begin{thebibliography}{}
-\bibitem[Padmanabhan et al.
-(2008)]{pad08} {Padmanabhan, N., Schlegel, D.~J., Finkbeiner, D.~P., Barentine, J.~C., Blanton, M.~R., Brewington, H.~J., Gunn, J.~E., Harvanek, M., Hogg, D.~W., Ivezi{\'c}, {\v Z}., Johnston, D., Kent, S.~M., Kleinman, S.~J., Knapp, G.~R., Krzesinski, J., Long, D., Neilsen, Jr., E.~H., Nitta, A., Loomis, C., Lupton, R.~H., Roweis, S., Snedden, S.~A., Strauss, M.~A., \& Tucker, D.~L., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj, 674, 1217}
-\bibitem[Windhorst et al.
-(2011)]{win11}{Windhorst, R.~A., Cohen, S.~H., Hathi, N.~P., McCarthy, P.~J., Ryan, Jr., R.~E., Yan, H., Baldry, I.~K., Driver, S.~P., Frogel, J.~A., Hill, D.~T., Kelvin, L.~S., Koekemoer, A.~M., Mechtley, M., O'Connell, R.~W., Robotham, A.~S.~G., Rutkowski, M.~J., Seibert, M., Straughn, A.~N.,Tuffs, R.~J., Balick, B., Bond, H.~E., Bushouse, H., Calzetti, D., Crockett, M., Disney, M.~J., Dopita, M.~A., Hall, D.~N.~B., Holtzman, J.~A., Kaviraj, S., Kimble, R.~A., MacKenty, J.~W., Mutchler, M., Paresce, F., Saha, A., Silk, J.~I., Trauger, J.~T., Walker, A.~R., Whitmore, B.~C., Young, E.~T., The Hubble Space Telescope Wide Field Camera 3 Early Release Science Data: Panchromatic Faint Object Counts for 0.2-2 $\upmu$m Wavelength, 2011, XX.}
+\bibitem[Padmanabhan et al. (2008)]{pad08} {Padmanabhan, N. et al., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj, 674, 1217}
+303	\bibitem[Windhorst et al.
+(2011)]{win11}{Windhorst, R.~A.
+et al., The Hubble Space Telescope Wide Field Camera 3 Early Release Science Data: Panchromatic Faint Object Counts for 0.2-2 $\upmu$m Wavelength, 2011, XX.}
 \end{thebibliography}

### 2011-11-29 paper/self_calibration_I.tex
euclid: cleaning up paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19958 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \section{Discussion}
-From the simulations presented in this paper, we are draw two firm conclusions about the self-calibration of large-scale imaging surveys.
+From the simulations presented in this paper\footnote{All of the simulations presented in this paper are publicly available at \url{http://trac.astrometry.net/browser/trunk/projects/euclid}.}, we are draw two firm conclusions about the self-calibration of large-scale imaging surveys.
 The first is that, in agreement with previous work, we show that this technique is very powerful, allowing the relative instrument response of an imaging instrument to be accurately retrospectively constrained from the science images alone.
@@
 
-All of the code used in this work is publicly available at XX.
-
 \acknowledgments
@@
 
-\appendix
-
-%% The reference list follows the main body and any appendices.
-%% Use LaTeX's thebibliography environment to mark up your reference list.
-%% Note \begin{thebibliography} is followed by an empty set of
-%% curly braces.  If you forget this, LaTeX will generate the error
-%% "Perhaps a missing \item?".
-%%
-%% thebibliography produces citations in the text using \bibitem-\cite
-%% cross-referencing. Each reference is preceded by a
-%% \bibitem command that defines in curly braces the KEY that corresponds
-%% to the KEY in the \cite commands (see the first section above).
-%% Make sure that you provide a unique KEY for every \bibitem or else the
-%% paper will not LaTeX. The square brackets should contain
-%% the citation text that LaTeX will insert in
-%% place of the \cite commands.
-
-%% We have used macros to produce journal name abbreviations.
-%% AASTeX provides a number of these for the more frequently-cited journals.
-%% See the Author Guide for a list of them.
-
-%% Note that the style of the \bibitem labels (in []) is slightly
-%% different from previous examples.  The natbib system solves a host
-%% of citation expression problems, but it is necessary to clearly
-%% delimit the year from the author name used in the citation.
-%% See the natbib documentation for more details and options.
-
 \begin{thebibliography}{}
-\bibitem[Padmanabhan et al. (2008)]{pad08} {Padmanabhan, N. et al., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj, 674, 1217}
-303	\bibitem[Windhorst et al.
+\bibitem[Padmanabhan et al. (2008)]{pad08} {Padmanabhan, N. et al., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj~674, p1217.}
+\bibitem[Windhorst et al.
 (2011)]{win11}{Windhorst, R.~A.
-et al., The Hubble Space Telescope Wide Field Camera 3 Early Release Science Data: Panchromatic Faint Object Counts for 0.2-2 $\upmu$m Wavelength, 2011, XX.}
+et al., The Hubble Space Telescope Wide Field Camera 3 Early Release Science Data: Panchromatic Faint Object Counts for 0.2-2 $\upmu$m Wavelength, 2011, \apjs~193, p27.}
 \end{thebibliography}
@@
 
-%% Use the figure environment and \plotone or \plottwo to include
-%% figures and captions in your electronic submission.
-%% To embed the sample graphics in
-%% the file, uncomment the \plotone, \plottwo, and
-%% \includegraphics commands
-%%
-%% If you need a layout that cannot be achieved with \plotone or
-%% \plottwo, you can invoke the graphicx package directly with the
-%% \includegraphics command or use \plotfiddle. For more information,
-%% please see the tutorial on "Using Electronic Art with AASTeX" in the
-%% documentation section at the AASTeX Web site, http://aastex.aas.org/
-%%
-%% The examples below also include sample markup for submission of
-%% supplemental electronic materials. As always, be sure to check
-%% the instructions to authors for the journal you are submitting to
-%% for specific submissions guidelines as they vary from
-%% journal to journal.
-
-%% This example uses \plotone to include an EPS file scaled to
-%% 80% of its natural size with \epsscale. Its caption
-%% has been written to indicate that additional figure parts will be
-%% available in the electronic journal.
-
-%% Here we use \plottwo to present two versions of the same figure,
-%% one in black and white for print the other in RGB color
-%% for online presentation. Note that the caption indicates
-%% that a color version of the figure will be available online.
-%%
-
-%% This figure uses \includegraphics to scale and rotate the still frame
-%% for an mpeg animation.
-
-%% If you are not including electonic art with your submission, you may
-%% mark up your captions using the \figcaption command. See the
-%% User Guide for details.
-%%
-%% No more than seven \figcaption commands are allowed per page,
-%% so if you have more than seven captions, insert a \clearpage
-%% after every seventh one.
-
-%% Tables should be submitted one per page, so put a \clearpage before
-%% each one.
-
-%% Two options are available to the author for producing tables:  the
-%% deluxetable environment provided by the AASTeX package or the LaTeX
-%% table environment.  Use of deluxetable is preferred.
-%%
-
-%% Three table samples follow, two marked up in the deluxetable environment,
-%% one marked up as a LaTeX table.
-
-%% In this first example, note that the \tabletypesize{}
-%% command has been used to reduce the font size of the table.
-%% We also use the \rotate command to rotate the table to
-%% landscape orientation since it is very wide even at the
-%% reduced font size.
-%%
-%% Note also that the \label command needs to be placed
-%% inside the \tablecaption.
-
-%% This table also includes a table comment indicating that the full
-%% version will be available in machine-readable format in the electronic
-%% edition.
-
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=\textwidth]{./single_image.pdf}
-\end{center}
-\caption{A single exposure of the synthetic sky.
+\begin{figure}
+\plotone{./single_image.pdf}
+\figcaption{A single exposure of the synthetic sky.
 Left: A plot of the bright sources within the synthetic sky used in the self-calibration procedure, with the focal plane footprint overlaid.
@@
 \begin{deluxetable}{cccc}
+\rotate
+\tablecaption{A summary of the four simple survey strategies considered in this paper. \label{tab:simple_surveys}}
 \tablewidth{0pt}
-\tablecaption{A summary of the four simple survey strategies considered in this paper. \label{tab:simple_surveys}}
 \tablehead{
@@
 
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=\textwidth]{simple_surveys.pdf}
-\end{center}
-\caption{Focal plane footprints projected onto the synthetic sky according to the four simple survey strategies described in Section \ref{sec:simple_surveys} and summarized in Table \ref{tab:simple_surveys}.
+\begin{figure}
+\plotone{simple_surveys.pdf}
+\figcaption{Focal plane footprints projected onto the synthetic sky according to the four simple survey strategies described in Section \ref{sec:simple_surveys} and summarized in Table \ref{tab:simple_surveys}.
 Surveys A, B and D have 1728 pointings and survey C has 1720 pointings.\label{fig:simple_surveys}}
@@
 
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=\textwidth]{A_8192_ff.pdf}
-\end{center}
-\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey A dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+\begin{figure}
+\plotone{A_8192_ff.pdf}
+\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey A dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
 The self-calibration procedure did not converged to the final solution and the results presented here correspond to the instrument response found after $8192$ iterations.
@@
 
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=\textwidth]{B_4160_ff.pdf}
-\end{center}
-\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey B dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+\begin{figure}
+\plotone{B_4160_ff.pdf}
+\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey B dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
 The self-calibration procedure converged to the final solution after 4160 iterations.
@@
 
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=\textwidth]{C_026_ff.pdf}
-\end{center}
-\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey C dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+\begin{figure}
+\plotone{C_026_ff.pdf}
+\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey C dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
 The self-calibration procedure converged to the final solution after 26 iterations.
@@
 
-\begin{figure}[ht]
-\begin{center}
-\includegraphics[width=\textwidth]{D_009_ff.pdf}
-\end{center}
-\caption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey D dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+\begin{figure}
+\plotone{D_009_ff.pdf}
+\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey D dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
 The self-calibration procedure converged to the final solution after 9 iterations.
@@
 \begin{deluxetable}{ccccc}
+\rotate
 \tablecaption{A summary of the quality of the final fit from the self-calibration procedure with the four simple survey strategies summarized in Table \ref{tab:simple_surveys}.
@@
 *Did not converge.\label{tab:simple_results}}
+\tablewidth{0pt}
 \tablehead{
@@
 
-%% If you use the table environment, please indicate horizontal rules using
-%% \tableline, not \hline.
-%% Do not put multiple tabular environments within a single table.
-%% The optional \label should appear inside the \caption command.
-
-%% If the table is more than one page long, the width of the table can vary
-%% from page to page when the default \tablewidth is used, as below.  The
-%% individual table widths for each page will be written to the log file; a
-%% maximum tablewidth for the table can be computed from these values.
-%% The \tablewidth argument can then be reset and the file reprocessed, so
-%% that the table is of uniform width throughout. Try getting the widths
-%% from the log file and changing the \tablewidth parameter to see how
-%% adjusting this value affects table formatting.
-
-%% The \dataset{} macro has also been applied to a few of the objects to
-%% show how many observations can be tagged in a table.
-
-%% Tables may also be prepared as separate files. See the accompanying
-%% sample file table.tex for an example of an external table file.
-%% To include an external file in your main document, use the \input
-%% command. Uncomment the line below to include table.tex in this
-%% sample file. (Note that you will need to comment out the \documentclass,
-%% \begin{document}, and \end{document} commands from table.tex if you want
-%% to include it in this document.)
-
-%% \input{table}
-
-%% The following command ends your manuscript. LaTeX will ignore any text
-%% that appears after it.
-
 \end{document}
-
-%%
-%% End of file `sample.tex'.

### 2011-11-29 paper/self_calibration_I.tex
euclid: add Wittman reference, kind of...
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19959 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 We self-calibrate the dataset generated in the survey simulations to recover a best possible estimate of the true instrument response applied and the true source fluxes.
-This self-calibration procedure has been successfully applied to ground-based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}\footnote{Here this procedure is dubbed ``uber-calibration.''}.
+This self-calibration procedure has been successfully applied to ground-based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}\footnote{Here this procedure is dubbed ``uber-calibration.''} and the Deep Lens Survey \citep{wit11}.
 This iterative procedure comprises two steps: (1) a refinement of the source flux estimates based on the latest instrument response model and (2) a refinement of the instrument response model based on the updated source flux estimates.
@@
 et al., The Hubble Space Telescope Wide Field Camera 3 Early Release Science Data: Panchromatic Faint Object Counts for 0.2-2 $\upmu$m Wavelength, 2011, \apjs~193, p27.}
+\bibitem[Wittman et al. (2011)]{wit11}{XX}
 \end{thebibliography}

### 2011-12-02 paper/self_calibration_I.tex
euclid: add temporal paragraph
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19986 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 
+As this work was completed in the context of space-based imaging surveys, we have not considered the temporal variation of the relative instrument response as a high priority.
+We concede that this will have a significant impact on ground-based surveys, where variations on nightly timescales can be expected.
+The self-calibration procedure discussed in this paper can be trivially modified to fit for a time evolving relative instrument response model.
+For this to be effective though it will be necessary to not only observe the same sources at very different focal plane positions, but also to observe them at very different times within the survey.
+To quantify this in more detail, we will include temporal variations in future simulations.
+
 In this paper, we have considered the design of survey strategies for calibrating the relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained from this method.
@@
 \section{Acknowledgments}
-We are grateful to
+DWH was partially supported by NASA (grant XXX), the NSF (grant XXX), and a research fellowship from the Alexander von Humboldt Foundation.
 
@@
 et al., The Hubble Space Telescope Wide Field Camera 3 Early Release Science Data: Panchromatic Faint Object Counts for 0.2-2 $\upmu$m Wavelength, 2011, \apjs~193, p27.}
-\bibitem[Wittman et al. (2011)]{wit11}{XX}
+\bibitem[Wittman et al. (2011)]{wit11}{Wittman, D., Ryan, R. and Thorman, P., Ubercalibration of the Deep Lens Survey, 2011, ArXiv e-prints, 1111.2058}
 \end{thebibliography}

### 2012-01-04 paper/self_calibration_I.tex
euclid: start adding HW comments
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@20088 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \begin{abstract}
-In this paper we show that, with a well chosen observing strategy, it is possible to accurately calibrate instruments used in large-scale imaging surveys with survey science data alone.
-We create end-to-end simulations of an imaging survey, which produces simulated datasets from mock observations of a synthetic sky according to a defined survey strategy.
-These catalog based simulations include realistic measurement uncertainties and a complex, position-dependent instrument response.
+In this paper we show that, with a well chosen observing strategy, it is possible to accurately calibrate the large-scale response of imaging instruments used in large-scale imaging surveys with survey science data alone.
+We create end-to-end simulations of an imaging survey, which produce simulated photometric catalogs from mock observations of a synthetic sky according to a defined survey strategy.
+These catalog-based simulations include realistic measurement uncertainties and a complex, position-dependent instrument response.
 We then use a self-calibration technique to recover the relative instrument response by fitting an model that best explains the survey dataset, based on the multiple observations of (non-varying) sources at different focal plane positions.

### 2012-01-16 paper/self_calibration_I.tex
euclid: fix title error in paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@20089 03a3cea6-2c03-0410-ac9b-9271b1c66c29
euclid: paper post HW comments
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@20142 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 
-\shorttitle{Designing Large Scale Imaging Surveys for Retrospective Relative Photometric Calibration}
+\shorttitle{Designing Large-Scale Imaging Surveys for Retrospective Relative Photometric Calibration}
 \shortauthors{Holmes et al.}
@@
 
-\title{Optimizing Large Scale Imaging Surveys for a Retrospective Relative Photometric Calibration}
+\title{Designing Large-Scale Imaging Surveys for a Retrospective Relative Photometric Calibration}
 
@@
 \begin{abstract}
-In this paper we show that, with a well chosen observing strategy, it is possible to accurately calibrate the large-scale response of imaging instruments used in large-scale imaging surveys with survey science data alone.
-We create end-to-end simulations of an imaging survey, which produce simulated photometric catalogs from mock observations of a synthetic sky according to a defined survey strategy.
+In this paper we show that, with a well chosen observing strategy, it is possible to accurately calibrate the large-scale response of instruments used in large-scale imaging surveys with survey science data alone.
+We create end-to-end simulations of an imaging survey, which produce simulated photometric catalogs from mock observations of a synthetic sky according to a defined survey strategies.
 These catalog-based simulations include realistic measurement uncertainties and a complex, position-dependent instrument response.
-We then use a self-calibration technique to recover the relative instrument response by fitting an model that best explains the survey dataset, based on the multiple observations of (non-varying) sources at different focal plane positions.
-By considering four simple survey strategies we find that, with appropriate redundancy built into the survey strategy, it is possible to accurately constrain the relative photometric response of an imaging instrument, and therefore the relative calibration of all the images and the relative fluxes of all the sources observed.
+We then use a self-calibration technique to recover the relative instrument response by fitting an model that best explains the survey catalog, based on (non-varying) sources that fall on different focal plane positions in multiple observations.
+By considering four simple survey strategies we find that, with appropriate redundancy built into the survey strategy, it is possible to accurately constrain the large-scale relative photometric response of an imaging instrument, and therefore the relative calibration of all the images and the relative fluxes of all the sources observed.
 The majority of the remaining post-self-calibration errors are due to the limitations in the basis used to model the relative instrument response.
-We find that returning the same sources to very different focal plane positions is the key property of any survey strategy designed for accurate calibration.
-The results of this work suggest the following advice for those considering the design of large scale imaging surveys: Do not use a regular tiling of the sky; instead return the same sources to very different focal plane positions.
+We find that returning the same sources to very different focal plane positions is the key property of any survey strategy designed for accurate retrospective calibration.
+The results of this work suggest the following advice for those considering the design of large-scale imaging surveys: Do not use a regular tiling of the sky; instead return the same sources to very different focal plane positions.
 \end{abstract}
@@
 \section{Introduction}
-Astronomers tend to think in terms of taking science data and calibration data separately.
+Astronomers tend to think conventionally in terms of taking science data and calibration data separately.
 The former is used for scientific measurements of astronomical sources and the latter is used to constrain instrument parameters, such as the instrument response, the dark currents and so-on.
@@
 Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS), much more calibration information was obtained from the science data than the calibration data \citep{pad08}.
-But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: Good redundancy is required in the data stream, and a redundancy of a very specific kind.
+But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: Good redundancy was required in the data stream, and a redundancy of a very specific kind.
 
@@
 
-We construct realistic survey datasets through mock observations of a synthetic sky according to a specified survey strategy.
+We construct realistic survey catalogs through mock observations of a synthetic sky according to a specified survey strategy.
 These simulations include a complex, position-dependent instrument response for the imaging instrument.
-Through the self-calibration procedure, we recover this instrument response by fitting a model that best describes the survey dataset.
+Through the self-calibration procedure, we recover this instrument response by fitting a model that best describes the survey catalog.
 By comparing the \textit{fitted} instrument response to the \textit{true} instrument response we are able to assess the performance of the self-calibration procedure with different survey strategies.
-There is a degeneracy in this problem: the self-calibration procedure is only able to fit for a \textit{relative} instrument response.
-We have no way to know, for example, if all the sources are uniformly fainter, or if the instrument sensitivity is uniformly lower.
 
 In this work we do not produced pixelated images; instead we concentrate on catalog-level simulations with realistic measurement uncertainties.
-Complex effects are included in the simulations that are not  precisely modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the dataset.
+Complex effects are included in the simulations that are not  precisely modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the catalog.
 
-In Section \ref{sec:survey_simulations} we introduce the simulation chain constructed to produce the realistic survey datasets.
+In Section \ref{sec:survey_simulations} we introduce the simulation chain constructed to produce the realistic survey catalogs.
 Section \ref{sec:self_cal} goes onto to detail the self-calibration procedure, with Section \ref{sec:metrics} summarizing the metrics we use to assess its performance.
@@
 
-\section{Survey Dataset Simulations}
+\section{Survey catalog Simulations}
 \label{sec:survey_simulations}
-We have constructed an end-to-end simulation chain that produces a realistic imaging dataset from a specified survey strategy.
-In this work we have kept the simulation parameters intentionally flexible, so that the sensitivity of the self-calibration procedure to different values can be investigated.
+We have constructed an end-to-end simulation chain that produces a realistic imaging catalog from a specified survey strategy.
+In this work we have kept the simulation parameters intentionally flexible, so that the sensitivity of the self-calibration procedure to different values can be investigated (the fiducial values are given in Table \ref{tab:parameters}).
 The simulations are split into a number of steps.
@@
 With a given pointing, single mock observations are performed on this sky.
-To build up a full-survey dataset, multiple single observations are performed according to the specified survey strategy.
+To build up a full-survey catalog, multiple single observations are performed according to the specified survey strategy.
 In this section, we detail the assumptions and methods used in each of these steps.
@@
 \subsection{The Synthetic Sky}
-We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and $10\upsigma{}$ limits of deep, space-based, near-infrared survey.
+We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and a $10\upsigma{}$ limits of deep, space-based, near-infrared survey.
 Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes $m$ distributed according to
@@
 \label{sec:single_exposure}
-With a camera pointing $(\alpha, \beta)$ and camera orientation $\theta$ this synthetic sky is transformed into focal plane image.
+With a camera pointing $(\alpha, \beta)$ and camera orientation $\theta$, we find the focal plane positions of the sources on the synthetic sky that fall within the instrument's field-of-view.
 In our simulations, we use a large instrument field-of-view of $0.76~ \text{deg} \times 0.72~\text{deg}$; a size consistent with the next generation of large survey imagers.
@@
 \subsubsection{Measured Count Rates}
-We do not consider pixelated images in these simulations; instead the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model.
-For a measurement $i$ the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s_{k,\true}$
+To generate photometric catalogs, the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model.
+For a measurement $i$ the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s_{k,\true}$ ,
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\vec{q}_\true$ are the parameters defining the \textit{true} instrument response, and $e_i$ is a noise contribution drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$.
+where $\vec{q}_\true$ are the parameters defining the \textit{true} instrument response (see Section \ref{sec:instrument_response_model}), and $e_i$ is a noise contribution drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$.
 
@@
 where $\alpha$ and $\eta$ are both constants and $\epsilon_i$ is a random number, drawn uniformly in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$.
-The $m = 22$ mag 10$\upsigma$ detection limit introduced previously and the 500 limit on the signal-to-noise ratio are used to set $\alpha = 0.1585$ and $\eta = 0.0017$.
+The $m = 22$ mag 10$\upsigma$ detection limit introduced previously and the limit S/N$<500$ are used to set $\alpha = 0.1585$ and $\eta = 0.0017$.
 The $\epsilon_i$ contribution is not taken into account in the analysis and therefore the uncertainty variances on the count rates are \textit{assumed} (incorrectly, but realistically) to be
@@
 \end{eqnarray*}
-where $\vec{x_i} = (x_i, y_i)$ is the focal plane position that the $k$th source falls at during the $i^\text{th}$ measurement and $\vec{q}_\true$ are the parameters defining the instrument response model.
+where $\vec{x_i} = (x_i, y_i)$ is the focal plane position that the $k$th source falls at during the $i\text{th}$ measurement and $\vec{q}_\true$ are the parameters defining the instrument response model.
+This response model subsumes various effects commonly discussed in terms of detector flat-fielding and optical vignetting.
 The large-scale instrument response $f_\text{large}(\vec{x_i} | \vec{q}_{\true, 1 \ldots 6})$ is modeled as a second order polynomial:
@@
 It is this instrument response that we try and recover with the self-calibration procedure.
+With this instrument response model, we have assumed that the variations in sensitivities on a pixel-to-pixel-scale have been calibrated with some other method, such as an internal flat-field calibration source.
+It is variations on larger scales that we aim to constrain with the self-calibration procedure.
 Our final assumption in these simulations is that the instrument response is temporally stable.
@@
 \subsection{The Complete Survey}
-In this work we are interested in simulating complete surveys, in order to identify the crucial characteristics of survey strategies that allow for the relative instrument response model to be accurately, retrospectively constrained from the resulting dataset.
+In this work we are interested in simulating complete surveys, in order to identify the crucial characteristics of survey strategies that allow for the relative instrument response model to be accurately, retrospectively constrained from the resulting catalog.
 We therefore apply the single exposure procedure introduced in the previous section for each pointing specified in a defined survey strategy.
-The resultant source measurements are collated into a survey-wide dataset.
+The resultant source measurements are collated into a survey-wide catalog.
 
-\section{Self-Calibrating the Survey-Wide Dataset}
+\section{Self-Calibrating the Survey-Wide catalog}
 \label{sec:self_cal}
-We self-calibrate the dataset generated in the survey simulations to recover a best possible estimate of the true instrument response applied and the true source fluxes.
+We self-calibrate the catalog generated in the survey simulations to recover an optimum estimate of the true instrument response applied and the true source fluxes.
 This self-calibration procedure has been successfully applied to ground-based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}\footnote{Here this procedure is dubbed ``uber-calibration.''} and the Deep Lens Survey \citep{wit11}.
@@
 
-\subsection{Fitted Measurement Model}
-To complicate the simulations, we use the self-calibration procedure to fit a model that is \textit{incomplete} in two ways.
+\subsection{Fitting the Instrument Response Model}
+To introduce a realistic level of ignorance into the simulations, we use the self-calibration procedure to fit a model that is \textit{incomplete} in two ways.
 Firstly, the \textit{fitted} instrument response is modeled as an eighth order polynomial, and not the second order polynomial superimposed with sine and cosine contributions used to model the \textit{true} instrument response.
 Secondly, the assumed measurement uncertainty variances do not include the additional random measurement error $\epsilon_{i}$ introduced in Subsection \ref{sec:noise}.
-The incomplete measurement model is
+We
+We therefore fit for the following, incomplete measurement model:
 \begin{eqnarray*}
@@
 \end{eqnarray*} 
-where $\alpha$ and $\eta$ are the parameters set by the instruments $10\upsigma$ and saturation limits.
+where $\alpha$ and $\eta$ are the parameters set by the instrument's $10\upsigma$ detection and the S/N $<500$ limits.
 The $\epsilon_i$ error contribution is intentionally not included in order to simulate systematic problems with the instrument noise model.
 
-We make one further assumption during this analysis phase: We assume that clearly variable sources have been removed from the source catalog, and that the remaining sources vary by less than $\epsilon_{max}$.
+We make one further assumption during this analysis phase, namely that clearly variable sources have been removed from the source catalog, and that the remaining sources vary by less than $\epsilon_{max}$.
 
@@
 \begin{eqnarray*}
-\chi'^2_k = \sum_{i \in \mathcal{O}(k)} \frac{(c_i-f_{i}(\vec{x_i} | \vec{q}) \, s'_{k})^2}{{\sigma_i}^2}  \quad .
+\chi'^2_k = \sum_{i \in \mathcal{O}(k)} \frac{(c_i-f_{i}(\vec{x_i} | \vec{q}) \, s'_{k})^2}{{\sigma_i'}^2}  \quad .
 \end{eqnarray*}
@@
 \begin{eqnarray*}
-\chi^2 =\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{(c_i- s'_{k} \sum_{l = 1}^L q_{l} g_l(\vec{x_i}))^2}{{\sigma_i}^2}   \quad .
+\chi^2 =\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{(c_i- s'_{k} \sum_{l = 1}^L q_{l} g_l(\vec{x_i}))^2}{{\sigma_i'}^2}   \quad .
 \end{eqnarray*}
@@
 \begin{eqnarray*}
-\frac{d\chi^2}{dq_{l}} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{-2 g_l(\vec{x_i}) s'_{k} (c_i- s'_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'}(\vec{x_i}))}{{\sigma_i}^2} = 0  \quad ,
+\frac{d\chi^2}{dq_{l}} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{-2 g_l(\vec{x_i}) s'_{k} (c_i- s'_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'}(\vec{x_i}))}{{\sigma_i'}^2} = 0  \quad ,
 \end{eqnarray*}
 \begin{eqnarray*}
-\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{{\sigma_i}^2} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l (\vec{x_i}) s'^2_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'} (\vec{x_i})} {{\sigma_i}^2}   \quad .
+\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{{\sigma_i'}^2} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l (\vec{x_i}) s'^2_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'} (\vec{x_i})} {{\sigma_i'}^2}   \quad .
 \end{eqnarray*}
@@
 \begin{eqnarray}
-b_l & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{\sigma_i^2}    \quad ,\\
-G_{ll'} & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{{s'}_{k}^2}{\sigma_i^2} g_l(x_i) g_{l'}(x_i)   \quad .
+b_l & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{\sigma_i'^2}    \quad ,\\
+G_{ll'} & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{{s'}_{k}^2}{\sigma_i'^2} g_l(x_i) g_{l'}(x_i)   \quad .
 \end{eqnarray}
@@
 \begin{eqnarray*}
-\vec{q'} \leftarrow G^{-1}  \cdot \vec{b}   \quad .
+\vec{q'} = G^{-1}  \cdot \vec{b}   \quad .
 \end{eqnarray*}
@@
 In this paper, we consider four simple but very different survey strategies.
-We use the end-to-end dataset simulation and self-calibration chain, introduced in Sections \ref{sec:survey_simulations} and \ref{sec:self_cal}, to evaluate the performance of the self-calibration procedure with these strategies.
+We use the end-to-end catalog simulation and self-calibration chain, introduced in Sections \ref{sec:survey_simulations} and \ref{sec:self_cal}, to evaluate the performance of the self-calibration procedure with these strategies.
 
@@
 These survey strategies are summarized in Table \ref{tab:simple_surveys} and are shown in Figure \ref{fig:simple_surveys}.
-Strategy A is the simplest strategy; the field is regularly tiled with small overlaps between adjacent pointings ($\sim 5$~\% and $\sim 8$~\% in the camera pointing directions $\alpha$ and $\beta$ respectively).
+Strategy A is the simplest strategy; the target field is regularly tiled with small overlaps between adjacent pointings ($\sim 5$~\% and $\sim 8$~\% in the camera pointing directions $\alpha$ and $\beta$ respectively).
 The pointings in the 12 passes over the same field are exactly aligned.
-Survey B is the same as A, but with each pass over the field the orientation of the telescope is rotated by $30^\circ$.
+Survey B is the same as A, but with each pass over the target field the orientation of the telescope is rotated by $30^\circ$.
 Survey C is more complex.
@@
 \subsection{Self-Calibration Performance}
-The iterative self-calibration procedure converges to a final fitted solution for the Survey B, C and D datasets.
-With the Survey A dataset it does not converge in any reasonable time period, even when the system is started close to the optimum fit.
-The fitted instrument response solutions from the Surveys A to D datasets are shown in Figures \ref{fig:A_result} to \ref{fig:D_result}.
+The iterative self-calibration procedure converges to a final fitted solution for the Survey B, C and D catalogs.
+With the Survey A catalog it does not converge in any reasonable time period, even when the system is started close to the optimum fit.
+The fitted instrument response solutions from the Surveys A to D catalogs are shown in Figures \ref{fig:A_result} to \ref{fig:D_result}.
 In these plots the final fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ is compared to the true instrument response $f_\true(\vec{x} | \vec{q_\true})$ and to the best possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$ (``best-in-basis'').
@@
 
-The self-calibration procedure performs best with the Survey D dataset, followed by the Survey C dataset, then the Survey B dataset.
-The self-calibration procedure failed completely with the Survey A dataset and no solution was obtained.
+The self-calibration procedure performs best with the Survey D catalog, followed by the Survey C catalog, then the Survey B catalog.
+The self-calibration procedure failed completely with the Survey A catalog and no solution was obtained.
 The number of iterations of the self-calibration procedure required to converge to the final solution also decreases as the quality of the final fit increases: Survey D required 9 iterations, Survey C 26, Survey B 4160 and the procedure was iterated 8192 times with Survey A without reaching a solution.
 
-With the Survey C and D datasets, the instrument response is accurately recovered.
-As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top) the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level.
+With the Survey C and D catalogs, the instrument response is accurately recovered.
+As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top), the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level.
 That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the model basis used.
 The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the best-in-basis badness $B_\basis$ compared to the true badness $B_\true$.
-The results from these two survey datasets are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the model used to represent the instrument response.
+The results from these two survey catalogs are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the model used to represent the instrument response.
 
-The Survey A and B datasets also give interesting results.
+The Survey A and B catalogs also give interesting results.
 With Survey A, the self-calibration procedure did not reach a solution.
-This -- in itself -- is an important result, as the naive tiling of the sky currently being considered for many large surveys will result in datasets that cannot be retrospectively self-calibrated.
+This -- in itself -- is an important result, as the naive tiling of the sky currently being considered for many large surveys will result in catalogs that cannot be retrospectively self-calibrated.
 Minor changes to survey strategies, such as those in Survey C, can drastically improve the performance of the self-calibration procedure.
@@
 
-This, along with the success of the procedure with the Surveys C and D datasets, gives an insight into the key property of a survey that makes it good for self-calibration.
+This, along with the success of the procedure with the Surveys C and D catalogs, gives an insight into the key property of a survey that makes it good for self-calibration.
 Namely, the survey should return the same sources to many, very different focal plane positions.
 This allows the self-calibration procedure to link the different parts of the focal plane with each other through the observations of the same sources.
-Surveys C and D do this, and the self-calibration procedure performs well with their dataset.
+Surveys C and D do this, and the self-calibration procedure performs well with their catalogs.
 Survey A does not do this; it only connects the outer edges of the focal plane to each other.
@@
 This conclusion in confirmed with Strategy B.
-Here, positions at the outside of the focal plane are well connected to other positions and therefore the self-calibration procedure can well fit the instrument response in these regions.
+Here, positions at the outside of the focal plane are well connected to other positions, and therefore the self-calibration procedure can well fit the instrument response in these regions.
 In the center of the focal plane, on which the same sources always fall, there are insufficient redundant measurements to allow for a good fit.
@@
 \section{Discussion}
-From the simulations presented in this paper\footnote{All of the simulations presented in this paper are publicly available at \url{http://trac.astrometry.net/browser/trunk/projects/euclid}.}, we are draw two firm conclusions about the self-calibration of large-scale imaging surveys.
-The first is that, in agreement with previous work, we show that this technique is very powerful, allowing the relative instrument response of an imaging instrument to be accurately retrospectively constrained from the science images alone.
+From the simulations presented in this paper\footnote{All of the simulations presented in this paper are publicly available at \url{http://trac.astrometry.net/browser/trunk/projects/euclid}.}, we are able to draw two firm conclusions about the self-calibration of large-scale imaging surveys.
+The first is that, in agreement with previous work, we show that this technique is very powerful, allowing the large-scale, relative instrument response of an imaging instrument to be accurately retrospectively constrained from the science images alone.
 The second is that there must be the correct redundancy built into the survey strategy.
@@
 The self-calibration procedure discussed in this paper can be trivially modified to fit for a time evolving relative instrument response model.
-For this to be effective though it will be necessary to not only observe the same sources at very different focal plane positions, but also to observe them at very different times within the survey.
+For this to be effective though, it will be necessary to not only observe the same sources at very different focal plane positions, but also to observe them at very different times within the survey.
 To quantify this in more detail, we will include temporal variations in future simulations.
 
-In this paper, we have considered the design of survey strategies for calibrating the relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained from this method.
+In this paper, we have considered the design of survey strategies for calibrating the large-scale, relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained from this method.
 For example, the optical distortion of an instrument or the point-spread-function variation could also be constrained with such a method, although we concede that the properties of a survey strategy that makes them good for one calibration, may not be the same as that for another; ultimately a trade-off would need to be performed between the different calibration concerns and the other goals of the survey, such as sky coverage, variability monitoring and so-on.
@@
 \begin{thebibliography}{}
-\bibitem[Padmanabhan et al. (2008)]{pad08} {Padmanabhan, N. et al., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj~674, p1217.}
+\bibitem[Padmanabhan et al. (2008)]{pad08}{Padmanabhan, N. et al., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj~674, p1217.}
 \bibitem[Windhorst et al.
@@
 \plotone{A_8192_ff.pdf}
-\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey A dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey A catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
 The self-calibration procedure did not converged to the final solution and the results presented here correspond to the instrument response found after $8192$ iterations.
-(a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response.
-(c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model.
 The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial).
-The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:A_result}}
+The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
+With this survey strategy, the center of the focal plane is never connected to other parts with repeat observations of the same sources.
+\label{fig:A_result}}
 \end{figure}
@@
 \plotone{B_4160_ff.pdf}
-\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey B dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey B catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
 The self-calibration procedure converged to the final solution after 4160 iterations.
-(a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response.
-(c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model.
-The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial).
-The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:B_result}}
+The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
+With this survey strategy, it is only regions towards the edge of the focal plane that are well interconnected with repeat observations of the same sources, and therefore the self-calibration procedure can only fit a reasonable instrument response model at these positions.\label{fig:B_result}}
 \end{figure}
@@
 \plotone{C_026_ff.pdf}
-\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey C dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey C catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
 The self-calibration procedure converged to the final solution after 26 iterations.
-(a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response.
-(c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model.
-The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial).
 The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:C_result}}
@@
 \plotone{D_009_ff.pdf}
-\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ obtained from self-calibrating the Survey D dataset compared to the true $f_\true(\vec{x} | \vec{q_\true})$ and best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey D catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (b) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
 The self-calibration procedure converged to the final solution after 9 iterations.
-(a) Contour plot of the fitted (black) compared to the \textit{true} (gray) instrument response.
-(c) Contour plot of the fitted (black) compared to the \textit{best-in-basis} (gray) instrument response model.
-The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial).
-The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:D_result}}
+The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
+With this quasi-random observing strategies, all regions of the focal plane well interconnected with each other.\label{fig:D_result}}
 \end{figure}

### 2012-02-16 paper/self_calibration_I.tex
euclid: update paper abstract
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@20264 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \begin{abstract}
-In this paper we show that, with a well chosen observing strategy, it is possible to accurately calibrate the large-scale response of instruments used in large-scale imaging surveys with survey science data alone.
-We create end-to-end simulations of an imaging survey, which produce simulated photometric catalogs from mock observations of a synthetic sky according to a defined survey strategies.
+In this paper, we investigate under which survey conditions the self-calibration of an instrument response can produce precise photometric catalogs for wide-field imaging surveys, drawing only on the science data themselves.
+We consider four basic survey strategies and create end-to-end simulations of an imaging survey, which produce simulated photometric catalogs from mock observations of a realistic, synthetic sky.
 These catalog-based simulations include realistic measurement uncertainties and a complex, position-dependent instrument response.
-We then use a self-calibration technique to recover the relative instrument response by fitting an model that best explains the survey catalog, based on (non-varying) sources that fall on different focal plane positions in multiple observations.
-By considering four simple survey strategies we find that, with appropriate redundancy built into the survey strategy, it is possible to accurately constrain the large-scale relative photometric response of an imaging instrument, and therefore the relative calibration of all the images and the relative fluxes of all the sources observed.
-The majority of the remaining post-self-calibration errors are due to the limitations in the basis used to model the relative instrument response.
-We find that returning the same sources to very different focal plane positions is the key property of any survey strategy designed for accurate retrospective calibration.
+We then use a self-calibration technique to recover the instrument response by fitting an model that best explains the survey catalog, based on the sources that fall on different focal plane positions in separate observations.
+We demonstrate that with a favorable survey strategy, a complex (but smooth) instrument response can be precisely calibrated using this self-calibration technique.
+We also find that returning the same sources to very different focal plane positions is the key property of any survey strategy designed for accurate retrospective calibration.
 The results of this work suggest the following advice for those considering the design of large-scale imaging surveys: Do not use a regular tiling of the sky; instead return the same sources to very different focal plane positions.
@@
 \section{Acknowledgments}
+Rory Holmes was funded by the Deutschen Zentrums f\"ur Luft- und Raumfahrt (DLR) through the grant 50QE1202.
+
 DWH was partially supported by NASA (grant XXX), the NSF (grant XXX), and a research fellowship from the Alexander von Humboldt Foundation.

### 2012-03-16 dev/god.py
euclid: dev update of self-cal code, now just need to get plotting working again
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@20411 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@ -18,25 +18,25 @@ import transformations as tran
 
 
 def flat_field_parameters():
-    np.random.seed(1)  # same seed => same FF
+    #np.random.seed(1)  # same seed => same FF
     return np.append(np.array([1, -0.02, 0.03, -0.2, 0.01, -0.5]),
                                 (1e-4) * np.random.uniform(size=256))
 
 
-def flat_field(params, x, y, par=flat_field_parameters()):
+def flat_field(p, x, y, par=flat_field_parameters()):
     # Large Scale
     ff = (par[0] + par[1] * x + par[2] * y \
                     + par[3] * x ** 2 + par[4] * x * y + par[5] * y ** 2)
     # Small Scale
+    FoV = p['FoV']
     k = 6
-    fov = params['FoV']
     no_loops = int(np.sqrt((len(par) - 6) / 4))
     for nx in range(no_loops):
-        kx = nx * np.pi / fov[0]
+        kx = nx * np.pi / FoV[0]
         ckx = np.cos(kx * x)
         skx = np.sin(kx * x)
         for ny in range(no_loops):
-            ky = ny * np.pi / fov[1]
+            ky = ny * np.pi / FoV[1]
             cky = np.cos(ky * y)
             sky = np.sin(ky * y)
             ff += par[k] * ckx * cky
@@ -105,15 +105,6 @@ def generate_magnitudes(p, number_sources, data_dir, plots=False):
         difference = power_law(A, mag[ii]) / (power_law(B, mag[ii]) * c)
 
     selected_sources = np.sort(mag)[0:int(number_sources)]
-    # Print out catalog for plotting
-    if plots:
-        survey_dic = {}
-        survey_dic['mag'] = selected_sources
-        survey_dic['all_sources'] = mag
-        survey_dic['fit_mag'] = np.arange(14, 26, 0.01)
-        survey_dic['fit_dens'] = power_law(A, np.arange(14, 26, 0.01)) * area
-        pickle_path = data_dir + '/source_catalog.p'
-        pickle.dump(survey_dic, open(pickle_path, "wb"))
     return selected_sources
 
 
@@ -133,22 +124,16 @@ class SourceCatalog:
                                                         size=number_sources)
 
 
-def create_catalog(p, data_dir, plots=False, verbose=False):
-    if verbose:
+def create_catalog(p):
+    if p["verbose"]:
         print("Generating God's Catalog...")
     # Calculate total number of stars in catalog
[…]

### 2012-03-16 paper/self_calibration_I.tex
more useful README; prepare for github; abstract, optimization, acks, symbols
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@20412 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
-\documentclass[manuscript]{aastex}
+\documentclass[preprint,pdftex]{aastex}
 \newcommand{\vdag}{(v)^\dagger}
@@
 \affil{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, Heidelberg, 69117, Germany.}
-\altaffiltext{1}{2 Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA.}
+\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA.}
 
 \begin{abstract}
-In this paper, we investigate under which survey conditions the self-calibration of an instrument response can produce precise photometric catalogs for wide-field imaging surveys, drawing only on the science data themselves.
-We consider four basic survey strategies and create end-to-end simulations of an imaging survey, which produce simulated photometric catalogs from mock observations of a realistic, synthetic sky.
-These catalog-based simulations include realistic measurement uncertainties and a complex, position-dependent instrument response.
-We then use a self-calibration technique to recover the instrument response by fitting an model that best explains the survey catalog, based on the sources that fall on different focal plane positions in separate observations.
-We demonstrate that with a favorable survey strategy, a complex (but smooth) instrument response can be precisely calibrated using this self-calibration technique.
-We also find that returning the same sources to very different focal plane positions is the key property of any survey strategy designed for accurate retrospective calibration.
-The results of this work suggest the following advice for those considering the design of large-scale imaging surveys: Do not use a regular tiling of the sky; instead return the same sources to very different focal plane positions.
+We study the self-calibration---determination of complex system
+response from science data alone---for precise photometric catalogs
+from wide-field imaging surveys. We create an artificial sky of stars
+and synthetically observe it under four basic survey strategies,
+creating an end-to-end simulation of an imaging survey for each. These
+catalog-level simulations include realistic measurement uncertainties
+and a complex, position-dependent instrument response. We then use a
+self-calibration technique to recover the instrument response by
+simultaneously fitting for all the star fluxes and the parameters of a
+position-dependent flat-field, optimizing a justified likelihood
+function.  For realism, we deliberately fit with a wrong noise model
+and a flat-field functional basis that does not include the model that
+generated the synthetic data.  We demonstrate that with a favorable
+survey strategy, a complex (but smooth) instrument response can be
+precisely self-calibrated. We also find that returning the same
+sources to very different focal plane positions is the key property of
+any survey strategy designed for accurate retrospective
+calibration. The results of this work suggest the following advice for
+those considering the design of large-scale imaging surveys: Do not
+use a regular tiling of the sky; instead return the same sources to
+very different focal plane positions.
 \end{abstract}
@@
 
-In this work we do not produced pixelated images; instead we concentrate on catalog-level simulations with realistic measurement uncertainties.
-Complex effects are included in the simulations that are not  precisely modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the catalog.
+In this work we do not produce pixelated images.
+Instead we create catalog-level simulations with realistic measurement uncertainties, including unmodeled stellar variability.
+Complex effects are included in the simulations that are (deliberately) not correctly modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the catalog.
 
@@
 \begin{eqnarray}
-\sigma_{i, \true}^{2} = (1 + \epsilon_i) \, \alpha^{2} + \eta^{2}\, [ f_\true(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} ]^2 \quad , \label{eqn:noise}
+\sigma_{i, \true}^{2} = (1 + \epsilon_i) \, \delta^{2} + \eta^{2}\, [ f_\true(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} ]^2 \quad , \label{eqn:noise}
 \end{eqnarray}
-where $\alpha$ and $\eta$ are both constants and $\epsilon_i$ is a random number, drawn uniformly in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$.
-The $m = 22$ mag 10$\upsigma$ detection limit introduced previously and the limit S/N$<500$ are used to set $\alpha = 0.1585$ and $\eta = 0.0017$.
+where $\delta$ and $\eta$ are both constants and $\epsilon_i$ is a random number, drawn uniformly in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$.
+The $m = 22$ mag 10$\upsigma$ detection limit introduced previously and the limit S/N$<500$ are used to set $\delta = 0.1585$ and $\eta = 0.0017$.
 The $\epsilon_i$ contribution is not taken into account in the analysis and therefore the uncertainty variances on the count rates are \textit{assumed} (incorrectly, but realistically) to be
 \begin{eqnarray*}
-\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2} \, c^{2}_i \quad 
+\sigma_{{i}}^{2} = \delta^{2} + \eta^{2} \, c^{2}_i \quad 
 \end{eqnarray*}
@@
 In this work we are interested in simulating complete surveys, in order to identify the crucial characteristics of survey strategies that allow for the relative instrument response model to be accurately, retrospectively constrained from the resulting catalog.
-We therefore apply the single exposure procedure introduced in the previous section for each pointing specified in a defined survey strategy.
+We therefore apply the single exposure procedure introduced in the previous section for each pointing specified in a defined survey strategy, which is a set of pointings $(\alpha, \beta)$ and orientations $(\theta)$.
 The resultant source measurements are collated into a survey-wide catalog.
@@
 We self-calibrate the catalog generated in the survey simulations to recover an optimum estimate of the true instrument response applied and the true source fluxes.
-This self-calibration procedure has been successfully applied to ground-based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}\footnote{Here this procedure is dubbed ``uber-calibration.''} and the Deep Lens Survey \citep{wit11}.
+This self-calibration procedure has been successfully applied to ground-based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}, where the procedure is dubbed ``uber-calibration'', and the Deep Lens Survey \citep{wit11}.
 This iterative procedure comprises two steps: (1) a refinement of the source flux estimates based on the latest instrument response model and (2) a refinement of the instrument response model based on the updated source flux estimates.
@@
 \begin{eqnarray*}
-\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2}\, c^{2}_i \quad ,
+\sigma_{{i}}^{2} = \delta^{2} + \eta^{2}\, c^{2}_i \quad ,
 \end{eqnarray*} 
-where $\alpha$ and $\eta$ are the parameters set by the instrument's $10\upsigma$ detection and the S/N $<500$ limits.
+where $\delta$ and $\eta$ are the parameters set by the instrument's $10\upsigma$ detection and the S/N $<500$ limits.
 The $\epsilon_i$ error contribution is intentionally not included in order to simulate systematic problems with the instrument noise model.
@@
 
+It is worthy of note that although this back-and-forth (Step 1, Step
+2) iteration scheme works well and is easy to understand, there is no
+reason in principle not to simply put the whole system into a giant
+non-linear optimizer.  A sufficiently clever general-purpose optimizer
+might outperform the bilinear iterations.  At the scale of the
+simulations and optimizations performed in this paper, there is no
+need to look for higher performance algorithms than the bilinear
+iterative solution presented here.
+
 \section{Metrics}
@@
 These survey strategies are summarized in Table \ref{tab:simple_surveys} and are shown in Figure \ref{fig:simple_surveys}.
-Strategy A is the simplest strategy; the target field is regularly tiled with small overlaps between adjacent pointings ($\sim 5$~\% and $\sim 8$~\% in the camera pointing directions $\alpha$ and $\beta$ respectively).
+Strategy A is the simplest strategy; the target field is regularly tiled with small overlaps between adjacent pointings ($\sim 5$~percent and $\sim 8$~percent in the camera pointing directions $\alpha$ and $\beta$ respectively).
 The pointings in the 12 passes over the same field are exactly aligned.
@@
 The pointing positions in Survey D are quasi-random: the pointings are the same as with Survey A, but each has a random offset within [-0.35,0.35)~deg applied in both the $\alpha$ and $\beta$ directions.
-By fixing the pointings within these $0.7~\text{deg} \times 0.7~\text{deg}$ boxes, we ensure that the quasi random strategy has a uniform coverage of the field.
+By fixing the pointings within these $0.7~\text{deg} \times 0.7~\text{deg}$ boxes, we ensure that the quasi random strategy has a uniform coverage of the field; that is, the fluctuations in coverage are less than they would be in a Poisson (totally random) process.
 The orientations of the pointings in Survey D are completely random.
@@
 With the Survey C and D catalogs, the instrument response is accurately recovered.
-As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top), the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level.
+As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top), the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~percent$ level.
 That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the model basis used.
@@
 \section{Discussion}
-From the simulations presented in this paper\footnote{All of the simulations presented in this paper are publicly available at \url{http://trac.astrometry.net/browser/trunk/projects/euclid}.}, we are able to draw two firm conclusions about the self-calibration of large-scale imaging surveys.
+From the simulations presented in this paper, we are able to draw two firm conclusions about the self-calibration of large-scale imaging surveys.
 The first is that, in agreement with previous work, we show that this technique is very powerful, allowing the large-scale, relative instrument response of an imaging instrument to be accurately retrospectively constrained from the science images alone.
@@
 \section{Acknowledgments}
-Rory Holmes was funded by the Deutschen Zentrums f\"ur Luft- und Raumfahrt (DLR) through the grant 50QE1202.
+It is a pleasure to thank Rob Fergus (NYU), Jason Kalirai (STScI),
+and Sam Roweis (deceased) for valuable conversations.
 
-DWH was partially supported by NASA (grant XXX), the NSF (grant XXX), and a research fellowship from the Alexander von Humboldt Foundation.
+RH was funded by the Deutschen Zentrums f\"ur Luft- und Raumfahrt
+(DLR) through the grant 50QE1202.  DWH was partially supported by NASA
+(grant NNX08AJ48G), the NSF (grants AST-0908357 and IIS-1124794), and
+a research fellowship from the Alexander von Humboldt Foundation.
+
+All of the code used in this paper is publicly available at
+\url{http://github.com/davidwhogg/SelfCalibration/}~.
 
@@
 Field-of-View (deg$^{2}$)& $0.76 \times 0.72$ \\
-Noise Model -- Eqn. \ref{eqn:noise} & $\alpha = 0.1585$, $\eta = 0.0017$, $\epsilon_\text{max} = 1.0$\\
+Noise Model -- Eqn. \ref{eqn:noise} & $\delta = 0.1585$, $\eta = 0.0017$, $\epsilon_\text{max} = 1.0$\\
 Fitted Instrument Response Model & 8$^{\text{th}}$ order polynomial\\
@@
 \tablehead{
-\colhead{Survey Name} & \colhead{Iterations} & \colhead{Source Error $S_\rms$(\%)} & \colhead{True Badness $B_\true$ (\%)} & \colhead{Best-in-Basis Badness $B_\basis$ (\%)} }
+\colhead{Survey Name} & \colhead{Iterations} & \colhead{Source Error}       & \colhead{True Badness}        & \colhead{Best-in-Basis} \\
+                      &                      & \colhead{$S_\rms$ (percent)} & \colhead{$B_\true$ (percent)} & \colhead{$B_\basis$ (percent)}}
 \startdata

### 2012-03-19 paper/self_calibration_I.tex
euclid: minor english editing to paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@20439 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 \begin{abstract}
-We study the self-calibration---determination of complex system
-response from science data alone---for precise photometric catalogs
-from wide-field imaging surveys. We create an artificial sky of stars
+We study the self-calibration -- the determination of a complex system
+response from science data alone -- for precise photometric catalogs
+from wide-field imaging surveys. We create an artificial sky of sources
 and synthetically observe it under four basic survey strategies,
@@
 This work focuses on the relative photometric calibration of a typical imaging instrument only, although similar techniques could be used to constrain many other calibration parameters, such as the optical distortion, point-spread-functions and so-on.
-The self-calibration technique utilizes the multiple measurements of (non-varying) sources, at different focal plane positions and at different times within the survey, to constrain the relative instrument response by requiring that the post-calibration measurements yield consistent flux values.
+The self-calibration technique utilizes the multiple measurements of sources, at different focal plane positions and at different times within the survey, to constrain the relative instrument response by requiring that the post-calibration measurements yield consistent flux values.
 Through end-to-end, catalog-level survey simulations, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.
@@
 In this work we do not produce pixelated images.
-Instead we create catalog-level simulations with realistic measurement uncertainties, including unmodeled stellar variability.
+Instead we create catalog-level simulations with realistic measurement uncertainties, including unmodeled source variability.
 Complex effects are included in the simulations that are (deliberately) not correctly modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the catalog.
@@
 \label{sec:single_exposure}
-With a camera pointing $(\alpha, \beta)$ and camera orientation $\theta$, we find the focal plane positions of the sources on the synthetic sky that fall within the instrument's field-of-view.
+With a telescope pointing $(\alpha, \beta)$ and camera orientation $\theta$, we find the focal plane positions of the sources on the synthetic sky that fall within the instrument's field-of-view.
 In our simulations, we use a large instrument field-of-view of $0.76~ \text{deg} \times 0.72~\text{deg}$; a size consistent with the next generation of large survey imagers.
@@
 To generate photometric catalogs, the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model.
-For a measurement $i$ the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s_{k,\true}$ ,
+For a measurement $i$, the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$ -- which is a function of focal plane position $\vec{x_i}$ -- and the source's true flux $s_{k,\true}$ ,
 \begin{eqnarray*}
@@
 With the Survey C and D catalogs, the instrument response is accurately recovered.
-As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top), the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~percent$ level.
+As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top), the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1$~percent level.
 That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the model basis used.

### 2012-03-22 paper/self_calibration_I.tex
euclid: fixed abstract sentence
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@20474 03a3cea6-2c03-0410-ac9b-9271b1c66c29

@@
 catalog-level simulations include realistic measurement uncertainties
-and a complex, position-dependent instrument response. We then use a
-self-calibration technique to recover the instrument response by
-simultaneously fitting for all the star fluxes and the parameters of a
-position-dependent flat-field, optimizing a justified likelihood
-function.  For realism, we deliberately fit with a wrong noise model
-and a flat-field functional basis that does not include the model that
-generated the synthetic data.  We demonstrate that with a favorable
-survey strategy, a complex (but smooth) instrument response can be
-precisely self-calibrated. We also find that returning the same
-sources to very different focal plane positions is the key property of
-any survey strategy designed for accurate retrospective
-calibration. The results of this work suggest the following advice for
-those considering the design of large-scale imaging surveys: Do not
-use a regular tiling of the sky; instead return the same sources to
-very different focal plane positions.
+and a complex, position-dependent instrument response. In the
+self-calibration step, we simultaneously fit for all the star fluxes
+and the parameters of a position-dependent flat-field.  For realism,
+we deliberately fit with a wrong noise model and a flat-field
+functional basis that does not include the model that generated the
+synthetic data.  We demonstrate that with a favorable survey strategy,
+a complex (but smooth) instrument response can be precisely
+self-calibrated. We also find that returning the same sources to very
+different focal plane positions is the key property of any survey
+strategy designed for accurate retrospective calibration. The results
+of this work suggest the following advice for those considering the
+design of large-scale imaging surveys: Do not use a regular tiling of
+the sky; instead return the same sources to very different focal plane
+positions.
 \end{abstract}

### 2012-03-27 paper/self_calibration_I.tex
paper: minor corrections after HW comments. SUBMITTED to PASP

@@
 catalog-level simulations include realistic measurement uncertainties
-and a complex, position-dependent instrument response. In the
+and a complex focal-plane dependence of the instrument response. In the
 self-calibration step, we simultaneously fit for all the star fluxes
@@
 a complex (but smooth) instrument response can be precisely
-self-calibrated. We also find that returning the same sources to very
+self-calibrated. We show that returning the same sources to very
 different focal plane positions is the key property of any survey
@@
 
-\keywords{Relative Photometric Calibration: Imaging Survey: Survey Strategy}
+\keywords{Catalogs --- methods: data analysis --- techniques: photometric}
 
@@
 Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS), much more calibration information was obtained from the science data than the calibration data \citep{pad08}.
-But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: Good redundancy was required in the data stream, and a redundancy of a very specific kind.
+But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: good redundancy was required in the data stream, and a redundancy of a very specific kind.
 
@@
 This work focuses on the relative photometric calibration of a typical imaging instrument only, although similar techniques could be used to constrain many other calibration parameters, such as the optical distortion, point-spread-functions and so-on.
-The self-calibration technique utilizes the multiple measurements of sources, at different focal plane positions and at different times within the survey, to constrain the relative instrument response by requiring that the post-calibration measurements yield consistent flux values.
+The self-calibration technique utilizes the multiple measurements of sources in the survey arising from overlapping pointings.
+If these redundant measurements are recorded at different focal plane positions or at different times within the survey, it is possible to constrain the relative instrument response by requiring that the post-calibration measurements yield consistent flux values.
 Through end-to-end, catalog-level survey simulations, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.
@@
 
-\section{Survey catalog Simulations}
+\section{Survey Catalog Simulations}
 \label{sec:survey_simulations}
@@
 
-\subsection{The Synthetic Sky}
+\subsection{Synthetic Sky}
 We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and a $10\upsigma{}$ limits of deep, space-based, near-infrared survey.
@@
 
-\subsubsection{The True Instrument Response Model}
+\subsubsection{True Instrument Response Model}
 \label{sec:instrument_response_model}
@@
 
-\subsection{The Complete Survey}
+\subsection{Complete Survey}
 In this work we are interested in simulating complete surveys, in order to identify the crucial characteristics of survey strategies that allow for the relative instrument response model to be accurately, retrospectively constrained from the resulting catalog.
@@
 
-\section{Self-Calibrating the Survey-Wide catalog}
+\section{Self-Calibrating the Survey-Wide Catalog}
 \label{sec:self_cal}
@@
 
-\section{Metrics}
+\section{Performance Metrics}
 \label{sec:metrics}
@@
 For this to be effective though, it will be necessary to not only observe the same sources at very different focal plane positions, but also to observe them at very different times within the survey.
-To quantify this in more detail, we will include temporal variations in future simulations.
 
 In this paper, we have considered the design of survey strategies for calibrating the large-scale, relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained from this method.
-For example, the optical distortion of an instrument or the point-spread-function variation could also be constrained with such a method, although we concede that the properties of a survey strategy that makes them good for one calibration, may not be the same as that for another; ultimately a trade-off would need to be performed between the different calibration concerns and the other goals of the survey, such as sky coverage, variability monitoring and so-on.
+For example, the optical distortion of an instrument or the point-spread-function variation could also be constrained with such a method, although it would need to be explored whether the properties of a survey strategy that makes it good for one calibration, are also suitable for other calibration aspects; ultimately a trade-off would need to be performed between the different calibration concerns and the other goals of the survey, such as sky coverage, variability monitoring and so-on.
 
@@
 \plotone{A_8192_ff.pdf}
-\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey A catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
+\figcaption{Survey Strategy A: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey A catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
 The self-calibration procedure did not converged to the final solution and the results presented here correspond to the instrument response found after $8192$ iterations.
@@
 \plotone{B_4160_ff.pdf}
-\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey B catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
+\figcaption{Survey Strategy B: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey B catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
 The self-calibration procedure converged to the final solution after 4160 iterations.
@@
 \plotone{C_026_ff.pdf}
-\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey C catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
+\figcaption{Survey Strategy C: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey C catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
 The self-calibration procedure converged to the final solution after 26 iterations.
@@
 \plotone{D_009_ff.pdf}
-\figcaption{A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey D catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (b) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
+\figcaption{Survey Strategy D: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey D catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (b) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
 The self-calibration procedure converged to the final solution after 9 iterations.

### 2012-04-15 paper/self_calibration_I.tex
paper: remove reference to code for ArXiv upload
minor changes following reviewer comments

@@
 With a telescope pointing $(\alpha, \beta)$ and camera orientation $\theta$, we find the focal plane positions of the sources on the synthetic sky that fall within the instrument's field-of-view.
-In our simulations, we use a large instrument field-of-view of $0.76~ \text{deg} \times 0.72~\text{deg}$; a size consistent with the next generation of large survey imagers.
+In our simulations, we use a large instrument field-of-view of $0.76~ \text{deg} \times 0.72~\text{deg}$; a size consistent with current large survey imagers.
 An example of a single pointing exposure is shown in Figure \ref{fig:single_image}.
@@
 Secondly, the assumed measurement uncertainty variances do not include the additional random measurement error $\epsilon_{i}$ introduced in Subsection \ref{sec:noise}.
-We
 We therefore fit for the following, incomplete measurement model:
@@
 \subsection{Survey Description}
-We label the four survey strategies -- which all cover the same patch of sky -- with the letters A to D.
+We label the four survey strategies -- which all cover the same 64~deg$^2$ patch of sky -- with the letters A to D.
 These survey strategies are summarized in Table \ref{tab:simple_surveys} and are shown in Figure \ref{fig:simple_surveys}.
@@
 
-All of the code used in this paper is publicly available at
-\url{http://github.com/davidwhogg/SelfCalibration/}~.
+%All of the code used in this paper is publicly available at \url{http://github.com/davidwhogg/SelfCalibration/}~.
 

### 2012-06-04 paper/referee_response.txt
response to all apart from one referee comment

excellent suggestions, improve the paper. Respond inline

	--Sec 2.2: "current" generation of large survey imagers would be more
	  accurate than "next" generation.

Done.

	--Sec 3.1: "We We"

Done.

	--Sec 5.1: please include some description of the density and/or total
	  number of sources used, either here or in Sec. 2.2.  Also insert "16
	  deg^2" before "patch of sky".

We explicitly put the source density into the last sentence of 2.1 and we re-point the reader to the parameter table in the last sentence prior to Section 5.1. We included the solid angle of the total patch, as requested.

	--suggestion for going a bit deeper in Sec 5.1: some surveys may find
	  it difficult to rotate their hardware randomly with every shot.
	  Could you do a survey E which has random pointing centers but no
	  rotation?

The referee is absolutely right that there are cameras that cannot be rotated with respect to the sky. That is one of the reasons why we designed Survey C without rotation. Since Survey C calibrates just as well as Survey D, we conclude that rotation is not necessary for good self-calibration. We agree that a Survey E would confirm that, but we believe that Survey C already makes this point well. We added a few sentences to the discussion of Survey C and D about this point in Sections 5.1 and 5.2.

	--suggestion for going a bit deeper in Sec 5.2 (and/or Sec 4): you
	  show the residuals in the instrument response, but more relevant
	  might be the residuals in the final survey.  For Survey A (and to a
	  large extent B), the residuals in the final survey are just as bad
	  as in the instrument response.  But in Surveys C and D, they get beat
	  down because the calibration error on a given piece of sky changes
	  from observation to observation.  It would be instructive to add a
	  performance metric which includes this effect.

In Addition to these changes, we also made the following small changes:
	- MeasuredCatalog.true_invvar - multiplied by flat-field one time too many??
	- Generate epsilon now per measurement and not per source

Neither made significant changes to content of papers

### 2012-06-04 paper/self_calibration_I.tex
response to all apart from one referee comment

@@
 \begin{abstract}
-We study the self-calibration -- the determination of a complex system
-response from science data alone -- for precise photometric catalogs
+We study the self-calibration---the determination of a complex system
+response from science data alone---for precise photometric catalogs
 from wide-field imaging surveys. We create an artificial sky of sources
@@
 The source magnitudes $m$ are related to the source fluxes $s$ simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies (nmgy).
-To limit the data, we only select the brightest sources within the survey area, up to a source density per unit solid angle $d$, for the self-calibration procedure.
+To limit the data, we only select the brightest sources within the survey area, up to a source density $d=100$~deg$^{-2}$, for the self-calibration procedure.
 
@@
 To generate photometric catalogs, the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model.
-For a measurement $i$, the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$ -- which is a function of focal plane position $\vec{x_i}$ -- and the source's true flux $s_{k,\true}$ ,
+For a measurement $i$, the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$---which is a function of focal plane position $\vec{x_i}$---and the source's true flux $s_{k,\true}$ ,
 \begin{eqnarray*}
@@
 We use the end-to-end catalog simulation and self-calibration chain, introduced in Sections \ref{sec:survey_simulations} and \ref{sec:self_cal}, to evaluate the performance of the self-calibration procedure with these strategies.
+The parameters of these simulations are summarized in Table \ref{tab:parameters}.
 
 \subsection{Survey Description}
-We label the four survey strategies -- which all cover the same 64~deg$^2$ patch of sky -- with the letters A to D.
+We label the four survey strategies---which all cover the same 64~deg$^2$ patch of sky---with the letters A to D.
 These survey strategies are summarized in Table \ref{tab:simple_surveys} and are shown in Figure \ref{fig:simple_surveys}.
@@
 Survey B is the same as A, but with each pass over the target field the orientation of the telescope is rotated by $30^\circ$.
+
 Survey C is more complex.
@@
 These three passes are then repeated four times.
+
 The pointing positions in Survey D are quasi-random: the pointings are the same as with Survey A, but each has a random offset within [-0.35,0.35)~deg applied in both the $\alpha$ and $\beta$ directions.
@@
 The orientations of the pointings in Survey D are completely random.
+
+Surveys A and C can be executed without rotating the camera with respect to the celestial coordinates. Surveys B and D can only be executed with systems that permit rotations of the camera.
 
@@
 
+Surveys C and D perform comparably well, although they are very different in their configuration.
+Two substantial advantages of Survey C over Survey D are: In Survey C the camera would not need to be rotated relative to the sky between pointings; indeed many survey imaging cameras have fixed celestial orientations.
+Survey D requires generating quasi-random numbers, whereas Survey C has a completely deterministic set of pointing centers.
+This is a conceptual advantage to Survey C, but possibly also an operational advantage.
+
 The Survey A and B catalogs also give interesting results.
 With Survey A, the self-calibration procedure did not reach a solution.
-This -- in itself -- is an important result, as the naive tiling of the sky currently being considered for many large surveys will result in catalogs that cannot be retrospectively self-calibrated.
+This---in itself---is an important result, as the naive tiling of the sky currently being considered for many large surveys will result in catalogs that cannot be retrospectively self-calibrated.
 Minor changes to survey strategies, such as those in Survey C, can drastically improve the performance of the self-calibration procedure.
@@
 
-In this paper, we have considered the design of survey strategies for calibrating the large-scale, relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained from this method.
+In this paper, we have considered the design of survey strategies for calibrating the large-scale, relative photometric response of an instrument, but there are many more calibration parameters that can be constrained from this method.
 For example, the optical distortion of an instrument or the point-spread-function variation could also be constrained with such a method, although it would need to be explored whether the properties of a survey strategy that makes it good for one calibration, are also suitable for other calibration aspects; ultimately a trade-off would need to be performed between the different calibration concerns and the other goals of the survey, such as sky coverage, variability monitoring and so-on.
@@
 \acknowledgments
-\section{Acknowledgments}
 It is a pleasure to thank Rob Fergus (NYU), Jason Kalirai (STScI),
@@
 \begin{thebibliography}{}
-\bibitem[Padmanabhan et al. (2008)]{pad08}{Padmanabhan, N. et al., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj~674, p1217.}
-\bibitem[Windhorst et al.
-(2011)]{win11}{Windhorst, R.~A.
+\bibitem[Padmanabhan et al.(2008)]{pad08}{Padmanabhan, N. et al., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj~674, p1217.}
+\bibitem[Windhorst et al.(2011)]{win11}{Windhorst, R.~A.
 et al., The Hubble Space Telescope Wide Field Camera 3 Early Release Science Data: Panchromatic Faint Object Counts for 0.2-2 $\upmu$m Wavelength, 2011, \apjs~193, p27.}
-\bibitem[Wittman et al. (2011)]{wit11}{Wittman, D., Ryan, R. and Thorman, P., Ubercalibration of the Deep Lens Survey, 2011, ArXiv e-prints, 1111.2058}
+\bibitem[Wittman et al.(2011)]{wit11}{Wittman, D., Ryan, R. and Thorman, P., Ubercalibration of the Deep Lens Survey, 2011, ArXiv e-prints, 1111.2058}
 \end{thebibliography}

### 2012-06-04 src/code_examples/paper_plots/parameters.py

@@ -22,7 +22,7 @@
     # Float: the maximum number of sources (all magnitude) per unit area
     # on the sky. If more than this number are produced by the magnitude
     # distribution (see below), only the brightest are selected.
-    'density_of_stars'    :     200.,
+    'density_of_stars'    :     100.,
     # Float array: The parameters describing the magnitude distribution
     # of the sources in the sky, according to
     # log10(dN/dm) = A + B * mag + C * mag ** 2

### 2012-07-27 paper/referee_response.txt
made all referee changes

@@
-excellent suggestions, improve the paper. Respond inline
+The review makes some excellent suggestions to improve the paper. Please see the inline responses below:
 
@@
 
+This is an excellent suggestion. We have added Figures 7 to 11, which show the residuals in the final survey.
+
 In Addition to these changes, we also made the following small changes:
-	- MeasuredCatalog.true_invvar - multiplied by flat-field one time too many??
-	- Generate epsilon now per measurement and not per source
-
-Neither made significant changes to content of papers
+	- We removed "Large-Scale" from the paper title
+	- We corrected a few minor bugs in the code, none of which made made significant changes to the contents of the paper.
+	- We have modified the form of the true instrument response (last equation in 2.2.3) to now include a (1+a+b)**-0.75 factor.
+This reduces the magnitude of the high frequency sine and cosine contributions.
+We did this so that the best-in-basis instrument response was closer to the true instrument response.
+	- The code used in this work is publicly available and we added the github links in Section 2 and the Acknowledgements 
+	- We re-ran the simulations with a higher source density (d = 300 not d = 100 deg^-2), as this makes the survey residual plots (Figure 7 - 11) clearer.
+	- For clarity, we have slightly reworded the abstract.
+	- Finally, we have also corrected a few typos.

### 2012-07-27 paper/self_calibration_I.tex
made all referee changes

@@
 \documentclass[preprint,pdftex]{aastex}
+%\documentclass[draft,pdftex]{aastex}
+
 \newcommand{\vdag}{(v)^\dagger}
@@
 \begin{document}
-
 %% LaTeX will automatically break titles if they run longer than
@@
 \begin{abstract}
-We study the self-calibration---the determination of a complex system
-response from science data alone---for precise photometric catalogs
-from wide-field imaging surveys. We create an artificial sky of sources
+In this paper, we investigate the impact of survey strategy on the performance of self-calibration when the goal is to produce accurate photometric catalogs
+from wide-field imaging surveys.
+This self-calibration technique utilizes multiple measurements of sources to constrain instruments' large-scale response (flat-field) from survey science data alone.
+We create an artificial sky of sources
 and synthetically observe it under four basic survey strategies,
@@
 synthetic data.  We demonstrate that with a favorable survey strategy,
-a complex (but smooth) instrument response can be precisely
+a complex instrument response can be precisely
 self-calibrated. We show that returning the same sources to very
 different focal plane positions is the key property of any survey
-strategy designed for accurate retrospective calibration. The results
+strategy designed for accurate retrospective calibration of this type. The results
 of this work suggest the following advice for those considering the
-design of large-scale imaging surveys: Do not use a regular tiling of
+design of large-scale imaging surveys: Do not use a regular, repeated tiling of
 the sky; instead return the same sources to very different focal plane
@@
 \label{sec:survey_simulations}
-We have constructed an end-to-end simulation chain that produces a realistic imaging catalog from a specified survey strategy.
+We have constructed an end-to-end simulation chain that produces a realistic imaging catalog from a specified survey strategy\footnote{All code used in this work is publicly available at \url{http://github.com/davidwhogg/SelfCalibration/}}.
 In this work we have kept the simulation parameters intentionally flexible, so that the sensitivity of the self-calibration procedure to different values can be investigated (the fiducial values are given in Table \ref{tab:parameters}).
@@
 \subsection{Synthetic Sky}
-We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and a $10\upsigma{}$ limits of deep, space-based, near-infrared survey.
+We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and a $10\upsigma{}$ limits of deep, space-based, near-infrared, imaging survey.
 Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes $m$ distributed according to
-\begin{eqnarray}
+\begin{eqnarray*}
 \log_{10} \frac{\dd N}{\dd m \, \dd \Omega} = a + b\,m + c\,m^2 \label{eqn:power_law} \quad , 
-\end{eqnarray}
+\end{eqnarray*}
 where $\dfrac{\dd N}{\dd m \, \dd  \Omega}$  is the density of sources $N$ per unit magnitude $m$ and per unit solid angle $\Omega$, and $a$, $b$ and $c$ are model parameters.
@@
 The source magnitudes $m$ are related to the source fluxes $s$ simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies (nmgy).
-To limit the data, we only select the brightest sources within the survey area, up to a source density $d=100$~deg$^{-2}$, for the self-calibration procedure.
+To limit the data, we only select the brightest sources within the survey area, up to a source density $d=300$~deg$^{-2}$, for the self-calibration procedure.
 
@@
 \begin{eqnarray*}
-f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7\ldots 260})  & = &  \sum_{a=0}^6 \sum_{b=0}^a \left[ q_{\true, 7+4b} \cos (k_x \, x_i) + q_{\true, 8+4b} \sin (k_x \, x_i) \right] \\
+f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7\ldots 260})  & = &  \sum_{a=0}^6 \sum_{b=0}^a \left(1 + a + b\right) ^ {-0.75} \left[ q_{\true, 7+4b} \cos (k_x \, x_i) + q_{\true, 8+4b} \sin (k_x \, x_i) \right] \\
 & & \qquad \qquad \times \left[ q_{\true, 9+4b} \cos (k_y \, y_i) + q_{\true, 10+4b} \sin (k_y \, y_i) \right] \quad ,
@@
 with the physical focal plane dimensions $X$ and $Y$.
+The $(1 + a +d)^{-0.75}$ factor reduces the magnitude of the higher order terms.
 In total, the instrument response model is parameterized with 260 parameters; an example can be seen in Figure \ref{fig:single_image}      (right).
@@
 With this instrument response model, we have assumed that the variations in sensitivities on a pixel-to-pixel-scale have been calibrated with some other method, such as an internal flat-field calibration source.
-It is variations on larger scales that we aim to constrain with the self-calibration procedure.
+It is only the variations on larger scales that we aim to constrain with the self-calibration procedure.
 Our final assumption in these simulations is that the instrument response is temporally stable.
@@
 \begin{eqnarray*}
-\vec{q'} = G^{-1}  \cdot \vec{b}   \quad .
+\vec{q'} \leftarrow G^{-1}  \cdot \vec{b}   \quad .
 \end{eqnarray*}
@@
 2) iteration scheme works well and is easy to understand, there is no
-reason in principle not to simply put the whole system into a giant
+reason in principle not to simply put the whole system into a
 non-linear optimizer.  A sufficiently clever general-purpose optimizer
-might outperform the bilinear iterations.  At the scale of the
+might outperform these bilinear iterations.  At the scale of the
 simulations and optimizations performed in this paper, there is no
@@
 
-The self-calibration procedure performs best with the Survey D catalog, followed by the Survey C catalog, then the Survey B catalog.
-The self-calibration procedure failed completely with the Survey A catalog and no solution was obtained.
-The number of iterations of the self-calibration procedure required to converge to the final solution also decreases as the quality of the final fit increases: Survey D required 9 iterations, Survey C 26, Survey B 4160 and the procedure was iterated 8192 times with Survey A without reaching a solution.
+The survey strategies, the resultant sky coverages and the errors in final self-calibration flux solutions are summarized for each survey individually in Figures \ref{fig:A_survey} to \ref{fig:D_survey}.
+With Surveys A and B---and to a lesser extent Survey C---the imprint of the strategies can be clearly seen in the coverage maps.
+It is important to note that---with sufficient re-visits to the field---the coverage of the quasi-random survey D is more uniform than the other strategies investigated.
+
+The self-calibration procedure performs best with the Survey C and Survey D catalogs.
+The performance of the self-calibration procedure is worse for the Survey B catalog and it fails completely with the Survey A catalog, where no solution was obtained.
+The number of iterations of the self-calibration procedure required to converge to the final solution also decreases as the quality of the final fit increases: Survey C required 163 iterations, Survey D 197, Survey B 3,746 and the procedure was iterated 10,000 times with Survey A without reaching a solution.
 
 With the Survey C and D catalogs, the instrument response is accurately recovered.
-As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top), the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1$~percent level.
+As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top), the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 0.5$~percent level.
 That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the model basis used.
@@
 Here, it is the quality of the fit over the focal plane that is interesting.
-From the residual plot against the best-in-basis instrument response $f_\basis(\vec{x_j} | \vec{q_\basis})$ (Fig \ref{fig:B_result} bottom right), we can see that the quality of the fit is rotationally symmetric, with the self-calibration procedure able to fit the instrument response well at the edge of the focal plane, but poorly at the center.
+From the residual plot against the best-in-basis instrument response $f_\basis(\vec{x_j} | \vec{q_\basis})$ (Figure \ref{fig:B_result}d), we can see that the quality of the fit is rotationally symmetric, with the self-calibration procedure able to fit the instrument response well at the edge of the focal plane, but poorly at the center.
+As a result, the fluxes of the sources that fall at the center of the focal plane during a measurement cannot be constrained as well as those that fall at the edges; this can clearly be seen in the post-self-calibration flux error map shown in Figure \ref{fig:B_survey}d.
 
@@
 
-RH was funded by the Deutschen Zentrums f\"ur Luft- und Raumfahrt
-(DLR) through the grant 50QE1202.  DWH was partially supported by NASA
+Rory Holmes was funded by the Deutschen Zentrums f\"ur Luft- und Raumfahrt
+(DLR) through the grant 50QE1202.  David W. Hogg was partially supported by NASA
 (grant NNX08AJ48G), the NSF (grants AST-0908357 and IIS-1124794), and
@@
 
-%All of the code used in this paper is publicly available at \url{http://github.com/davidwhogg/SelfCalibration/}~.
+The code used in this work is publicly available at \url{github.com/davidwhogg/SelfCalibration/}~.
 
@@
 \begin{figure}
-\plotone{./single_image.pdf}
+\plotone{./camera_image.pdf}
 \figcaption{A single exposure of the synthetic sky.
@@
 Survey Area (deg$^2$)& $8 \times 8$ \\
-Source Density (deg$^{-2}$) & $ d = 100 $ \\
+Source Density (deg$^{-2}$) & $ d = 300 $ \\
 Saturation Limit (mag) & $m_\text{min} = 17$\\
@@
 \figcaption{Survey Strategy A: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey A catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
-The self-calibration procedure did not converged to the final solution and the results presented here correspond to the instrument response found after $8192$ iterations.
+The self-calibration procedure did not converged to the final solution and the results presented here correspond to the instrument response found after 8,192 iterations.
 The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial).
@@
 \begin{figure}
-\plotone{B_4160_ff.pdf}
-\figcaption{Survey Strategy B: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey B catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models.
-The self-calibration procedure converged to the final solution after 4160 iterations.
+\plotone{B_3746_ff.pdf}
+\figcaption{Survey Strategy B: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey B catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
+The self-calibration procedure converged to the final solution after 3,746 iterations.
 The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
@@
 \begin{figure}
-\plotone{C_026_ff.pdf}
+\plotone{C_163_ff.pdf}
 \figcaption{Survey Strategy C: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey C catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
-The self-calibration procedure converged to the final solution after 26 iterations.
+The self-calibration procedure converged to the final solution after 163 iterations.
 The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:C_result}}
@@
 \begin{figure}
-\plotone{D_009_ff.pdf}
+\plotone{D_197_ff.pdf}
 \figcaption{Survey Strategy D: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey D catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (b) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
-The self-calibration procedure converged to the final solution after 9 iterations.
+The self-calibration procedure converged to the final solution after 197 iterations.
 The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
 With this quasi-random observing strategies, all regions of the focal plane well interconnected with each other.\label{fig:D_result}}
+\end{figure}
+
+\begin{figure}
+\plotone{A_survey.pdf}
+\figcaption{A summary of Survey Strategy A and the resultant self-calibration results: (a) shows the survey footprint diagram (as shown in Figure \ref{fig:simple_surveys}), (b) shows the source coverage for this survey, with (c) showing a histogram of the number of source observations and (d) shows the flux errors for each of the sources after the self-calibration procedure.
+The imprint of the survey strategy can be clearly seen in the coverage map.
+With this survey strategy, the self-calibration procedure is not able to converge to a solution and therefore there are strong and regular residuals in the flux error map.
+\label{fig:A_survey}}
+\end{figure}
+
+\begin{figure}
+\plotone{B_survey.pdf}
+\figcaption{A summary of Survey Strategy B and the resultant self-calibration results: (a) shows the survey footprint diagram (as shown in Figure \ref{fig:simple_surveys}), (b) shows the source coverage for this survey, with (c) showing a histogram of the number of source observations and (d) shows the flux errors for each of the sources after the self-calibration procedure.
+The imprint of the survey strategy can be clearly seen in the coverage map.
+With this survey strategy, the self-calibration procedure can well calibrate the instrument response at the edge of the focal plane, but not in the center.
+As a result, the residuals in the flux error map are higher at the center of each pointing, where the sources fall on a less well contained part of the focal plane.
+\label{fig:B_survey}}
+\end{figure}
+
+\begin{figure}
+\plotone{C_survey.pdf}
+\figcaption{A summary of Survey Strategy C and the resultant self-calibration results: (a) shows the survey footprint diagram (as shown in Figure \ref{fig:simple_surveys}), (b) shows the source coverage for this survey, with (c) showing a histogram of the number of source observations and (d) shows the flux errors for each of the sources after the self-calibration procedure.
+\label{fig:C_survey}}
+\end{figure}
+
+\begin{figure}
+\plotone{D_survey.pdf}
+\figcaption{A summary of Survey Strategy D and the resultant self-calibration results: (a) shows the survey footprint diagram (as shown in Figure \ref{fig:simple_surveys}), (b) shows the source coverage for this survey, with (c) showing a histogram of the number of source observations and (d) shows the flux errors for each of the sources after the self-calibration procedure.
+As can be seen from (c), this quasi-random survey offers a more uniform coverage of the field than the other survey investigated.
+\label{fig:D_survey}}
 \end{figure}
@@
 \startdata
-A*  & 8192 & 2.964 & 3.101 & 3.090 \\
-B  & 4160 & 0.299 & 0.379 & 0.232 \\
-C  & 26 & 0.199 & 0.305 & 0.043 \\
-D  & 9 & 0.113 & 0.304 & 0.032 \\
+A*  & - & 1.092 & 1.490 & 1.450 \\
+B  & 3746 & 0.129 & 0.163 & 0.082 \\
+C  & 163 & 0.113 & 0.106 & 0.013 \\
+D  & 197 & 0.147 & 0.108 & 0.005 \\
 \enddata

### 2012-07-27 src/code_examples/paper_plots/additional_paper_plot.py

@@ -1,0 +1,86 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+# Rory Holmes
+# Feb 2012
+
+# This script contains the 2x2 survey comparison figure from the paper. All
+# other figures are made with the standard plot.py script
+
+# Make Python 3 compatible
+from __future__ import division, print_function
+
+import matplotlib.pyplot as plt
+import glob
+import pickle
+import numpy as np
+import sys
+import os
+
+sys.path.append('./../..')  # add simulator modules to Python path
+import transformations
+
+old_figures = glob.glob('*.png')
+old_figures += glob.glob('*.pdf')
+for path in old_figures:
+    os.remove(path)
+
+dirs = ['A', 'B', 'C', 'D']
+plot_suffix = ".pdf"
+verbose = True
+
+plt.rcParams['font.family'] = 'Computer Modern'
+plt.rcParams['text.usetex'] = True
+
+fig = plt.figure(figsize=(7, 7))
+
+middle = [0.525, 0.525]
+size = [0.45, 0.45]
+ax = [
+    fig.add_axes([middle[0] - size[0], middle[1], size[0], size[1]]),
+    fig.add_axes([middle[0], middle[1], size[0], size[1]]),
+    fig.add_axes([middle[0] - size[0], middle[1] - size[1], size[0], size[1]]),
+    fig.add_axes([middle[0], middle[1] - size[1], size[0], size[1]])
+    ]
+
+ax[0].set_xticklabels([''])
+ax[1].set_yticklabels([''])
+ax[3].set_yticklabels([''])
+
+fig.text(0.5, 0.025, r'Sky Position $\alpha$ (deg$^2$)',
+                                    va='center', ha='center')
+fig.text(0.025, 0.5, r'Sky Position $\beta$ (deg$^2$)',
+                                    va='center', ha='center', rotation=90)
+
+for indx in range(len(dirs)):
+
+    params = pickle.load(open('{0}/parameters.p'.format(dirs[indx])))
+    FoV = params['FoV']
+    sky_limits = params['sky_limits']
+
+    x = 0.5 * np.array([-FoV[0], -FoV[0], FoV[0], FoV[0], -FoV[0]])
+    y = 0.5 * np.array([-FoV[1], FoV[1], FoV[1], -FoV[1], -FoV[1]])
+
+    survey = np.loadtxt((dirs[indx] + '/survey.txt'))
+
+    for image in survey:
[…]

### 2012-07-27 src/code_examples/paper_plots/paper_plots.py

@@ -1,139 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-# Rory Holmes
-# Feb 2012
-
-# This file contains the functions to create the figures from the
-# self-calibration simulations.
-
-# Make Python 3 compatible
-from __future__ import division, print_function
-
-import matplotlib.pyplot as plt
-import glob
-import pickle
-import numpy as np
-import os
-import sys
-
-sys.path.append('./../..')  # add simulator modules to Python path
-import transformations
-from plot import camera_image, flat_fields
-
-GOLDEN_RATIO = 1.61803399
-
-def plot_grid(xlabel, ylabel):
-    fig = plt.figure(figsize=(figure_width, figure_width))
-    fig.clf()
-
-    ax1 = fig.add_axes([0.075, 0.5, 0.425, 0.425])
-    ax2 = fig.add_axes([0.5, 0.5, 0.425, 0.425])    
-    ax3 = fig.add_axes([0.075, 0.075, 0.425, 0.425])
-    ax4 = fig.add_axes([0.5, 0.075, 0.425, 0.425])
-    
-    ax1.set_xticklabels([''])
-    ax2.set_xticklabels([''])
-    ax2.set_yticklabels([''])
-    ax4.set_yticklabels([''])
-    
-    fig.text(0.5, 0.025, xlabel, va='center', ha='center')
-    fig.text(0.025, 0.5, ylabel, va='center', ha='center', rotation=90)
-    
-    return [fig, [ax1, ax2, ax3, ax4]]
-
-def paper_surveys(dirs):
-    xlabel = r'Sky Position $\alpha$ (deg$^2$)'
-    ylabel = r'Sky Position $\beta$ (deg$^2$)'
-    fig, axes = plot_grid(xlabel, ylabel)
-    for indx in range(len(dirs)):
-        params = pickle.load(open('{0}/parameters.p'.format(dirs[indx])))
-        single_survey(fig, axes[indx], (dirs[indx] + '/survey.txt'), 
-                            params['FoV'], params['sky_limits'],
-                            fig_width=8.3, suffix='.png', verbose=False)
-       
-    fig.savefig('simple_surveys.pdf')
-    fig.clf() 
-
-def single_survey(fig, ax, survey_file, FoV, sky_limits, fig_width=8.3,
-                                                suffix='.png', verbose=False):
-
-    if verbose:
[…]

### 2012-07-27 src/code_examples/paper_plots/parameters.py
forgot to include the new paper paramter file, here it is

@@ -64,8 +64,8 @@
     'seed'                :     [1.],
     # Float: The stop condition for the self-calibration procedure and the
     # best-in-basis fitting (stop when difference is less than 2 times this)
-    'stop_condition'	    :   1e-12,
+    'stop_condition'	    :   1e-14,
     # Integer: The maximum number of iterations in the self-calibration
     # procedure and the best-in-basis fitting
-    'max_iterations'	    :     16385
+    'max_iterations'	    :    10001
     }

### 2012-08-02 paper/self_calibration_I.tex
minor paper edits

@@
 \label{sec:instrument_response_model}
-We construct a complex, position independent instrument response model $f_\true(\vec{x_i} | \vec{q}_{\true})$ from a superposition of large and small-scale variations:
+We construct a complex, position independent instrument response model $f_\true(\vec{x_i} | \vec{q}_{\true})$ from a superposition of large- and small-scale variations:
 \begin{eqnarray*}
@@
 \tablewidth{0pt}
-\tablecaption{A summary of the tuneable parameters in these simulations and their fiducial values.\label{tab:parameters}}
+\tablecaption{A summary of the tunable parameters in the simulations and their fiducial values.\label{tab:parameters}}
 \tablehead{
@@
 The self-calibration procedure converged to the final solution after 163 iterations.
-The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.\label{fig:C_result}}
+The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
+With this regular observing strategies, all regions of the focal plane well interconnected with each other.\label{fig:C_result}}
 \end{figure}

### 2012-08-13 src/code_examples/paper_plots/additional_paper_plot.py
update RMS calculation so that now only sources within predefined survey used. Also add survey generation files

@@ -65,7 +65,7 @@ for indx in range(len(dirs)):
 
     for image in survey:
         alpha, beta = transformations.fp2sky(x, y, image[1:3], image[3])
-        ax[indx].plot(alpha, beta, 'k-', alpha=0.2)
+        ax[indx].plot(alpha, beta, 'k-', alpha=0.4)
 
     min_sky = np.min(sky_limits) \
                         - 0.02 * (np.max(sky_limits) - np.min(sky_limits))

### 2012-08-13 src/code_examples/paper_plots/generate_surveys.py

@@ -1,0 +1,110 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+# Rory Holmes / David Hogg
+# August 2012
+
+# This script is used to generate the four survey strategies investigated in
+# the self-calibration paper.
+
+from __future__ import division, print_function
+
+import numpy as np
+
+
+def uniform_center_list(sky_limits, FoV, nx, ny):
+    x_step_size = (sky_limits[1] - sky_limits[0] - FoV[0]) / (nx - 1.)
+    y_step_size = (sky_limits[3] - sky_limits[2] - FoV[1]) / (ny - 1.)
+
+    x_center_list = sky_limits[0] + 0.5 * FoV[0] + x_step_size * np.arange(nx)
+    y_center_list = sky_limits[2] + 0.5 * FoV[1] + y_step_size * np.arange(ny)
+    return x_center_list, y_center_list
+
+
+def generate_uniform_survey(sky_limits, FoV, Ncovering,
+                                                rotate=False, offset=False):
+
+    theta = 0.
+
+    nx = np.ceil((sky_limits[1] - sky_limits[0]) / (0.975 * FoV[0]))\
+                                                                .astype(int)
+    ny = np.ceil((sky_limits[3] - sky_limits[2]) / (0.975 * FoV[1]))\
+                                                                .astype(int)
+
+    x_center_list, y_center_list = uniform_center_list(sky_limits, FoV, nx, ny)
+    if offset:
+        x_center_list_1, y_center_list_1 =\
+                    uniform_center_list(sky_limits, FoV, nx + 1, ny - 1)
+        x_center_list_2, y_center_list_2 =\
+                    uniform_center_list(sky_limits, FoV, nx - 1, ny + 1)
+
+    x = np.zeros((Ncovering * nx * ny, 4))
+    indx = 0
+    for covering in range(Ncovering):
+        xcl = x_center_list
+        ycl = y_center_list
+        tnx = nx
+        tny = ny
+        if offset:
+            if covering % 3 == 1:
+                xcl = x_center_list_1
+                ycl = y_center_list_1
+                tnx = nx + 1
+                tny = ny - 1
+            if covering % 3 == 2:
+                xcl = x_center_list_2
+                ycl = y_center_list_2
[…]

### 2012-08-13 src/code_examples/paper_plots/paper_script.py

@@ -16,6 +16,7 @@ from __future__ import division, print_function
 import os
 import sys
 import numpy as np
+import pickle
 
 sys.path.append('./../..')  # add simulator modules to Python path
 
@@ -26,10 +27,10 @@ import analysis
 # Multiprocessing Flag:
 # False - then does not use multiprocessing
 # int - uses that many separate processes
-multi_proc = False
+multi_proc = 4
 
 # The four survey directories
-survey_files = ['D'] #['A', 'B', 'C', 'D']
+survey_files = ['A', 'B', 'C', 'D']
 
 # Load the default parameters
 dic = eval(open('parameters.py').read())
@@ -60,3 +61,48 @@ else:
     results = []
     for params in parameter_dictionaries:
         results.append(simulation.run_sim(params))
+
+#  Since there are lots of souces at the edge of Survey D that are only
+#  measured a few times, the RMS 
+f = open('rms_results.txt', 'w')
+survey_area = [-4., 4., -4., 4.]
+header = 'Source RMS Error with {0} survey area:'.format(survey_area)
+f.write(header + '\n')
+print(header)
+
+for path in survey_files:
+	
+	# Load the catalogs from the simulation run
+	fitted_cat = pickle.load(open((path + '/fitted_catalog.p'), mode='rb'))
+	true_cat = pickle.load(open((path + '/source_catalog.p'), mode='rb'))
+	
+	if true_cat.k.max() > fitted_cat.k.max():
+		k_max = true_cat.k.max()
+	else:
+		k_max = fitted_cat.k.max()
+	
+	error = np.array([])
+	norm_error = np.array([])
+	
+	true_flux = np.array([])
+	fitted_flux = np.array([])
+	
+	inside = ((true_cat.alpha > survey_area[0])
+	                * (true_cat.alpha < survey_area[1])
+	                * (true_cat.beta > survey_area[2])
+	                * (true_cat.beta < survey_area[3]))
+
+	for indx in range(k_max):
+		indx_t = np.where(true_cat.k[inside] == indx)[0]
+		indx_f = np.where(fitted_cat.k == indx)[0]
+	
+		if (len(indx_t == 1) and len(indx_f == 1)):
+			true_flux = np.append(true_flux, true_cat.flux[inside][indx_t])
+			fitted_flux = np.append(fitted_flux, fitted_cat.flux[indx_f])
+	
[…]

### 2012-08-13 src/code_examples/paper_plots/parameters.py

@@ -67,5 +67,5 @@
     'stop_condition'	    :   1e-14,
     # Integer: The maximum number of iterations in the self-calibration
     # procedure and the best-in-basis fitting
-    'max_iterations'	    :    10001
+    'max_iterations'	    :    3#10001
     }

### 2012-08-13 src/default_parameters.py

@@ -64,8 +64,8 @@
     'seed'                :     [1.],
     # Float: The stop condition for the self-calibration procedure and the
     # best-in-basis fitting (stop when difference is less than 2 times this)
-    'stop_condition'	    :     1e-5,
+    'stop_condition'	  :     1e-5,
     # Integer: The maximum number of iterations in the self-calibration
     # procedure and the best-in-basis fitting
-    'max_iterations'	    :     1049
+    'max_iterations'	  :     1049
     }

### 2012-08-13 src/plot.py

@@ -272,7 +272,6 @@ def survey_coverage(ax, source_filename, measurement_filename,
 
     ax.xaxis.set_zorder(3000)
     ax.yaxis.set_zorder(3000)
-    print(ax.get_clip_on())
 
     return [colors, colorbar_nobs]
 

### 2012-08-13 src/self_calibration.py

@@ -19,7 +19,7 @@ import analysis
 
 def self_calibration(obs_cat, sky_catalog, order, FoV,
                         ff_samples, stop_condition, max_iterations,
-                        best_fit_parameters, data_dir=False, verbose=False):
+                        best_fit_parameters,data_dir=False, verbose=False):
     ''' This functions performs the self-calibration procedure on the survey.
 
     Parameters
@@ -76,7 +76,9 @@ def self_calibration(obs_cat, sky_catalog, order, FoV,
         q, q_invvar, chi2 = q_step(obs_cat, s, order, FoV, ff_samples)
         old_chi2 = temp_chi2
 
+        
         indx = [s != 0]
+
         rms = analysis.rms_error(s[indx], sky_catalog.flux[indx], verbose)
         bdness = analysis.badness(q, FoV, ff_samples, verbose)
         bdness_bestfitff = analysis.best_in_basis(q, FoV, ff_samples,
@@ -90,8 +92,9 @@ def self_calibration(obs_cat, sky_catalog, order, FoV,
                     obs_cat.size))
 
         if data_dir:
-            if (count == next_plot_iteration) or \
-                                (abs(chi2 - old_chi2) < stop_condition):
+            if ((count == next_plot_iteration)
+                        or (abs(chi2 - old_chi2) < stop_condition)
+                        or (count == max_iterations - 1)):
                 save_out.fitted_flat_field(q, FoV, ff_samples, count,
                                                             data_dir, verbose)
                 next_plot_iteration *= 2

### 2012-08-13 src/code_examples/paper_plots/parameters.py
update number of iterations in parameter file

@@ -67,5 +67,5 @@
     'stop_condition'	    :   1e-14,
     # Integer: The maximum number of iterations in the self-calibration
     # procedure and the best-in-basis fitting
-    'max_iterations'	    :    3#10001
+    'max_iterations'	    :    10001
     }

### 2012-08-20 whitepaper/pixel_level.tex
start

\documentclass[12pt]{article}
\begin{document}
Hello world.
\end{document}

### 2012-08-20 whitepaper/pixel_level.tex
some text

@@
-\documentclass[12pt]{article}
+\documentclass[12pt,preprint,dvips]{aastex}
+
+\newcommand{\foreign}[1]{\textsl{#1}}
+\newcommand{\etal}{\foreign{et~al.}}
+
 \begin{document}
-Hello world.
+
+\paragraph{Can we infer the pixel-level flat-field without an internal calibration source?}
+
+We have shown with large surveys that the redundant information coming
+from the multiple observations of stars is sufficient to infer the
+large-scale sensitivity (effective area or flat-field or vignetting)
+of an imaging device (\citealt{ubercal}, \citealt{holmes}).  In these
+experiments, the photometry has been thought of as being well sampled,
+and the flat-field function has had only thousands of parameters. We
+have not attempted to obtain millions of pixel-level parameters from
+the data and we have not considered arbitrarily non-smooth functions
+of two-dimensional position.
+
+With observations of billions of sources on the sky, however, it might
+be possible to infer millions of sensitivity parameters, even as many
+or more sensitivity parameters than there are pixels in the imaging
+device.  After all, we know a lot about the sky, many sources are
+observed multiple times, and every source touches multiple pixels even
+in a single visit.
+
+\begin{thebibliography}{70}
+\bibitem[Padmanabhan \etal(2008)]{ubercal}
+Padmanabhan,~N., Schlegel,~D.~J., Finkbeiner,~D.~P., \etal, 2008, \apj, 674, 1217
+\bibitem[Holmes \etal(2012)]{2012arXiv1203.6255H}
+Holmes,~R., Hogg,~D.~W., \& Rix,~H.-W., 2012, arXiv:1203.6255 
+\end{thebibliography}
+
 \end{document}

### 2012-08-20 whitepaper/pixel_level.tex
holmes ref, number of parameters

--- 2012-08-20 whitepaper/pixel_level.tex
@@
-\documentclass[12pt]{article}
+\documentclass[12pt,preprint,dvips]{aastex}
+
+\newcommand{\foreign}[1]{\textsl{#1}}
+\newcommand{\etal}{\foreign{et~al.}}
+\newcommand{\project}[1]{\textsl{#1}}
+\newcommand{\SDSS}{\project{SDSS}}
+
 \begin{document}
-Hello world.
+
+\section*{Can we infer the pixel-level flat-field without an internal calibration source?}
+
+\noindent
+David W. Hogg (NYU)
+\\
+\textsl{Max-Planck-Institut f\"ur Astronomie, Heidelberg, Germany}
+
+We have shown with large optical imaging surveys that the redundant
+information coming from the multiple observations of stars is
+sufficient to infer the large-scale sensitivity (effective area or
+flat-field or vignetting) of an astronomical imaging device
+(\citealt{ubercal}, \citealt{holmes}).  In these self-calibration
+experiments, the sensitivity function has had only hundreds or
+thousands of parameters.  We have not attempted to obtain millions of
+pixel-level parameters from the data and we have not considered
+arbitrarily non-smooth two-dimensional sensitivity functions.
+
+With observations of billions of sources on the sky, however, it might
+be possible to infer millions of sensitivity parameters, even as many
+or more sensitivity parameters than there are pixels in the imaging
+device.  After all, we know a lot about the sky, many sources are
+observed multiple times, and every source touches multiple pixels even
+in a single visit.
+
+In detail, the sensitivity of a pixel can be a function of spectral
+energy distribution (different pixels will be differently sensitive to
+sources with different colors or spectra) and a function of
+intra-pixel illumination pattern (the pixel might have a more
+sensitive region and some less sensitive regions).  Also, diffuse
+illumination, such as uniform sky background, might illuminate the
+pixel from a different mixture of angles than compact illumination,
+such as from a star, so there might be, effectively, different pixel
+sensitivities for the sky and for sources.  Indeed, reflections in the
+\SDSS\ camera led to problems of this kind, which invalidated the
+sky-based flat-fields (\citealt{ubercal}).  For all these reasons we
+might actually have many sensitivity parameters per pixel; this makes
+our task harder quantitatvely, but perhaps not qualitatively.
+
+\begin{thebibliography}{70}
+\bibitem[Padmanabhan \etal(2008)]{ubercal}
+Padmanabhan,~N., Schlegel,~D.~J., Finkbeiner,~D.~P., \etal, 2008, \apj, 674, 1217
+\bibitem[Holmes \etal(2012)]{holmes}
+Holmes,~R., Hogg,~D.~W., \& Rix,~H.-W., 2012, arXiv:1203.6255 
+\end{thebibliography}
+
 \end{document}

### 2012-08-20 whitepaper/pixel_level.tex
pixel signal models

@@
 \newcommand{\SDSS}{\project{SDSS}}
+
+\newcommand{\expectation}[1]{\tilde{#1}}
+\newcommand{\sky}{^{(\mathrm{sky})}}
 
@@
 flat-field or vignetting) of an astronomical imaging device
-(\citealt{ubercal}, \citealt{holmes}).  In these self-calibration
-experiments, the sensitivity function has had only hundreds or
-thousands of parameters.  We have not attempted to obtain millions of
-pixel-level parameters from the data and we have not considered
-arbitrarily non-smooth two-dimensional sensitivity functions.
+(\citealt{ubercal}, \citealt{schlafly}, \citealt{holmes}).  In these
+self-calibration experiments, the sensitivity function has had only
+hundreds or thousands of parameters.  We have not attempted to obtain
+millions of pixel-level parameters from the data and we have not
+considered arbitrarily non-smooth two-dimensional sensitivity
+functions.
 
@@
 
+\paragraph{varying levels of dumbness:}
+There are various thought experiments or considerations that suggest
+that a pixel-level self-calibration is possible.  One observation is
+that if a pixel is unusually low in sensitivity, it's mean output
+signal will be low on average for the whole data set.  This statement
+could be made more precise for the sky flat by considering only
+observations in which the pixel is not near any detectable source.  It
+could be made even more precise if each observation is divided by the
+mean pixel value, to account for variations in sky brightness.  These
+kinds of operations or considerations could lead to a very good sky
+flat for any device, and is not unlike what is usually done.
+
+As we said, however, each pixel might have a different response to
+compact sources than it does to the diffuse sky, so slightly less-dumb
+methods might be called for.  For example, consider all overlapping
+$3\times 3$-pixel patches in a full data set.  Most of these patches
+would be far from any detectable source and could be labeled ``sky''.
+A fraction would be near or on top of compact, detected sources and
+could be labeled ``source''.  This labeling could be
+non-comprehensive; there could be intermediate patches labeled
+``neither''; this is just an engineering consideration.  Less dumb
+still would be to make bigger patches (say $5\times 5$ or larger) or
+to make use of the residuals away from a physically motivated model of
+the observed images.
+
+\paragraph{pixel signal models}
+Sticking with the intermediate level of dumbness for now: In a set of
+$3\times 3$ sky patches, the expected signal
+$\expectation{D}_{ij}$---the expected number of electrons or data
+numbers read out---in pixel $i$ of exposure $j$ might be a combination
+of zero-level, dark-current, and sky-photon signals
+\begin{equation}
+\expectation{D}_{ij} = z_i + d_i\,t_j + f\sky_i\,S\sky_j\,t_j
+\end{equation}
+where $i$ labels the device pixel ($i$ maps back to a pixel location
+in the original imaging device), $j$ labels the exposure, each of
+which might have a slightly different sky level, $z_i$ is a zero-level
+or bias for pixel $i$, $d_i$ is a dark current rate or signal per time
+in the absence of illumination, $t_j$ is the exposure time for
+exposure $j$, $f\sky_i$ is the sensitivity of pixel $i$ to diffuse sky
+illumination, and $S\sky_j$ is the intensity of the sky in exposure
+$j$ (interpolated to pixel $i$ if there are gradients).
+
+In a $3\times 3$ source patch, the expectation is similar...
+\begin{equation}
+\expectation{D}_{ij} = z_i + d_i\,t_j + f\sky_i\,S\sky_j\,t_j
+\end{equation}
+
 \begin{thebibliography}{70}
+\bibitem[Holmes \etal(2012)]{holmes}
+Holmes,~R., Hogg,~D.~W., \& Rix,~H.-W., 2012, arXiv:1203.6255 
 \bibitem[Padmanabhan \etal(2008)]{ubercal}
 Padmanabhan,~N., Schlegel,~D.~J., Finkbeiner,~D.~P., \etal, 2008, \apj, 674, 1217
-\bibitem[Holmes \etal(2012)]{holmes}
-Holmes,~R., Hogg,~D.~W., \& Rix,~H.-W., 2012, arXiv:1203.6255 
+\bibitem[Schlafly \etal(2012)]{schlafly}
+Schlafly,~E.~F., Finkbeiner,~D.~P., Juric,~M., \etal, 2012, arXiv:1201.2208 
 \end{thebibliography}

### 2012-08-20 whitepaper/pixel_level.tex
noise models, source sensitivity, blind decon, sampling

@@
 \newcommand{\sky}{^{(\mathrm{sky})}}
+\newcommand{\source}{^{(\mathrm{src})}}
 
@@
 David W. Hogg (NYU)
+\\{\footnotesize
+\textsl{Center for Cosmology and Particle Physics, Department of Physics, New York University, New York, USA}
 \\
-\textsl{Max-Planck-Institut f\"ur Astronomie, Heidelberg, Germany}
+\textsl{Max-Planck-Institut f\"ur Astronomie, Heidelberg, Germany}}
+
+\textsl{This document is a draft dated 2012-08-20.  It is based in
+  part on conversations with Rob Fergus (NYU), Rory Holmes (MPIA), and
+  Bernhard Sch\"olkopf (MPI-IS).}
 
@@
 
-\paragraph{pixel signal models}
-Sticking with the intermediate level of dumbness for now: In a set of
-$3\times 3$ sky patches, the expected signal
-$\expectation{D}_{ij}$---the expected number of electrons or data
-numbers read out---in pixel $i$ of exposure $j$ might be a combination
-of zero-level, dark-current, and sky-photon signals
-\begin{equation}
+\paragraph{pixel signal models:}
+Sticking with the intermediate level of dumbness for now: In a
+$3\times 3$ sky patch (a patch far from any detectable sources), the
+expected signal $\expectation{D}_{ij}$---the expected number of
+electrons or data numbers read out---in pixel $i$ of exposure $j$
+might be a combination of zero-level, dark-current, and sky-photon
+signals
+\begin{eqnarray}
 \expectation{D}_{ij} = z_i + d_i\,t_j + f\sky_i\,S\sky_j\,t_j
-\end{equation}
+\quad ,
+\end{eqnarray}
 where $i$ labels the device pixel ($i$ maps back to a pixel location
@@
 
-In a $3\times 3$ source patch, the expectation is similar...
-\begin{equation}
+In a $3\times 3$ source patch (a patch containing some detectable
+source or part thereof), the expectation is similar but there is now
+source flux
+\begin{eqnarray}
 \expectation{D}_{ij} = z_i + d_i\,t_j + f\sky_i\,S\sky_j\,t_j
-\end{equation}
+                     + f\source_i\,S\source_{ij}\,t_j
+\quad ,
+\end{eqnarray}
+where now $f\source_i$ is the sensitivity of pixel $i$ to compact
+source flux, and $S\source_{ij}$ is the source flux in exposure $j$
+interpolated to or evaluated at the location of pixel $i$.  This
+interpolation to pixel $i$ is where the $3\times 3$ pixel patch comes
+in: We are going to use the 8 pixels surrounding pixel $i$ to
+interpolate the source flux probabilistically to pixel $i$.  This also
+shows that using a bigger patch, say $5\times 5$, would be
+substantially less dumb.  Even the sky patches (as noted momentarily
+above) might require interpolation to pixel $i$, since there might be
+sky gradients.
+
+In the case that there are also \emph{dark}
+observations---observations with the shutter closed---or \emph{zero}
+observations---observations with zero exposure times---there can be
+additional constraints on the $z_i$ and the $d_i$ beyond just the
+science-data constraints.  These considerations also suggest that a
+survey that varies its exposure time will thereby obtain much more
+information about the calibration information (and also vet the
+shutter!).
+
+There might be an additional dependence of the source sensitivity
+$f\source_i$ on color or spectral energy distribution.  There might
+also be variations of all the calibration parameters on time, possibly
+even with break-points or sharp transitions (think episodic radiation
+damage and the like).  There will also be many, many pixels affected
+by cosmic rays.  All of these issues make the real-world execution of
+this program more complicated than the cartoon given here.  For now we
+might try assuming that we can work with only patches that are
+obviously \emph{not} affected by cosmic rays, and work only with
+time-slices of the data within which no significant variations are
+expected.
+
+The likelihood of any model for the data will be related to the idea
+that the observed or read-out data $D_{ij}$ differs from the model
+expectation $\expectation{D}_{ij}$ by the contribution of some kind of
+noise, which might be Poisson or Gaussian or some combination,
+depending on the device and our precision needs.
+
+\paragraph{blind deconvolution:}
+This problem looks very much like a blind deconvolution problem,
+because there is a true but unknown scene multiplied by some unknown
+set of pixel responses before noisy detection.  If the astronomical
+fields $(S\sky_j, S\source_{ij})$ are known, the pixel response
+parameters $(z_i, d_i, f\sky_i, f\source_i)$ are related to the signal
+by a pure linear operation and \foreign{vice versa}.  If the noise in
+the data is Gaussian (or can be approximated as such), the inference
+can proceed by iterated weighted linear least squares.  If the imager
+is well-behaved, $S\sky_j$ will be a very smooth function of pixel
+position $i$ or no function at all.  If the imaging is well-sampled,
+$S\source_{ij}$ will be expressable as a linear combination of
+band-limited basis functions, like sincs or Gaussians.
+
+If we aren't satisfied with point estimates---and we shouldn't
+be---then we can propagate the uncertainties in both the calibration
+quantities and the astronomical scene as a set of samples or explicit
+analytic descriptions of the posterior PDF.  The latter might be
+possible in the case of pure Gaussian noise.
 

### 2012-08-24 paper/self_calibration_I.tex
correct survey strategy bug, text tweeks, plotting tweaks

@@
 from wide-field imaging surveys.
-This self-calibration technique utilizes multiple measurements of sources to constrain instruments' large-scale response (flat-field) from survey science data alone.
+This self-calibration technique utilizes multiple measurements of sources at different focal-plane positions to constrain instruments' large-scale response (flat-field) from survey science data alone.
 We create an artificial sky of sources
@@
 self-calibrated. We show that returning the same sources to very
-different focal plane positions is the key property of any survey
+different focal-plane positions is the key property of any survey
 strategy designed for accurate retrospective calibration of this type. The results
@@
 design of large-scale imaging surveys: Do not use a regular, repeated tiling of
-the sky; instead return the same sources to very different focal plane
+the sky; instead return the same sources to very different focal-plane
 positions.
@@
 But typically far more photons, or readout electrons, are collected during science exposures themselves; are these not incredibly constraining on the calibration?
-Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS), much more calibration information was obtained from the science data than the calibration data \citep{pad08}.
+Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey imaging data (SDSS), much more calibration information was obtained from the science data than the calibration data \citep{pad08}.
 But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: good redundancy was required in the data stream, and a redundancy of a very specific kind.
+Self-calibration has been used to obtain the most precise photometric calibration of the PanSTARRS imaging data \citep{sch12}.
 
@@
 The self-calibration technique utilizes the multiple measurements of sources in the survey arising from overlapping pointings.
-If these redundant measurements are recorded at different focal plane positions or at different times within the survey, it is possible to constrain the relative instrument response by requiring that the post-calibration measurements yield consistent flux values.
+If these redundant measurements are recorded at different focal-plane positions or at different times within the survey, it is possible to constrain the relative instrument response by requiring that the post-calibration measurements yield consistent flux values.
 Through end-to-end, catalog-level survey simulations, we aim to identify the important properties of survey strategies that makes them advantageous for this kind of self-calibration.
@@
 \label{sec:single_exposure}
-With a telescope pointing $(\alpha, \beta)$ and camera orientation $\theta$, we find the focal plane positions of the sources on the synthetic sky that fall within the instrument's field-of-view.
-In our simulations, we use a large instrument field-of-view of $0.76~ \text{deg} \times 0.72~\text{deg}$; a size consistent with current large survey imagers.
+With a telescope pointing $(\alpha, \beta)$ and camera orientation $\theta$, we find the focal-plane positions of the sources on the synthetic sky that fall within the instrument's field-of-view.
+In our simulations, we use a large instrument field-of-view of $0.75~ \text{deg} \times 0.75~\text{deg}$; a size consistent with current large survey imagers.
 An example of a single pointing exposure is shown in Figure \ref{fig:single_image}.
@@
 To generate photometric catalogs, the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model.
-For a measurement $i$, the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$---which is a function of focal plane position $\vec{x_i}$---and the source's true flux $s_{k,\true}$ ,
+For a measurement $i$, the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$---which is a function of focal-plane position $\vec{x_i}$---and the source's true flux $s_{k,\true}$ ,
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $\vec{x_i} = (x_i, y_i)$ is the focal plane position that the $k$th source falls at during the $i\text{th}$ measurement and $\vec{q}_\true$ are the parameters defining the instrument response model.
+where $\vec{x_i} = (x_i, y_i)$ is the focal-plane position that the $k$th source falls at during the $i\text{th}$ measurement and $\vec{q}_\true$ are the parameters defining the instrument response model.
 This response model subsumes various effects commonly discussed in terms of detector flat-fielding and optical vignetting.
@@
 \end{eqnarray*}
-with the physical focal plane dimensions $X$ and $Y$.
+with the physical focal-plane dimensions $X$ and $Y$.
 The $(1 + a +d)^{-0.75}$ factor reduces the magnitude of the higher order terms.
@@
 \end{eqnarray*}
-where $c_i$ is the recorded count rate, $f(\vec{x_i} | \vec{q})$ is the fitted instrument response model at a focal plane position $\vec{x}_i$, $\vec{q}$ is a vector parameterizing the eighth order polynomial instrument response model, $s_k$ is the model source flux estimate and the error $e_i$ is drawn from the Normal Distribution $N(e|0,{\sigma_i}^2)$, such that
+where $c_i$ is the recorded count rate, $f(\vec{x_i} | \vec{q})$ is the fitted instrument response model at a focal-plane position $\vec{x}_i$, $\vec{q}$ is a vector parameterizing the eighth order polynomial instrument response model, $s_k$ is the model source flux estimate and the error $e_i$ is drawn from the Normal Distribution $N(e|0,{\sigma_i}^2)$, such that
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $c_i$ are the measured count rates, $f(\vec{x_i} | \vec{q})$ is the fitted instrument response model at a focal plane position $\vec{x}_i$ and $\sigma_i$ is the assumed noise model.
+where $c_i$ are the measured count rates, $f(\vec{x_i} | \vec{q})$ is the fitted instrument response model at a focal-plane position $\vec{x}_i$ and $\sigma_i$ is the assumed noise model.
 A new estimate of the model source flux $s'_{k}$ is then found by minimizing the error function with respect to the old model source flux $s_{k}$:
@@
 This, along with the success of the procedure with the Surveys C and D catalogs, gives an insight into the key property of a survey that makes it good for self-calibration.
-Namely, the survey should return the same sources to many, very different focal plane positions.
+Namely, the survey should return the same sources to many, very different focal-plane positions.
 This allows the self-calibration procedure to link the different parts of the focal plane with each other through the observations of the same sources.
@@
 In the center of the focal plane, on which the same sources always fall, there are insufficient redundant measurements to allow for a good fit.
-Returning the same sources to very different focal plane positions is critical for the self-calibration to work effectively.
+Returning the same sources to very different focal-plane positions is critical for the self-calibration to work effectively.
 
@@
 The second is that there must be the correct redundancy built into the survey strategy.
-The same sources must be imaged at very different focal plane positions; regular tiling of the sky is useless for this kind of calibration.
+The same sources must be imaged at very different focal-plane positions; regular tiling of the sky is useless for this kind of calibration.
 
@@
 The self-calibration procedure discussed in this paper can be trivially modified to fit for a time evolving relative instrument response model.
-For this to be effective though, it will be necessary to not only observe the same sources at very different focal plane positions, but also to observe them at very different times within the survey.
+For this to be effective though, it will be necessary to not only observe the same sources at very different focal-plane positions, but also to observe them at very different times within the survey.
 
@@
 \bibitem[Padmanabhan et al.(2008)]{pad08}{Padmanabhan, N. et al., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj~674, p1217.}
+\bibitem[Schlafly et al.(2012)]{sch12} Schlafly, E.~F., 
+Finkbeiner, D.~P., Juric, M., et al., Photometric Calibration of the First 1.5 Years of the Pan-STARRS1 Survey, 2012, arXiv:1201.2208 
 \bibitem[Windhorst et al.(2011)]{win11}{Windhorst, R.~A.
@@
 \figcaption{A single exposure of the synthetic sky.
-Left: A plot of the bright sources within the synthetic sky used in the self-calibration procedure, with the focal plane footprint overlaid.
+Left: A plot of the bright sources within the synthetic sky used in the self-calibration procedure, with the focal-plane footprint overlaid.
 Right: The resultant distribution of the sources on the instrument's focal plane.
@@
 10$\upsigma$ Detection Limit (mag)& $m_\text{max} = 22$\\
-Field-of-View (deg$^{2}$)& $0.76 \times 0.72$ \\
+Field-of-View (deg$^{2}$)& $0.75 \times 0.75$ \\
 Noise Model -- Eqn. \ref{eqn:noise} & $\delta = 0.1585$, $\eta = 0.0017$, $\epsilon_\text{max} = 1.0$\\
@@
 \plotone{simple_surveys.pdf}
-\figcaption{Focal plane footprints projected onto the synthetic sky according to the four simple survey strategies described in Section \ref{sec:simple_surveys} and summarized in Table \ref{tab:simple_surveys}.
+\figcaption{Focal-plane footprints projected onto the synthetic sky according to the four simple survey strategies described in Section \ref{sec:simple_surveys} and summarized in Table \ref{tab:simple_surveys}.
 Surveys A, B and D have 1728 pointings and survey C has 1720 pointings.\label{fig:simple_surveys}}

### 2012-08-24 src/code_examples/paper_plots/generate_surveys.py

@@ -10,9 +10,13 @@
 from __future__ import division, print_function
 
 import numpy as np
+import parameters
 
 
 def uniform_center_list(sky_limits, FoV, nx, ny):
+    assert nx * FoV[0] > sky_limits[1] - sky_limits[0]
+    assert ny * FoV[1] > sky_limits[3] - sky_limits[2]
+    
     x_step_size = (sky_limits[1] - sky_limits[0] - FoV[0]) / (nx - 1.)
     y_step_size = (sky_limits[3] - sky_limits[2] - FoV[1]) / (ny - 1.)
 
@@ -25,11 +29,8 @@ def generate_uniform_survey(sky_limits, FoV, Ncovering,
                                                 rotate=False, offset=False):
 
     theta = 0.
-
-    nx = np.ceil((sky_limits[1] - sky_limits[0]) / (0.975 * FoV[0]))\
-                                                                .astype(int)
-    ny = np.ceil((sky_limits[3] - sky_limits[2]) / (0.975 * FoV[1]))\
-                                                                .astype(int)
+    nx = 12
+    ny = 12
 
     x_center_list, y_center_list = uniform_center_list(sky_limits, FoV, nx, ny)
     if offset:
@@ -93,9 +94,10 @@ def generate_random_survey(sky_limits, FoV, N):
 
 if __name__ == "__main__":
 
+    dic = eval(open('parameters.py').read())
     sky_limits = [-4., 4., -4., 4.]
     number_passes = 12
-    FoV = [0.76, 0.72]
+    FoV = dic['FoV']
 
     xA = generate_uniform_survey(sky_limits, FoV, number_passes)
     np.savetxt('A.txt', xA)

### 2012-08-24 src/code_examples/paper_plots/parameters.py

@@ -34,7 +34,7 @@
     # Instrument Parameters
     # =====================
     # Float Array: The simulate imager's field-of-view in degrees [alpha, beta]
-    'FoV'                 :     [0.76, 0.72],
+    'FoV'                 :     [0.75, 0.75],
     # Float: The saturation limit of the simulated imager
     'm_min'               :     17.,
     # Float: The 10-sigma detection limit of the simulated imager
@@ -67,5 +67,5 @@
     'stop_condition'	    :   1e-14,
     # Integer: The maximum number of iterations in the self-calibration
     # procedure and the best-in-basis fitting
-    'max_iterations'	    :    10001
+    'max_iterations'	    :    1
     }

### 2012-08-24 src/plot.py

@@ -307,7 +307,7 @@ def survey_coverage_colorbar(ax_cb, colors, colorbar_nobs):
     return None
 
 
-def survey_coverage_histogram(ax, measurement_filename, xlim=None, ylim=None):
+def survey_coverage_histogram(ax, measurement_filename, xlim, ylim=None):
     ''' This function plots a histogram of the number of source observations
 
     Input
@@ -324,22 +324,27 @@ def survey_coverage_histogram(ax, measurement_filename, xlim=None, ylim=None):
 
     measurement_catalog = pickle.load(open(measurement_filename))
 
+    print(measurement_catalog.k.size)
+
     source_ID = np.unique(measurement_catalog.k)
+    print(source_ID.size)
     nobs = np.zeros(source_ID.size)
 
     for i, sid in enumerate(source_ID):
         found = np.where(measurement_catalog.k == sid)[0]
         nobs[i] = found.size
 
-    hist = ax.hist(nobs, color='k', bins=np.arange(nobs.max()) + 0.5,
-                                                                histtype='bar')
+    print(np.sum(nobs))
+    print(np.mean(nobs))
+    assert np.sum(nobs).astype(int) == measurement_catalog.k.size
+    hist = ax.hist(nobs, color='k',
+                        bins=np.arange(int(xlim[1]) + 5) - 0.5,
+                        histtype='step')
 
-    if xlim is None:
-        ax.set_xlim(1, nobs.max() + 2)
-    else:
-        ax.set_xlim(xlim)
+    
+    ax.set_xlim(xlim)
     if ylim is None:
-        ax.set_ylim(1, 1.1 * np.max(hist[0]))
+        ax.set_ylim(-0.1 * np.max(hist[0]), 1.1 * np.max(hist[0]))
     else:
         ax.set_ylim(ylim)
 
@@ -620,8 +625,7 @@ def survey(source_filename, measurement_filename,
     if verbose:
         print('Plotting measurement histogram from {0}...'
                                                 .format(measurement_filename))
-    survey_coverage_histogram(ax3, measurement_filename,
-                                            xlim=(1, nobs_plot_lim))
+    survey_coverage_histogram(ax3, measurement_filename, (0, nobs_plot_lim))
     if verbose:
         print('...done...')
 
@@ -851,7 +855,7 @@ if __name__ == "__main__":
 
     verbose = True
     mult_proc = True
-    plot_suffix = ".pdf"
[…]

### 2012-08-24 paper/self_calibration_I.tex
add analysis region

@@
 In this paper, we have considered the design of survey strategies for calibrating the large-scale, relative photometric response of an instrument, but there are many more calibration parameters that can be constrained from this method.
-For example, the optical distortion of an instrument or the point-spread-function variation could also be constrained with such a method, although it would need to be explored whether the properties of a survey strategy that makes it good for one calibration, are also suitable for other calibration aspects; ultimately a trade-off would need to be performed between the different calibration concerns and the other goals of the survey, such as sky coverage, variability monitoring and so-on.
+For example, the optical distortion of an instrument or the point-spread-function variation could also be constrained with such a method, although it would need to be explored whether the properties of a survey strategy that makes it good for one calibration, are also suitable for other calibration aspects.
+
+HOGG write van Dokkum paragraph
+Ultimately a trade-off would need to be performed between the different calibration concerns and the other goals of the survey, such as sky coverage, variability monitoring and so-on.
 
 \acknowledgments
-It is a pleasure to thank Rob Fergus (NYU), Jason Kalirai (STScI),
-and Sam Roweis (deceased) for valuable conversations.
+It is a pleasure to thank Rob Fergus (NYU), Jason Kalirai (STScI), Sam Roweis (deceased), and Pieter van Dokkum (Yale) for valuable conversations.
 

### 2012-08-24 src/analysis.py

@@ -172,7 +172,7 @@ def best_in_basis(q, FoV, ff_samples, best_fit_params, verbose=False):
     return best_in_basis_badness
 
 
-def rms_error(flux, true_flux, verbose=False):
+def rms_error(flux, true_flux, weight, verbose=False):
     ''' Calculates the root-mean-square error between the estimated source
     fluxes and the true source fluxes. This only applies to the bright sources
     selected for the self-calibration procedure and not necessarily all the
@@ -184,6 +184,8 @@ def rms_error(flux, true_flux, verbose=False):
         Containing all the source flux estimates
     true_flux   :   numpy array
         Containing all the true fluxes for the sources
+    weight      :   numpy array
+        Weights to use
     verbose     :   Boolean
         Set to true to run simulations in verbose mode
 
@@ -196,7 +198,9 @@ def rms_error(flux, true_flux, verbose=False):
     if verbose:
         print("Calculating RMS Error of fitted sources...")
 
-    rms = 100 * np.sqrt(np.mean(((flux - true_flux) / true_flux) ** 2))
+    w = weight.astype(float)
+    ms = np.sum(w * ((flux - true_flux) / true_flux) ** 2) / np.sum(w)
+    rms = 100 * np.sqrt(ms)
 
     if verbose:
         print("...done!")

### 2012-08-24 src/code_examples/multiprocessing_example/parameters.py

@@ -30,6 +30,9 @@
     # Float Array: The area of sky to generate sources in
     # [alpha_min, alpha_max, beta_min, beta_max]
     'sky_limits'          :     [-4.8, 4.8, -4.8, 4.8],
+    # Float Array: The area of sky to use for analyses like rms
+    # [alpha_min, alpha_max, beta_min, beta_max]
+    'analysis_limits'          :     [-3.8, 3.8, -3.8, 3.8],
 
     # Instrument Parameters
     # =====================

### 2012-08-24 src/code_examples/paper_plots/paper_script.py

@@ -27,10 +27,10 @@ import analysis
 # Multiprocessing Flag:
 # False - then does not use multiprocessing
 # int - uses that many separate processes
-multi_proc = 4
+multi_proc = False
 
 # The four survey directories
-survey_files = ['A', 'B', 'C', 'D']
+survey_files = ['D']#['A', 'B', 'C', 'D']
 
 # Load the default parameters
 dic = eval(open('parameters.py').read())

### 2012-08-24 src/default_parameters.py

@@ -30,7 +30,7 @@
     # Float Array: The area of sky to generate sources in
     # [alpha_min, alpha_max, beta_min, beta_max]
     'sky_limits'          :     [-4.8, 4.8, -4.8, 4.8],
-    # Float Array: The area of sky to generate sources in
+    # Float Array: The area of sky to use for analyses like rms
     # [alpha_min, alpha_max, beta_min, beta_max]
     'analysis_limits'          :     [-3.8, 3.8, -3.8, 3.8],
 

### 2012-08-24 src/plot.py

@@ -326,17 +326,18 @@ def survey_coverage_histogram(ax, measurement_filename, xlim, ylim=None):
 
     print(measurement_catalog.k.size)
 
-    source_ID = np.unique(measurement_catalog.k)
+    k = measurement_catalog.k[measurement_catalog.is_in_analysis_region]
+    source_ID = np.unique(k)
     print(source_ID.size)
     nobs = np.zeros(source_ID.size)
 
     for i, sid in enumerate(source_ID):
-        found = np.where(measurement_catalog.k == sid)[0]
+        found = np.where(k == sid)[0]
         nobs[i] = found.size
 
     print(np.sum(nobs))
     print(np.mean(nobs))
-    assert np.sum(nobs).astype(int) == measurement_catalog.k.size
+    assert np.sum(nobs).astype(int) == k.size
     hist = ax.hist(nobs, color='k',
                         bins=np.arange(int(xlim[1]) + 5) - 0.5,
                         histtype='step')

### 2012-08-24 src/self_calibration.py

@@ -78,7 +78,9 @@ def self_calibration(obs_cat, sky_catalog, order, FoV,
 
         indx = [s != 0]
 
-        rms = analysis.rms_error(s[indx], sky_catalog.flux[indx], verbose)
+        rms = analysis.rms_error(s[indx], sky_catalog.flux[indx],
+                                    sky_catalog.is_in_analysis_region[indx],
+                                    verbose)
         bdness = analysis.badness(q, FoV, ff_samples, verbose)
         bdness_bestfitff = analysis.best_in_basis(q, FoV, ff_samples,
                                         best_fit_parameters, verbose)

### 2012-08-24 src/simulation.py

@@ -87,7 +87,8 @@ def run_sim(dic):
                                                     dic['density_of_stars'],
                                                     dic['m_min'],
                                                     dic['m_max'],
-                                                    dic['powerlaw_constants'])
+                                                    dic['powerlaw_constants'],
+                                                    dic['analysis_limits'])
         if dic['verbose']:
             print('...{0} sources generated!'.format(sky_catalog.size))
     if dic['data_dir']:

### 2012-08-24 src/survey.py

@@ -96,6 +96,7 @@ class CameraCatalog:
         self.mag = sky_catalog.mag
         self.alpha = sky_catalog.alpha
         self.beta = sky_catalog.beta
+        self.is_in_analysis_region = sky_catalog.is_in_analysis_region
         self.x, self.y = \
             transformations.sky2fp(sky_catalog.alpha, sky_catalog.beta,\
                                                         pointing, orientation)
@@ -126,6 +127,7 @@ class MeasuredCatalog:
         self.k = camera_catalog.k[inside_FoV]
         self.alpha = camera_catalog.alpha[inside_FoV]
         self.beta = camera_catalog.beta[inside_FoV]
+        self.is_in_analysis_region = camera_catalog.is_in_analysis_region[inside_FoV]
         self.x = camera_catalog.x[inside_FoV]
         self.y = camera_catalog.y[inside_FoV]
         flat = true_functions.flat_field(self.x, self.y, FoV)
@@ -156,6 +158,8 @@ class MeasuredCatalog:
         self.k = np.append(self.k, other.k)
         self.alpha = np.append(self.alpha, other.alpha)
         self.beta = np.append(self.beta, other.beta)
+        self.is_in_analysis_region = np.append(self.is_in_analysis_region,
+                                                other.is_in_analysis_region)
         self.x = np.append(self.x, other.x)
         self.y = np.append(self.y, other.y)
         self.counts = np.append(self.counts, other.counts)

### 2012-08-24 src/true_functions.py

@@ -78,7 +78,8 @@ def flat_field(x, y, FoV, par=flat_field_parameters()):
 
 
 class SourceCatalog:
-    def __init__(self, sky_limits, density_of_stars, m_min, m_max, A):
+    def __init__(self, sky_limits, density_of_stars,
+                                            m_min, m_max, A, analysis_limits):
         ''' This class generates the Source Catalog object
 
         Input
@@ -96,6 +97,8 @@ class SourceCatalog:
         A                   :   numpy array
             The parameters describing the magnitude distribution of the sources
             in the sky, according to: log10(dN/dm) = A + B * mag + C * mag ** 2
+        analysis_limits     :   list
+            Sky region for rms calculation.
 
         '''
 
@@ -109,6 +112,10 @@ class SourceCatalog:
                         high=sky_limits[3], size=self.size)
         self.mag = self.generate_magnitudes(m_min, m_max, A, area, self.size)
         self.flux = transformations.mag2flux(self.mag)
+        self.is_in_analysis_region = ((self.alpha > analysis_limits[0])
+                                    * (self.alpha < analysis_limits[1])
+                                    * (self.beta > analysis_limits[2])
+                                    * (self.beta < analysis_limits[3])).astype(bool)
 
     def power_law(self, A, m):
         ''' This function calculates the value of the source magnitude

### 2012-08-24 paper/self_calibration_I.tex
text changes, histogram fix, rady to rerun on everything

@@
 To do this we defined three quantities.
-The first is the root-mean-squared (RMS) error $S_\text{RMS}$ in the final fitted source fluxes $s_{k,\fit}$ compared to the true source fluxes $s_{k,\true}$ for the $K$ sources used in the self-calibration procedure:
+The first is the root-mean-squared (RMS) error $S_\rms$ in the final fitted source fluxes $s_{k,\fit}$ compared to the true source fluxes $s_{k,\true}$ for the $K$ sources used in the self-calibration procedure:
 \begin{equation}
-S_\text{RMS} = \sqrt{\dfrac{1}{K} \sum_k^K \left( \dfrac{s_{k,\fit} - s_{k,\true}}{s_{k,\true}} \right)^2}   \quad .
+S_\rms = \sqrt{\dfrac{1}{K} \sum_k^K \left( \dfrac{s_{k,\fit} - s_{k,\true}}{s_{k,\true}} \right)^2}   \quad .
 \end{equation}
-The other two metrics, called ``badnesses'', are defined as the RMS error between the final fitted instrument response and a reference instrument response sampled on a regular $300 \times 300$ grid across the focal plane.
+In detail, we compute this $S_\rms$ excluding sources near the survey boundary.
+The other two metrics, called ``badnesses'', are defined as the error between the final fitted instrument response and a reference instrument response sampled on a regular $300 \times 300$ grid across the focal plane.
 For the ``true badness'' $B_\true$, the \textit{fitted} instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ is compared to the \textit{true} instrument response $f_\true(\vec{x} | \vec{q_\true})$ at the $J$ sample points
@@
 
+\newcommand{\surveyfigcaption}{and the self-calibration results: (a) shows the survey footprint diagram (as shown in Figure \ref{fig:simple_surveys}), (b) shows the source coverage for this survey, with (c) showing a histogram of the number of source observations (excluding sources very near the survey boundary) and (d) shows the flux errors for each of the sources after the self-calibration procedure.
+In panels (b) and (d) the white gaps are \emph{not} gaps in the survey coverage but rather sky positions at which there do not happen to be observed stars}
+
 \begin{figure}
 \plotone{A_survey.pdf}
-\figcaption{A summary of Survey Strategy A and the resultant self-calibration results: (a) shows the survey footprint diagram (as shown in Figure \ref{fig:simple_surveys}), (b) shows the source coverage for this survey, with (c) showing a histogram of the number of source observations and (d) shows the flux errors for each of the sources after the self-calibration procedure.
+\figcaption{A summary of Survey Strategy A \surveyfigcaption.
 The imprint of the survey strategy can be clearly seen in the coverage map.
@@
 \plotone{B_survey.pdf}
-\figcaption{A summary of Survey Strategy B and the resultant self-calibration results: (a) shows the survey footprint diagram (as shown in Figure \ref{fig:simple_surveys}), (b) shows the source coverage for this survey, with (c) showing a histogram of the number of source observations and (d) shows the flux errors for each of the sources after the self-calibration procedure.
+\figcaption{A summary of Survey Strategy B \surveyfigcaption.
 The imprint of the survey strategy can be clearly seen in the coverage map.
@@
 \plotone{C_survey.pdf}
-\figcaption{A summary of Survey Strategy C and the resultant self-calibration results: (a) shows the survey footprint diagram (as shown in Figure \ref{fig:simple_surveys}), (b) shows the source coverage for this survey, with (c) showing a histogram of the number of source observations and (d) shows the flux errors for each of the sources after the self-calibration procedure.
+\figcaption{A summary of Survey Strategy C \surveyfigcaption.
 \label{fig:C_survey}}
@@
 \plotone{D_survey.pdf}
-\figcaption{A summary of Survey Strategy D and the resultant self-calibration results: (a) shows the survey footprint diagram (as shown in Figure \ref{fig:simple_surveys}), (b) shows the source coverage for this survey, with (c) showing a histogram of the number of source observations and (d) shows the flux errors for each of the sources after the self-calibration procedure.
-As can be seen from (c), this quasi-random survey offers a more uniform coverage of the field than the other survey investigated.
+\figcaption{A summary of Survey Strategy D \surveyfigcaption. As can be seen in panel (c), this quasi-random survey offers a more uniform coverage of the field than the other survey investigated.
 \label{fig:D_survey}}
@@
 \tablecaption{A summary of the quality of the final fit from the self-calibration procedure with the four simple survey strategies summarized in Table \ref{tab:simple_surveys}.
-The self-calibration procedure was run with only the brightest sources within the survey area; the source error $S_\text{RMS}$ corresponds to the measurements of these sources only and not all those within the survey.
+The self-calibration procedure was run with only the brightest sources within the survey area; the source error $S_\rms$ corresponds to the measurements of these sources only and not all those within the survey.
 *Did not converge.\label{tab:simple_results}}

### 2012-08-24 src/code_examples/paper_plots/paper_script.py

@@ -27,10 +27,10 @@ import analysis
 # Multiprocessing Flag:
 # False - then does not use multiprocessing
 # int - uses that many separate processes
-multi_proc = False
+multi_proc = 4
 
 # The four survey directories
-survey_files = ['D']#['A', 'B', 'C', 'D']
+survey_files = ['A', 'B', 'C', 'D']
 
 # Load the default parameters
 dic = eval(open('parameters.py').read())
@@ -61,48 +61,5 @@ else:
     results = []
     for params in parameter_dictionaries:
         results.append(simulation.run_sim(params))
-
-#  Since there are lots of souces at the edge of Survey D that are only
-#  measured a few times, the RMS 
-f = open('rms_results.txt', 'w')
-survey_area = [-4., 4., -4., 4.]
-header = 'Source RMS Error with {0} survey area:'.format(survey_area)
-f.write(header + '\n')
-print(header)
-
-for path in survey_files:
-	
-	# Load the catalogs from the simulation run
-	fitted_cat = pickle.load(open((path + '/fitted_catalog.p'), mode='rb'))
-	true_cat = pickle.load(open((path + '/source_catalog.p'), mode='rb'))
-	
-	if true_cat.k.max() > fitted_cat.k.max():
-		k_max = true_cat.k.max()
-	else:
-		k_max = fitted_cat.k.max()
-	
-	error = np.array([])
-	norm_error = np.array([])
-	
-	true_flux = np.array([])
-	fitted_flux = np.array([])
-	
-	inside = ((true_cat.alpha > survey_area[0])
-	                * (true_cat.alpha < survey_area[1])
-	                * (true_cat.beta > survey_area[2])
-	                * (true_cat.beta < survey_area[3]))
-
-	for indx in range(k_max):
-		indx_t = np.where(true_cat.k[inside] == indx)[0]
-		indx_f = np.where(fitted_cat.k == indx)[0]
-	
-		if (len(indx_t == 1) and len(indx_f == 1)):
-			true_flux = np.append(true_flux, true_cat.flux[inside][indx_t])
-			fitted_flux = np.append(fitted_flux, fitted_cat.flux[indx_f])
-	
-	rms = analysis.rms_error(fitted_flux, true_flux)
-	result = 'Survey {0}:\t\tRMS = {1:0.7f} %'.format(path, rms)
-	f.write(result + '\n')
-	print(result)
 
 print('Simulation Complete!')

### 2012-08-24 src/code_examples/paper_plots/parameters.py

@@ -70,5 +70,5 @@
     'stop_condition'	    :   1e-14,
     # Integer: The maximum number of iterations in the self-calibration
     # procedure and the best-in-basis fitting
-    'max_iterations'	    :    1
+    'max_iterations'	    :    10001
     }

### 2012-08-24 src/plot.py
remove debug print

@@ -335,8 +335,6 @@ def survey_coverage_histogram(ax, measurement_filename, xlim, ylim=None):
         found = np.where(k == sid)[0]
         nobs[i] = found.size
 
-    print(np.sum(nobs))
-    print(np.mean(nobs))
     assert np.sum(nobs).astype(int) == k.size
     hist = ax.hist(nobs, color='k',
                         bins=np.arange(int(xlim[1]) + 5) - 0.5,

### 2012-08-24 paper/self_calibration_I.tex
minor text

@@
 \plotone{D_survey.pdf}
-\figcaption{A summary of Survey Strategy D \surveyfigcaption. As can be seen in panel (c), this quasi-random survey offers a more uniform coverage of the field than the other survey investigated.
+\figcaption{A summary of Survey Strategy D \surveyfigcaption.
+As can be seen in panel (c), this quasi-random survey offers a more uniform coverage of the field than the other surveys; it also has much more uniformity than would a Poisson distribution of field centers.
 \label{fig:D_survey}}

### 2012-09-12 paper/self_calibration_I.tex
final simulation run

@@
 The performance of the self-calibration procedure is worse for the Survey B catalog and it fails completely with the Survey A catalog, where no solution was obtained.
-The number of iterations of the self-calibration procedure required to converge to the final solution also decreases as the quality of the final fit increases: Survey C required 163 iterations, Survey D 197, Survey B 3,746 and the procedure was iterated 10,000 times with Survey A without reaching a solution.
+The number of iterations of the self-calibration procedure required to converge to the final solution also decreases as the quality of the final fit increases: Survey D required 65 iterations, Survey C 333, Survey B 2,477 and the procedure was iterated 10,000 times with Survey A without reaching a solution.
 
@@
 \begin{figure}
-\plotone{./camera_image.pdf}
+\plotone{./camera_image}
 \figcaption{A single exposure of the synthetic sky.
@@
 \begin{figure}
-\plotone{A_8192_ff.pdf}
+\plotone{A_10000_ff}
 \figcaption{Survey Strategy A: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey A catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
-The self-calibration procedure did not converged to the final solution and the results presented here correspond to the instrument response found after 8,192 iterations.
+The self-calibration procedure did not converged to the final solution and the results presented here correspond to the instrument response found after 10,000 iterations.
 The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial).
@@
 \begin{figure}
-\plotone{B_3746_ff.pdf}
+\plotone{B_2477_ff}
 \figcaption{Survey Strategy B: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey B catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
-The self-calibration procedure converged to the final solution after 3,746 iterations.
+The self-calibration procedure converged to the final solution after 2,447 iterations.
 The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
@@
 \begin{figure}
-\plotone{C_163_ff.pdf}
+\plotone{C_333_ff}
 \figcaption{Survey Strategy C: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey C catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
-The self-calibration procedure converged to the final solution after 163 iterations.
+The self-calibration procedure converged to the final solution after 333 iterations.
 The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
@@
 \begin{figure}
-\plotone{D_197_ff.pdf}
+\plotone{D_65_ff}
 \figcaption{Survey Strategy D: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey D catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (b) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
-The self-calibration procedure converged to the final solution after 197 iterations.
+The self-calibration procedure converged to the final solution after 65 iterations.
 The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
@@
 \begin{figure}
-\plotone{A_survey.pdf}
+\plotone{A_survey}
 \figcaption{A summary of Survey Strategy A \surveyfigcaption.
@@
 \begin{figure}
-\plotone{B_survey.pdf}
+\plotone{B_survey}
 \figcaption{A summary of Survey Strategy B \surveyfigcaption.
@@
 \begin{figure}
-\plotone{C_survey.pdf}
+\plotone{C_survey}
 \figcaption{A summary of Survey Strategy C \surveyfigcaption.
@@
 \begin{figure}
-\plotone{D_survey.pdf}
+\plotone{D_survey}
 \figcaption{A summary of Survey Strategy D \surveyfigcaption.
@@
 \startdata
-A*  & - & 1.092 & 1.490 & 1.450 \\
-B  & 3746 & 0.129 & 0.163 & 0.082 \\
-C  & 163 & 0.113 & 0.106 & 0.013 \\
-D  & 197 & 0.147 & 0.108 & 0.005 \\
+A*  & - & 1.115 & 1.517 & 1.485 \\
+B  & 2477 & 0.107 & 0.114 & 0.024 \\
+C  & 333 & 0.106 & 0.106 & 0.010 \\
+D  & 65 & 0.103 & 0.108 & 0.005 \\
 \enddata

### 2012-09-12 src/plot.py

@@ -323,12 +323,8 @@ def survey_coverage_histogram(ax, measurement_filename, xlim, ylim=None):
     '''
 
     measurement_catalog = pickle.load(open(measurement_filename))
-
-    print(measurement_catalog.k.size)
-
     k = measurement_catalog.k[measurement_catalog.is_in_analysis_region]
     source_ID = np.unique(k)
-    print(source_ID.size)
     nobs = np.zeros(source_ID.size)
 
     for i, sid in enumerate(source_ID):

### 2012-09-13 paper/self_calibration_I.tex
last paragraph done

@@
 
-HOGG write van Dokkum paragraph
-Ultimately a trade-off would need to be performed between the different calibration concerns and the other goals of the survey, such as sky coverage, variability monitoring and so-on.
+Ultimately, survey design requires trade-offs between calibration
+requirements and other requirements related to cadence, telescope
+overheads, and observing constraints (like alt, az, airmass, and
+field-rotation constraints).  One requirement that is often
+over-emphasized, however, is conceptual or apparent ``uniformity'';
+these these different survey strategies have different uniformities of
+coverage, each with possibly non-obvious consequences.  Many
+astronomers will see survey strategy A as being ``more uniform'' in
+its coverage (especially relative to survey D).  This is not true, if
+uniformity is defined by the variance or skewness or any other simple
+statistic in exposure times; strategy D is extremely uniform (more
+uniform than Poisson, for example).  In any survey, past and future,
+variations in exposure time have been valuable for checking systematic
+and random errors, and don't---in themselves---make it difficult to
+obtain a uniform survey in properties like brightness (since samples
+can be cut on any well-measured property).  In general, in the
+presence of real astronomical variations in distributions of
+luminosity, distance, metallicity, and (more importantly) extinction
+by dust, there is no way to make a survey uniformly sensitive to the
+objects of interest.  As a community we should be prepared to adjust
+our analysis for the non-uniformity of the survey rather than adjust
+(cut) our data to match the uniformity of unrealistically simplified
+analyses.  This is already standard practice in the precise
+cosmological measurement experiments, and will be required practice
+for the next generation of massively-multiple-epoch imaging surveys.
 

### 2012-09-13 paper/self_calibration_I.tex
typo

@@
 over-emphasized, however, is conceptual or apparent ``uniformity'';
-these these different survey strategies have different uniformities of
+these different survey strategies have different uniformities of
 coverage, each with possibly non-obvious consequences.  Many

### 2012-09-15 paper/self_calibration_I.tex
it's a strategy in this context
change number of passes over field to 9

@@
 \begin{eqnarray*}
-f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7\ldots 260})  & = &  \sum_{a=0}^6 \sum_{b=0}^a \left(1 + a + b\right) ^ {-0.75} \left[ q_{\true, 7+4b} \cos (k_x \, x_i) + q_{\true, 8+4b} \sin (k_x \, x_i) \right] \\
+f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7\ldots 260})  & = &  \sum_{a=0}^6 \sum_{b=0}^a \left(1 + a + b\right) ^ {-0.5} \left[ q_{\true, 7+4b} \cos (k_x \, x_i) + q_{\true, 8+4b} \sin (k_x \, x_i) \right] \\
 & & \qquad \qquad \times \left[ q_{\true, 9+4b} \cos (k_y \, y_i) + q_{\true, 10+4b} \sin (k_y \, y_i) \right] \quad ,
@@
 with the physical focal-plane dimensions $X$ and $Y$.
-The $(1 + a +d)^{-0.75}$ factor reduces the magnitude of the higher order terms.
+The $(1 + a +d)^{-0.5}$ factor reduces the magnitude of the higher order terms.
 In total, the instrument response model is parameterized with 260 parameters; an example can be seen in Figure \ref{fig:single_image} (right).
@@
 Strategy A is the simplest strategy; the target field is regularly tiled with small overlaps between adjacent pointings ($\sim 5$~percent and $\sim 8$~percent in the camera pointing directions $\alpha$ and $\beta$ respectively).
-The pointings in the 12 passes over the same field are exactly aligned.
+The pointings in the 9 passes over the same field are exactly aligned.
 Survey B is the same as A, but with each pass over the target field the orientation of the telescope is rotated by $30^\circ$.
@@
 In the third pass over the field this change is reversed and the field is measured on a $11 \times 13$ grid.
-These three passes are then repeated four times.
+These three passes are then repeated three times.
 
@@
 \startdata
-A  & Uniform Grid ($12\times12$) & $0^\circ$ & 1728 \\
-B      & Uniform Grid ($12\times12$)     & Each Pass: $\theta +30^\circ$& 1728 \\
-C   & Pass 1:  Uniform Grid ($12\times12$)  &  & 1720\\
+A  & Uniform Grid ($12\times12$) & $0^\circ$ & 1296 \\
+B      & Uniform Grid ($12\times12$)     & Each Pass: $\theta +30^\circ$& 1296 \\
+C   & Pass 1:  Uniform Grid ($12\times12$)  &  & 1290\\
  & Pass 2:  Uniform Grid ($13§\times11$) & $0^\circ$&\\
   & Pass 3:  Uniform Grid ($11§\times13$) & &\\
-D   & Quasi-Random  & Random & 1728\\
+D   & Quasi-Random  & Random & 1296\\
 \enddata
@@
 \figcaption{Focal-plane footprints projected onto the synthetic sky according to the four simple survey strategies described in Section \ref{sec:simple_surveys} and summarized in Table \ref{tab:simple_surveys}.
-Surveys A, B and D have 1728 pointings and survey C has 1720 pointings.\label{fig:simple_surveys}}
+Surveys A, B and D have 1296 pointings and survey C has 1290 pointings.\label{fig:simple_surveys}}
 \end{figure}

### 2012-09-15 src/code_examples/paper_plots/generate_surveys.py

@@ -96,7 +96,7 @@ if __name__ == "__main__":
 
     dic = eval(open('parameters.py').read())
     sky_limits = [-4., 4., -4., 4.]
-    number_passes = 12
+    number_passes = 9
     FoV = dic['FoV']
 
     xA = generate_uniform_survey(sky_limits, FoV, number_passes)

### 2012-09-15 src/true_functions.py

@@ -69,10 +69,10 @@ def flat_field(x, y, FoV, par=flat_field_parameters()):
             cky = np.cos(ky * y)
             sky = np.sin(ky * y)
             # XX Magic Numbers: ** 0.75 to kill higher order terms
-            ff += par[k + 0] * ckx * cky / (1 + nx + ny) ** 0.75
-            ff += par[k + 1] * ckx * sky / (1 + nx + ny) ** 0.75
-            ff += par[k + 2] * skx * cky / (1 + nx + ny) ** 0.75
-            ff += par[k + 3] * skx * sky / (1 + nx + ny) ** 0.75
+            ff += par[k + 0] * ckx * cky / (1 + nx + ny) ** 0.5
+            ff += par[k + 1] * ckx * sky / (1 + nx + ny) ** 0.5
+            ff += par[k + 2] * skx * cky / (1 + nx + ny) ** 0.5
+            ff += par[k + 3] * skx * sky / (1 + nx + ny) ** 0.5
             k += 4
     return ff
 

### 2012-09-20 paper/referee_response.txt
update RMS calculation so that now only sources within predefined survey used. Also add survey generation files
change density of sources to 300

@@
 
-This is an excellent suggestion. We have added Figures 7 to 11, which show the residuals in the final survey.
+This is an excellent suggestion.
+We have added Figures 7 to 11, which show the residuals in the final survey.
+We comment on these figures in Section 5.2, and we feel that this led to a substantial improvement of the paper.
 
@@
 	- We corrected a few minor bugs in the code, none of which made made significant changes to the contents of the paper.
-	- We have modified the form of the true instrument response (last equation in 2.2.3) to now include a (1+a+b)**-0.75 factor.
+	- We have modified the form of the true instrument response (last equation in 2.2.3) to now include a (1+a+b)**-0.5 factor.
 This reduces the magnitude of the high frequency sine and cosine contributions.

### 2012-09-20 src/code_examples/paper_plots/parameters.py

@@ -22,7 +22,7 @@
     # Float: the maximum number of sources (all magnitude) per unit area
     # on the sky. If more than this number are produced by the magnitude
     # distribution (see below), only the brightest are selected.
-    'density_of_stars'    :     200.,
+    'density_of_stars'    :     300.,
     # Float array: The parameters describing the magnitude distribution
     # of the sources in the sky, according to
     # log10(dN/dm) = A + B * mag + C * mag ** 2

### 2012-09-20 src/plot.py

@@ -612,7 +612,7 @@ def survey(source_filename, measurement_filename,
     colors, colorbar_nobs = survey_coverage(ax2, source_filename,
                                                 measurement_filename,
                                                 nobs_plot_lim=nobs_plot_lim,
-                                                zero_level='0.9', ms=10)
+                                                zero_level='0.9', ms=6)
     survey_coverage_colorbar(ax_cb2, colors, colorbar_nobs)
     if verbose:
         print('...done...')
@@ -629,7 +629,7 @@ def survey(source_filename, measurement_filename,
                     .format(source_filename, fitted_filename))
     colors, errors = survey_source_error(ax4, source_filename, fitted_filename,
                             lower_limit=error_lim[0], upper_limit=error_lim[1],
-                            lower_color=0.9, upper_color=0., ms=10)
+                            lower_color=0.9, upper_color=0., ms=6)
     survey_source_error_colorbar(ax_cb4, colors, errors)
     if verbose:
         print('...done...')

### 2012-09-20 paper/self_calibration_I.tex
increase strength of sine and cosines so that see difference better between B and C

@@
 \begin{eqnarray*}
-f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7\ldots 260})  & = &  \sum_{a=0}^6 \sum_{b=0}^a \left(1 + a + b\right) ^ {-0.5} \left[ q_{\true, 7+4b} \cos (k_x \, x_i) + q_{\true, 8+4b} \sin (k_x \, x_i) \right] \\
+f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7\ldots 260})  & = &  \sum_{a=0}^6 \sum_{b=0}^a \left(1 + a + b\right) ^ {-0.25} \left[ q_{\true, 7+4b} \cos (k_x \, x_i) + q_{\true, 8+4b} \sin (k_x \, x_i) \right] \\
 & & \qquad \qquad \times \left[ q_{\true, 9+4b} \cos (k_y \, y_i) + q_{\true, 10+4b} \sin (k_y \, y_i) \right] \quad ,
@@
 with the physical focal-plane dimensions $X$ and $Y$.
-The $(1 + a +d)^{-0.5}$ factor reduces the magnitude of the higher order terms.
+The $(1 + a +d)^{-0.25}$ factor reduces the magnitude of the higher order terms.
 In total, the instrument response model is parameterized with 260 parameters; an example can be seen in Figure \ref{fig:single_image} (right).
@@
 astronomers will see survey strategy A as being ``more uniform'' in
-its coverage (especially relative to survey D).  This is not true, if
+its coverage (especially relative to strategy D).  This is not true, if
 uniformity is defined by the variance or skewness or any other simple

### 2012-09-20 src/plot.py

@@ -828,9 +828,9 @@ def flat_fields(filename, FoV, ff_samples, best_fit_params, output_path=False,
     true_residual = 100 * (dic['fitted_ff'] - true_ff) / true_ff
     best_residual = 100 * (dic['fitted_ff'] - best_ff) / best_ff
     fp = (-FoV[0] / 2, FoV[0] / 2, -FoV[1] / 2, FoV[1] / 2)
-    a = ax4.imshow(true_residual, extent=fp, vmin=-0.5, vmax=0.5,
+    a = ax4.imshow(true_residual, extent=fp, vmin=-2., vmax=2.,
                                                     cmap='gray', aspect='auto')
-    ax3.imshow(best_residual, extent=fp, vmin=-0.5, vmax=0.5,
+    ax3.imshow(best_residual, extent=fp, vmin=-2., vmax=2.,
                                                     cmap='gray', aspect='auto')
 
     cbar = fig.colorbar(a, ax_cb, orientation='vertical')

### 2012-09-20 src/true_functions.py

@@ -69,10 +69,10 @@ def flat_field(x, y, FoV, par=flat_field_parameters()):
             cky = np.cos(ky * y)
             sky = np.sin(ky * y)
             # XX Magic Numbers: ** 0.75 to kill higher order terms
-            ff += par[k + 0] * ckx * cky / (1 + nx + ny) ** 0.5
-            ff += par[k + 1] * ckx * sky / (1 + nx + ny) ** 0.5
-            ff += par[k + 2] * skx * cky / (1 + nx + ny) ** 0.5
-            ff += par[k + 3] * skx * sky / (1 + nx + ny) ** 0.5
+            ff += par[k + 0] * ckx * cky / (1 + nx + ny) ** 0.25
+            ff += par[k + 1] * ckx * sky / (1 + nx + ny) ** 0.25
+            ff += par[k + 2] * skx * cky / (1 + nx + ny) ** 0.25
+            ff += par[k + 3] * skx * sky / (1 + nx + ny) ** 0.25
             k += 4
     return ff
 

### 2012-09-25 paper/self_calibration_I.tex
add final final final results to paper

@@
 The pointings in the 9 passes over the same field are exactly aligned.
-Survey B is the same as A, but with each pass over the target field the orientation of the telescope is rotated by $30^\circ$.
+Survey B is the same as A, but with each pass over the target field the orientation of the telescope is rotated by $40^\circ$.
 
@@
 The performance of the self-calibration procedure is worse for the Survey B catalog and it fails completely with the Survey A catalog, where no solution was obtained.
-The number of iterations of the self-calibration procedure required to converge to the final solution also decreases as the quality of the final fit increases: Survey D required 65 iterations, Survey C 333, Survey B 2,477 and the procedure was iterated 10,000 times with Survey A without reaching a solution.
+The number of iterations of the self-calibration procedure required to converge to the final solution also decreases as the quality of the final fit increases: Survey D required 11 iterations, Survey C 24, Survey B 1,114 and the procedure was iterated 5,000 times with Survey A without reaching a solution.
 
@@
 
-Ultimately, survey design requires trade-offs between calibration
-requirements and other requirements related to cadence, telescope
-overheads, and observing constraints (like alt, az, airmass, and
-field-rotation constraints).  One requirement that is often
-over-emphasized, however, is conceptual or apparent ``uniformity'';
-these different survey strategies have different uniformities of
-coverage, each with possibly non-obvious consequences.  Many
-astronomers will see survey strategy A as being ``more uniform'' in
-its coverage (especially relative to strategy D).  This is not true, if
-uniformity is defined by the variance or skewness or any other simple
-statistic in exposure times; strategy D is extremely uniform (more
-uniform than Poisson, for example).  In any survey, past and future,
-variations in exposure time have been valuable for checking systematic
-and random errors, and don't---in themselves---make it difficult to
-obtain a uniform survey in properties like brightness (since samples
-can be cut on any well-measured property).  In general, in the
-presence of real astronomical variations in distributions of
-luminosity, distance, metallicity, and (more importantly) extinction
-by dust, there is no way to make a survey uniformly sensitive to the
-objects of interest.  As a community we should be prepared to adjust
-our analysis for the non-uniformity of the survey rather than adjust
-(cut) our data to match the uniformity of unrealistically simplified
-analyses.  This is already standard practice in the precise
-cosmological measurement experiments, and will be required practice
-for the next generation of massively-multiple-epoch imaging surveys.
+Ultimately, survey design requires trade-offs between calibration requirements and other requirements related to cadence, telescope overheads, and observing constraints (like alt, az, airmass, and field-rotation constraints).
+One requirement that is often over-emphasized, however, is conceptual or apparent ``uniformity''; these different survey strategies have different uniformities of coverage, each with possibly non-obvious consequences.
+Many astronomers will see survey strategy A as being ``more uniform'' in its coverage (especially relative to strategy D).
+This is not true, if uniformity is defined by the variance or skewness or any other simple statistic in exposure times; strategy D is extremely uniform (more
+uniform than Poisson, for example).
+In any survey, past and future, variations in exposure time have been valuable for checking systematic and random errors, and don't---in themselves---make it difficult to obtain a uniform survey in properties like brightness (since samples can be cut on any well-measured property).
+In general, in the presence of real astronomical variations in distributions of luminosity, distance, metallicity, and (more importantly) extinction by dust, there is no way to make a survey uniformly sensitive to the objects of interest.
+As a community we should be prepared to adjust our analysis for the non-uniformity of the survey rather than adjust (cut) our data to match the uniformity of unrealistically simplified analyses.
+This is already standard practice in the precise cosmological measurement experiments, and will be required practice for the next generation of massively-multiple-epoch imaging surveys.
 
@@
 A  & Uniform Grid ($12\times12$) & $0^\circ$ & 1296 \\
-B      & Uniform Grid ($12\times12$)     & Each Pass: $\theta +30^\circ$& 1296 \\
+B      & Uniform Grid ($12\times12$)     & Each Pass: $\theta +40^\circ$& 1296 \\
 C   & Pass 1:  Uniform Grid ($12\times12$)  &  & 1290\\
@@
 \begin{figure}
-\plotone{A_10000_ff}
+\plotone{A_5000_ff}
 \figcaption{Survey Strategy A: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey A catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
-The self-calibration procedure did not converged to the final solution and the results presented here correspond to the instrument response found after 10,000 iterations.
+The self-calibration procedure did not converged to the final solution and the results presented here correspond to the instrument response found after 5,000 iterations.
 The best-in-basis instrument response is the best fit to the true instrument response possible with the basis used to model the instrument response in the self-calibration procedure (in this case an eighth order polynomial).
@@
 \begin{figure}
-\plotone{B_2477_ff}
+\plotone{B_1114_ff}
 \figcaption{Survey Strategy B: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey B catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
-The self-calibration procedure converged to the final solution after 2,447 iterations.
+The self-calibration procedure converged to the final solution after 1,114 iterations.
 The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
@@
 \begin{figure}
-\plotone{C_333_ff}
+\plotone{C_24_ff}
 \figcaption{Survey Strategy C: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey C catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (c) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
-The self-calibration procedure converged to the final solution after 333 iterations.
+The self-calibration procedure converged to the final solution after 24 iterations.
 The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
@@
 \begin{figure}
-\plotone{D_65_ff}
+\plotone{D_11_ff}
 \figcaption{Survey Strategy D: A comparison of the fitted instrument response model $f_\fit(\vec{x_j} | \vec{q_\fit})$ (black) obtained from self-calibrating the Survey D catalog compared to (a) the true $f_\true(\vec{x} | \vec{q_\true})$ and (b) the best-in-basis $f_\basis(\vec{x_j} | \vec{q_\basis})$ instrument response models (gray).
-The self-calibration procedure converged to the final solution after 65 iterations.
+The self-calibration procedure converged to the final solution after 11 iterations.
 The plots (b) and (d) show the residuals between the two instrument response models plotted in (a) and (c) respectively.
@@
 \startdata
-A*  & - & 1.115 & 1.517 & 1.485 \\
-B  & 2477 & 0.107 & 0.114 & 0.024 \\
-C  & 333 & 0.106 & 0.106 & 0.010 \\
-D  & 65 & 0.103 & 0.108 & 0.005 \\
+A*  & - & 1.172 & 1.680 & 1.527 \\
+B  & 1140 & 0.201 & 0.507 & 0.207 \\
+C  & 24 & 0.164 & 0.361 & 0.019 \\
+D  & 11 & 0.136 & 0.364 & 0.010 \\
 \enddata

### 2012-09-27 paper/self_calibration_I.tex
fix minor bug in generation of survey D

@@
 These survey strategies are summarized in Table \ref{tab:simple_surveys} and are shown in Figure \ref{fig:simple_surveys}.
-Strategy A is the simplest strategy; the target field is regularly tiled with small overlaps between adjacent pointings ($\sim 5$~percent and $\sim 8$~percent in the camera pointing directions $\alpha$ and $\beta$ respectively).
+Strategy A is the simplest strategy; the target field is regularly tiled with small overlaps between adjacent pointings ($\sim 12$~percent in both of the camera pointing directions).
 The pointings in the 9 passes over the same field are exactly aligned.
@@
 C  & 24 & 0.164 & 0.361 & 0.019 \\
-D  & 11 & 0.136 & 0.364 & 0.010 \\
+D  & 11 & 0.130 & 0.321 & 0.012 \\
 \enddata

### 2012-09-27 paper/self_calibration_I.tex
remove large-scale from paper title

@@
 
-\shorttitle{Designing Large-Scale Imaging Surveys for Retrospective Relative Photometric Calibration}
+\shorttitle{Designing Imaging Surveys for Retrospective Relative Photometric Calibration}
 \shortauthors{Holmes et al.}
@@
 
-\title{Designing Large-Scale Imaging Surveys for a Retrospective Relative Photometric Calibration}
+\title{Designing Imaging Surveys for a Retrospective Relative Photometric Calibration}
 

### 2012-09-27 paper/self_calibration_I.tex
edit format of tex file
change text for minor bug in generation of survey D

@@
 \begin{abstract}
-In this paper, we investigate the impact of survey strategy on the performance of self-calibration when the goal is to produce accurate photometric catalogs
-from wide-field imaging surveys.
+In this paper, we investigate the impact of survey strategy on the performance of self-calibration when the goal is to produce accurate photometric catalogs from wide-field imaging surveys.
 This self-calibration technique utilizes multiple measurements of sources at different focal-plane positions to constrain instruments' large-scale response (flat-field) from survey science data alone.
-We create an artificial sky of sources
-and synthetically observe it under four basic survey strategies,
-creating an end-to-end simulation of an imaging survey for each. These
-catalog-level simulations include realistic measurement uncertainties
-and a complex focal-plane dependence of the instrument response. In the
-self-calibration step, we simultaneously fit for all the star fluxes
-and the parameters of a position-dependent flat-field.  For realism,
-we deliberately fit with a wrong noise model and a flat-field
-functional basis that does not include the model that generated the
-synthetic data.  We demonstrate that with a favorable survey strategy,
-a complex instrument response can be precisely
-self-calibrated. We show that returning the same sources to very
-different focal-plane positions is the key property of any survey
-strategy designed for accurate retrospective calibration of this type. The results
-of this work suggest the following advice for those considering the
-design of large-scale imaging surveys: Do not use a regular, repeated tiling of
-the sky; instead return the same sources to very different focal-plane
-positions.
+We create an artificial sky of sources and synthetically observe it under four basic survey strategies, creating an end-to-end simulation of an imaging survey for each.
+These catalog-level simulations include realistic measurement uncertainties
+and a complex focal-plane dependence of the instrument response.
+In the self-calibration step, we simultaneously fit for all the star fluxes and the parameters of a position-dependent flat-field.
+For realism, we deliberately fit with a wrong noise model and a flat-field functional basis that does not include the model that generated the synthetic data.
+We demonstrate that with a favorable survey strategy, a complex instrument response can be precisely self-calibrated.
+We show that returning the same sources to very
+different focal-plane positions is the key property of any survey strategy designed for accurate retrospective calibration of this type.
+The results of this work suggest the following advice for those considering the design of large-scale imaging surveys: Do not use a regular, repeated tiling of the sky; instead return the same sources to very different focal-plane positions.
 \end{abstract}
@@
 
-The pointing positions in Survey D are quasi-random: the pointings are the same as with Survey A, but each has a random offset within [-0.35,0.35)~deg applied in both the $\alpha$ and $\beta$ directions.
-By fixing the pointings within these $0.7~\text{deg} \times 0.7~\text{deg}$ boxes, we ensure that the quasi random strategy has a uniform coverage of the field; that is, the fluctuations in coverage are less than they would be in a Poisson (totally random) process.
+The pointing positions in Survey D are quasi-random: the pointings are the same as with Survey A, but each has a random offset within [-0.375,0.375]~deg applied in both the $\alpha$ and $\beta$ directions.
+By fixing the pointings within these $0.75~\text{deg} \times 0.75~\text{deg}$ boxes, we ensure that the quasi random strategy has a uniform coverage of the field; that is, the fluctuations in coverage are less than they would be in a Poisson (totally random) process.
 The orientations of the pointings in Survey D are completely random.

### 2013-02-03 proposal/pixel_level.tex
rearrangement

--- 2012-08-20 whitepaper/pixel_level.tex
@@
 
+\paragraph{tests with real data}
+We have identified the \project{HAWK-I} camera on \project{Paranal} as
+a possible analog.  It appears to have flat-field morphology similar
+to that expected for the \project{Euclid} infrared camera, and it is
+the same detector technology.  The next steps are
+\begin{itemize}
+\item Find and download a range of science data taken in one chosen
+  filter with a range of exposure times.
+\item Choose a small patch of the detector where there is known to be
+  good or interesting flat-field structure.
+\item Make source and sky patches from those data, and do some
+  analysis.
+\end{itemize}
+
 \begin{thebibliography}{70}

### 2013-02-03 paper/self_calibration_I.tex

@@
 With the Survey C and D catalogs, the instrument response is accurately recovered.
-As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top), the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 0.5$~percent level.
+As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top), the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$.
 That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the model basis used.
-The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the best-in-basis badness $B_\basis$ compared to the true badness $B_\true$.
+The residuals in this comparison are low; this can also be clearly seen in the significant improvement in the best-in-basis badness $B_\basis$ compared to the true badness $B_\true$.
 The results from these two survey catalogs are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the model used to represent the instrument response.
